### 2016-01-02 pdocs/thesis/abstract.tex
pdocs started

\section*{Abstract}

\textbf{\myname}\\
\textbf{\mytitle}\\
Lappeenranta \myyear\\
\pageref{LastPage} pages\\
Acta Universitatis Lappeenrantaensis XXXXXX\\
Diss. Lappeenranta University of Technology\\
ISBN \myisbn, ISBN \mypdfisbn (PDF), ISSN 1456-4491\\

In this work suitable methods for handling of plasticity in physics engine are presented.
Handling of plasticity basically means that objects involved may absorb energy and possibly break.

Methods are verified by extending Bullet Physics Library.

Keywords: Physics engine, simulation, dynamics, plasticity, deformation.

### 2016-01-02 pdocs/thesis/introduction.tex
pdocs started

\section{Introduction}

\subsection{Physics Engines}

\cite{Obrien:1999:GMA}.

\subsection{Motivation and objectives of the study}

\subsection{Scientific contribution}

\cleardoublepage

### 2016-01-02 pdocs/thesis/main.tex
pdocs started

% Simo.Nikula@gmail.com
% based on latex template at http://www.lut.fi/web/en/library/dissertations1 
\def\mytitle{Handling of plasticity in physics engine}
\def\myisbn{978-952-265-XXX-X}
\def\mypdfisbn{978-952-265-XXX-X}
\def\myyear{2017}
\def\mymonth{May}
\def\myname{Simo Nikula}
\documentclass[a4paper,twoside,12pt,notitlepage,openright]{article}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{lastpage}   % page count
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[english]{babel}
\usepackage[intoc]{nomencl} % including table of content
\usepackage{ifthen}
\usepackage[hmargin=3.0cm,vmargin=3.6cm]{geometry} % setting marginals
\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
\usepackage{times}  % to change font to times
\usepackage{setspace} % for linespacing

\newcommand{\nomunit}[1]{\renewcommand{\nomentryend}{\hspace*{\fill}#1}} % Inserts units on the right at symbol list
\renewcommand{\nomgroup}[1]{%
 \ifthenelse{\equal{#1}{C}}{\item[\textbf{Latin alphabet}]\item}{%
 \ifthenelse{\equal{#1}{G}}{\item\item[\textbf{Greek alphabet}]\item}{}}{%
 \ifthenelse{\equal{#1}{L}}{\item\item[\textbf{Subscripts}]\item}{}}{%
 \ifthenelse{\equal{#1}{H}}{\item\item[\textbf{Superscripts}]\item}{}}{%
 \ifthenelse{\equal{#1}{W}}{\item\item[\textbf{Abbreviations}]\item}{}}{}}

\singlespacing

\pagestyle{fancy}
\fancyhf{}% clearing the header and footer
\fancyhead[LE,RO]{\bfseries\thepage}    % page number to header
\fancyhead[LO]{\nouppercase{\bfseries\rightmark}}     %
\fancyhead[RE]{\nouppercase{\bfseries\leftmark}}%

\renewcommand\sectionmark[1]
 {\markboth{\thesection\ #1}{}}         % section name to header
\renewcommand\subsectionmark[1]
 {\markright{\thesubsection\ #1}}       % subsection name to header
\renewcommand{\headrulewidth}{0.5pt}    % ruler thickness between head and body
\renewcommand{\footrulewidth}{0pt}      % no ruler between body and footer

\setlength{\nomitemsep}{-\parsep}   % removing default extra skip between entries at nomenclature

\numberwithin{equation}{section}    % equation numbers with section numbers
\numberwithin{table}{section}       % table numbers with section numbers
\numberwithin{figure}{section}      % figure numbers with section numbers

% makeindex command needs to run at command prompt to create nomenclature list file
\makenomenclature % makeindex main_v2.nlo -s nomencl.ist -o main_v2.nls
%\bibpunct{(}{)}{;}{a}{,}{,}%

\begin{document}

\pagestyle{empty}
\include{nimio}
\include{abstract}
\addtocontents{toc}{\contentsline {section}{Abstract}{}}

\include{acknowledgments}
\addtocontents{toc}{\contentsline {section}{Acknowledgments}{}}%
\input{preface}
\cleardoublepage%
\thispagestyle{empty}%
\addtocontents{toc}{\contentsline{section}{Contents}{}}%

\tableofcontents%
\thispagestyle{empty}%
\cleardoublepage%
\pagestyle{fancy}%

\cleardoublepage%
\printnomenclature[2.0cm]%

\include{introduction}

%\def\refname{what ever}% to change References title
\addcontentsline{toc}{section}{References}%

\fancyhead[LO]{\nouppercase{\bfseries\firstleftxmark}}%
\fancyhead[RE]{\nouppercase{\bfseries\lastrightxmark}}%
\extramarks{}{References}   % this will remove the "References" to be appearing at the header at first page of bibliography
\extramarks{References}{References}
\bibliography{ref}%
\bibliographystyle{LUTapa}%

%\cleardoublepage %

\end{document}

### 2016-01-02 pdocs/thesis/nimio.tex
pdocs started

\thispagestyle{empty} \setlength{\parindent}{0pt}
\begin{figure}
\includegraphics[width=65mm]{./figs/Merkki_Logo_CMYK}\\
\end{figure}
~\\

\vspace{60mm}

{\sffamily\large Simo Nikula\\
\\
\MakeUppercase{\Large \mytitle
}}\\

%\vspace{4mm}
{\sffamily\footnotesize Thesis for the degree of Doctor of Science (Technology) to be presented with\\
due permission for public examination and criticism in the Auditorium ...\\
at Lappeenranta University of Technology, Lappeenranta, Finland on the\\
..th of \mymonth, \myyear, at noon.}\\

\vspace{\stretch{1}}
{\sffamily\normalsize Acta Universitatis\\
Lappeenrantaensis XXX}

\pagebreak

\thispagestyle{empty}

\begin{tabular}{l p{11.0cm}}Supervisor& Professor Timo Björk\\
&Laboratory of Steel Structures\\
&School of Energy Systems\\
&Lappeenranta University of Technology\\
&Finland\\
\\
Reviewers& Title Name\\
&Department of \\
&University of \\
&Country\\
\\
& Title Name\\
&Department of \\
&University\\
&Country\\

\\
Opponent& PhD Name\\
&Institution\\
&Country\\
\end {tabular}

\vspace{\stretch{1}}

\begin{center}
ISBN \myisbn\\
ISBN \mypdfisbn (PDF)\\
ISSN 1456-4491\\
Lappeenrannan teknillinen yliopisto\\
Yliopistopaino \myyear

 \end{center}
\pagebreak

### 2016-01-03 pdocs/thesis/discussion.tex
pdocs

\section{Discussion}

\subsection{Future work}
\label{subsec:futureWork}

\subsubsection{Large displacements}

\subsubsection{Material nonlinearity}

\cleardoublepage

### 2016-01-03 pdocs/thesis/introduction.tex
pdocs

@@
 
-\subsection{Physics Engines}
+\subsection{Physics engines}
 
-\cite{Obrien:1999:GMA}.
+Physics engines is active research area on many research and industry areas.
+Especially gaming and movie industry are powerful drivers.
+Robotics and solid mechanics are typical research areas that use physics engines.
+
+\cite{erleben.thesis} provides extensive introduction to physics engines. 
+He uses term Physics-based Animation and does detailed analysis on many areas like
+\begin{itemize}
+\item preferring fast solutions rather than exact solutions
+\item various methods for solving differential equations
+\item processing of constraints
+\end{itemize}
+
+\subsection{Structural analysis}
+
+Physics engines are not commonly used in structural analysis but using structural analysis in physics engines is not rare.
+
+Structural analysis for advanced structures is nowadays typically done using finite element method and 
+analysis time is typically hours. Finite element method is already used in physics engines and usage will come 
+more widespread in future as more computing capacity and memory will be available.
+
+E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
 
@@
 
+Target of this work is to find out suitable methods for handling of plasticity in bulletphysics, \cite{bulletphysics}. 
+Handling of plasticity basically means that objects involved may absorb kinetic energy and break.
+
+There are two major areas of applications for these new features.
+
+Possibility to provide simulated demostrations and verifications of new theories or exercises  would be targeted for 
+studies of mechanics and strength of material. Implementation of computer programs could be done as part of computer graphics.
+
+Methods presented in this work can be used in gaming industry to provide objects that behave more realistically than currently without significant extra work.
+
+\subsection{Scope}
+
+In this work one solution to handling of plasticity is presented and verified with few examples.
+
 \subsection{Scientific contribution}
 
+This work opens wide area of further work of combining structural analysis and plasticity with physics engines.
+Few examples are shown in sections~\ref{sec:results} and \ref{subsec:futureWork}
+
 \cleardoublepage

### 2016-01-06 pdocs/thesis/introduction.tex
pdocs

@@
 \section{Introduction}
+
+This work tries to invite structural engineers to take active part in development of more realistic physis engines.
+Current computer systems provide enough computing performance and efficient algorithms so collision detection and 
+rigid body dynamics can be handled in real time in most scenarios.
+Soft body dynamics in which  object are deformable is in many cases still too complicated for real time applications.
+Finite element method will probably replace other methods by providing general solution at some time but not in very near future.
+In this work focus is on method where rigid bodies are connected by constraints.
 
@@
 Especially gaming and movie industry are powerful drivers.
-Robotics and solid mechanics are typical research areas that use physics engines.
+Robotics is typical research areas that uses physics engines.
 
@@
 \end{itemize}
+
+\cite{cornell.cs3152}
 
@@
 This work opens wide area of further work of combining structural analysis and plasticity with physics engines.
-Few examples are shown in sections~\ref{sec:results} and \ref{subsec:futureWork}
+Few examples are shown in sections~\ref{sec:results} and possible areas of further work are discussed in \ref{subsec:futureWork}.
 

### 2016-01-07 pdocs/thesis/introduction.tex
pdocs

@@
 Robotics is typical research areas that uses physics engines.
+In \lut simulation of working machines has been active research area, e.g. \cite{moisio.thesis}.
 
@@
 
-\cite{cornell.cs3152}
+E.g. \cite{cornell.cs3152} and \cite{cornell.cs5643} are good examples of modern university cources
+about physics engines. They also provide additional resources on used technologies.
 
@@
 
-Target of this work is to find out suitable methods for handling of plasticity in bulletphysics, \cite{bulletphysics}. 
+Target of this work is to find out suitable methods for handling of plasticity and verify them 
+in commonly used physics engine bulletphysics, \cite{bulletphysics}. 
 Handling of plasticity basically means that objects involved may absorb kinetic energy and break.

### 2016-01-07 pdocs/thesis/nimio.tex
pdocs

@@
 due permission for public examination and criticism in the Auditorium ...\\
-at Lappeenranta University of Technology, Lappeenranta, Finland on the\\
+at \lut, Lappeenranta, Finland on the\\
 ..th of \mymonth, \myyear, at noon.}\\

### 2016-01-07 pdocs/thesis/introduction.tex
pdocs

@@
 
-\subsection{Physics engines}
+\begin{tikzpicture}
+  \begin{scope}[blend group = soft light]
+    \fill[red!30!white]   ( 90:3.2) circle (5);
+    \fill[green!30!white] (210:3.2) circle (5);
+    \fill[blue!30!white]  (330:3.2) circle (5);
+  \end{scope}
+  \node at ( 90:4)   [align=left] {structural analysis\\
+material plasticity};
+  \node at ( 210:4)   [align=left] {physics engine\\
+rigid bodies\\
+constraints};
+  \node at ( 330:4)   [align=left] {programming\\
+extending of bulletphysics\\
+verification
+};
+  \node {\begin{varwidth}{2cm}plasticity in constraints\end{varwidth}};
+\end{tikzpicture}
+
+\subsection{Physics engine}
 

### 2016-01-09 pdocs/thesis/discussion.tex
pdocs

@@
 
+\subsubsection{Acceptance to current physics engines}
+
+Required work depends on selected dynamics engine. Availability of these features makes similar 
+implementation possible.
+
+\begin{itemize}
+\item generic spring or motor constraint is already available 
+\item possibility to update constraints after each simulation step
+\end{itemize}
+
+Current implementation is done based on \bullet but it still needs refactoring and testing before it can be 
+accepted to main stream. Interfaces should be reafctored so that handling of additional features like large 
+displacements and material nonlinearity can be implemented without further changes.
+
 \subsubsection{Large displacements}
+Large displacements are typical if plasticity is involved and
+large displacements can also take place due to rigid body motion and should be taken into account.
+
+Update of constraints can be done in  many ways but using callbacks allows clean integration i.e. no changes
+to solver core are needed.
+In \bullet there are already three per step callbacks.
+
+\begin{itemize}
+\item updateActions
+\item updateActivationState
+\item btInternalTickCallback
+\end{itemize}
+
+\demolisher already uses updateActions and that seems 
+good way to do also other updates. Special care should be taken to avoid unnecessary updates.
 
 \subsubsection{Material nonlinearity}
+Update of constraint properties can be done in callbacks in similar way as for large displacements.
+Material definition should be defined so that stress-strain curve can be defined as array of pairs.
 

### 2016-01-09 pdocs/thesis/introduction.tex
pdocs

--- 2016-01-07 pdocs/thesis/introduction.tex
@@
 
-\subsection{Physics engines}
+Figure \ref{fig:areas} summarizes branches involved in this work.
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+  \begin{scope}[blend group = soft light]
+    \fill[red!30!white]   ( 90:3.2) circle (5);
+    \fill[green!30!white] (210:3.2) circle (5);
+    \fill[blue!30!white]  (330:3.2) circle (5);
+  \end{scope}
+  \node at ( 90:5)   [align=left] {solid mechanics\\
+structural analysis\\
+material plasticity};
+  \node at ( 210:4)   [align=left] {mechanics\\
+physics engine\\
+rigid bodies\\
+constraints};
+  \node at ( 330:5.5)   [align=left] {computer sciences\\
+software engineering\\
+extending of bulletphysics\\
+verification
+};
+  \node {\begin{varwidth}{2cm}handling of plasticity in physics\\
+engine\end{varwidth}};
+\end{tikzpicture}
+\caption{Multidisciplinary approach to plasticity in physics engine.}
+\label{fig:areas}
+\end{figure}
+
+\subsection{Structural analysis}
+
+Structural analysis for advanced structures is nowadays typically done using finite element method and 
+analysis time is typically hours. Finite element method is already used in physics engines and usage will come 
+more widespread in future as more computing capacity and memory will be available.
+
+In this work all theories in area of structural analysis are quite old and e.g. 
+\cite{timoshenko} is typical reference. Main focus is on elastic and 
+linear plastic parts of stress-strain curve and elastic and plastic section modulus.
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
+\draw[->] (0,0) -- (0,8) node[above] {\large{$\sigma$}};
+\draw(0,0) -- (2,4) -- (4,4) .. controls (7,6) .. (10,5);
+\draw[dashed](0,4) -- (2,4);
+\node at (-0.2,4) [align=right] {$f_y$};
+\end{tikzpicture}
+\caption{Stress-strain curve of ductile steel (not to scale).}
+\label{fig:areas}
+\end{figure}
+
+\subsection{Physics engine}
 
@@
 In \lut simulation of working machines has been active research area, e.g. \cite{moisio.thesis}.
+
+Physics engines cannot used in serious structural analysis but using structural analysis in physics engines is not rare.  
+E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
 
@@
 
-\subsection{Structural analysis}
-
-Physics engines are not commonly used in structural analysis but using structural analysis in physics engines is not rare.
-
-Structural analysis for advanced structures is nowadays typically done using finite element method and 
-analysis time is typically hours. Finite element method is already used in physics engines and usage will come 
-more widespread in future as more computing capacity and memory will be available.
-
-E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
+\subsection{software engineering}
 
@@
 This work opens wide area of further work of combining structural analysis and plasticity with physics engines.
-Few examples are shown in sections~\ref{sec:results} and possible areas of further work are discussed in \ref{subsec:futureWork}.
+Few examples are shown in sections~\ref{sec:results} and possible areas of further work are discussed in section \ref{subsec:futureWork}.
 

### 2016-01-10 pdocs/thesis/introduction.tex
pdocs

@@
 
+Software engineering is needed to implement suggested idea.
+
+Few decades ago software engineering and structural analysis were highly coupled
+and e.g. \cite{bathe} and \cite{cook} contain very detailed instructions on
+computer implementation. Those books provided good baseline for many computer programs created in \lut for for analysis of 
+thin-walled structures \cite{agifap,fsm,vtb}.
+
+Getting and compiling complicated software packages is currently quite easy. 
+Understanding and getting enough knowledge to extend them requires a fair amount of hard work.
+
 \subsection{Motivation and objectives of the study}

### 2016-01-11 examples/CharpyDemo/CharpyDemo.cpp
fu replaced by fy

@@ -111,8 +111,8 @@ btScalar initialE(200E9); // Steel
 btScalar E(initialE);
 btScalar nu(0.3); // Steel
 btScalar G(initialE / (2 * (1 + nu)));
-btScalar initialFu(400e6);
-btScalar fu(initialFu);
+btScalar initialFy(400e6);
+btScalar fy(initialFy);
 btScalar initialDamping(0.2);
 btScalar damping(initialDamping);
 btScalar initialFrequencyRatio(10);
@@ -462,7 +462,7 @@ public:
 				break;
 			case 4: // cutting
 				hammerDraft = btScalar(0);
-				fu = btScalar(250e6);
+				fy = btScalar(250e6);
 				spaceBetweenAnvils = btScalar(0.022);
 				break;
 			case 5: // sidestep
@@ -472,7 +472,7 @@ public:
 				break;
 			case 6: // big and soft
 				hammerDraft = btScalar(0);
-				fu = btScalar(15e3);
+				fy = btScalar(15e3);
 				w = 0.2;
 				h = 0.2;
 				l = 0.4;
@@ -483,7 +483,7 @@ public:
 				variableTimeStep = false;
 				break;
 			case 7: // Integration instability
-				fu = btScalar(800e6);
+				fy = btScalar(800e6);
 			}
 		}
 
@@ -544,10 +544,10 @@ public:
 		E = tv*1e9;
 		restartHandler(control);
 	}
-	void setFu(Gwen::Controls::Base* control){
-		btScalar tv(fu / 1e6);
+	void setFy(Gwen::Controls::Base* control){
+		btScalar tv(fy / 1e6);
 		setScalar(control, &tv);
-		fu = tv*1e6;
+		fy = tv*1e6;
 restartHandler(control);
 	}
 	void setL(Gwen::Controls::Base* control){
@@ -697,16 +697,16 @@ restartHandler(control);
 		gy += gyInc;
 		gc->onReturnPressed.Add(pPage, &CharpyDemo::setE);
 	}
-	void addFu(){
-		addLabel("fu [MPa]");
+	void addfy(){
+		addLabel("fy [MPa]");
 		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
[…]

### 2016-01-11 examples/CharpyDemo/DemolisherDemo.cpp

@@ -810,7 +810,7 @@ public:
 		int loopSize = ha.size() - 1;
 		btScalar E(200E9);
 		btScalar G(80E9);
-		btScalar fu(200E6);
+		btScalar fy(200E6);
 		bool limitIfNeeded = true;
 		btScalar damping(0.1);
 		btScalar l4s = lsx / lpc;
@@ -825,10 +825,10 @@ public:
 		btScalar k0(E*steelArea/ 0.2); 
 		btScalar k1(E*steelArea*lsy/2.5);
 		btScalar k2(E*steelArea*lsz/2.5);
-		btScalar m(fu / E);
+		btScalar m(fy / E);
 		btScalar w0(k0*m);
-		btScalar w1(k1*fu/E);
-		btScalar w2(k2*fu/E);
+		btScalar w1(k1*fy/E);
+		btScalar w2(k2*fy/E);
 		for (int i = 0; i < loopSize; i++){
 			bt6DofElasticPlastic2Constraint *sc =
 				new bt6DofElasticPlastic2Constraint(*ha[i], *ha[i + 1],

### 2016-01-11 pdocs/thesis/introduction.tex
pdocs

@@
 linear plastic parts of stress-strain curve and elastic and plastic section modulus.
+Straing hardening is taken into account mainly by assuming that plasticity in bending expands.
+This can be seen e.g. by bending paperclip. It does not break low angles but can take few full bends.
 
@@
 \begin{tikzpicture}
+\coordinate (Y) at (0.2,4);
 \draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
-\draw[->] (0,0) -- (0,8) node[above] {\large{$\sigma$}};
-\draw(0,0) -- (2,4) -- (4,4) .. controls (7,6) .. (10,5);
-\draw[dashed](0,4) -- (2,4);
+\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
+\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
+\draw[dashed](0,4) -- (Y);
 \node at (-0.2,4) [align=right] {$f_y$};
@@
 \end{figure}
+
+E.g. \cite{camp} is good example of university cource providing needed information about plasticity.
 
@@
 
-E.g. \cite{cornell.cs3152} and \cite{cornell.cs5643} are good examples of modern university cources
+E.g. \cite{cornell.cs3152} and \cite{cornell.cs5643} are good examples of  university cources
 about physics engines. They also provide additional resources on used technologies.
+
+Few physics engines which have public source
+
+\begin{itemize}
+\item \bullet - \cite{\bullet} which is used e.g. in Blender and Erwin Coumans
+ has received Tehnical Academic Award for creation of the Bullet physics library, \cite{erwin.oscar}
+\item Box2D - \cite\{box2d} which is used e.g. in Angry Birds
+\item RigsOfRods - \cite{ror} which uses large amount of rod elements to simulate deformation and breaking of vehicles
+\end{itemize}
 
@@
 
-Software engineering is needed to implement suggested idea.
+Main requirements for software implementation can be summarized
+\begin{itemize}
+\item solution can be integrated easily to existing dynamics engines
+\item additional calculation time for each timestep may not exceed few milliseconds 
+\end{itemize}
 
@@
 and e.g. \cite{bathe} and \cite{cook} contain very detailed instructions on
-computer implementation. Those books provided good baseline for many computer programs created in \lut for for analysis of 
+computer implementation. Those books provided good baseline for many computer 
+standalone programs created in \lut for for analysis of 
 thin-walled structures \cite{agifap,fsm,vtb}.  
-
 Getting and compiling complicated software packages is currently quite easy. 
 Understanding and getting enough knowledge to extend them requires a fair amount of hard work.
+
+Using calculation methods that have initially planned to be hand calculated gives good starting point for 
+getting fast processing. Lack of clean boundary conditions will require addional logic or accepting larger errors.
 
@@
 Target of this work is to find out suitable methods for handling of plasticity and verify them 
-in commonly used physics engine bulletphysics, \cite{bulletphysics}. 
+in commonly used physics engine \bullet, \cite{bulletphysics}. 
 Handling of plasticity basically means that objects involved may absorb kinetic energy and break.

### 2016-01-11 pdocs/thesis/introduction.tex
pdocs

@@
 \begin{itemize}
-\item \bullet - \cite{\bullet} which is used e.g. in Blender and Erwin Coumans
+\item \bullet - \cite{bullet} which is used e.g. in Blender and Erwin Coumans
  has received Tehnical Academic Award for creation of the Bullet physics library, \cite{erwin.oscar}
-\item Box2D - \cite\{box2d} which is used e.g. in Angry Birds
+\item Box2D - \cite{box2d} which is used e.g. in Angry Birds
 \item RigsOfRods - \cite{ror} which uses large amount of rod elements to simulate deformation and breaking of vehicles
@@
 \begin{itemize}
-\item solution can be integrated easily to existing dynamics engines
+\item plastic deformation takes place if force or moment exceeds given limit
+\item plastic deformation absorbs enegry
+\item after predefined limit of displacement is exceeded constraint is deactivated 
+\item results should be in same magnitude as theoretical and test results
+\item solution can be integrated to existing dynamics engines
 \item additional calculation time for each timestep may not exceed few milliseconds 
@@
 Target of this work is to find out suitable methods for handling of plasticity and verify them 
-in commonly used physics engine \bullet, \cite{bulletphysics}. 
+in commonly used physics engine \bullet. 
 Handling of plasticity basically means that objects involved may absorb kinetic energy and break.
@@
 Possibility to provide simulated demostrations and verifications of new theories or exercises  would be targeted for 
-studies of mechanics and strength of material. Implementation of computer programs could be done as part of computer graphics.
+studies of mechanics and strength of material.
 
 Methods presented in this work can be used in gaming industry to provide objects that behave more realistically than currently without significant extra work.
+Efficient simulation algorithms can also help in saving energy.
+Saving even few wats or percent for single user makes big difference as gaming PCs are estimated to have consumed 75 TWh/year of electricity globally in 2012, \cite{gaming.energy}.
 

### 2016-01-13 examples/CharpyDemo/PlateDemo.cpp
PlateDemo started

@@ -14,9 +14,9 @@ subject to the following restrictions:
 */
 
 /**
-Based on ForkLiftDemo by Simo Nikula 2015-
+Based on DemolisherDemo by Simo Nikula 2016-
 */
-#include "DemolisherDemo.h"
+#include "PlateDemo.h"
 #include "LinearMath/btQuickprof.h"
 #include <stdio.h> 
 #include <string>
@@ -31,11 +31,8 @@ Based on ForkLiftDemo by Simo Nikula 2015-
 #include "bt6DofElasticPlastic2Constraint.h"
 #include "../plasticity/PlasticityExampleBrowser.h"
 
-class btVehicleTuning;
-struct btVehicleRaycaster;
 class btCollisionShape;
 
-#include "BulletDynamics/Vehicle/btRaycastVehicle.h"
 #include "BulletDynamics/ConstraintSolver/btHingeConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btSliderConstraint.h"
 
@@ -49,9 +46,9 @@ class btCollisionShape;
 #include "../ExampleBrowser/GwenGUISupport/gwenUserInterface.h"
 #include "../ExampleBrowser/GwenGUISupport/gwenInternalData.h"
 
-const char * PROFILE_DEMOLISHER_SLEEP = "DemolisherDemo::Sleep";
+const char * PROFILE_PLATE_SLEEP = "PlateDemo::Sleep";
 
-class DemolisherDemo : public Gwen::Event::Handler, public CommonRigidBodyBase
+class PlateDemo : public Gwen::Event::Handler, public CommonRigidBodyBase
 {
 public:
 	class btDiscreteDynamicsWorld* m_dynamicsWorld;
@@ -81,16 +78,11 @@ public:
 	btVector3*	m_vertices;
 	btScalar	maxEngineForce;//this should be engine/velocity dependent
 	btScalar	defaultBreakingForce;
-	btScalar	wheelFriction;
 	btScalar lsx, lsy, lsz;
 	int lpc;
 	btScalar breakingImpulseThreshold;
 	btScalar density;
 	btScalar carMass;
-	btScalar suspensionStiffness, suspensionMaxForce;
-	btScalar suspensionDamping;
-	btScalar suspensionCompression;
-	btScalar suspensionRestLength;
 	btScalar steelArea;
 	btScalar maxPlasticRotation;
[…]

### 2016-01-13 examples/CharpyDemo/PlateDemo.h

@@ -1,0 +1,38 @@
+/*
+Bullet Continuous Collision Detection and Physics Library
+Copyright (c) 2003-2006 Erwin Coumans  http://continuousphysics.com/Bullet/
+
+This software is provided 'as-is', without any express or implied warranty.
+In no event will the authors be held liable for any damages arising from the use of this software.
+Permission is granted to anyone to use this software for any purpose, 
+including commercial applications, and to alter it and redistribute it freely, 
+subject to the following restrictions:
+
+1. The origin of this software must not be misrepresented; you must not claim that you wrote the original software. If you use this software in a product, an acknowledgment in the product documentation would be appreciated but is not required.
+2. Altered source versions must be plainly marked as such, and must not be misrepresented as being the original software.
+3. This notice may not be removed or altered from any source distribution.
+*/
+#ifndef PLATE_DEMO_H
+#define PLATE_DEMO_H
+#include "../CommonInterfaces/CommonExampleInterface.h"
+#include "../CommonInterfaces/CommonGUIHelperInterface.h"
+#include "../CommonInterfaces/CommonRenderInterface.h"
+#include "../CommonInterfaces/CommonWindowInterface.h"
+#include "../CommonInterfaces/CommonGraphicsAppInterface.h"
+#include "../CommonInterfaces/CommonRigidBodyBase.h"
+
+#include "LinearMath/btAlignedObjectArray.h"
+
+class btBroadphaseInterface;
+class btCollisionShape;
+class btOverlappingPairCache;
+class btCollisionDispatcher;
+class btConstraintSolver;
+struct btCollisionAlgorithmCreateFunc;
+class btDefaultCollisionConfiguration;
+
+class CommonExampleInterface*    PlateDemoCreateFunc(struct CommonExampleOptions& options);
+
+#endif // PLATE_DEMO_H
+
+

### 2016-01-13 examples/plasticity/PlasticityExampleBrowser.cpp

@@ -580,7 +580,7 @@ bool PlasticityExampleBrowser::init(int argc, char* argv[])
 #endif
 #define BLEN 256
 	char buf[BLEN];
-	sprintf_s(buf, BLEN, "0.7.0 - %d-%02d-%02d - %s",
+	sprintf_s(buf, BLEN, "0.8.0 - %d-%02d-%02d - %s",
 		tm.tm_year + 1900, tm.tm_mon + 1, tm.tm_mday,
 		btScalarType);
 #if defined (_DEBUG) || defined (DEBUG)

### 2016-01-13 examples/plasticity/PlasticityExampleEntries.cpp

@@ -5,6 +5,7 @@
 #include "LinearMath/btAlignedObjectArray.h"
 #include "../CharpyDemo/CharpyDemo.h"
 #include "../CharpyDemo/DemolisherDemo.h"
+#include "../CharpyDemo/PlateDemo.h"
 #include "../ExampleBrowser/ExampleEntries.h"
 
 #ifdef ENABLE_LUA
@@ -82,18 +83,31 @@ cutting case, large objects, very low fu",
 very stiff constraints cause instability",
 	CharpyDemoCreateFunc, 70102),
 	ExampleEntry(0, "Demolisher"),
-	ExampleEntry(1, "None",
+	ExampleEntry(1, "Demolisher:None",
 	"Only mass resistance",
 	DemolisherDemoCreateFunc, 0),
-	ExampleEntry(1, "Rigid",
+	ExampleEntry(1, "Demolisher:Rigid",
 	"Not moveable",
 	DemolisherDemoCreateFunc, 1),
-	ExampleEntry(1, "Impulse",
-	"Constrains breaks based on impulse",
+	ExampleEntry(1, "Demolisher:Impulse",
+	"Constraints break based on impulse",
 	DemolisherDemoCreateFunc, 2),
-	ExampleEntry(1, "ElasticPlastic",
-	"ElasticPlastic constraint which breaks based on work",
+	ExampleEntry(1, "Demolisher:ElasticPlastic",
+	"ElasticPlastic constraints which break based on work",
 	DemolisherDemoCreateFunc, 3),
+	ExampleEntry(0, "Plate"),
+	ExampleEntry(1, "Plate:None",
+	"Only mass resistance",
+	PlateDemoCreateFunc, 0),
+	ExampleEntry(1, "Plate:Rigid",
+	"Not moveable",
+	PlateDemoCreateFunc, 1),
+	ExampleEntry(1, "Plate:Impulse",
+	"Constraints breaks based on impulse",
+	PlateDemoCreateFunc, 2),
+	ExampleEntry(1, "Plate:ElasticPlastic",
+	"ElasticPlastic constraints which break based on work",
+	PlateDemoCreateFunc, 3),
 };
 struct ExampleEntriesInternalData
 {

### 2016-01-17 examples/CharpyDemo/PlateDemo.cpp
disableCollisionsBetweenLinked bodies added to demolisher

@@ -640,7 +640,7 @@ public:
 	}
 	void prepareAndAdd(bt6DofElasticPlastic2Constraint *sc){
 		btScalar damping(0.1);
-		m_dynamicsWorld->addConstraint(sc, true);
+		m_dynamicsWorld->addConstraint(sc, disableCollisionsBetweenLinkedBodies);
 		for (int i = 0; i < 6; i++)
 		{
 			sc->enableSpring(i, true);
@@ -804,7 +804,8 @@ void PlateDemo::setGameBindings(Gwen::Controls::Base* control){
 		static_cast<Gwen::Controls::CheckBox*>(control);
 	demo->gameBindings = cb->IsChecked();
 }
-void PlateDemo::setDisableCollisionsBetweenLinkedBodies(Gwen::Controls::Base* control){
+void PlateDemo::setDisableCollisionsBetweenLinkedBodies
+	(Gwen::Controls::Base* control){
 	Gwen::Controls::CheckBox* cb =
 		static_cast<Gwen::Controls::CheckBox*>(control);
 	demo->disableCollisionsBetweenLinkedBodies = cb->IsChecked();

### 2016-01-17 examples/CharpyDemo/PlateDemo.cpp
activate load after raise in PlateDemo

@@ -772,6 +772,7 @@ public:
 		m_loadBody->setCenterOfMassTransform(getLoadRaisePoint());
 		m_loadBody->setAngularVelocity(zero);
 		m_loadBody->setLinearVelocity(zero);
+		m_loadBody->activate();
 		raiseLoad = false;
 	}
 
@@ -787,16 +788,14 @@ void PlateDemo::handleRaiseLoad(Gwen::Controls::Base* control){
 	demo->raiseLoad = true;
 }
 void PlateDemo::updateLoadRaiseX(btScalar increment){
-	PlasticityExampleBrowser::setPauseSimulation(true);
 	demo->raiseLoad = true;
 	demo->loadRaiseX += increment;
-	updateLoadRaiseXGc();
+	demo->updateLoadRaiseXGc();
 }
 void PlateDemo::updateLoadRaiseZ(btScalar increment){
-	PlasticityExampleBrowser::setPauseSimulation(true);
 	demo->raiseLoad = true;
 	demo->loadRaiseZ += increment;
-	updateLoadRaiseZGc();
+	demo->updateLoadRaiseZGc();
 }
 
 void PlateDemo::setGameBindings(Gwen::Controls::Base* control){
@@ -1206,6 +1205,12 @@ bool	PlateDemo::keyboardCallback(int key, int state)
 	{
 		switch (key) 
 		{
+		case ' ':
+			{
+				handled = true;
+				demo->raiseLoad = true;
+				break;
+			}
 		case 'a':if (!gameBindings){break;}
 		case B3G_LEFT_ARROW :
 			{

### 2016-01-19 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp
pdocs and debugDraw for ElasticPlastic2Constraint

@@ -42,6 +42,7 @@ http://gimpact.sf.net
 #include "bt6DofElasticPlastic2Constraint.h"
 #include "BulletDynamics/Dynamics/btRigidBody.h"
 #include "LinearMath/btTransformUtil.h"
+#include "LinearMath/btIDebugDraw.h"
 #include <new>
 #include "../plasticity/PlasticityData.h"
 
@@ -1258,6 +1259,13 @@ void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
 }
 void bt6DofElasticPlastic2Constraint::debugDraw(btIDebugDraw* debugDrawer)
 {
+	btVector3 color(1, 0, 0);
+	btVector3 from = m_rbA.getCenterOfMassPosition();
+	btVector3 to = from + m_frameInA.getOrigin();
+	debugDrawer->drawLine(from, to, color);
+	from = m_rbB.getCenterOfMassPosition();
+	to = from + m_frameInB.getOrigin();
+	debugDrawer->drawLine(from, to, color);
 }
 
 

### 2016-01-19 examples/CharpyDemo/PlateDemo.cpp
object indexing fixed in plate

@@ -612,7 +612,7 @@ public:
 			for (int i = 0; i < cx - 1; i++){
 				for (int j = 0; j < cz; j++){
 					btGeneric6DofConstraint *sc =
-						new btGeneric6DofConstraint(*ha[cx*i + j], *ha[cx*(i + 1) + j],
+						new btGeneric6DofConstraint(*ha[cz*i + j], *ha[cz*(i + 1) + j],
 						tra, trb, true);
 					prepareAndAdd(sc);
 				}
@@ -631,7 +631,7 @@ public:
 			for (int i = 0; i < cx ; i++){
 				for (int j = 0; j < cz-1; j++){
 					btGeneric6DofConstraint *sc =
-						new btGeneric6DofConstraint(*ha[cx*i + j], *ha[cx*i + j+1],
+						new btGeneric6DofConstraint(*ha[cz*i + j], *ha[cz*i + j+1],
 						tra, trb, true);
 					prepareAndAdd(sc);
 				}
@@ -677,7 +677,7 @@ public:
 				for (int j = 0; j < cz; j++){
 					bt6DofElasticPlastic2Constraint *sc =
 						new bt6DofElasticPlastic2Constraint
-						(*ha[cx*i + j], *ha[cx*(i + 1) + j],
+						(*ha[cz*i + j], *ha[cz*(i + 1) + j],
 						tra, trb);
 					sc->setMaxPlasticRotation(maxPlasticRotation);
 					sc->setMaxPlasticStrain(maxPlasticStrain);
@@ -717,7 +717,7 @@ public:
 				for (int j = 0; j < cz - 1; j++){
 					bt6DofElasticPlastic2Constraint *sc =
 						new bt6DofElasticPlastic2Constraint
-						(*ha[cx*i + j], *ha[cx*i + j+1],
+						(*ha[cz*i + j], *ha[cz*i + j+1],
 						tra, trb);
 					sc->setMaxPlasticRotation(maxPlasticRotation);
 					sc->setMaxPlasticStrain(maxPlasticStrain);
@@ -1097,24 +1097,21 @@ void PlateDemo::renderScene()
 	if (raiseLoad){
 		raiseLoadAction();
 	}
-	m_guiHelper->syncPhysicsToGraphics(m_dynamicsWorld);
 	updatePitch();
 	updateYaw();
 	updatePauseButtonText();
 	setDumpFilename();
[…]

### 2016-01-19 examples/plasticity/PlasticityExampleBrowser.cpp

@@ -748,12 +748,8 @@ void PlasticityExampleBrowser::update(float deltaTime)
 		s_instancingRenderer->init();
         DrawGridData dg;
         dg.upAxis = s_app->getUpAxis();
-
         {
             BT_PROFILE("Update Camera and Light");
-
-	
-
             s_instancingRenderer->updateCamera(dg.upAxis);
         }
 
@@ -763,7 +759,6 @@ void PlasticityExampleBrowser::update(float deltaTime)
 			glPolygonOffset(3.0, 3);
 			glEnable(GL_POLYGON_OFFSET_FILL);
             s_app->drawGrid(dg);
-			
         }
 		static int frameCount = 0;
 		frameCount++;
@@ -799,9 +794,9 @@ void PlasticityExampleBrowser::update(float deltaTime)
                 sCurrentDemo->renderScene();
             }
             {
-				
 				glPolygonMode( GL_FRONT_AND_BACK, GL_FILL );
-                sCurrentDemo->physicsDebugDraw(gDebugDrawFlags);
+				BT_PROFILE("physicsDebugDraw");
+				sCurrentDemo->physicsDebugDraw(gDebugDrawFlags);
             }
 		}
 

### 2016-01-20 examples/CharpyDemo/PlateDemo.cpp
automatic load size for plate to match object size

@@ -755,6 +755,22 @@ public:
 	btScalar xhl = 0.5, zhl = 0.5;
 	btScalar yhl;
 	btBoxShape * getLoadShape(){
+		switch (cz % 2){
+		case 0:
+			zhl = lsz /cz;
+			break;
+		case 1:
+			zhl = lsz /cz/2;
+			break;
+		}
+		switch (cx % 2){
+		case 0:
+			xhl = lsx /cz;
+			break;
+		case 1:
+			xhl = lsx /cx/2;
+			break;
+		}
 		yhl = loadMass / density / 2 / (2 * xhl) / (2 * zhl);
 		return new btBoxShape(btVector3(xhl, yhl, zhl));
 	}

### 2016-01-20 examples/CharpyDemo/PlateDemo.cpp
conversion to solid steel plate started

@@ -80,10 +80,10 @@ public:
 	btScalar breakingImpulseThreshold;
 	btScalar density;
 	btScalar loadMass, loadRaise, loadRaiseX, loadRaiseZ;
-	btScalar steelArea;
 	btScalar maxPlasticRotation;
 	btScalar maxPlasticStrain;
 	bool gameBindings = true;
+	bool solidPlate = true;
 	bool disableCollisionsBetweenLinkedBodies = true;
 	bool dumpPng = false;
 	char *logDir = _strdup("d:/wrk");
@@ -278,10 +278,10 @@ public:
 	void setLoadRaiseZ(Gwen::Controls::Base* control);
 	void setDefaultBreakingForce(Gwen::Controls::Base* control);
 	void setMaxEngineForce(Gwen::Controls::Base* control);
-	void setSteelArea(Gwen::Controls::Base* control);
 	void setMaxPlasticStrain(Gwen::Controls::Base* control);
 	void setMaxPlasticRotation(Gwen::Controls::Base* control);
 	void setGameBindings(Gwen::Controls::Base* control);
+	void setSolidPlate(Gwen::Controls::Base* control);
 	void setDisableCollisionsBetweenLinkedBodies(Gwen::Controls::Base* control);
 	void setLogDir(Gwen::Controls::Base* control);
 	void setDumpPng(Gwen::Controls::Base* control);
@@ -305,6 +305,15 @@ public:
 		gc->SetChecked(gameBindings);
 		gy += gyInc;
 		gc->onCheckChanged.Add(pPage, &PlateDemo::setGameBindings);
+	}
+	void addSolidPlate(){
+		Gwen::Controls::Label* label = addLabel("solidPlate");
+		Gwen::Controls::CheckBox* gc = new Gwen::Controls::CheckBox(pPage);
+		gc->SetToolTip("solid steel plate instead of enforced concrete");
+		gc->SetPos(gxi, gy);
+		gc->SetChecked(solidPlate);
+		gy += gyInc;
+		gc->onCheckChanged.Add(pPage, &PlateDemo::setSolidPlate);
 	}
 	Gwen::Controls::CheckBox* dumpPngGc;
 	void addDumpPng(){
@@ -448,15 +457,6 @@ public:
 		place(gc);
[…]

### 2016-01-30 examples/CharpyDemo/DemolisherDemo.cpp
wheel parameters in demolisher (dark gray and 1 m radius)

@@ -123,8 +123,8 @@ public:
 	float	m_minCameraDistance;
 	float	m_maxCameraDistance;
 	bool useMCLPSolver = true;
-	float	wheelRadius = 0.5f;
-	float	wheelWidth = 0.4f;
+	float	wheelRadius = 1;
+	float	wheelWidth = 0.6;
 	float	rollInfluence = 0.1f;//1.0f;
 
 	DemolisherDemo(CommonExampleOptions & options);
@@ -1191,7 +1191,7 @@ tr.setOrigin(btVector3(0,-3,0));
 
 	const float position[4]={0,10,10,0};
 	const float quaternion[4]={0,0,0,1};
-	const float color[4]={0,1,0,1};
+	const float color[4]={0.4,0.4,0.4,1};
 	const float scaling[4] = {1,1,1,1};
 
 	for (int i=0;i<4;i++)
@@ -1255,7 +1255,7 @@ tr.setOrigin(btVector3(0,-3,0));
 		m_vehicle->setCoordinateSystem(rightIndex, upIndex, forwardIndex);
 
 		btVector3 connectionPointCS0(xhl-(0.3*wheelWidth),
-			connectionHeight,zhl-wheelRadius);
+			connectionHeight,zhl-2*wheelRadius);
 		btScalar caster = 0.05;
 		btScalar camber = 0.05;
 		btVector3 wheelDirectionCS0(camber, -1, caster);
@@ -1265,14 +1265,14 @@ tr.setOrigin(btVector3(0,-3,0));
 			suspensionRestLength,wheelRadius,m_tuning,isFrontWheel);
 
 		connectionPointCS0 = btVector3(-xhl+(0.3*wheelWidth),
-			connectionHeight,zhl-wheelRadius);
+			connectionHeight,zhl-2*wheelRadius);
 		wheelDirectionCS0 = btVector3(-camber, -1, caster);
 		// right front
 		m_vehicle->addWheel(connectionPointCS0,wheelDirectionCS0,wheelAxleCS,
 			suspensionRestLength,wheelRadius,m_tuning,isFrontWheel);
 
 		connectionPointCS0 = btVector3(-xhl+(0.3*wheelWidth),
-			connectionHeight,-zhl+wheelRadius);
+			connectionHeight,-zhl+2*wheelRadius);
 		wheelDirectionCS0 = btVector3(-camber, -1, -caster);
 		isFrontWheel = false;
 		// right rear
[…]

### 2016-01-30 examples/CharpyDemo/bt6DofElasticPlasticConstraint.h
ElasticPlasticPlate constraint initialization

@@ -75,7 +75,7 @@ public:
 	void setStiffness(int index, btScalar stiffness);
 	void setDamping(int index, btScalar damping);
 	// bcc
-	void setMaxForce(int index, btScalar damping);
+	void setMaxForce(int index, btScalar value);
 	void setMaxPlasticStrain(btScalar value);
 	void setMaxPlasticRotation(btScalar value);
 	void scalePlasticity(btScalar scale);

### 2016-01-30 examples/CharpyDemo/btElasticPlasticPlate.cpp

@@ -15,10 +15,12 @@ subject to the following restrictions:
 /*
 Simo Nikula 2016- while studying plasticity
 */
+#ifdef _MSC_VER
+#define _CRT_SECURE_NO_WARNINGS
+#endif
 
 #include "btElasticPlasticPlate.h"
 #include "btBulletDynamicsCommon.h"
-
 /**
 It is assumed that rbA and rbB have same size and link is wanted between them
 */
@@ -37,12 +39,65 @@ void btElasticPlasticPlate::setMaterial(btElasticPlasticMaterial* material) {
 	m_material = material;
 }
 
+void btElasticPlasticPlate::initStiffnesses
+(btScalar* v, btScalar l, btScalar b, btScalar t){
+	btScalar E(m_material->getE());
+	btScalar G(m_material->getG());
+	btScalar A(b*t);
+	btScalar I1(b*t*t*t / 12);
+	btScalar I2(t*b*b*b / 12);
+	btScalar It(b*t*t*t/7);
+	btScalar k0(E*A/l);
+	btScalar k1(12 * E*I1 / l / l / l);
+	btScalar k2(12 * E*I2 / l / l/ l);
+	v[0] = k0;
+	v[1] = k1;
+	v[2] = k2;
+	v[3] = G*It / l;
+	v[4] = 3 * E*I1 / l;
+	v[5] = 3 * E*I2 / l;
+}
+
+void btElasticPlasticPlate::initMaxForces
+(btScalar* v, btScalar l, btScalar b, btScalar t){
+	btScalar fy = m_material->getFy();
+	btScalar w1 = fy*b*t*t / 4;
+	btScalar w2(fy*b*b*t / 4);
+	v[0] = fy*b*t;
+	v[1] = v[0]/2;
+	v[2] = v[0] / 2;
+	v[3] = w1;
+	v[4] = w1;
+	v[5] = w2;
+}
+
+
+void btElasticPlasticPlate::initMaterialBasedLimits(){
+	btScalar mps = m_material->getMaxPlasticStrain();
+	m_maxPlasticStrain = m_ml / m_mc*mps;
+	m_maxPlasticRotation = mps * 10;
+}
+
+void btElasticPlasticPlate::initStiffnesses(){
+	initStiffnesses(m_stiffnessL, btScalar(m_ll / m_lc), 
+			btScalar(m_ml / m_mc), m_thickness);
[…]

### 2016-01-30 examples/CharpyDemo/btElasticPlasticPlate.h

@@ -36,9 +36,13 @@ protected:
 	int m_mc; // how many objects for middle dimension
 	btScalar m_ll,m_ml,m_thickness;
 	int m_maxAxis, m_middleAxis, m_minAxis;
-	btBoxShape *m_subShape=0;
+	btScalar m_maxForceL[6], m_maxForceM[6];
+	btScalar m_stiffnessL[6], m_stiffnessM[6];
+	btScalar m_maxPlasticRotation, m_maxPlasticStrain;
+	btBoxShape *m_subShape = 0;
 	btTransform m_mainTransform;
-	bool disableCollisionsBetweenLinkedBodies;
+	bool m_disableCollisionsBetweenLinkedBodies;
+	bool m_limitIfNeedeed;
 	btScalar damping=0.1;
 	/**
 	* store all sequentially m-direction in inner loop
@@ -47,10 +51,17 @@ protected:
 	btAlignedObjectArray<bt6DofElasticPlastic2Constraint*> m_constraints;
 	btElasticPlasticMaterial* m_material;
 	virtual void initSubShape();
+	virtual void initStiffnesses
+		(btScalar* v, btScalar l, btScalar b, btScalar t);
+	virtual void initMaxForces
+		(btScalar* v, btScalar l, btScalar b, btScalar t);
+	virtual void initStiffnesses(), initMaxForces(),initMaterialBasedLimits();
 	virtual void initRigidBodies(btDiscreteDynamicsWorld* dw);
 	virtual void initConstraints(btDiscreteDynamicsWorld* dw);
-	virtual const btTransform getConnectingFrame(btRigidBody& rbA, btRigidBody& rbB);
-	virtual void updateConstraint(bt6DofElasticPlastic2Constraint &constraint);
+	virtual const btTransform getConnectingFrame
+		(btRigidBody& rbA, btRigidBody& rbB);
+	virtual void updateConstraint
+		(bt6DofElasticPlastic2Constraint &constraint);
 	virtual void prepareAndAdd
 		(bt6DofElasticPlastic2Constraint *sc, btDiscreteDynamicsWorld* dw);
 public:
@@ -62,6 +73,8 @@ public:
 	int getLongCount(){ return m_lc; }
 	int getMiddleCount(){ return m_mc; }
 	virtual void join(btDiscreteDynamicsWorld* dw);
+	bool getLimitIfNeeded(){ return m_limitIfNeedeed; }
[…]

### 2016-02-01 examples/CharpyDemo/btElasticPlasticPlate.cpp
fixes for ElasticPlasticPlate initialization

@@ -139,12 +139,12 @@ void btElasticPlasticPlate::initSubShape(){
 			m_mc = 1;
 		}
 	}
-	m_ll = v[m_maxAxis];
-	m_ml = v[m_middleAxis];
+	m_ll = 2*v[m_maxAxis];
+	m_ml = 2*v[m_middleAxis];
 	btVector3 sv=v;
 	sv[m_maxAxis] /= m_lc;
 	sv[m_middleAxis] /= m_mc;
-	m_thickness = v[m_minAxis];
+	m_thickness = 2*v[m_minAxis];
 	m_subShape = new btBoxShape(sv);
 }
 
@@ -181,6 +181,7 @@ void btElasticPlasticPlate::prepareAndAdd
 	sc->setMaxPlasticStrain(m_maxPlasticStrain);
 	updateConstraint(*sc);
 	dw->addConstraint(sc, m_disableCollisionsBetweenLinkedBodies);
+	m_constraints.push_back(sc);
 	for (int i = 0; i < 6; i++)
 	{
 		sc->enableSpring(i, true);

### 2016-02-03 pdocs/thesis/bullet-plasticity.tex
bullet-plasticity started based on google doc page

\section{Simulation of Charpy impact test using \bullet}
\label{sec:bullet-plasticity}

In following sections various methods for handling plasticity are described.
As physics engines have not been traditional area of research for plasticity, 
well known Charpy impact test was selected to be used as primary scenario
although it is against first two tips in \bullet Manual.

\begin{itemize}
\item Avoid very small and very larger collision shapes
\item Avoid large mass ratios
\end{itemize}

\subsection{Introduction to constraint processing in \bullet}
Constraint processing in \bullet is based on \cite{ode}.
 Mathematical background and detailed examples are available in \cite{ode.joints}.
Main parameters are described in table.

\subsection{Common data}

\subsubsection{Material}

Material is steel. Density is 7800 $kg\over{m^{3}}$. Young’s modulus is 200 GPa. Yield stress is initially 400 MPa but it can be modified.

\subsubsection{Coordinate system}

X axis is horizontal, positive to direction where hammer comes from (left)\\
Y axis is vertical, positive up\\
Z axis is horizontal and Z=0 is symmetry plane

\subsubsection{Specimen}
Specimen dimensions can modified. Basic measures are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations. Specimen is positioned symmetrically around z=0, bottom at y=0.2 m and backside at x=0. 
Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
ultimate tensile strength of specimen. For hinge angle of 1.9 radians and ultimate tensile strength of 400 MPa expected energy 
loss is 122 J. Mass is about 40 g and in most cases it is modelled as two 20 g items. In laboratory tests variation for similar 
specimens is roughly about 10 %; see e.g. http://www.npl.co.uk/upload/pdf/cop06.pdf
Specimen can be subdivided to multiple parts.

\subsubsection{Support anvils}
Support anvils initially have 40 mm open space between them. Their width is 40 mm. 
If specimen bends about 1.9 radians (108 degrees) it will go between anvils. Space between anvils can be changed.

\subsubsection{Hammer}
Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m. Mass is 19.5 kg. Hammer and hinge are positioned so that 
impact is horizontal  (global x-direction). Hammer arm is 40x40x500 mm. Hammer thickness can be changed.
Setting it to zero causes hammer not to get added to model. Hammer has modifiable draft. Default value is 0.04 m. 
Arm and draft are not taken into account for mass and inertia calculation.

\subsubsection{Energy calculation}

Energy for whole system is calculated so that potential energy and kinetic energy are calculated in updateEnergy. 
When hammer is resting in low position, energy is 64 J.

\subsubsection{Breaking of constraints between objects}
For standard \bullet constraints breakingImpulseThreshold (BITh) can be defined. If impulse is larger than set limit constraint is activated. 
This allows object breaking to be simulated with very cheap calculation. 
This method is too simplified for ductile material and developing more precise methods in main objective of this work.

\subsubsection{Timestep}

Usually bullet simulations are done using fixed time step of 1/60 s i.e. 16.67 ms. 
For this case 17 ms is too large timestep even to keep system stable. For impact time much smaller timestep is needed. 
Real time simulation was tried as option but was removed from code as it did not provide good results.  
Default time step was selected to be 5 ms outside impact time and 0.1 ms during impact. 
Automatic time stepping routine changes timestep so that at angles higher than 0.2 5 ms time step is
used and adjusts it linearly to selected time step between angles of 0.2 and 0.05. If specimen is no
longer at anvil larger timestep is also used.

\subsection{Single rigid body}
This is basic reference case without plasticity. Specimen should stop hammer.

\subsection{SpringConstraint}
This shows how btGeneric6DofSpringConstraints act in case like this. 
Specimen can absorb very little elastic energy so rotational spring constants are calculated using plastic state i.e.
$W=bh^2/4$ where h=b-0.002m and yield stress is used instead of Young’s modulus for rotational springs.

\subsection{Constraint with zero limits}
In this case btGeneric6DofConstraint is used with high and low limits set to zero. 
Constraint forces are obtained using JointFeedback feature. Plastic moment is 64 Nm for default dimensions

\subsection{Spring2Constraint}
This basically same as SpringConstraint but uses newer btGeneric6DofSpring2Constraint which has recently appeared in
\bullet. 
Automatic stiffness limitation helps to avoid instability but may make constraint too soft.

\subsection{Hinge constraint with motors}
In this case btHingeConstraint is used.
Idea is to enable angular motor in which has target velocity of 0 and maximum motor impulse to 
be plastic moment multiplied by timestep so that motor resists any rotation.

\subsection{PlasticHingeConstraint}
In this case new btPlasticHingeConstraint is used.
It is modification of btHingeConstraint and was is developed during this work.
PlasticHingeConstraint has additional fields for plasticMoment and previousHingeAngle and 
getAbsorbedEnergy method which returns product of plasticMoment and hingeAngle change.

\subsubsection{MaxPlasticRotation}
MaxPlasticRotation is new term which was introduced to allow efficient way to express ultimate strain in bending.
Implemented constraint has additional fields for maxPlasticRotation and currentPlasticRotation. 
After each step in which rotation angle changes, absolute value of 
angle change is accumulated to currentPlasticRotation. As currentPlasticRotation 
reaches maxPlasticRotation constraint is inactivated and specimen breaks. 
For ductile behaviour maxPlasticRotation could be set to e.g. 3-6 (radians) and for brittle
 behaviour to e.g.0.01-0.1 radians. MaxPlasticRotation is basically rotational stretch under bending moment.
 Basic sample can be tried using paperclip. It can typically handle two bend overs before breaking. 
Sewing machine needle or tooth pick which has same diameter is much harder to bend and it 
often breaks before plastic deformation takes place.

\subsubsection{Limiting maximum moment in constraint}
btPlasticHingeConstraint::getInfo2InternalUsingFrameOffset is modified so that in case of lostop=histop, 
plasticMoment*timeStep is used as upper and lower limit instead of SIMD\_INFINITY.

\subsection{ElasticPlasticConstraint}
In this case new bt6DofElasticPlasticConstraint is used.
It is based on DofSpringConstraints with some modifications

\begin{itemize}
\item internalUpdateSprings is modified so that nonlinear relation between displacements and forces can be defined. 
 If force is smaller than maximum force or moment for given degree of freedom,
  elastic behavior is handled in similar way as spring works. 
  If force is larger, maximum force is used. InternalUpdateSprings is called by getInfo2.
\item constraint breaks if active forces are high enough and plastic reserve has been used
\item frequency ratio of lowest mode and integration is used to limit spring functionality to avoid stability issues. 
 Ratio defines how many steps are needed for one period. 
 If number of integration steps is too low, spring is modified to ignore elastic part.
\end{itemize}

\subsection{ElasticPlastic2Constraint}
This is basically same as elasticPlasticConstraint but uses new bt6DofElasticPlastic2Constraint 
which is based on btGeneric6DofSpring2Constraint.
Get\_limit\_motor\_info2, setAngularLimits and setLinearLimits which are called by getInfo2 are modified. 
SIMD\_INFINITY values are replaced by maximum force values which are transformed to
corresponding impulsed by multiplying forces by integration interval.
btGeneric6DofSpring2Constraint has option limitIfNeeded for stiffness and damping to avoid issues. 
Spring is softened so that 4 steps are used for one step. This is not realistic for typical steel structures. 
In elasticPlasticConstraint2 plasticity code is activated if frequency is too high for stable solution.
If plasticity is activated due to large spring force or due to high frequency forces constraint error is 
calculated in similar way as for limited case (m\_currentLimit==3).

\cleardoublepage

### 2016-02-04 pdocs/thesis/bullet-plasticity.tex
bullet-plasticity constraint math

@@
 \subsection{Introduction to constraint processing in \bullet}
-Constraint processing in \bullet is based on \cite{ode}.
- Mathematical background and detailed examples are available in \cite{ode.joints}.
-Main parameters are described in table.
+Constraint processing in \bullet is based on ODE, \cite{ode}.
+Mathematical background and detailed examples are available by \cite{ode.joints}.
+Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
+\ref{eq:lambdaHigh} 
+are created for each constraint.
+
+\begin{equation} \label{eq:constraintEquation}
+J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
+\end{equation}
+
+\begin{equation} \label{eq:lambdaLow}
+\lambda \geq l
+\end{equation}
+
+\begin{equation} \label{eq:lambdaHigh}
+\lambda \leq h
+\end{equation}
+
+Main parameters  and corresponding fields in \bullet 
+ are described in table \ref{tab:constraintParameters}.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|c| l| l|}
+\hline
+{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
+$J_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
+$J_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
+$v$ & linear velocity & \\ \hline
+$\omega$ & angular velocity & \\ \hline
+$c$        &  right side vector   & m\_constraintError \\ \hline
+$C$  & constraint force mixing & cfm \\  \hline
+$\lambda$ & constraint force &  \\ \hline
+$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
+$h$ & high limit for constraint force & m\_upperLimit \\ \hline
+\end {tabular}
+\end{center}
+\caption {Constraint parameters} \label{tab:constraintParameters} 
+\end {table}
 

### 2016-02-06 pdocs/thesis/bullet-plasticity.tex
bullet-plasticity docs tuning

@@
 In this case btGeneric6DofConstraint is used with high and low limits set to zero. 
-Constraint forces are obtained using JointFeedback feature. Plastic moment is 64 Nm for default dimensions
+Constraint is made breakable by setting breakingImpulseThreshold.
+This is common technique to provide breakable objects in games and provides visually accetable results for brittle materials.
+Constraint forces are obtained using JointFeedback feature. Plastic moment is 64 Nm for default dimensions.
 
@@
 In this case btHingeConstraint is used.
-Idea is to enable angular motor in which has target velocity of 0 and maximum motor impulse to 
+Angular motor is enabled with target velocity of 0 and maximum motor impulse is set to
 be plastic moment multiplied by timestep so that motor resists any rotation.
@@
 In this case new bt6DofElasticPlasticConstraint is used.
-It is based on DofSpringConstraints with some modifications
+It is based on SpringConstraint with some modifications
 
@@
   elastic behavior is handled in similar way as spring works. 
-  If force is larger, maximum force is used. InternalUpdateSprings is called by getInfo2.
+  If force is larger, maximum force is used. 
 \item constraint breaks if active forces are high enough and plastic reserve has been used
@@
 Spring is softened so that 4 steps are used for one step. This is not realistic for typical steel structures. 
-In elasticPlasticConstraint2 plasticity code is activated if frequency is too high for stable solution.
+In bt6DofElasticPlastic2Constraint plasticity code is activated if frequency is too high for stable solution.
 If plasticity is activated due to large spring force or due to high frequency forces constraint error is 
-calculated in similar way as for limited case (m\_currentLimit==3).
+calculated in similar way as for case where upper limit equals lower limit. (m\_currentLimit==3).
 

### 2016-02-07 pdocs/thesis/bullet-plasticity.tex
pdocs introduction tuning

@@
 If plasticity is activated due to large spring force or due to high frequency forces constraint error is 
-calculated in similar way as for case where upper limit equals lower limit. (m\_currentLimit==3).
+calculated in similar way as for case where upper limit equals lower limit, (m\_currentLimit==3).
 

### 2016-02-07 pdocs/thesis/introduction.tex
pdocs and debugDraw for ElasticPlastic2Constraint
typo fix in thesis/introduction.tex
pdocs introduction tuning

@@
 linear plastic parts of stress-strain curve and elastic and plastic section modulus.
+
+Typical stress strain curve of ductile steel is shown in \ref{fig:areas}.
+Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
 Straing hardening is taken into account mainly by assuming that plasticity in bending expands.
-This can be seen e.g. by bending paperclip. It does not break low angles but can take few full bends.
+This can be seen e.g. by bending paperclip. It does not break at low angles but can take few full bends.
+
+Difference between elastic and plastic modulus is shown in \ref{fig:wp}. 
+If stress is below yield limit, stress and strain are linear within cross section.
+If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
+so plastic modulus is higher than elastic modulus.
+
+Elastic part is often ignored in this work as displacements due to to elastic deformation are so small and 
+related frequencies are so high.
 
@@
 
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (S) at (2.5,5);
+\draw (0,5) -- (4,5) ;
+\draw (0,0) -- (4,0) ;
+\draw (2,0) -- (2,5) ;
+\draw (1.5,0) -- (S); 
+\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
+\end{tikzpicture}
+\hspace{1cm}
+\begin{tikzpicture}
+\coordinate (S) at (3,5);
+\draw (0,5) -- (4,5) ;
+\draw (0,0) -- (4,0) ;
+\draw (2,0) -- (2,5) ;
+\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
+\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
+\end{tikzpicture}
+\caption{Stress distribution under elastic and plastic loads.}
+\label{fig:wp}
+\end{figure}
+
 E.g. \cite{camp} is good example of university cource providing needed information about plasticity.
@@
 E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
+Already \cite{cg1988} used quite similar principles as in this work to handle plasticity.
 
@@
 
-\subsection{software engineering}
+\subsection{Software engineering}
 
@@
 Using calculation methods that have initially planned to be hand calculated gives good starting point for 
-getting fast processing. Lack of clean boundary conditions will require addional logic or accepting larger errors.
+getting fast processing. Lack of clean boundary conditions will require additional logic or accepting larger errors.
 

### 2016-02-08 pdocs/thesis/bullet-plasticity.tex
pdocs charpy-results

@@
 \section{Simulation of Charpy impact test using \bullet}
-\label{sec:bullet-plasticity}
+\label{sec:bullet-charpy}
 
@@
 
+\subsection{Constraint with zero limits}
+In this case btGeneric6DofConstraint is used with high and low limits set to zero. 
+Constraint is made breakable by setting breakingImpulseThreshold.
+This is common technique to provide breakable objects in games and provides visually accetable results for brittle materials.
+
 \subsection{SpringConstraint}
@@
 $W=bh^2/4$ where h=b-0.002m and yield stress is used instead of Young’s modulus for rotational springs.
-
-\subsection{Constraint with zero limits}
-In this case btGeneric6DofConstraint is used with high and low limits set to zero. 
-Constraint is made breakable by setting breakingImpulseThreshold.
-This is common technique to provide breakable objects in games and provides visually accetable results for brittle materials.
-Constraint forces are obtained using JointFeedback feature. Plastic moment is 64 Nm for default dimensions.
 
@@
 
+\section{Charpy impact test results using \bullet}
+\label{sec:bullet-charpy-results}
+
+\subsection{Results for single rigid body}
+Timestep of 0.1 ms provides expeced results and specimen bounces hammer back.
+Larger timesteps produce various kinds of unexpected results like
+\begin{itemize}
+\item Specimen goes trough anvil and penetrates hammer
+\item Specimen moves sideways and stays on anvil. This can be seen e.g. by setting timestep to 17 ms and leaving hammer draft away.
+\end{itemize}
+
+\subsection{Results for constraint with zero limits}
+This technique provides effective and stable way to simulate breaking into two objects. 
+Required energy to break constraint can be controlled using
+breakingImpulseThreshold.
+
+\subsection{Results for springConstraint}
+Energy loss is in right scale but visual feedback is not realistic as specimen gets back to initial shape.
+Unstable simulation can be seen by setting yield stress to e.g. 800 MPa.
+
+\subsection{Results for spring2Constraint}
+Results are quite similar as for springConstraint but automatic parameter tuning helps in avoiding unstable simulations.
+
 \cleardoublepage

### 2016-02-10 pdocs/thesis/bullet-plasticity.tex
pdocs charpy-results

@@
 
+\subsection{Results for hinge constraint with motors}
+Results are very good although sensitive to parameters changes. Energy loss is in right scale. 
+Specimen bends and keeps deformed shape.
+
+\subsection{Results for plasticHingeConstraint}
+Results are quite similar as for hinge constraint with motors. 
+Breaking point can be configured based on rotation angle which makes material behaviour more realistic.
+
+\subsection{Results for elasticPlasticConstraint}
+Results are quite similar as for plasticHingeConstraint as this case is dominated by plastic bending.
+Additional logic helps to avoid stablity issues.
+
+\subsection{Results for elasticPlastic2Constraint}
+Results are quite similar as for elasticPlasticConstraint and this constraint seems to best candidate for further development.
+Table \ref{tab:ep2ts} shows energy loss for few timesteps as example to demonstrate sensitivity of solution.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{| c| c|}
+\hline
+{\bf Timestep[ms]} & {\bf Energy loss [J]}\\  \hline
+ 0.100001 &  120 \\ \hline
+ 0.1 &  110 \\ \hline
+ 0.09999 &  140 \\ \hline
+ 0.08 &  170 \\ \hline
+ 0.05 &  215 \\ \hline
+\end {tabular}
+\end{center}
+\caption {Energy loss for few timesteps} \label{tab:ep2ts} 
+\end {table}
+
 \cleardoublepage

### 2016-02-10 pdocs/thesis/discussion.tex
pdocs
pdocs charpy-results

@@
 \label{subsec:futureWork}
+
+\subsubsection{Implementation using GPU}
+\bullet\ can already utilize GPU for many operations. 
+Analysis on possible issues in used interfaces should be done before integrations.
 
@@
 
-Current implementation is done based on \bullet but it still needs refactoring and testing before it can be 
+Current implementation is done based on \bullet\ but it still needs refactoring and testing before it can be 
 accepted to main stream. Interfaces should be reafctored so that handling of additional features like large 
@@
 to solver core are needed.
-In \bullet there are already three per step callbacks.
+In \bullet\ there are already three per step callbacks.
 
@@
 Material definition should be defined so that stress-strain curve can be defined as array of pairs.
+This is probably not significant feature for simulations in gaming area but interesting feature for research and teaching area.
+
+\subsection{Object subdivision just in time}
+Just in time activation could avoid extra overhead from multiple objects and constraints.
+Impacts causing plasticity are usually quite brief and they happen at distinct times. 
+Simulation could allocate processing units for handling these and release them after impact has been processed.
 

### 2016-02-13 pdocs/thesis/bullet-plasticity.tex
f1 and f2 results

@@
 Material is steel. Density is 7800 $kg\over{m^{3}}$. Young’s modulus is 200 GPa. Yield stress is initially 400 MPa but it can be modified.
+Coefficient of restitution can be modified and has value of 0 unless otherwise staten.
 
@@
 \subsection{Results for single rigid body}
-Timestep of 0.1 ms provides expeced results and specimen bounces hammer back.
-Larger timesteps produce various kinds of unexpected results like
-\begin{itemize}
-\item Specimen goes trough anvil and penetrates hammer
-\item Specimen moves sideways and stays on anvil. This can be seen e.g. by setting timestep to 17 ms and leaving hammer draft away.
-\end{itemize}
+
+Timestep of 0.1 ms provides expeced results and specimen bounces hammer back as seen in figure \ref{fig:f1-ok-results}.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=3cm]{figs/f1-18-01-1}
+\includegraphics[height=3cm]{figs/f1-18-01-2}
+\caption{Specimen bounces hammer back with timestep of 0.1 ms}
+\label{fig:f1-ok-results}
+\end{figure}
+
+Larger timesteps produce various kinds of unrealistic results seen in figure \ref{fig:f1-nok-results}.
+In these cases timestep is 17 ms. Specimen penetrates hammer and anvils.
+Without draft specimen stays on anvil.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=3cm]{figs/f1-18-17-1}
+\includegraphics[height=3cm]{figs/f1-18-17-2}
+\includegraphics[height=3cm]{figs/f1-18-17-3}
+\caption{Results with timestep of 17 ms}
+\label{fig:f1-nok-results}
+\end{figure}
 
 \subsection{Results for constraint with zero limits}
-This technique provides effective and stable way to simulate breaking into two objects. 
+This technique provides effective and stable way to simulate breaking of brittle material 
+but does not provide realistic results for ductile material as can be seen in figure \ref{fig:f3-results}.
 Required energy to break constraint can be controlled using
 breakingImpulseThreshold.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=2.5cm]{figs/f3-18-01-1}
+\includegraphics[height=2.5cm]{figs/f3-18-01-2}
+\includegraphics[height=2.5cm]{figs/f3-18-01-3}
+\caption{Specimen breaks in brittle way}
+\label{fig:f3-results}
+\end{figure}
 

### 2016-02-14 pdocs/thesis/bullet-plasticity.tex
charpy results

@@
 Material is steel. Density is 7800 $kg\over{m^{3}}$. Young’s modulus is 200 GPa. Yield stress is initially 400 MPa but it can be modified.
-Coefficient of restitution can be modified and has value of 0 unless otherwise staten.
+Coefficient of restitution can be modified and has value of 0 (fully plastic) unless otherwise staten.
 
@@
 Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
-ultimate tensile strength of specimen. For hinge angle of 1.9 radians and ultimate tensile strength of 400 MPa expected energy 
+ultimate tensile strength of specimen. For hinge angle of 1.9 radians and yields stress of 400 MPa expected energy 
 loss is 122 J. Mass is about 40 g and in most cases it is modelled as two 20 g items. In laboratory tests variation for similar 
-specimens is roughly about 10 %; see e.g. http://www.npl.co.uk/upload/pdf/cop06.pdf
+specimens is roughly about 10 %; see e.g. \cite{ct2000}.
 Specimen can be subdivided to multiple parts.
@@
 This allows object breaking to be simulated with very cheap calculation. 
-This method is too simplified for ductile material and developing more precise methods in main objective of this work.
+This method is too simplified for ductile material and developing more precise and realistic methods is main objective of this work.
 
@@
 
-\subsubsection{MaxPlasticRotation}
-MaxPlasticRotation is new term which was introduced to allow efficient way to express ultimate strain in bending.
+\subsubsection{Maximum plastic rotation}
+Maximum plastic rotation is new term which was introduced to allow efficient way to express ultimate strain in bending.
 Implemented constraint has additional fields for maxPlasticRotation and currentPlasticRotation. 
@@
 reaches maxPlasticRotation constraint is inactivated and specimen breaks. 
-For ductile behaviour maxPlasticRotation could be set to e.g. 3-6 (radians) and for brittle
- behaviour to e.g.0.01-0.1 radians. MaxPlasticRotation is basically rotational stretch under bending moment.
+For ductile behaviour maxPlasticRotation could be set to e.g. 1-6 (radians) and for brittle
+ behaviour to e.g.0.01-0.1 radians. Maxium plastic rotation is basically rotational stretch under bending moment.
  Basic sample can be tried using paperclip. It can typically handle two bend overs before breaking. 
@@
 \label{sec:bullet-charpy-results}
+Results are quite sensitive and obtained in February 2016 using following processes
+\begin{itemize}
+\item single precision debug build is used
+\item sequential impulse constraint solver with 10 iterations is used
+\item energy loss values are rounded to nearest 10 joules 
+\end{itemize}
 
@@
 \subsection{Results for springConstraint}
-Energy loss is in right scale but visual feedback is not realistic as specimen gets back to initial shape.
+Energy loss is about 50 J.
+Visual feedback is not realistic as specimen gets back to initial shape as seen in figure \ref{fig:f3-results}.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=2.5cm]{figs/f2-18-01-1}
+\includegraphics[height=2.5cm]{figs/f2-18-01-2}
+\caption{Spring constraint brings specimen back to initial shape}
+\label{fig:f3-results}
+\end{figure}
+
 Unstable simulation can be seen by setting yield stress to e.g. 800 MPa.
+Few frames of such simulation can be seen in figure \ref{fig:f3-us-results}.
+Unstable behaviour starts immediately at start of simulation and is triggered without impact from hammer.
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=2.2cm]{figs/f2-18-01-us-1}
+\includegraphics[height=2.2cm]{figs/f2-18-01-us-2}
+\includegraphics[height=2.2cm]{figs/f2-18-01-us-3}
+\caption{Unstable simulation with stiff spring contstaints}
+\label{fig:f3-us-results}
+\end{figure}
 
 \subsection{Results for spring2Constraint}
-Results are quite similar as for springConstraint but automatic parameter tuning helps in avoiding unstable simulations.
+Results are quite similar as for springConstraint but automatic parameter tuning prevents unstable simulations.
+Energy loss is about 40 J.
 
 \subsection{Results for hinge constraint with motors}
-Results are very good although sensitive to parameters changes. Energy loss is in right scale. 
-Specimen bends and keeps deformed shape.
+Energy loss is about 110 J.
+Visual feedback is realistic as specimen keeps deformed shape as seen in figure \ref{fig:f5-results}.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=2.5cm]{figs/f5-18-01-1}
+\includegraphics[height=2.5cm]{figs/f5-18-01-2}
+\includegraphics[height=2.5cm]{figs/f5-18-01-3}
+\caption{Hinge constraint with motors}
+\label{fig:f5-results}
+\end{figure}
 
@@
 Breaking point can be configured based on rotation angle which makes material behaviour more realistic.
+In figure \ref{fig:f6-results} maximum plastic rotation is limited to 1.5 and energy loss is about 95 J.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=2.5cm]{figs/f6-18-01-1}
+\includegraphics[height=2.5cm]{figs/f6-18-01-2}
+\includegraphics[height=2.5cm]{figs/f6-18-01-3}
+\caption{PlasticHingeConstraint with maximum plastic rotation set to 1.5}
+\label{fig:f6-results}
+\end{figure}
 
@@
 {\bf Timestep[ms]} & {\bf Energy loss [J]}\\  \hline
- 0.100001 &  120 \\ \hline
+ 0.1001 &  100 \\ \hline
  0.1 &  110 \\ \hline
- 0.09999 &  140 \\ \hline
- 0.08 &  170 \\ \hline
- 0.05 &  215 \\ \hline
+ 0.099 &  90 \\ \hline
+ 0.08 &  130 \\ \hline
+ 0.05 &  150 \\ \hline
 \end {tabular}
@@
 
+\subsection{Summary and evaluation of results}
+Suggested methods for handling plasticity are well suited for cases where simulation of ductile failure is desired.
+It requires only slightly more computational effort than methods that are used for simulation of brittle failures.
+Hinge constraint with motors method can be used without introducing any additional code to most existing dynamic engine
+as it is based on widely used motor-codes. 
+Additional features are quite simple to implement and make spring based constraints even more usable than currently by 
+allowing stiff constraints to be used without instability.
+
+Suggested methods cannot at least currently be used if exact results are needed but are well suited for educational and entertainment scenarios.
+
 \cleardoublepage

### 2016-02-14 pdocs/thesis/introduction.tex
more precise terms in introduction

@@
 
-Difference between elastic and plastic modulus is shown in \ref{fig:wp}. 
-If stress is below yield limit, stress and strain are linear within cross section.
-If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
-so plastic modulus is higher than elastic modulus.
-
-Elastic part is often ignored in this work as displacements due to to elastic deformation are so small and 
-related frequencies are so high.
-
 \begin{figure}[htb!]
@@
 \end{figure}
+
+Difference between elastic and plastic section modulus is shown in \ref{fig:wp}.
 
@@
 \end{figure}
+
+If stress is below yield limit, stress and strain are linear within cross section.
+If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
+so plastic section modulus is higher than elastic section modulus.
+
+Elastic part is often ignored in this work as displacements due to to elastic deformation are small and 
+related frequencies are high.
 

### 2016-02-15 examples/CharpyDemo/DemolisherDemo.cpp
chassis tuning in demolisher

@@ -233,16 +233,19 @@ public:
 	btVector3 speedometerLocation = lastLocation;
 	float updateInterval = 0.3;
 	int updateViewCount = 0;
+	void resetClocks(){
+		driveClock.reset();
+		speedometerUpdated = 0;
+	}
 	void updateView(){
 		updateViewCount++;
 		float now = driveClock.getTimeSeconds();
-		float timeDelta = now - lastClock;
 		lastClock = currentClock;
 		currentClock = now;
 		lastLocation = currentLocation;
 		currentLocation = m_carChassis->getCenterOfMassPosition();
 		updateCameraLocation();
-		timeDelta = now - speedometerUpdated;
+		float timeDelta = now - speedometerUpdated;
 		if (timeDelta < updateInterval){
 			return;
 		}
@@ -802,6 +805,7 @@ public:
 		exitPhysics();
 		PlasticityExampleBrowser::getRenderer()->removeAllInstances();
 		initParameterUi();
+		resetClocks();
 		initPhysics();
 	}
 	void addFixedConstraint(btAlignedObjectArray<btRigidBody*> ha){
@@ -1027,10 +1031,10 @@ void DemolisherDemo::reinit(){
 	steelArea = 0.001;
 	maxPlasticStrain = 0.1;
 	maxPlasticRotation = 1;
-	driveClock.reset();
-	speedometerUpdated = 0;
 	gameBindings = true;
+	resetClocks();
 }
+
 void DemolisherDemo::resetHandler(Gwen::Controls::Base* control){
 	demo->reinit();
 	restartHandler(control);
@@ -1177,13 +1181,18 @@ tr.setOrigin(btVector3(0,-3,0));
 
 	btCompoundShape* compound = new btCompoundShape();
 	m_collisionShapes.push_back(compound);
+	btScalar sy = suspensionRestLength - 10 * carMass / suspensionStiffness / 4;
+	if (sy < 0){
+		sy = 0.06;
+	}
 	btTransform localTrans;
 	localTrans.setIdentity();
-	//localTrans effectively shifts the center of mass with respect to the chassis
-	localTrans.setOrigin(btVector3(0,suspensionRestLength+yhl,0));
-	compound->addChildShape(localTrans,chassisShape);
-	tr.setOrigin(btVector3(0,0.f,0));
-	m_carChassis = localCreateRigidBody(carMass,tr,compound);//chassisShape);	
[…]

### 2016-02-15 examples/CharpyDemo/DemolisherDemo.cpp
bridge started in demolisher

@@ -85,6 +85,7 @@ public:
 	btScalar lsx, lsy, lsz;
 	int lpc;
 	btScalar breakingImpulseThreshold;
+	btScalar bridgeSteelScale=10,steelScale;
 	btScalar density;
 	btScalar carMass;
 	btScalar suspensionStiffness, suspensionMaxForce;
@@ -822,7 +823,7 @@ public:
 			btGeneric6DofConstraint *sc =
 				new btGeneric6DofConstraint(*ha[i], *ha[i + 1],
 				tra, trb, true);
-			sc->setBreakingImpulseThreshold(breakingImpulseThreshold);
+			sc->setBreakingImpulseThreshold(steelScale*breakingImpulseThreshold);
 			m_dynamicsWorld->addConstraint(sc, disableCollisionsBetweenLinkedBodies);
 			for (int i = 0; i < 6; i++){
 				sc->setLimit(i, 0, 0); // make fixed
@@ -845,9 +846,9 @@ public:
 		btVector3 cpos(halfLength, 0, 0);
 		tra.setOrigin(cpos);
 		trb.setOrigin(-cpos);
-		btScalar k0(E*steelArea/ 0.2); 
-		btScalar k1(E*steelArea*lsy/2.5);
-		btScalar k2(E*steelArea*lsz/2.5);
+		btScalar k0(E*steelArea*steelScale / 0.2);
+		btScalar k1(E*steelArea*steelScale*lsy / 2.5);
+		btScalar k2(E*steelArea*steelScale*lsz / 2.5);
 		btScalar m(fy / E);
 		btScalar w0(k0*m);
 		btScalar w1(k1*fy/E);
@@ -1017,7 +1018,7 @@ void DemolisherDemo::reinit(){
 	defaultBreakingForce = 1000;
 	wheelFriction = 0.8;
 	lpc = 5;
-	lsx = 10;
+	lsx = 30;
 	lsy = 3;
 	lsz = 2;
 	density = 2000;
@@ -1211,6 +1212,7 @@ tr.setOrigin(btVector3(0,-3,0));
 
 	/// create load parts
 	{
+		steelScale = btScalar(1);
 		btAlignedObjectArray<btRigidBody*> ha;
 		btScalar xlen = lsx / lpc;
 		btCollisionShape* loadShape = new btBoxShape(btVector3(xlen/2, lsy / 2, lsz / 2));
[…]

### 2016-03-03 examples/CharpyDemo/DemolisherDemo.cpp
demolisher gateSupport refactored

@@ -92,7 +92,8 @@ public:
 	// vehicle body measures
 	btScalar yhl,xhl,zhl;
 	int lpc;
-	btScalar breakingImpulseThreshold;
+	btScalar breakingImpulseThreshold, breakingSpeed;
+	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
 	btScalar bridgeSteelScale,gateSteelScale,fenceSteelScale;
 	btScalar density;
 	btScalar defaultCarMass = 50000;
@@ -422,7 +423,7 @@ public:
 	void setLsy(Gwen::Controls::Base* control);
 	void setLsz(Gwen::Controls::Base* control);
 	void setDensity(Gwen::Controls::Base* control);
-	void setBreakingImpulseThreshold(Gwen::Controls::Base* control);
+	void setBreakingSpeed(Gwen::Controls::Base* control);
 	void setCarMass(Gwen::Controls::Base* control);
 	void setDefaultBreakingForce(Gwen::Controls::Base* control);
 	void setMaxEngineForce(Gwen::Controls::Base* control);
@@ -628,14 +629,14 @@ public:
 		place(gc);
 		gc->onReturnPressed.Add(pPage, &DemolisherDemo::setDensity);
 	}
-	void addBreakingImpulseThreshold(){
-		addLabel("breakingImpulse");
+	void addBreakingSpeed(){
+		addLabel("breakingSpeed");
 		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
-		std::string text = uif(breakingImpulseThreshold, "%.0f");
-		gc->SetToolTip("breakingImpulseThreshold for load parts [Ns]");
+		std::string text = uif(breakingSpeed, "%.2f");
+		gc->SetToolTip("breakingSpeed [m/s]");
 		gc->SetText(text);
 		place(gc);
-		gc->onReturnPressed.Add(pPage, &DemolisherDemo::setBreakingImpulseThreshold);
+		gc->onReturnPressed.Add(pPage, &DemolisherDemo::setBreakingSpeed);
 	}
 	void addMaxPlasticStrain(){
 		addLabel("maxPlasticStrain");
@@ -803,7 +804,10 @@ public:
 		switch (constraintType){
 		case Impulse:
 			addCollisionBetweenLinkedBodies();
[…]

### 2016-03-04 examples/CharpyDemo/DemolisherDemo.cpp
debugging helpers to demolisher

@@ -79,7 +79,8 @@ public:
 	class btDefaultCollisionConfiguration* m_collisionConfiguration;
 
 	class btTriangleIndexVertexArray*	m_indexVertexArrays;
-
+	btDantzigSolver* mlcp=0;
+	btMLCPSolver* sol=0;
 	btVector3*	m_vertices;
 	btScalar	maxEngineForce;//this should be engine/velocity dependent
 	btScalar	defaultBreakingForce;
@@ -91,6 +92,8 @@ public:
 	btScalar tolerance;
 	// vehicle body measures
 	btScalar yhl,xhl,zhl;
+	long stepCount;
+	btScalar stepTime;
 	int lpc;
 	btScalar breakingImpulseThreshold, breakingSpeed;
 	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
@@ -445,6 +448,7 @@ public:
 	Gwen::Controls::Button* pauseButton;
 	Gwen::Controls::Label *db11, *db12, *db13, *db14, *db15, *db16;
 	Gwen::Controls::Label *db21, *db22, *db23;
+	Gwen::Controls::Label *db31, *db32;
 	int fps = 0;
 	int kmph = 0;
 	int mps = 0;
@@ -468,6 +472,11 @@ public:
 		sprintf_s(buffer, UIF_SIZE, "%9.0f ", -dragForce);
 		str = std::string(buffer);
 		db23->SetText(str);
+		if (useMCLPSolver && sol){
+			sprintf_s(buffer, UIF_SIZE, "%5d ", sol->getNumFallbacks());
+			str = std::string(buffer);
+			db32->SetText(str);
+		}
 	}
 	void addCollisionBetweenLinkedBodies(){
 		Gwen::Controls::Label* label = addLabel("disableCollisions");
@@ -517,6 +526,8 @@ public:
 		db21 = new Gwen::Controls::Label(pPage);
 		db22 = new Gwen::Controls::Label(pPage);
 		db23 = new Gwen::Controls::Label(pPage);
+		db31 = new Gwen::Controls::Label(pPage);
+		db32 = new Gwen::Controls::Label(pPage);
 		updateDashboards();
 		db11->SizeToContents();
 		db11->SetPos(gx, gy);
@@ -540,6 +551,14 @@ public:
 		db22->SetPos(gx + wxi, gy);
 		db23->SizeToContents();
 		db23->SetPos(gx + 2 * wxi, gy);
+		gy += gyInc;
+		if (useMCLPSolver){
[…]

### 2016-03-05 pdocs/thesis/bullet-plasticity.tex
small fixes in pdocs

@@
 
-Main parameters  and corresponding fields in \bullet 
+Main parameters  and corresponding fields in \bullet\  
  are described in table \ref{tab:constraintParameters}.
@@
 {\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
-$J_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
+$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
+$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
 $v$ & linear velocity & \\ \hline

### 2016-03-05 examples/CharpyDemo/DemolisherDemo.cpp
vehicle is optional at code level in demolisher

@@ -61,7 +61,7 @@ public:
 	{
 		return m_dynamicsWorld;
 	}
-	btRigidBody* m_carChassis;
+	btRigidBody* m_carChassis=0;
 	btRigidBody* localCreateRigidBody(btScalar mass, const btTransform& worldTransform, btCollisionShape* colSape);
 
 	GUIHelperInterface* m_guiHelper;
@@ -135,9 +135,9 @@ public:
 	enum CarPosition { Gate, Fence, Bridge};
 	Constraint constraintType;
 	btRaycastVehicle::btVehicleTuning	m_tuning;
-	btVehicleRaycaster*	m_vehicleRayCaster;
-	btRaycastVehicle*	m_vehicle;
-	btCollisionShape*	m_wheelShape;
+	btVehicleRaycaster*	m_vehicleRayCaster=0;
+	btRaycastVehicle*	m_vehicle=0;
+	btCollisionShape*	m_wheelShape=0;
 
 	float		m_cameraHeight;
 
@@ -1251,6 +1251,88 @@ public:
 			break;
 		}
 	}
+	void addVehicle(){
+		btCollisionShape* chassisShape = new btBoxShape(btVector3(xhl, yhl, zhl));
+		m_collisionShapes.push_back(chassisShape);
+		btCompoundShape* compound = new btCompoundShape();
+		m_collisionShapes.push_back(compound);
+		btTransform localTrans;
+		localTrans.setIdentity();
+		localTrans.setOrigin(btVector3(0, yhl, 0));
+		compound->addChildShape(localTrans, chassisShape);
+		m_carChassis = localCreateRigidBody(carMass, getCarTransform(), compound);
+		m_wheelShape = new btCylinderShapeX(btVector3(wheelWidth, wheelRadius, wheelRadius));
+		m_guiHelper->createCollisionShapeGraphicsObject(m_wheelShape);
+		int wheelGraphicsIndex = m_wheelShape->getUserIndex();
+		const float position[4] = { 0, 10, 10, 0 };
+		const float quaternion[4] = { 0, 0, 0, 1 };
+		const float color[4] = { 0.4, 0.4, 0.4, 1 };
+		const float scaling[4] = { 1, 1, 1, 1 };
+		for (int i = 0; i<4; i++)
+		{
+			m_wheelInstances[i] = m_guiHelper->registerGraphicsInstance
+				(wheelGraphicsIndex, position, quaternion, color, scaling);
+		}
[…]

### 2016-03-06 examples/CharpyDemo/DemolisherDemo.cpp
vehicle and objects can be disabled by setting friction or steel%  <0

@@ -264,7 +264,12 @@ public:
 		lastClock = currentClock;
 		currentClock = now;
 		lastLocation = currentLocation;
-		currentLocation = m_carChassis->getCenterOfMassPosition();
+		if (m_carChassis){
+			currentLocation = m_carChassis->getCenterOfMassPosition();
+		}
+		else{
+			currentLocation = getCarTransform().getOrigin();
+		}
 		updateCameraLocation();
 		float timeDelta = now - speedometerUpdated;
 		if (timeDelta < updateInterval){
@@ -1138,7 +1143,6 @@ public:
 		}
 	}
 	void addBridge(){
-		bridgeSupportY = 1.2*lsy;
 		btAlignedObjectArray<btRigidBody*> ha;
 		btScalar xlen = bridgeLsx / lpc;
 		btCollisionShape* partShape = 
@@ -1569,14 +1573,14 @@ void DemolisherDemo::exitPhysics()
 		delete shape;
 	}
 	m_collisionShapes.clear();
-
-	delete m_indexVertexArrays;
-	delete m_vertices;
-
-	//delete dynamics world
-	delete m_dynamicsWorld;
-	m_dynamicsWorld=0;
-
+	if (m_indexVertexArrays){
+		delete m_indexVertexArrays;
+		m_indexVertexArrays = 0;
+	}
+	if (m_vertices){
+		delete m_vertices;
+		m_vertices = 0;
+	}
 	if (m_vehicleRayCaster){
 		delete m_vehicleRayCaster;
 		m_vehicleRayCaster = 0;
@@ -1587,21 +1591,26 @@ void DemolisherDemo::exitPhysics()
 		delete m_wheelShape;
 		m_wheelShape = 0;
 	}
-	//delete solver
-	delete m_constraintSolver;
-	m_constraintSolver=0;
-
-	//delete broadphase
-	delete m_overlappingPairCache;
-	m_overlappingPairCache=0;
-
-	//delete dispatcher
-	delete m_dispatcher;
-	m_dispatcher=0;
-
-	delete m_collisionConfiguration;
-	m_collisionConfiguration=0;
-
+	if (m_constraintSolver){
+		delete m_constraintSolver;
+		m_constraintSolver = 0;
+	}
+	if (m_dynamicsWorld){
+		delete m_dynamicsWorld;
[…]

### 2016-03-30 pdocs/thesis/introduction.tex
literature references for pdocs

@@
 E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
-Already \cite{cg1988} used quite similar principles as in this work to handle plasticity.
+
+Already \cite{cg1988} used quite similar principles as in this work to handle plasticity and there are quite a few further
+papers that cite \cite{cg1988}. E.g. following papers provide alternative solutions to similar kind of simulations as 
+are handled in this work 
+\begin{itemize}
+\item \cite{gibson1997survey} presents quite extensive study on deformable modelling in computer graphics
+\item \cite{o2002graphical} provides excellent paper on ductile behaviour of material
+\item \cite{muller2004point} presents a method for modeling and animating of 
+elastic and plastic objects in realtime using point based animation
+\item \cite{irving2004invertible} introduces algortithm for  finite element simulation of elastoplastic solids with very 
+large deformations in reasonable time 
+\end{itemize}
 

### 2016-03-31 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.h
remove elasticPlastic2constraint after disabling it

@@ -48,7 +48,8 @@ http://gimpact.sf.net
 #include "BulletDynamics/Dynamics/btActionInterface.h"
 #include "BulletDynamics/Dynamics/btRigidBody.h"
 class btRigidBody;
-
+#include "BulletDynamics/Dynamics/btDynamicsWorld.h"
+class btDynamicsWorld;
 
 #ifdef BT_USE_DOUBLE_PRECISION
 #define bt6DofElasticPlastic2ConstraintData2		bt6DofElasticPlastic2ConstraintDoubleData2
@@ -299,8 +300,11 @@ public:
 	///btActionInterface interface
 	virtual void updateAction(btCollisionWorld* collisionWorld, btScalar step)
 	{
-		(void)collisionWorld;
 		updatePlasticity(jointFeedback);
+		if (!isEnabled()){
+			btDynamicsWorld *dw = (btDynamicsWorld *)collisionWorld;
+			dw->removeConstraint(this);
+		}
 	}
 	///btActionInterface interface
 	void	debugDraw(btIDebugDraw* debugDrawer);

### 2016-03-31 examples/CharpyDemo/DemolisherDemo.cpp
more profiling in Demolisher and elasticPlastic2constraint

@@ -1869,12 +1869,16 @@ void DemolisherDemo::renderScene()
 	updatePauseButtonText();
 	setDumpFilename();
 	{
-		BT_PROFILE("CharpyDemo::showMessage");
+		BT_PROFILE("DemolisherDemo::showMessage");
 		showMessage();
 	}
 	updateDashboards();
-	m_guiHelper->render(m_dynamicsWorld);
+	{
+		BT_PROFILE("m_guiHelper::render");
+		m_guiHelper->render(m_dynamicsWorld);
+	}
 	if (stepCount > syncedStep){
+		BT_PROFILE("m_guiHelper::syncPhysicsToGraphics");
 		syncWheelsToGraphics();
 		m_guiHelper->syncPhysicsToGraphics(m_dynamicsWorld);
 		syncedStep = stepCount;

### 2016-03-31 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp

@@ -43,6 +43,7 @@ http://gimpact.sf.net
 #include "BulletDynamics/Dynamics/btRigidBody.h"
 #include "LinearMath/btTransformUtil.h"
 #include "LinearMath/btIDebugDraw.h"
+#include "LinearMath/btQuickprof.h"
 #include <new>
 #include "../plasticity/PlasticityData.h"
 
@@ -492,6 +493,7 @@ void bt6DofElasticPlastic2Constraint::getInfo1 (btConstraintInfo1* info)
 
 void bt6DofElasticPlastic2Constraint::getInfo2 (btConstraintInfo2* info)
 {
+	BT_PROFILE("bt6DofElasticPlastic2Constraint::getInfo2");
 	const btTransform& transA = m_rbA.getCenterOfMassTransform();
 	const btTransform& transB = m_rbB.getCenterOfMassTransform();
 	const btVector3& linVelA = m_rbA.getLinearVelocity();
@@ -673,6 +675,7 @@ int bt6DofElasticPlastic2Constraint::get_limit_motor_info2(
 	const btTransform& transA,const btTransform& transB,const btVector3& linVelA,const btVector3& linVelB,const btVector3& angVelA,const btVector3& angVelB,
 	btConstraintInfo2 *info, int row, btVector3& ax1, int rotational,int rotAllowed, btScalar maxForce)
 {
+	BT_PROFILE("bt6DofElasticPlastic2Constraint::get_limit_motor_info2");
 	bool useBcc = true;
 	int count = 0;
 	int srow = row * info->rowskip;
@@ -1241,6 +1244,7 @@ int bt6DofElasticPlastic2Constraint::getMaxRatioDof(){
 	return m_maxRatioDof;
 }
 void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
+	BT_PROFILE("updatePlasticity");
 	if (!isEnabled()){
 		return;
 	}

### 2016-04-01 examples/CharpyDemo/DemolisherDemo.cpp
starting point near Bridge tuned in Demolisher

@@ -1110,7 +1110,7 @@ public:
 			{
 				btQuaternion q(SIMD_HALF_PI,0,0);
 				tr.setRotation(q);
-				xStart = -bridgeLsx / 2 ;
+				xStart = -bridgeLsx / 2-zhl ;
 				zStart = bridgeZ;
 				yStart += bridgeSupportY + bridgeLsy;
 			}

### 2016-06-10 pdocs/thesis/article-introduction.tex
pdocs/article

\section{Introduction}

Physics engine is computer software that provides an approximate simulation of physical system. 
E.g. ODE - \cite{ode}, \bullet - \cite{bullet}, Box2D - \cite{box2d} are open source physics engines that are  widely used in 
areas where fast solutions are required. Typical industries are games and film productions.

Computational methods used in physics engines are divided to modules that handle collision detection and modules that handle simulation.
In this work focus is on solving new types of contraints between rigid objects.
Detailed description of various components can be found in e.g. \cite{erleben.thesis}.

Plasticity is not typically taken into account in gaming solutions. 
Breaking of various objects typically takes place based on collision or impulse.
Theory for handling of plasticity has been presented already in \cite{cg1988}. \cite{muller2004point} presents a method for modeling and animating of elastic and plastic objects in realtime using point based animation but it is not widely used.

Goal of this work is to find method that can be used in widely used physics engines to make simulation more realistic
by connecting rigid objects by  spring based constraints which have additional properties for handling plasticity.
Main feature is that plastic deformation takes place if force or moment exceeds given limit and deformation absorbs energy.
This is done by reusing motor code that is already available in most physics engines.

Methods presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work. 
This kind of metodology also opens quite large area of combining old structural analysis methods to modern simulation frameworks.

### 2016-06-11 pdocs/thesis/article-introduction.tex
pdocs/article

@@
 Computational methods used in physics engines are divided to modules that handle collision detection and modules that handle simulation.
-In this work focus is on solving new types of contraints between rigid objects.
+Simulation can further be subdivided into time control, motion solver, constraint solver and collision solver modules.
+Velocity-based formulation is used in constraint based rigid body simulation frameworks as collisions cannot be handled easily in acceleration based formulations.
+Friction is typically taken into account.
+Joints are handled by constraint equations.
 Detailed description of various components  can be found in e.g. \cite{erleben.thesis}.
+In this work focus is on solving new types of joints between rigid objects.
 
 Plasticity is not typically taken into account in gaming solutions. 
-Breaking of various objects typically takes place based on collision or impulse.
+Breaking of various objects typically takes place based on collision or impulse and breaking of steel or 
+reinforced concrete structures does not look realistic.
 Theory for handling of plasticity has been presented already in \cite{cg1988}. 
@@
 by connecting rigid objects by  spring based constraints which have additional properties for handling plasticity.
-Main feature is that plastic deformation takes place if force or moment exceeds given limit and deformation absorbs energy.
-This is done by reusing motor code that is already available in most physics engines.
+Main target is that plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and 
+joint breaks if plastic capacity is exeeced.
 
 Methods presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work. 
+For gaming purposes presented method works best in scenarios where connected parts are relatively heavy. 
+This allows normal integration timestep to be used without stability issues.
 This kind of metodology also opens quite large area of combining old structural analysis methods to modern simulation frameworks.

### 2016-06-12 pdocs/thesis/article-introduction.tex
pdocs/article

@@
 Simulation can further be subdivided into time control, motion solver, constraint solver and collision solver modules.
-Velocity-based formulation is used in constraint based rigid body simulation frameworks as collisions cannot be handled easily in acceleration based formulations.
-Friction is typically taken into account.
-Joints are handled by constraint equations.
+Velocity-based formulation is typically used in constraint based rigid body simulation 
+frameworks as collisions cannot be handled easily in acceleration based formulations. 
+Friction is typically taken into account. Joints are handled by constraint equations.
 Detailed description of various components  can be found in e.g. \cite{erleben.thesis}.
-In this work focus is on solving new types of joints between rigid objects.
 
@@
 Theory for handling of plasticity has been presented already in \cite{cg1988}. 
-\cite{muller2004point} presents a method for modeling and animating of elastic and plastic objects in realtime using point based animation but it is not widely used.
+\cite{muller2004point} and \cite{muller2005meshless} present a method for 
+modeling and animating of elastic and plastic objects in realtime using point based animation but it is not widely used.
+On major issue is collision handling of deformable objects.
 
-Goal of this work is to find method that can be used in widely used physics engines to make simulation more realistic
-by connecting rigid objects by  spring based constraints which have additional properties for handling plasticity.
+Goal of this work is to find method that can be used in widely used rigid body based physics engines to make simulation more realistic.
 Main target is that plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and 
-joint breaks if plastic capacity is exeeced.
+joint breaks if plastic capacity is exceeced. Suggested method is based on using joint motors to model plasticity. 
+\cite[p.~90]{erleben.thesis} suggests similar method for modelling friction in joints.
+Adjacent objects are connected by motors. Motor limits are calculated based on plastic section modulus.
+Joint breaking is simulated by summing plastic deformation and comparing it to predefined material based limit.
+Elastic part of deformation is modelled by spring based simulation which is based on modification of existing constraint in \bullet.
 

### 2016-06-14 examples/CharpyDemo/DemolisherDemo.cpp
pole started in demolisher

@@ -89,7 +89,8 @@ public:
 	btScalar bridgeLsx, bridgeLsy, bridgeLsz, bridgeSupportY;
 	bool twoSupportsForGate = false;
 	btScalar gateLsx, gateLsy, gateLsz;
-	btScalar bridgeZ = 40, gateZ = -20, fenceZ=5, bridgeSupportX=1;
+	btScalar poleLsx, poleLsy;
+	btScalar poleZ=60, bridgeZ = 40, gateZ = -20, fenceZ=5, bridgeSupportX=1;
 	btScalar tolerance;
 	// vehicle body measures
 	btScalar yhl,xhl,zhl;
@@ -100,7 +101,7 @@ public:
 	int lpc;
 	btScalar breakingImpulseThreshold, breakingSpeed;
 	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
-	btScalar bridgeSteelScale,gateSteelScale,fenceSteelScale;
+	btScalar bridgeSteelScale,gateSteelScale,fenceSteelScale, poleSteelScale;
 	btScalar density;
 	btScalar defaultCarMass = 50000;
 	btScalar carMass;
@@ -487,6 +488,7 @@ public:
 	void setFenceSteelScale(Gwen::Controls::Base* control);
 	void setBridgeSteelScale(Gwen::Controls::Base* control);
 	void setGateSteelScale(Gwen::Controls::Base* control);
+	void setPoleSteelScale(Gwen::Controls::Base* control);
 	void setMaxPlasticStrain(Gwen::Controls::Base* control);
 	void setMaxPlasticRotation(Gwen::Controls::Base* control);
 	void setGameBindings(Gwen::Controls::Base* control);
@@ -788,6 +790,15 @@ public:
 		place(gc);
 		gc->onReturnPressed.Add(pPage, &DemolisherDemo::setBridgeSteelScale);
 	}
+	void addPoleSteelScale(){
+		addLabel("pole steel%");
+		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
+		std::string text = uif(poleSteelScale * 100, "%.3f");
+		gc->SetToolTip("Area of enforcement steel [%] [0-100]");
+		gc->SetText(text);
+		place(gc);
+		gc->onReturnPressed.Add(pPage, &DemolisherDemo::setPoleSteelScale);
+	}
 	void addCarMass(){
[…]

### 2016-06-16 pdocs/thesis/article-introduction.tex
pdocs/article changes based on Aki's comments

--- 2016-06-10 pdocs/thesis/article-introduction.tex
@@
 
-Physics engine is computer software that provides an approximate simulation of physical system. 
-E.g. ODE - \cite{ode}, \bullet - \cite{bullet}, Box2D - \cite{box2d} are open source physics engines that are  widely used in 
-areas where fast solutions are required. Typical industries are games and film productions.
+The application of modern computer game techniques enables the description of complex dynamic systems such as military vehicles with a high level of detail while still solving the equations in real-time.
+Film production and war games, in particular, is a key area that have benefited from simulation technology.
+In practice, games are often accomplished using an open-source platform such like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
 
-Computational methods used in physics engines are divided to modules that handle collision detection and modules that handle simulation.
-In this work focus is on solving new types of contraints between rigid objects.
-Detailed description of various components can be found in e.g. \cite{erleben.thesis}.
+Computational methods used in physics engines are divided to modules that handle collision detection and contact description and modules that handle solution of equations in real-time.
+Equations need to be solved can further be subdivided to be associated to motion, constraints and collisions.
+Velocity-based formulation is typically used in constraint based rigid body simulation frameworks as collisions cannot be handled easily in acceleration based formulations.
+Friction is typically taken into account and mechanical joints are handled by constraint equations.
+Detailed description of various components can be found in e.g.
+\cite{erleben.thesis}.
 
@@
 Breaking of various objects typically takes place based on collision or impulse.
+Nevertheless, breaking of steel or reinforced concrete structures using this approach is not appropriate making a simulation to look unrealistic.
 Theory for handling of plasticity has been presented already in \cite{cg1988}.
-\cite{muller2004point} presents a method for modeling and animating of elastic and plastic objects in realtime using point based animation but it is not widely used.
+\cite{muller2004point} and \cite{muller2005meshless} present a method for modeling and animating of elastic and plastic objects in real-time using point based animation.
+This approach is not been widely used in simulation applications.
+On major issue is collision handling of deformable objects.
 
-Goal of this work is to find method that can be used in widely used physics engines to make simulation more realistic
-by connecting rigid objects by  spring based constraints which have additional properties for handling plasticity.
-Main feature is that plastic deformation takes place if force or moment exceeds given limit and deformation absorbs energy.
-This is done by reusing motor code that is already available in most physics engines.
+This study will introduce an approach to account plastic deformation in game applications.
+In the introduced method, the plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced.
+The approach is based on using joint motors to model plasticity.
+\cite[p.~90]{erleben.thesis} suggests similar method for modelling friction in joints.
+Adjacent objects are connected by motors.
+Motor power production limits are estimated based on plastic section modulus.
+Joint breaking is accounted by summing plastic deformation and comparing it to predefined material based limit.
+Elastic part of deformation is modelled by employing spring description which is based on modification of existing constraint in \bullet.
 
-Methods presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work. 
-This kind of metodology also opens quite large area of combining old structural analysis methods to modern simulation frameworks.
+Approach presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work.
+For gaming purposes presented method works best in scenarios where connected parts are relatively heavy.
+This allows normal integration timestep to be used without stability issues.
+This kind of metodology also opens large area of combining old structural analysis methods to modern simulation frameworks.

### 2016-06-17 pdocs/thesis/article-introduction.tex
line breaks added
section 2 for article

@@
 
-The application of modern computer game techniques enables the description of complex dynamic systems such as military vehicles with a high level of detail while still solving the equations in real-time.
+The application of modern computer game techniques enables the description of complex dynamic systems 
+such as military vehicles with a high level of detail while still solving the equations in real-time.
 Film production and war games, in particular, is a key area that have benefited from simulation technology. 
-In practice, games are often accomplished using an open-source platform such like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
+In practice, games are often accomplished using an open-source platform such 
+like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
 
-Computational methods used in physics engines are divided to modules that handle collision detection and contact description and modules that handle solution of equations in real-time.
-Equations need to be solved can further be subdivided to be associated to motion, constraints and collisions.
-Velocity-based formulation is typically used in constraint based rigid body simulation frameworks as collisions cannot be handled easily in acceleration based formulations.
+Computational methods used in physics engines are divided to modules that handle collision detection and 
+contact description and modules that handle solution of equations in real-time. Equations need to be 
+solved can further be subdivided to be associated to motion, constraints and collisions. 
+Velocity-based formulation is typically used in constraint based rigid body simulation. 
 Friction is typically taken into account and mechanical joints are handled by constraint equations.
-Detailed description of various components can be found in e.g.
-\cite{erleben.thesis}.
+Detailed description of various components can be found in e.g. \cite{erleben.thesis}.
 
@@
 Breaking of various objects typically takes place based on collision or impulse.
-Nevertheless, breaking of steel or reinforced concrete structures using this approach is not appropriate making a simulation to look unrealistic.
-Theory for handling of plasticity has been presented already in \cite{cg1988}.
-\cite{muller2004point} and \cite{muller2005meshless} present a method for modeling and animating of elastic and plastic objects in real-time using point based animation.
-This approach is not been widely used in simulation applications.
+Nevertheless, breaking of steel or reinforced concrete structures using this approach 
+is not appropriate making a simulation to look unrealistic. Theory for handling of plasticity 
+has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} 
+present a method for modeling and animating of elastic and plastic objects in real-time using 
+point based animation. This approach is not been widely used in simulation applications.  
 On major issue is collision handling of deformable objects.
@@
 This study will introduce an approach to account plastic deformation in game applications.   
-In the introduced method, the plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced.
-The approach is based on using joint motors to model plasticity.
-\cite[p.~90]{erleben.thesis} suggests similar method for modelling friction in joints.
-Adjacent objects are connected by motors.
+In the introduced method, the plastic deformation takes place if force or moment exceeds given 
+limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced. 
+The approach is based on using joint motors to model plasticity. \cite[p.~90]{erleben.thesis} 
+suggests similar method for modelling friction in joints. Adjacent objects are connected by motors. 
 Motor power production limits are estimated based on plastic section modulus. 
-Joint breaking is accounted by summing plastic deformation and comparing it to predefined material based limit.
-Elastic part of deformation is modelled by employing spring description which is based on modification of existing constraint in \bullet.
+Joint breaking is accounted by summing plastic deformation and comparing it to 
+predefined material based limit. Elastic part of deformation is modelled by employing 
+spring description which is based on modification of existing constraint in \bullet.
 
-Approach presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work.
-For gaming purposes presented method works best in scenarios where connected parts are relatively heavy.
-This allows normal integration timestep to be used without stability issues.
-This kind of metodology also opens large area of combining old structural analysis methods to modern simulation frameworks.
+Approach presented in this work can be used in gaming industry to provide more realistic 
+simulations without significant extra work. For gaming purposes presented method works 
+best in scenarios where connected parts are relatively heavy. This allows normal 
+integration timestep to be used without stability issues. 
+This kind of metodology also opens large area of combining old structural analysis
+methods to modern simulation frameworks.
+
+\section{Using physics engine for structural analysis}
+
+Velocity-based formulation is not typically used in structural or mechanical engineering.
+ \cite[p.~45]{erleben.thesis} provides reasoning and theoretical details. 
+In structural finite element analysis solution method is selected based on needed features.
+In most cases static small displacement solution using displacement based boudary conditions is used.
+For large displacement static analysis analysis loading is applied in substeps and 
+displacements are used to update element mesh.
+Further enhancements are material nonlinearity and dynamic analysis.
+In scope of this work physics engine provides dynamics analysis 
+with large displacements but elements are rigid objects connected by flexible joints.
+
+Material plasticity has not been taken into account in games and some basics are provided here.
+Typical stress strain curve of ductile steel is shown in \ref{fig:areas}.
+Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
+Straing hardening is taken into account mainly by assuming that plasticity in bending expands.
+This can be seen e.g. by bending paperclip. It does not break at low angles but can take few full bends.
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (Y) at (0.2,4);
+\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
+\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
+\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
+\draw[dashed](0,4) -- (Y);
+\node at (-0.2,4) [align=right] {$f_y$};
+\end{tikzpicture}
+\caption{Stress-strain curve of ductile steel (not to scale).}
+\label{fig:areas}
+\end{figure}
+
+Difference between elastic and plastic section modulus is shown in \ref{fig:wp}. 
+If stress is below yield limit, stress and strain are linear within cross section.
+If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
+so plastic section modulus is higher than elastic section modulus.
+Elastic part is often ignored in this work as displacements due to to elastic deformation are very small.
+Basic idea in this work can be tested with any framework having motors and hinge constraints.
+Setting target velocity of motor to zero and limiting maximum motor impulse to plastic moment 
+multiplied by timestep.
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (S) at (2.5,5);
+\draw (0,5) -- (4,5) ;
+\draw (0,0) -- (4,0) ;
+\draw (2,0) -- (2,5) ;
+\draw (1.5,0) -- (S); 
+\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
+\end{tikzpicture}
+\hspace{1cm}
+\begin{tikzpicture}
+\coordinate (S) at (3,5);
+\draw (0,5) -- (4,5) ;
+\draw (0,0) -- (4,0) ;
+\draw (2,0) -- (2,5) ;
+\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
+\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
+\end{tikzpicture}
+\caption{Stress distribution under elastic and plastic loads.}
+\label{fig:wp}
+\end{figure}
+
+Constraint processing in \bullet\ is based on ODE, \cite{ode}.
+Mathematical background and detailed examples are available by \cite{ode.joints}.
+Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
+\ref{eq:lambdaHigh} 
+are created for each constraint.
+
+\begin{equation} \label{eq:constraintEquation}
+J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
+\end{equation}
+
+\begin{equation} \label{eq:lambdaLow}
+\lambda \geq l
+\end{equation}
+
+\begin{equation} \label{eq:lambdaHigh}
+\lambda \leq h
+\end{equation}
+
+Main parameters  and corresponding fields in \bullet\  
+ are described in table \ref{tab:constraintParameters}.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|c| l| l|}
+\hline
+{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
+$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
+$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
+$v$ & linear velocity & \\ \hline
+$\omega$ & angular velocity & \\ \hline
+$c$        &  right side vector   & m\_constraintError \\ \hline
+$C$  & constraint force mixing & cfm \\  \hline
+$\lambda$ & constraint force &  \\ \hline
+$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
+$h$ & high limit for constraint force & m\_upperLimit \\ \hline
+\end {tabular}
+\end{center}
+\caption {Constraint parameters} \label{tab:constraintParameters} 
+\end {table}

### 2016-06-18 pdocs/thesis/article-introduction.tex
first draft of section 2 for article

--- 2016-06-16 pdocs/thesis/article-introduction.tex
@@
 This study will introduce an approach to account plastic deformation in game applications.   
-In the introduced method, the plastic deformation takes place if force or moment exceeds given limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced.
-The approach is based on using joint motors to model plasticity.
-\cite[p.~90]{erleben.thesis} suggests similar method for modelling friction in joints.
-Adjacent objects are connected by motors.
+In the introduced method, the plastic deformation takes place if force or moment exceeds given 
+limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced. 
+The approach is based on using joint motors to model plasticity. \citet[p.~90]{erleben.thesis} 
+suggests similar method for modelling friction in joints. Adjacent objects are connected by motors. 
 Motor power production limits are estimated based on plastic section modulus. 
-Joint breaking is accounted by summing plastic deformation and comparing it to predefined material based limit.
-Elastic part of deformation is modelled by employing spring description which is based on modification of existing constraint in \bullet.
+Joint breaking is accounted by summing plastic deformation and comparing it to 
+predefined material based limit. Elastic part of deformation is modelled by employing 
+spring description which is based on modification of existing constraint in \bullet.
 
-Approach presented in this work can be used in gaming industry to provide more realistic simulations without significant extra work.
-For gaming purposes presented method works best in scenarios where connected parts are relatively heavy.
-This allows normal integration timestep to be used without stability issues.
-This kind of metodology also opens large area of combining old structural analysis methods to modern simulation frameworks.
+Approach presented in this work can be used in gaming industry to provide more realistic 
+simulations without significant extra work. For gaming purposes presented method works 
+best in scenarios where connected parts are relatively heavy. This allows normal 
+integration timestep to be used without stability issues. 
+This kind of metodology also opens large area of combining old structural analysis
+methods to modern simulation frameworks.

### 2016-06-18 pdocs/thesis/article-section-2.tex
first draft of section 2 for article

\section{Using physics engine for structural analysis}

Velocity-based formulation is not typically used in structural or mechanical engineering.
 \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details why it is so popular in 
constraint-based rigid body simulation. 
Main reason is that collision handling can be done without additional procedures.
In structural finite element analysis solution method is selected carefully based on needed features.
In most cases static small displacement solution using displacement based boudary conditions is used.
For large displacement static analysis analysis loading is applied in substeps and 
displacements are used to update element mesh.
Further enhancements are material nonlinearity and dynamic analysis.
Physics engine provides dynamic analysis with large displacements.

Material plasticity has typically taken into account in games by using suitable coefficient of restitution.
This provides reasonable means to simulate loss of energy in collisions.
Simulation of breaking of objects made of ductile material can be made more realistic by splitting rigid objects
to multiple parts which are connected by energy absorbing joints.
As stress-strain curve may not be familiar to all readers some basics are provided here.
Typical stress strain curve of ductile steel is shown in \ref{fig:areas}.
Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
Straing hardening is taken into account mainly by assuming that plasticity in bending expands.
Material that starts to yield first is hardened and yielding moves slightly.
This can be seen e.g. by bending paperclip. It does not break at low angles but can take few full bends.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (Y) at (0.2,4);
\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
\draw[dashed](0,4) -- (Y);
\node at (-0.2,4) [align=right] {$f_y$};
\end{tikzpicture}
\caption{Stress-strain curve of ductile steel (not to scale).}
\label{fig:areas}
\end{figure}

Difference between elastic and plastic section modulus is shown in \ref{fig:wp}. 
If stress is below yield limit, stress and strain are linear within cross section.
If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
so plastic section modulus is higher than elastic section modulus.
Elastic part is often ignored in this work as displacements due to to elastic deformation are very small.

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (S) at (2.5,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1.5,0) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
\end{tikzpicture}
\hspace{1cm}
\begin{tikzpicture}
\coordinate (S) at (3,5);
\draw (0,5) -- (4,5) ;
\draw (0,0) -- (4,0) ;
\draw (2,0) -- (2,5) ;
\draw (1,0) -- (1,2.5) -- (3,2.5) -- (S); 
\node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
\end{tikzpicture}
\caption{Stress distribution under elastic and plastic loads.}
\label{fig:wp}
\end{figure}

Basic idea in this work can be tested with any framework having motors and hinge constraints.
This can be done by setting target velocity of motor to zero and limiting maximum motor impulse to plastic moment 
multiplied by timestep.

Further enhancements were created and tested by forking \bullet\ source code
and adding new constraints \cite{pbullet}.
Constraint processing in \bullet\ is based on ODE, \cite{ode}.
Mathematical background and detailed examples are available by \cite{ode.joints}.
Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
\ref{eq:lambdaHigh} 
are created for each constraint.

\begin{equation} \label{eq:constraintEquation}
J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
\end{equation}

\begin{equation} \label{eq:lambdaLow}
\lambda \geq l
\end{equation}

\begin{equation} \label{eq:lambdaHigh}
\lambda \leq h
\end{equation}

Main parameters  and corresponding fields in \bullet\  
 are described in table \ref{tab:constraintParameters}.

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$v$ & linear velocity & \\ \hline
$\omega$ & angular velocity & \\ \hline
$c$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\lambda$ & constraint force &  \\ \hline
$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
$h$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\caption {Constraint parameters} \label{tab:constraintParameters} 
\end {table}

### 2016-06-19 examples/CharpyDemo/AxisMapper.cpp
Pole converted to Rod supporting Bridge in Demolisher

@@ -1,0 +1,31 @@
+#include "AxisMapper.h"
+void AxisMapper::init(){
+	if (m_cpos.getX() > 0){
+		initX();
+	}
+	else if (m_cpos.getY()){
+		initY();
+	}
+	else{
+		btAssert(true);
+	}
+}
+
+void AxisMapper::initX(){
+	for (int i = 0; i < 6; i++){
+		m_ami[i] = i;
+	}
+	m_h = m_ylen;
+	m_b = m_zlen;
+}
+
+void AxisMapper::initY(){
+	m_ami[0] = 1;
+	m_ami[1] = 0;
+	m_ami[2] = 2;
+	m_ami[3] = 3;
+	m_ami[4] = 4;
+	m_ami[5] = 5;
+	m_h = m_xlen;
+	m_b = m_zlen;
+}

### 2016-06-19 examples/CharpyDemo/AxisMapper.h

@@ -1,0 +1,35 @@
+#ifndef AXIS_MAPPER_H
+#define AXIS_MAPPER_H
+
+#include "LinearMath/btVector3.h"
+#include "btBulletDynamicsCommon.h"
+
+/**
+* Provides Helper for constraint generation.
+Currently supports x and y axis alignment
+*/
+ATTRIBUTE_ALIGNED16(class) AxisMapper
+{
+protected:
+	btScalar m_xlen, m_ylen, m_zlen;
+	btScalar m_h, m_len, m_b;
+	btVector3 m_cpos;
+	int m_ami[6];
+	void init();
+	void initX();
+	void initY();
+public:
+	BT_DECLARE_ALIGNED_ALLOCATOR();
+	AxisMapper(btScalar xlen, btScalar ylen, btScalar zlen, btVector3& cpos) :
+		m_xlen(xlen),
+		m_ylen(ylen),
+		m_zlen(zlen),
+		m_cpos(cpos){
+		init();
+	};
+	~AxisMapper(){};
+	btScalar getH(){ return m_h; }
+	btScalar getB(){ return m_b; }
+	int* getIndexes(){ return m_ami; }
+};
+#endif //AXIS_MAPPER_H

### 2016-06-19 examples/CharpyDemo/DemolisherDemo.cpp

@@ -18,6 +18,7 @@ Based on ForkLiftDemo by Simo Nikula 2015-
 */
 #define _CRT_SECURE_NO_WARNINGS
 #include "DemolisherDemo.h"
+#include "AxisMapper.h"
 #include "LinearMath/btQuickprof.h"
 #include <stdio.h> 
 #include <string>
@@ -89,8 +90,9 @@ public:
 	btScalar bridgeLsx, bridgeLsy, bridgeLsz, bridgeSupportY;
 	bool twoSupportsForGate = false;
 	btScalar gateLsx, gateLsy, gateLsz;
-	btScalar poleLsx, poleLsy;
-	btScalar poleZ=60, bridgeZ = 40, gateZ = -20, fenceZ=5, bridgeSupportX=1;
+	btScalar rodLsx, rodLsy, rodLsz;
+	btScalar bridgeZ = 40, gateZ = -20, fenceZ=5, bridgeSupportX=1;
+	btScalar rodZ = bridgeZ;
 	btScalar tolerance;
 	// vehicle body measures
 	btScalar yhl,xhl,zhl;
@@ -101,7 +103,7 @@ public:
 	int lpc;
 	btScalar breakingImpulseThreshold, breakingSpeed;
 	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
-	btScalar bridgeSteelScale,gateSteelScale,fenceSteelScale, poleSteelScale;
+	btScalar bridgeSteelScale,gateSteelScale,fenceSteelScale, rodSteelScale;
 	btScalar density;
 	btScalar defaultCarMass = 50000;
 	btScalar carMass;
@@ -137,7 +139,7 @@ public:
 	CommonWindowInterface* window;
 	int m_option;
 	enum Constraint { None = 0, Rigid = 1, Impulse = 2, ElasticPlastic = 3 };
-	enum CarPosition { Gate, Fence, Bridge};
+	enum CarPosition { Gate, Fence, Bridge, Rod};
 	Constraint constraintType;
 	btRaycastVehicle::btVehicleTuning	m_tuning;
 	btVehicleRaycaster*	m_vehicleRayCaster=0;
@@ -488,7 +490,7 @@ public:
 	void setFenceSteelScale(Gwen::Controls::Base* control);
 	void setBridgeSteelScale(Gwen::Controls::Base* control);
 	void setGateSteelScale(Gwen::Controls::Base* control);
[…]

### 2016-06-19 pdocs/thesis/article-section-2.tex
typo fixes and other small tunings

@@
 Main reason is that collision handling can be done without additional procedures.
-In structural finite element analysis solution method is selected carefully based on needed features.
-In most cases static small displacement solution using displacement based boudary conditions is used.
+In structural analysis solution method is selected carefully based on needed features.
+For most complex scenarios finite element method is used.
+In most cases static small displacement solution using displacement based boundary conditions is used.
 For large displacement static analysis analysis loading is applied in substeps and 

### 2016-06-19 pdocs/thesis/article-section-2.tex
article tuning for section 2

@@
-\section{Using physics engine for structural analysis}
+\section{Adding plasticity to physics engine}
+
+In this section some central terms are explained and main differences between traditional structural 
+analysis and physics engines are reviewed.
 
@@
 Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
-Straing hardening is taken into account mainly by assuming that plasticity in bending expands.
+Straing hardening is taken into account in this work mainly by assuming that plasticity in bending expands.
 Material that starts to yield first is hardened and yielding moves slightly.

### 2016-06-30 pdocs/thesis/article-section-2.tex
material model selection clarified in article

@@
 As stress-strain curve may not be familiar to all readers some basics are provided here.
-Typical stress strain curve of ductile steel is shown in \ref{fig:areas}.
+Typical engineering stress-strain curve of ductile steel is shown in \ref{fig:sscurve}.
+E.g. \citet{dowling} provides detailed descriptions of engineering and true stress-strain curves.
 Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
-Straing hardening is taken into account in this work mainly by assuming that plasticity in bending expands.
+
+In this work elastic-fully plastic material model is used in most scenarios.
+It allows realistic simulations for most scenarios.
+Having elastic part allows elastic displacements for slender structures. 
+Elastic part is ignored in method suggested in this work if deformation is related
+to higher frequency than integration stability would allow.
+Strain hardening part of stress-straing curve could probably be taken into account but it has not been tried in this work.
+It should be noted that geometry
+of objects is not updated during analysis and thus engineering stress-strain properties should
+be used even with strain hardening.
+
+Strain hardening is taken into account in this work mainly by assuming that plasticity in bending expands, \citet[p.~672]{dowling}.
 Material that starts to yield first is hardened and yielding moves slightly.
@@
 \end{tikzpicture}
-\caption{Stress-strain curve of ductile steel (not to scale).}
-\label{fig:areas}
+\caption{Engineering stress-strain curve of ductile steel (not to scale).}
+\label{fig:sscurve}
 \end{figure}
@@
 so plastic section modulus is higher than elastic section modulus.
-Elastic part is often ignored in this work as displacements due to to elastic deformation are very small.
 

### 2016-07-01 examples/CharpyDemo/CharpyDemo.cpp
mode7 to elastic case in charpy, noDisplayWait if set<2 ms

@@ -459,6 +459,7 @@ public:
 			case 3: // slender
 				l = btScalar(4.1);
 				spaceBetweenAnvils = btScalar(4.0);
+				setTimeStep = 0.001;
 				break;
 			case 4: // cutting
 				hammerDraft = btScalar(0);
@@ -1094,8 +1095,10 @@ public:
 		return hammerBody->getCenterOfMassTransform().getRotation().getAngle();
 	}
 	void tuneDisplayWait(){
-		if (timeStep<1e-3){
-			displayWait = 2;
+		long smallStepWait = 2;
+		btScalar smallStepLimit(1e-3);
+		if (timeStep<smallStepLimit && setDisplayWait>smallStepWait){
+			displayWait = smallStepWait;
 		}
 		else{
 			displayWait = setDisplayWait;

### 2016-07-01 examples/plasticity/PlasticityExampleEntries.cpp

@@ -62,6 +62,12 @@ static ExampleEntry gDefaultExamples[]=
 	"two objects and 6DofElasticPlastic2Constraint",
 	CharpyDemoCreateFunc, 8),
 	ExampleEntry(2, "elastic",
+	"10 objects using 6DofElasticPlasticConstraint, \
+MLCP solver, \
+no hammer, \
+elastic deflection, l=4 m",
+	CharpyDemoCreateFunc, 102030507),
+	ExampleEntry(2, "elastic2",
 	"10 objects using 6DofElasticPlastic2Constraint, \
 MLCP solver, \
 no hammer, \

### 2016-07-01 examples/CharpyDemo/CharpyDemo.cpp
SI as default in mode7 elastic case in charpy

@@ -48,7 +48,7 @@ bool firstRun = true;
 bool hammerHitsSpecimen;
 btScalar initialStartAngle(1.8);
 btScalar startAngle(initialStartAngle);
-long  initialDisplayWait = 50;
+long  initialDisplayWait = 0;
 long displayWait = initialDisplayWait;
 long setDisplayWait = displayWait;
 btScalar ccdMotionThreshHold(0.001);
@@ -460,6 +460,7 @@ public:
 				l = btScalar(4.1);
 				spaceBetweenAnvils = btScalar(4.0);
 				setTimeStep = 0.001;
+				variableTimeStep = false;
 				break;
 			case 4: // cutting
 				hammerDraft = btScalar(0);

### 2016-07-01 examples/plasticity/PlasticityExampleEntries.cpp

@@ -63,10 +63,10 @@ static ExampleEntry gDefaultExamples[]=
 	CharpyDemoCreateFunc, 8),
 	ExampleEntry(2, "elastic",
 	"10 objects using 6DofElasticPlasticConstraint, \
-MLCP solver, \
+SI solver, \
 no hammer, \
 elastic deflection, l=4 m",
-	CharpyDemoCreateFunc, 102030507),
+	CharpyDemoCreateFunc, 2030507),
 	ExampleEntry(2, "elastic2",
 	"10 objects using 6DofElasticPlastic2Constraint, \
 MLCP solver, \

### 2016-07-02 examples/CharpyDemo/DemolisherDemo.cpp
bt6DofElasticPlasticConstraint added to demolisher

@@ -30,7 +30,9 @@ Based on ForkLiftDemo by Simo Nikula 2015-
 #include "BulletDynamics/MLCPSolvers/btDantzigSolver.h"
 #include "BulletDynamics/MLCPSolvers/btSolveProjectedGaussSeidel.h"
 #include "BulletDynamics/MLCPSolvers/btMLCPSolver.h"
+#include "bt6DofElasticPlasticConstraint.h"
 #include "bt6DofElasticPlastic2Constraint.h"
+#include "btElasticPlasticConstraint.h"
 #include "../plasticity/PlasticityExampleBrowser.h"
 #include "../plasticity/PlasticityData.h"
 #include "../plasticity/PlasticityStatistics.h"
@@ -138,7 +140,7 @@ public:
 	Gwen::Controls::Canvas* canvas;
 	CommonWindowInterface* window;
 	int m_option;
-	enum Constraint { None = 0, Rigid = 1, Impulse = 2, ElasticPlastic = 3 };
+	enum Constraint { None = 0, Rigid = 1, Impulse = 2, ElasticPlastic = 3, ElasticPlastic2 = 4 };
 	enum CarPosition { Gate, Fence, Bridge, Rod};
 	Constraint constraintType;
 	btRaycastVehicle::btVehicleTuning	m_tuning;
@@ -922,6 +924,7 @@ public:
 			addBreakingSpeed();
 			break;
 		case ElasticPlastic:
+		case ElasticPlastic2:
 			addCollisionBetweenLinkedBodies();
 			addMaxPlasticRotation();
 			addMaxPlasticStrain();
@@ -1013,9 +1016,78 @@ public:
 		addElasticPlasticConstraint(ha, cpos, xlen, ylen, zlen, steelScale);
 	}
 	/**
-	TODO, use cpos for selecting global dofs
 	*/
 	void addElasticPlasticConstraint(btAlignedObjectArray<btRigidBody*> ha,
+		btVector3& cpos,
+		btScalar xlen, btScalar ylen, btScalar zlen, btScalar steelScale){
+		int loopSize = ha.size() - 1;
+		btScalar E(200E9);
+		btScalar G(80E9);
+		btScalar fy(200E6);
+		bool limitIfNeeded = true;
+		btScalar damping(0.1);
+		btTransform tra;
+		btTransform trb;
+		tra.setIdentity();
[…]

### 2016-07-02 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp

@@ -73,7 +73,7 @@ bt6DofElasticPlastic2Constraint::bt6DofElasticPlastic2Constraint(btRigidBody& rb
 	initPlasticity();
 }
 
-int idCounter = 0;
+int bt6DofElasticPlastic2Constraint::idCounter; 
 void bt6DofElasticPlastic2Constraint::resetIdCounter(){
 	idCounter = 0;
 }

### 2016-07-02 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.h

@@ -46,6 +46,7 @@ http://gimpact.sf.net
 #include "BulletDynamics/ConstraintSolver/btTypedConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btGeneric6DofSpring2Constraint.h"
 #include "BulletDynamics/Dynamics/btActionInterface.h"
+#include "btElasticPlasticConstraint.h"
 #include "BulletDynamics/Dynamics/btRigidBody.h"
 class btRigidBody;
 #include "BulletDynamics/Dynamics/btDynamicsWorld.h"
@@ -59,7 +60,8 @@ class btDynamicsWorld;
 #define bt6DofElasticPlastic2ConstraintDataName	"bt6DofElasticPlastic2ConstraintData"
 #endif //BT_USE_DOUBLE_PRECISION
 
-ATTRIBUTE_ALIGNED16(class) bt6DofElasticPlastic2Constraint : public btTypedConstraint, public btActionInterface
+ATTRIBUTE_ALIGNED16(class) bt6DofElasticPlastic2Constraint : 
+public btTypedConstraint, public btActionInterface, public btElasticPlasticConstraint
 {
 protected:
 
@@ -276,6 +278,7 @@ public:
 	virtual void setParam(int num, btScalar value, int axis = -1);
 	virtual btScalar getParam(int num, int axis = -1) const;
 	// bcc
+	static int idCounter;
 	static void resetIdCounter();
 	int id;
 	int getId(){ return id; }
@@ -292,12 +295,12 @@ public:
 	void setMaxPlasticStrain(btScalar value);
 	void setMaxPlasticRotation(btScalar value);
 	void scalePlasticity(btScalar scale);
-	btScalar getMaxPlasticStrain();
-	btScalar getMaxPlasticRotation();
-	btScalar getCurrentPlasticStrain();
-	btScalar getCurrentPlasticRotation();
-	btScalar getMaxRatio();
-	int getMaxRatioDof();
+	virtual btScalar getMaxPlasticStrain();
+	virtual btScalar getMaxPlasticRotation();
+	virtual btScalar getCurrentPlasticStrain();
+	virtual btScalar getCurrentPlasticRotation();
+	virtual btScalar getMaxRatio();
+	virtual int getMaxRatioDof();
 	void initPlasticity();
 	void updatePlasticity(btJointFeedback& forces);
 	///btActionInterface interface

### 2016-07-02 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp

@@ -16,6 +16,8 @@ subject to the following restrictions:
 #include "bt6DofElasticPlasticConstraint.h"
 #include "BulletDynamics/Dynamics/btRigidBody.h"
 #include "LinearMath/btTransformUtil.h"
+#include "LinearMath/btIDebugDraw.h"
+#include "LinearMath/btQuickprof.h"
 
 
 bt6DofElasticPlasticConstraint::bt6DofElasticPlasticConstraint(btRigidBody& rbA, btRigidBody& rbB, const btTransform& frameInA, const btTransform& frameInB ,bool useLinearReferenceFrameA)
@@ -31,6 +33,11 @@ bt6DofElasticPlasticConstraint::bt6DofElasticPlasticConstraint(btRigidBody& rbB,
     init();
 }
 
+// bcc
+int bt6DofElasticPlasticConstraint::idCounter;
+void bt6DofElasticPlasticConstraint::resetIdCounter(){
+	idCounter = 0;
+}
 
 void bt6DofElasticPlasticConstraint::init()
 {
@@ -48,6 +55,8 @@ void bt6DofElasticPlasticConstraint::init()
 		m_maxPlasticRotation = btScalar(0.f);
 		m_currentPlasticRotation = btScalar(0.f);
 	}
+	id = ++idCounter;
+	setJointFeedback(&jointFeedback);
 }
 
 
@@ -265,6 +274,12 @@ void bt6DofElasticPlasticConstraint::scalePlasticity(btScalar scale){
 	m_maxPlasticRotation *= scale;
 }
 void bt6DofElasticPlasticConstraint::updatePlasticity(btJointFeedback& forces){
+	BT_PROFILE("updatePlasticity");
+	if (!isEnabled()){
+		return;
+	}
+	m_maxRatio = 0;
+	m_maxRatioDof = -1;
 	int i;
 	for (i = 0; i < 3; i++)
 	{
@@ -273,7 +288,13 @@ void bt6DofElasticPlasticConstraint::updatePlasticity(btJointFeedback& forces){
 			btScalar currPos = m_calculatedLinearDiff[i];
 			btScalar delta = currPos - m_equilibriumPoint[i];
 			btScalar force = forces.m_appliedForceBodyA[i];
-			if (btFabs(force)>m_maxForce[i]){
+			btScalar absForce = btFabs(force);
+			btScalar ratio = absForce / m_maxForce[i];
[…]

### 2016-07-02 examples/CharpyDemo/bt6DofElasticPlasticConstraint.h

@@ -23,6 +23,12 @@ April 2015 based on btGeneric6DofSpringConstraint
 #include "LinearMath/btVector3.h"
 #include "BulletDynamics/ConstraintSolver/btTypedConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btGeneric6DofConstraint.h"
+#include "BulletDynamics/Dynamics/btActionInterface.h"
+#include "btElasticPlasticConstraint.h"
+#include "BulletDynamics/Dynamics/btRigidBody.h"
+class btRigidBody;
+#include "BulletDynamics/Dynamics/btDynamicsWorld.h"
+class btDynamicsWorld;
 
 #ifdef BT_USE_DOUBLE_PRECISION
 #define bt6DofElasticPlasticConstraintData2		bt6DofElasticPlasticConstraintDoubleData2
@@ -44,7 +50,8 @@ April 2015 based on btGeneric6DofSpringConstraint
 /// 4 : rotation Y (2nd Euler rotational around new position of Y axis, range [-PI/2+epsilon, PI/2-epsilon] )
 /// 5 : rotation Z (1st Euler rotational around Z axis, range [-PI+epsilon, PI-epsilon] )
 
-ATTRIBUTE_ALIGNED16(class) bt6DofElasticPlasticConstraint : public btGeneric6DofConstraint
+ATTRIBUTE_ALIGNED16(class) bt6DofElasticPlasticConstraint : 
+public btGeneric6DofConstraint, public btActionInterface, public btElasticPlasticConstraint
 {
 protected:
 	bool		m_springEnabled[6];
@@ -52,6 +59,10 @@ protected:
 	btScalar	m_springStiffness[6];
 	btScalar	m_springDamping[6];
 	// bcc, new fields for plasticity
+	static int idCounter;
+	static void resetIdCounter();
+	int id;
+	int getId(){ return id; }
 	btScalar	m_maxForce[6];
 	btScalar	m_fpsLimit[6];
 	// how many integration steps is needed for single period before
@@ -62,6 +73,8 @@ protected:
 	btScalar    m_maxPlasticStrain;
 	btScalar    m_maxPlasticRotation = 3;
 	btScalar    m_currentPlasticRotation = 0;
+	btScalar m_maxRatio;
+	int m_maxRatioDof;
 	// bcc end of new fields for plasticity
 	void init();
 	void internalUpdateSprings(btConstraintInfo2* info);
@@ -79,14 +92,17 @@ public:
 	void setMaxPlasticStrain(btScalar value);
[…]

### 2016-07-02 examples/CharpyDemo/btElasticPlasticConstraint.h

@@ -1,0 +1,23 @@
+/*
+Simo Nikula, 2016/07
+*/
+
+#ifndef _BT_ELASTIC_PLASTIC_CONSTRAINT_INTERFACE_H
+#define _BT_ELASTIC_PLASTIC_CONSTRAINT_INTERFACE_H
+
+#include "LinearMath/btScalar.h"
+
+///Basic interface to allow common actions for various elasticPlasticConstraints
+class btElasticPlasticConstraint
+{
+public:
+	virtual btScalar getMaxPlasticRotation()=0;
+	virtual btScalar getCurrentPlasticRotation() = 0;
+	virtual btScalar getMaxPlasticStrain() = 0;
+	virtual btScalar getCurrentPlasticStrain() = 0;
+	virtual btScalar getMaxRatio() = 0;
+	virtual int getMaxRatioDof() = 0;
+};
+
+#endif //_BT_ELASTIC_PLASTIC_CONSTRAINT_INTERFACE_H
+

### 2016-07-02 examples/plasticity/PlasticityData.h

@@ -6,7 +6,8 @@
 
 enum btPlasticConstraintType
 {
-	BPT_EP2=21
+	BPT_EP=0x21,
+	BPT_EP2=0x22
 };
 
 using namespace std;

### 2016-07-02 examples/plasticity/PlasticityExampleEntries.cpp

@@ -101,6 +101,9 @@ very stiff constraints cause instability",
 	ExampleEntry(1, "Demolisher:ElasticPlastic",
 	"ElasticPlastic constraints which break based on work",
 	DemolisherDemoCreateFunc, 3),
+	ExampleEntry(1, "Demolisher:ElasticPlastic2",
+	"ElasticPlastic2 constraints which break based on work",
+	DemolisherDemoCreateFunc, 4),
 	ExampleEntry(0, "Plate"),
 	ExampleEntry(1, "Plate:None",
 	"Only mass resistance",

### 2016-07-15 pdocs/thesis/article-section-2.tex
typo fix
improvements based on aki's comments on section 2 of article

@@
 
-In this section some central terms are explained and main differences between traditional structural 
-analysis and physics engines are reviewed.
+In this section, key concepts related to the introduced model are explained. Main differences between 
+traditional structural analysis and physics engines are reviewed and discussed.
 
-Velocity-based formulation is not typically used in structural or mechanical engineering.
- \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details why it is so popular in 
-constraint-based rigid body simulation. 
+Velocity-based formulation is not often used in structural or mechanical engineering while, 
+however, it is popular within physics based game developers and film production teams.
+ \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details on why 
+velocity-based formulation is  popular in constraint-based rigid body simulation. 
 Main reason is that collision handling can be done without additional procedures.
-In structural analysis solution method is selected carefully based on needed features.
+Acceleration-based formulations need to stop at collision and switch to an impulse based method.
+In structural analysis, a formulation and associated numerical solution procedure are selected 
+based on needed features.
 For most complex scenarios finite element method is used.
-In most cases static small displacement solution using displacement based boundary conditions is used.
-For large displacement static analysis analysis loading is applied in substeps and 
-displacements are used to update element mesh.
+In most cases, static solution with assumption of linear strain-displacement relation small 
+displacement solution using displacement based boundary conditions is used.
+\citet{bathe-1975} provides description for handling of various nonlinearities.
+In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
+Lagrangian formulation where initial configuration is used.
 Further enhancements are material nonlinearity and dynamic analysis.
-Physics engine provides dynamic analysis with large displacements.
+Physics engine provides dynamic analysis with large large reference translations and rotations.
 
@@
 This provides reasonable means to simulate loss of energy in collisions.
-Simulation of breaking of objects made of ductile material can be made more realistic by splitting rigid objects
-to multiple parts which are connected by energy absorbing joints.
-As stress-strain curve may not be familiar to all readers some basics are provided here.
+Simulation of breaking of objects made of ductile material can be made more realistic by splitting rigid bodies
+to multiple bodies which are connected by energy absorbing joints.
 Typical engineering stress-strain curve of ductile steel is shown in \ref{fig:sscurve}.
@@
 to higher frequency than integration stability would allow.
-Strain hardening part of stress-straing curve could probably be taken into account but it has not been tried in this work.
+Strain hardening part of stress-strain curve could probably be taken into account but it has not been tried in this work.
 It should be noted that geometry
@@
 Difference between elastic and plastic section modulus is shown in \ref{fig:wp}. 
-If stress is below yield limit, stress and strain are linear within cross section.
+If stress is below yield limit, stress and strain are linear within material.
 If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
@@
 Mathematical background and detailed examples are available by \cite{ode.joints}.
+In following section, these equations will be clarified by simple example.
 Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and

### 2016-07-17 examples/CharpyDemo/CharpyDemo.cpp
added single step button to charpy

@@ -8,6 +8,9 @@ missing graphics
 or structural analysis background should be able to follow.
 
 It also helps myself as this work has had many long pauses
+This example cannot be used as template for copying due to
+having non class related variables.
+Other examples are better suited for usage as template.
 */
 
 
@@ -124,6 +127,7 @@ float maxEnergy;
 btDynamicsWorld* dw;
 btVector3 y_up(0.01, 1., 0.01);
 btRigidBody *specimenBody, *hammerBody=0, *specimenBody2;
+long stepCount, maxStepCount, syncedStep;
 btHingeConstraint *hammerHinge;
 btJointFeedback hammerHingeJointFeedback;
 btAlignedObjectArray<btJointFeedback*> specimenJointFeedback;
@@ -225,6 +229,7 @@ static Gwen::Controls::Canvas* canvas;
 int gx = 10; // for labels
 int gxi = 120; // for inputs elements
 int wxi = 60; // width
+int swxi = 30; // short width
 int gy;
 int gyInc=25;
 bool restartRequested = false;
@@ -442,6 +447,7 @@ public:
 	void initOptions(){
 		int option = m_option;
 		reinit();
+		maxStepCount = LONG_MAX;
 		m_guiHelper->resetCamera(0.5, -70, 15,	0, 0.2, 0.0);
 		m_mode = option % 100;
 		option /= 100;
@@ -626,7 +632,7 @@ public:
 		gc->SetPos(gx, gy);
 		return gc;
 	}
-	Gwen::Controls::Button* pauseButton;
+	Gwen::Controls::Button* pauseButton, *singleStepButton;
 	void addPauseSimulationButton(){
 		Gwen::Controls::Button* gc = new Gwen::Controls::Button(pPage);
 		pauseButton = gc;
@@ -635,17 +641,25 @@ public:
 		gc->SetSize(wxi - 4, gyInc - 4);
 		gc->onPress.Add(pPage, &CharpyDemo::handlePauseSimulation);
 	}
+	void addSingleStepButton(){
+		Gwen::Controls::Button* gc = new Gwen::Controls::Button(pPage);
+		gc->SetText(L"SS");
+		gc->SetToolTip(L"Single Step");
+		gc->SetPos(gx + wxi, gy);
+		gc->SetSize(swxi - 4, gyInc - 4);
[…]

### 2016-07-17 examples/CharpyDemo/TensionDemo.cpp
TensionDemo started

@@ -1,0 +1,1536 @@
+/*
+Bullet Continuous Collision Detection and Physics Library
+Copyright (c) 2003-2006 Erwin Coumans  http://continuousphysics.com/Bullet/
+
+This software is provided 'as-is', without any express or implied warranty.
+In no event will the authors be held liable for any damages arising from the use of this software.
+Permission is granted to anyone to use this software for any purpose, 
+including commercial applications, and to alter it and redistribute it freely, 
+subject to the following restrictions:
+
+1. The origin of this software must not be misrepresented; you must not claim that you wrote the original software. If you use this software in a product, an acknowledgment in the product documentation would be appreciated but is not required.
+2. Altered source versions must be plainly marked as such, and must not be misrepresented as being the original software.
+3. This notice may not be removed or altered from any source distribution.
+*/
+
+/**
+Based on DemolisherDemo by Simo Nikula 2016-
+*/
+#define _CRT_SECURE_NO_WARNINGS
+#include "TensionDemo.h"
+#include "AxisMapper.h"
+#include "LinearMath/btQuickprof.h"
+#include <stdio.h> 
+#include <string>
+#ifdef _WIN32
+#include <windows.h>
+#endif
+#include "btBulletDynamicsCommon.h"
+#include "BulletCollision/CollisionShapes/btHeightfieldTerrainShape.h"
+#include "BulletDynamics/MLCPSolvers/btDantzigSolver.h"
+#include "BulletDynamics/MLCPSolvers/btSolveProjectedGaussSeidel.h"
+#include "BulletDynamics/MLCPSolvers/btMLCPSolver.h"
+#include "bt6DofElasticPlasticConstraint.h"
+#include "bt6DofElasticPlastic2Constraint.h"
+#include "btElasticPlasticConstraint.h"
+#include "../plasticity/PlasticityExampleBrowser.h"
+#include "../plasticity/PlasticityData.h"
+#include "../plasticity/PlasticityStatistics.h"
+
+#include "BulletDynamics/ConstraintSolver/btSliderConstraint.h"
+
[…]

### 2016-07-17 examples/CharpyDemo/TensionDemo.h

@@ -1,0 +1,38 @@
+/*
+Bullet Continuous Collision Detection and Physics Library
+Copyright (c) 2003-2006 Erwin Coumans  http://continuousphysics.com/Bullet/
+
+This software is provided 'as-is', without any express or implied warranty.
+In no event will the authors be held liable for any damages arising from the use of this software.
+Permission is granted to anyone to use this software for any purpose, 
+including commercial applications, and to alter it and redistribute it freely, 
+subject to the following restrictions:
+
+1. The origin of this software must not be misrepresented; you must not claim that you wrote the original software. If you use this software in a product, an acknowledgment in the product documentation would be appreciated but is not required.
+2. Altered source versions must be plainly marked as such, and must not be misrepresented as being the original software.
+3. This notice may not be removed or altered from any source distribution.
+*/
+#ifndef TENSION_DEMO_H
+#define TENSION_DEMO_H
+#include "../CommonInterfaces/CommonExampleInterface.h"
+#include "../CommonInterfaces/CommonGUIHelperInterface.h"
+#include "../CommonInterfaces/CommonRenderInterface.h"
+#include "../CommonInterfaces/CommonWindowInterface.h"
+#include "../CommonInterfaces/CommonGraphicsAppInterface.h"
+#include "../CommonInterfaces/CommonRigidBodyBase.h"
+
+#include "LinearMath/btAlignedObjectArray.h"
+
+class btBroadphaseInterface;
+class btCollisionShape;
+class btOverlappingPairCache;
+class btCollisionDispatcher;
+class btConstraintSolver;
+struct btCollisionAlgorithmCreateFunc;
+class btDefaultCollisionConfiguration;
+
+class CommonExampleInterface*    TensionDemoCreateFunc(struct CommonExampleOptions& options);
+
+#endif // TENSION_DEMO_H
+
+

### 2016-07-17 examples/plasticity/PlasticityExampleEntries.cpp

@@ -6,6 +6,7 @@
 #include "../CharpyDemo/CharpyDemo.h"
 #include "../CharpyDemo/DemolisherDemo.h"
 #include "../CharpyDemo/PlateDemo.h"
+#include "../CharpyDemo/TensionDemo.h"
 #include "../ExampleBrowser/ExampleEntries.h"
 
 #ifdef ENABLE_LUA
@@ -189,6 +190,32 @@ very stiff constraints cause instability",
 40 t load\
 ",
 	PlateDemoCreateFunc, 400606203),
+	ExampleEntry(0, "Tension"),
+	ExampleEntry(1, "Tension:None",
+	"no constraint",
+	TensionDemoCreateFunc, 0),
+	ExampleEntry(1, "Tension:Rigid",
+	"rigid constraint using zero limits",
+	TensionDemoCreateFunc, 1),
+	ExampleEntry(1, "Tension:Spring",
+	"springConstraint",
+	TensionDemoCreateFunc, 2),
+	ExampleEntry(1, "Tension:Impulse",
+	"breakable constraint with zero limits",
+	TensionDemoCreateFunc, 3),
+	ExampleEntry(1, "Tension:Spring2",
+	"spring2Constraint",
+	TensionDemoCreateFunc, 4),
+	ExampleEntry(1, "Tension:Slider",
+	"SliderConstraint",
+	TensionDemoCreateFunc, 5),
+	ExampleEntry(1, "Tension:ElasticPlastic",
+	"6DofElasticPlasticConstraint",
+	TensionDemoCreateFunc, 6),
+	ExampleEntry(2, "Tension:ElasticPlastic2",
+	"6DofElasticPlastic2Constraint",
+	TensionDemoCreateFunc, 7),
+
 };
 struct ExampleEntriesInternalData
 {

### 2016-07-25 examples/CharpyDemo/AxisMapper.cpp
TensionDemo refactored to BlockDemo which contains also Bending

@@ -1,9 +1,9 @@
 #include "AxisMapper.h"
 void AxisMapper::init(){
-	if (m_cpos.getX() > 0){
+	if (m_cpos.getX()!=BT_ZERO){
 		initX();
 	}
-	else if (m_cpos.getY()){
+	else if (m_cpos.getY() != BT_ZERO){
 		initY();
 	}
 	else{

### 2016-07-25 examples/CharpyDemo/BlockDemo.cpp

@@ -17,7 +17,7 @@ subject to the following restrictions:
 Based on DemolisherDemo by Simo Nikula 2016-
 */
 #define _CRT_SECURE_NO_WARNINGS
-#include "TensionDemo.h"
+#include "BlockDemo.h"
 #include "AxisMapper.h"
 #include "LinearMath/btQuickprof.h"
 #include <stdio.h> 
@@ -50,9 +50,9 @@ Based on DemolisherDemo by Simo Nikula 2016-
 #include "../CommonInterfaces/CommonGraphicsAppInterface.h"
 #include "../ExampleBrowser/GwenGUISupport/gwenUserInterface.h"
 #include "../ExampleBrowser/GwenGUISupport/gwenInternalData.h"
-const char * PROFILE_TENSION_SLEEP = "TensionDemo::Sleep";
+const char * PROFILE_TENSION_SLEEP = "BlockDemo::Sleep";
 
-class TensionDemo : public Gwen::Event::Handler, public CommonRigidBodyBase
+class BlockDemo : public Gwen::Event::Handler, public CommonRigidBodyBase
 {
 public:
 	class btDiscreteDynamicsWorld* m_dynamicsWorld;
@@ -63,7 +63,6 @@ public:
 	btRigidBody* localCreateRigidBody(btScalar mass, const btTransform& worldTransform, btCollisionShape* colSape);
 
 	GUIHelperInterface* m_guiHelper;
-	int m_wheelInstances[4];
 
 	bool m_useDefaultCamera;
 	btAlignedObjectArray<btCollisionShape*> m_collisionShapes;
@@ -97,7 +96,7 @@ public:
 	btScalar damping2 = 0.1; // 0= no damping, 1=critically damped, >1 overdamped
 	bool hasFullGravity;
 	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
-	btScalar rodSteelScale;
+	btScalar blockSteelScale;
 	btScalar density;
 	btScalar maxPlasticRotation;
 	btScalar maxPlasticStrain;
@@ -120,14 +119,18 @@ public:
 		Impulse = 3, Spring2 = 4, Slider=5,
 		ElasticPlastic = 6, ElasticPlastic2 = 7 };
 	Constraint constraintType;
+	enum Orientation {
+		X = 0, Y = 1
+	};
+	Orientation orientationType;
 	float		m_cameraHeight;
 
 	float	m_minCameraDistance;
[…]

### 2016-07-25 examples/CharpyDemo/BlockDemo.h

@@ -31,7 +31,7 @@ class btConstraintSolver;
 struct btCollisionAlgorithmCreateFunc;
 class btDefaultCollisionConfiguration;
 
-class CommonExampleInterface*    TensionDemoCreateFunc(struct CommonExampleOptions& options);
+class CommonExampleInterface*    BlockDemoCreateFunc(struct CommonExampleOptions& options);
 
 #endif // TENSION_DEMO_H
 

### 2016-07-25 examples/plasticity/PlasticityExampleEntries.cpp

@@ -6,7 +6,7 @@
 #include "../CharpyDemo/CharpyDemo.h"
 #include "../CharpyDemo/DemolisherDemo.h"
 #include "../CharpyDemo/PlateDemo.h"
-#include "../CharpyDemo/TensionDemo.h"
+#include "../CharpyDemo/BlockDemo.h"
 #include "../ExampleBrowser/ExampleEntries.h"
 
 #ifdef ENABLE_LUA
@@ -193,28 +193,53 @@ very stiff constraints cause instability",
 	ExampleEntry(0, "Tension"),
 	ExampleEntry(1, "Tension:None",
 	"no constraint",
-	TensionDemoCreateFunc, 0),
+	BlockDemoCreateFunc, 0),
 	ExampleEntry(1, "Tension:Rigid",
 	"rigid constraint using zero limits",
-	TensionDemoCreateFunc, 1),
+	BlockDemoCreateFunc, 11),
 	ExampleEntry(1, "Tension:Spring",
 	"springConstraint",
-	TensionDemoCreateFunc, 2),
+	BlockDemoCreateFunc, 12),
 	ExampleEntry(1, "Tension:Impulse",
 	"breakable constraint with zero limits",
-	TensionDemoCreateFunc, 3),
+	BlockDemoCreateFunc, 13),
 	ExampleEntry(1, "Tension:Spring2",
 	"spring2Constraint",
-	TensionDemoCreateFunc, 4),
+	BlockDemoCreateFunc, 14),
 	ExampleEntry(1, "Tension:Slider",
 	"SliderConstraint",
-	TensionDemoCreateFunc, 5),
+	BlockDemoCreateFunc, 15),
 	ExampleEntry(1, "Tension:ElasticPlastic",
 	"6DofElasticPlasticConstraint",
-	TensionDemoCreateFunc, 6),
+	BlockDemoCreateFunc, 16),
 	ExampleEntry(2, "Tension:ElasticPlastic2",
 	"6DofElasticPlastic2Constraint",
-	TensionDemoCreateFunc, 7),
+	BlockDemoCreateFunc, 17),
+	ExampleEntry(0, "Bending"),
+	ExampleEntry(1, "Bending:None",
+	"no constraint",
+	BlockDemoCreateFunc, 0),
+	ExampleEntry(1, "Bending:Rigid",
+	"rigid constraint using zero limits",
+	BlockDemoCreateFunc, 1),
[…]

### 2016-07-26 examples/CharpyDemo/BlockDemo.cpp
Bending:Hinge and Tension:Slider fixes

@@ -116,7 +116,7 @@ public:
 	CommonWindowInterface* window;
 	int m_option;
 	enum Constraint { None = 0, Rigid = 1, Spring = 2, 
-		Impulse = 3, Spring2 = 4, Slider=5,
+		Impulse = 3, Spring2 = 4, SingleDof=5,
 		ElasticPlastic = 6, ElasticPlastic2 = 7 };
 	Constraint constraintType;
 	enum Orientation {
@@ -695,7 +695,7 @@ public:
 		case Spring:
 		case Impulse:
 		case Spring2:
-		case Slider:
+		case SingleDof:
 			addBlockSteelScale();
 			break;
 		case ElasticPlastic:
@@ -760,12 +760,22 @@ public:
 	btTypedConstraint* addSliderConstraint(btRigidBody* rb, btVector3& cpos){
 		btTransform tr;
 		tr.setIdentity();
-		tr.setOrigin(cpos);
+		tr.setRotation(btQuaternion(0, 0,SIMD_HALF_PI));
+		tr.setOrigin(cpos); //debug drawing is easily seen
 		btSliderConstraint *sc =
 			new btSliderConstraint(*rb, tr, true);
 		m_dynamicsWorld->addConstraint(sc, disableCollisionsBetweenLinkedBodies);
 		sc->setPoweredLinMotor(true);
 		sc->setMaxLinMotorForce(axisMapper->getMaxForce(1));
+		sc->setTargetLinMotorVelocity(0);
+		return sc;
+	}
+	btTypedConstraint* addHingeConstraint(btRigidBody* rb, btVector3& pivot, btVector3 & axis){
+		btHingeConstraint *sc =
+			new btHingeConstraint(*rb, pivot, axis);
+		m_dynamicsWorld->addConstraint(sc, disableCollisionsBetweenLinkedBodies);
+		sc->enableAngularMotor(true,0,axisMapper->getMaxForce(5)*m_fixedTimeStep);
+		sc->setLimit(-SIMD_PI, SIMD_PI);
 		return sc;
 	}
 	btTypedConstraint* addSpring2Constraint(btRigidBody* rb, btVector3& cpos){
@@ -938,8 +948,15 @@ public:
 		case Spring2:
 			m_constraint = addSpring2Constraint(m_body, cpos);
 			break;
-		case Slider:
-			m_constraint = addSliderConstraint(m_body, cpos);
+		case SingleDof:
+			switch (orientationType){
+			case X:
[…]

### 2016-07-26 examples/plasticity/PlasticityExampleEntries.cpp

@@ -38,28 +38,28 @@ option is decoded in constructors
 static ExampleEntry gDefaultExamples[]=
 {
 	ExampleEntry(0, "Charpy Impact"),
-	ExampleEntry(1, "F1",
+	ExampleEntry(1, "Charpy:Rigid",
 	"single rigid object",
 	CharpyDemoCreateFunc, 1),
-	ExampleEntry(1, "F2",
+	ExampleEntry(1, "Charpy:Spring",
 	"two objects and springConstraints",
 	CharpyDemoCreateFunc, 2),
-	ExampleEntry(1, "F3",
+	ExampleEntry(1, "Charpy:Impulse",
 	"two objects and constraints with zero limits",
 	CharpyDemoCreateFunc, 3),
-	ExampleEntry(1, "F4",
+	ExampleEntry(1, "Charpy:Spring2",
 	"two objects and spring2Constraints",
 	CharpyDemoCreateFunc, 4),
-	ExampleEntry(1, "F5",
+	ExampleEntry(1, "Charpy:Hinge",
 	"two objects and hingeConstraint",
 	CharpyDemoCreateFunc, 5),
-	ExampleEntry(1, "F6",
+	ExampleEntry(1, "Charpy:PlasticHinge",
 	"two objects and plasticHingeConstraint",
 	CharpyDemoCreateFunc, 6),
-	ExampleEntry(1, "F7",
+	ExampleEntry(1, "Charpy:ElasticPlastic",
 	"two objects and 6DofElasticPlasticConstraint",
 	CharpyDemoCreateFunc, 7),
-	ExampleEntry(2, "F8",
+	ExampleEntry(2, "Charpy:ElasticPlastic2",
 	"two objects and 6DofElasticPlastic2Constraint",
 	CharpyDemoCreateFunc, 8),
 	ExampleEntry(2, "elastic",
@@ -193,7 +193,7 @@ very stiff constraints cause instability",
 	ExampleEntry(0, "Tension"),
 	ExampleEntry(1, "Tension:None",
 	"no constraint",
-	BlockDemoCreateFunc, 0),
+	BlockDemoCreateFunc, 10),
 	ExampleEntry(1, "Tension:Rigid",
 	"rigid constraint using zero limits",
 	BlockDemoCreateFunc, 11),
@@ -231,8 +231,8 @@ very stiff constraints cause instability",
 	ExampleEntry(1, "Bending:Spring2",
[…]

### 2016-07-26 examples/plasticity/PlasticityExampleEntries.cpp
example reordering, tension and bending moved to first

@@ -37,6 +37,56 @@ option is decoded in constructors
 */
 static ExampleEntry gDefaultExamples[]=
 {
+	ExampleEntry(0, "Tension"),
+	ExampleEntry(1, "Tension:None",
+	"no constraint",
+	BlockDemoCreateFunc, 10),
+	ExampleEntry(1, "Tension:Rigid",
+	"rigid constraint using zero limits",
+	BlockDemoCreateFunc, 11),
+	ExampleEntry(1, "Tension:Spring",
+	"springConstraint",
+	BlockDemoCreateFunc, 12),
+	ExampleEntry(1, "Tension:Impulse",
+	"breakable constraint with zero limits",
+	BlockDemoCreateFunc, 13),
+	ExampleEntry(1, "Tension:Spring2",
+	"spring2Constraint",
+	BlockDemoCreateFunc, 14),
+	ExampleEntry(1, "Tension:Slider",
+	"SliderConstraint",
+	BlockDemoCreateFunc, 15),
+	ExampleEntry(1, "Tension:ElasticPlastic",
+	"6DofElasticPlasticConstraint",
+	BlockDemoCreateFunc, 16),
+	ExampleEntry(2, "Tension:ElasticPlastic2",
+	"6DofElasticPlastic2Constraint",
+	BlockDemoCreateFunc, 17),
+	ExampleEntry(0, "Bending"),
+	ExampleEntry(1, "Bending:None",
+	"no constraint",
+	BlockDemoCreateFunc, 0),
+	ExampleEntry(1, "Bending:Rigid",
+	"rigid constraint using zero limits",
+	BlockDemoCreateFunc, 1),
+	ExampleEntry(1, "Bending:Spring",
+	"springConstraint",
+	BlockDemoCreateFunc, 2),
+	ExampleEntry(1, "Bending:Impulse",
+	"breakable constraint with zero limits",
+	BlockDemoCreateFunc, 3),
+	ExampleEntry(1, "Bending:Spring2",
+	"spring2Constraint",
+	BlockDemoCreateFunc, 4),
+	ExampleEntry(1, "Bending:Hinge",
+	"HingeConstraint",
+	BlockDemoCreateFunc, 5),
+	ExampleEntry(1, "Bending:ElasticPlastic",
+	"6DofElasticPlasticConstraint",
+	BlockDemoCreateFunc, 6),
+	ExampleEntry(2, "Bending:ElasticPlastic2",
[…]

### 2016-07-27 pdocs/thesis/article-section-3.tex
article section 3 started

\section{Processing plasticity}

In this section, changes to constraint formulation are clarified by simple numerical examples.

System has only one dynamic rigid body which is three meters high bocks. Cross section is one square meter.
Density is $2000 kg \over m^3$. Gravity is $10 m\over s^2$. Simulation step is $1\over 60 s$ and 10 
iterations are done for single step.

Single simulation step is done using following substeps. 
In this work, changes are made only to constraing solving and update actions.
\begin{enumerate}
\item Apply gravity
\item Predict unconstrained motion
\item Predict contacts
\item Perform collision detection
\item Calculate simulation islands(groups). Objetcs that are near each other or connected with constraints are grouped in same group.
\item Solve constraints. Both contact and other constraints are processed in this step.
\item Integrate transforms
\item Update actions (callbacks) are called
\item Activation state of bodies is updated. Bodies that are not active
 are typically marked as sleeping and they are not processed.
\end{enumerate}

Equation \ref{eq:constraintEquation} is simplified
to \ref{eq:fixedConstraint} in this case.
\begin{itemize}
\item No rotation takes place. $\omega_1$ and $\omega_2$ are zeros.
\item Constraint force mixing can be ignored.
\item Only vertical velocity is handled.
\item Other involved body is rigid and it does not move.
\end{itemize}

\begin{equation} \label{eq:fixedConstraint}
m v_y = c 
\end{equation}

Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
in \bullet\ was used to pick up values.

\begin{description}
\item[velError] is calculated using velocities and external impulses of connected objects. 
 In this case, it is dominant contributor for constraint's impulse after initial phase.
\item[posError] is calculated by constraint. It is significant factor in designing stable constraints. 
 In fixed case, value is about 12 times actual position error. Factor 12 is based on time step (60) 
 and default value of error reduction parameter (0.2).
\end{description}

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Time} & 
{\bf velError} & {\bf posError} & {\bf rhs} &
{\bf vertical velocity} & 
{\bf constraint impulse [Ns]} \\  \hline
0.017 &  & & & -0.167 & 0 \\  \\hline
0.033 &  0.33 & -0.033 & -2200 & 0.0333 & 2200 \\  \\hline
0.050 &  -0.13 & -0.0266 & -960 & 0.0266 & 960 \\  \\hline
0.067 &  -0.14 & -0.021 & -970 & 0.021 & 970 \\  \\hline
0.35 &  -0.17 & \~0 & -1000 &0.0 & 1000 \\  \\hline
\end {tabular}
\end{center}
\caption {Constraint parameter values for fixed constraint} \label{tab:fixedBlockValues} 
\end {table}

Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
\ref{eq:lambdaHigh} 
are created for each constraint.

\begin{equation} \label{eq:constraintEquation}
J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
\end{equation}

\begin{equation} \label{eq:lambdaLow}
\lambda \geq l
\end{equation}

\begin{equation} \label{eq:lambdaHigh}
\lambda \leq h
\end{equation}

Main parameters  and corresponding fields in \bullet\  
 are described in table \ref{tab:constraintParameters}.

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$v$ & linear velocity & \\ \hline
$\omega$ & angular velocity & \\ \hline
$c$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\lambda$ & constraint force &  \\ \hline
$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
$h$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\caption {Constraint parameters} \label{tab:constraintParameters} 
\end {table}

### 2016-07-27 pdocs/thesis/article-section-3.tex
article section 3 started

@@
 \end {table}
-
-Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
-\ref{eq:lambdaHigh} 
-are created for each constraint.
-
-\begin{equation} \label{eq:constraintEquation}
-J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
-\end{equation}
-
-\begin{equation} \label{eq:lambdaLow}
-\lambda \geq l
-\end{equation}
-
-\begin{equation} \label{eq:lambdaHigh}
-\lambda \leq h
-\end{equation}
-
-Main parameters  and corresponding fields in \bullet\  
- are described in table \ref{tab:constraintParameters}.
-
-\begin {table}[htb!]
-\begin{center}
-\begin{tabular}{|c| l| l|}
-\hline
-{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
-$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
-$v$ & linear velocity & \\ \hline
-$\omega$ & angular velocity & \\ \hline
-$c$        &  right side vector   & m\_constraintError \\ \hline
-$C$  & constraint force mixing & cfm \\  \hline
-$\lambda$ & constraint force &  \\ \hline
-$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
-$h$ & high limit for constraint force & m\_upperLimit \\ \hline
-\end {tabular}
-\end{center}
-\caption {Constraint parameters} \label{tab:constraintParameters} 
-\end {table}

### 2016-07-28 pdocs/thesis/article-section-3.tex
article-section-3

@@
 
-In this section, changes to constraint formulation are clarified by simple numerical examples.
+In this section, changes to constraint formulation are clarified by doing few simulation steps of rigid and
+elastic-plastic examples.
 
-System has only one dynamic rigid body which is three meters high bocks. Cross section is one square meter.
-Density is $2000 kg \over m^3$. Gravity is $10 m\over s^2$. Simulation step is $1\over 60 s$ and 10 
+System has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
+Cross section is one square meter.
+Concrete density is $2000 \frac{kg}{m^3}$ and steel density id $7800 \frac{kg}{m^3}$. 
+Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
+Gravity is $10 \frac{m}{s^2}$. Simulation step is $\frac{1}{60} s$ and 10 
 iterations are done for single step.
@@
 \item Perform collision detection
-\item Calculate simulation islands(groups). Objetcs that are near each other or connected with constraints are grouped in same group.
+\item Calculate simulation islands(groups). Bodies that are near each other or connected with constraints are grouped in same group.
 \item Solve constraints. Both contact and other constraints are processed in this step.
 \item Integrate transforms
-\item Update actions (callbacks) are called
+\item Update actions (callbacks) are called. 
+In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
 \item Activation state of bodies is updated. Bodies that are not active
@@
 Equation \ref{eq:constraintEquation} is simplified
-to \ref{eq:fixedConstraint} in this case.
+to \ref{eq:fixedConstraint} in these cases.
 \begin{itemize}
@@
 
+Equations \ref{eq:lambdaLow} and \ref{eq:lambdaHigh} are not active for fixed case.
+For elastic-plastic case maximum impulse is set to product of yield stress, area of steel enforcement and time step (1330).
+
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ was used to pick up values.
+in \bullet\ was used to pick up values. Actual values for fixed constraint are shown in table
+\ref{tab:fixedBlockValues} and values for elastic-plastic case are shown in tables \ref{tab:epBlockValues} and 
+ \ref{tab:ep2BlockValues}. There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
+in this work elastic-plastic versions of both of them were developed. 
+Constraint activation of older spring and fixed constraints is implemented so that 
+it can be activated explicitly at start of simulation or it can be automatically activated when constraint is violated.
+In this case constraint is not activated explicitly and thus block drops freely during first simulation step.
 
 \begin{description}
-\item[velError] is calculated using velocities and external impulses of connected objects. 
- In this case, it is dominant contributor for constraint's impulse after initial phase.
+\item[velError] is calculated using velocities and external impulses of connected bodies. 
+ In this case, it is dominant contributor for constraint's impulse.
 \item[posError] is calculated by constraint. It is significant factor in designing stable constraints.
- In fixed case, value is about 12 times actual position error. Factor 12 is based on time step (60) 
+ \begin{description}
+ \item[In fixed case,] value is about 12 times actual position error. Factor 12 is based on time step (60) 
  and default value of error reduction parameter (0.2).
+ \item[In elastic plastic case,]  value is set to zero if impulse would be larger than maximum impulse or
+spring simulation cannot be done in stable way.
+ \end{description}
+\item[rhs] is calculated by velError jInv + posError jInv
+\item[jInv] is calculated using masses and constraint geometry. In this case, it is mass of body (6000).
 \end{description}
@@
 \begin{center}
-\begin{tabular}{|c| l| l|}
+\begin{tabular}{|l| l| l|l|l|l|}
 \hline
@@
 {\bf velError} & {\bf posError} & {\bf rhs} &
-{\bf vertical velocity} & 
-{\bf constraint impulse [Ns]} \\  \hline
-0.017 &  & & & -0.167 & 0 \\  \\hline
-0.033 &  0.33 & -0.033 & -2200 & 0.0333 & 2200 \\  \\hline
-0.050 &  -0.13 & -0.0266 & -960 & 0.0266 & 960 \\  \\hline
-0.067 &  -0.14 & -0.021 & -970 & 0.021 & 970 \\  \\hline
-0.35 &  -0.17 & \~0 & -1000 &0.0 & 1000 \\  \\hline
+{\bf velocity} & 
+{\bf impulse} \\  \hline
+0.017 &  & & & -0.17 & 0 \\  \hline
+0.033 &  0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
+0.050 &  -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
+0.067 &  -0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
+0.35... &  -0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
 \end {tabular}
@@
 \end {table}
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l| l| l|l|l|l|l|}
+\hline
+{\bf Time} & 
+{\bf velError} & {\bf posError} & {\bf rhs} &
+{\bf velocity} & 
+{\bf impulse} & 
+{\bf plastic strain} \\  \hline
+0.017 &          &    &          & -0.17 & 0 & 0 \\  \hline
+0.033 &  -0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
+0.050 &  -0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
+0.067 &  -0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
+0.083... &  -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
+\end {tabular}
+\end{center}
+\caption {Constraint parameter values for elastic-plastic constraint} \label{tab:epBlockValues} 
+\end {table}
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l| l| l|l|l|l|l|}
+\hline
+{\bf Time} & 
+{\bf velError} & {\bf posError} & {\bf rhs} &
+{\bf velocity} & 
+{\bf impulse} & 
+{\bf plastic strain} \\  \hline
+0.017... & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
+\end {tabular}
+\end{center}
+\caption {Constraint parameter values for elastic-plastic constraint 2} \label{tab:ep2BlockValues} 
+\end {table}

### 2016-07-28 examples/CharpyDemo/BlockDemo.cpp

@@ -89,9 +89,7 @@ public:
 	btScalar stepTime,gravityRampUpTime;
 	btVector3 m_gravity;
 	btScalar breakingImpulseThreshold;
-	btScalar E=200E9;
-	btScalar G=80E9;
-	btScalar fy=200E6;
+	btScalar E,fy,G;
 	btScalar damping=0.9; // 1= no damping, 0=full damping
 	btScalar damping2 = 0.1; // 0= no damping, 1=critically damped, >1 overdamped
 	bool hasFullGravity;
@@ -399,6 +397,8 @@ public:
 	void setLsy(Gwen::Controls::Base* control);
 	void setLsz(Gwen::Controls::Base* control);
 	void setDensity(Gwen::Controls::Base* control);
+	void setE(Gwen::Controls::Base* control);
+	void setFy(Gwen::Controls::Base* control);
 	void setBlockSteelScale(Gwen::Controls::Base* control);
 	void setMaxPlasticStrain(Gwen::Controls::Base* control);
 	void setMaxPlasticRotation(Gwen::Controls::Base* control);
@@ -621,7 +621,25 @@ public:
 		place(gc);
 		gc->onReturnPressed.Add(pPage, &BlockDemo::setDensity);
 	}
-	btScalar steelDensity=7800;
+	void addE(){
+		addLabel("E [MPa]");
+		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
+		std::string text = uif(E/1e6, "%.0f");
+		gc->SetToolTip("Young's modulus");
+		gc->SetText(text);
+		place(gc);
+		gc->onReturnPressed.Add(pPage, &BlockDemo::setE);
+	}
+	void addFy(){
+		addLabel("fy [MPa]");
+		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
+		std::string text = uif(fy / 1e6, "%.0f");
+		gc->SetToolTip("yield stress");
+		gc->SetText(text);
+		place(gc);
+		gc->onReturnPressed.Add(pPage, &BlockDemo::setFy);
+	}
+	btScalar steelDensity = 7800;
 	btScalar getDensity(btScalar steelScale){
 		return btScalar(steelScale*steelDensity+(1-steelScale)*density);
 	}
[…]

### 2016-07-28 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp

@@ -1274,13 +1274,13 @@ void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
 				if (btFabs(elasticPart)<btFabs(currPos)){
 					newVal += (currPos > 0 ? -elasticPart : elasticPart);
 				}
-				btScalar plasticDelta = btFabs(delta);
+				btScalar plasticDelta = btFabs(delta)-elasticPart;
 				/* direction changes */
 				if ((m_linearLimits.m_equilibriumPoint[i] * currPos)<0){
-					plasticDelta -= elasticPart * 2;
-					if (plasticDelta < 0){
-						plasticDelta = 0;
-					}
+					plasticDelta -= elasticPart;
+				}
+				if (plasticDelta < 0){
+					plasticDelta = 0;
 				}
 				setEquilibriumPoint(i, newVal);
 				m_currentPlasticStrain += plasticDelta;
@@ -1310,13 +1310,13 @@ void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
 				if (btFabs(elasticPart)<btFabs(currPos)){
 					newVal += (currPos > 0 ? -elasticPart : elasticPart);
 				}
-				btScalar plasticDelta = btFabs(delta);
+				btScalar plasticDelta = btFabs(delta)-elasticPart;
 				/* direction changes */
 				if ((m_angularLimits[i].m_equilibriumPoint*currPos)<0){
-					plasticDelta -= elasticPart * 2;
-					if (plasticDelta < 0){
-						plasticDelta = 0;
-					}
+					plasticDelta -= elasticPart;
+				}
+				if (plasticDelta < 0){
+					plasticDelta = 0;
 				}
 				setEquilibriumPoint(i + 3, newVal);
 				m_currentPlasticRotation += plasticDelta;

### 2016-07-28 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp

@@ -340,13 +340,13 @@ void bt6DofElasticPlasticConstraint::updatePlasticity(btJointFeedback& forces,
 				if (btFabs(elasticPart)<btFabs(currPos)){
 					newVal += (currPos > 0 ? -elasticPart : elasticPart);
 				}
-				btScalar plasticDelta = btFabs(delta);
+				btScalar plasticDelta = btFabs(delta)-elasticPart;
 				/* direction changes */
 				if ((m_equilibriumPoint[i] * currPos)<0){
-					plasticDelta -= elasticPart * 2;
-					if (plasticDelta < 0){
-						plasticDelta = 0;
-					}
+					plasticDelta -= elasticPart;
+				}
+				if (plasticDelta < 0){
+					plasticDelta = 0;
 				}
 				setEquilibriumPoint(i, newVal);
 				m_currentPlasticStrain += plasticDelta;
@@ -376,13 +376,13 @@ void bt6DofElasticPlasticConstraint::updatePlasticity(btJointFeedback& forces,
 				if (btFabs(elasticPart)<btFabs(currPos)){
 					newVal += (currPos > 0 ? -elasticPart : elasticPart);
 				}
-				btScalar plasticDelta = btFabs(delta);
+				btScalar plasticDelta = btFabs(delta)-elasticPart;
 				/* direction changes */
 				if ((m_equilibriumPoint[i+3]*currPos)<0){
-					plasticDelta -= elasticPart * 2;
-					if (plasticDelta < 0){
-						plasticDelta = 0;
-					}
+					plasticDelta -= elasticPart;
+				}
+				if (plasticDelta < 0){
+					plasticDelta = 0;
 				}
 				setEquilibriumPoint(i + 3, newVal);
 				m_currentPlasticRotation += plasticDelta;

### 2016-07-29 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp
code cleanup

@@ -166,32 +166,21 @@ void bt6DofElasticPlasticConstraint::internalUpdateSprings(btConstraintInfo2* in
 			// spring force is (delta * m_stiffness) according to Hooke's Law
 			btScalar force = delta * m_springStiffness[i];
 			bool overMaxForce=false;
-			bool slow = false;
 			// bcc
 			if (btFabs(force)>=m_maxForce[i]){
 				force = (delta > 0 ? m_maxForce[i] : -m_maxForce[i]);
-				// overMaxForce = true;
+				overMaxForce = true;
 			}
 			btScalar ratio(info->fps/m_fpsLimit[i]);
 			btScalar velFactor;
-			if (ratio>m_frequencyRatio){
+			if (!overMaxForce && ratio>m_frequencyRatio){
 				velFactor = m_springDamping[i] /
 					btScalar(info->m_numIterations);
 			}
 			else{
-				if (overMaxForce) {
-					slow = true;
-				}
-				else {
-					velFactor = 0;
-				}
+				velFactor = 0;
 			}
-			if (slow){
-				m_linearLimits.m_targetVelocity[i] = (delta > 0 ? SIMD_EPSILON : -SIMD_EPSILON);
-			}
-			else{
-				m_linearLimits.m_targetVelocity[i] = info->fps*velFactor * force;
-			}
+			m_linearLimits.m_targetVelocity[i] = info->fps*velFactor * force;
 			m_linearLimits.m_maxMotorForce[i] = btFabs(force) / info->fps;
 		}
 	}
@@ -208,31 +197,20 @@ void bt6DofElasticPlasticConstraint::internalUpdateSprings(btConstraintInfo2* in
 			btScalar force = -delta * m_springStiffness[i+3];
 			// bcc
 			bool overMaxForce = false;
-			bool slow = false;
 			if (btFabs(force) >= m_maxForce[i + 3]){
 				force = (delta > 0 ? -m_maxForce[i+3] : m_maxForce[i+3]);
-				//overMaxForce = true;
+				overMaxForce = true;
 			}
 			btScalar ratio(info->fps/m_fpsLimit[i + 3]);
 			btScalar velFactor;
-			if (ratio>m_frequencyRatio){
+			if (!overMaxForce && ratio>m_frequencyRatio){
 				velFactor = m_springDamping[i + 3] / 
 					btScalar(info->m_numIterations);
 			}
 			else{
-				if (overMaxForce) {
[…]

### 2016-07-29 examples/CharpyDemo/bt6DofElasticPlasticConstraint.h

@@ -97,6 +97,7 @@ public:
 	virtual btScalar getCurrentPlasticStrain();
 	virtual btScalar getCurrentPlasticRotation();
 	virtual btScalar getMaxRatio(){ return m_maxRatio; }
+	btScalar getFpsLimit(int index){ return m_fpsLimit[i]; }
 	virtual int getMaxRatioDof(){ return m_maxRatioDof; }
 	void updatePlasticity(btJointFeedback& forces, btCollisionWorld* collisionWorld, btScalar step);
 	void calculateFpsLimit(int index);

### 2016-07-29 examples/CharpyDemo/BlockDemo.cpp
frequencyRatio + codeCleanup

@@ -88,7 +88,7 @@ public:
 	long stepCount,maxStepCount,syncedStep;
 	btScalar stepTime,gravityRampUpTime;
 	btVector3 m_gravity;
-	btScalar breakingImpulseThreshold;
+	btScalar frequencyRatio;
 	btScalar E,fy,G;
 	btScalar damping=0.9; // 1= no damping, 0=full damping
 	btScalar damping2 = 0.1; // 0= no damping, 1=critically damped, >1 overdamped
@@ -258,6 +258,12 @@ public:
 			b_y_velocity = m_body->getLinearVelocity().y();
 		}
 	}
+	void resetView(){
+		c_impulse = 0;
+		c_percent = 0;
+		b_y_location = 0;
+		b_y_velocity = 0;
+	}
 	float pitchDelta = 0;
 	void setPitchDelta(float delta){
 		pitchDelta = delta;
@@ -398,6 +404,7 @@ public:
 	void setLsz(Gwen::Controls::Base* control);
 	void setDensity(Gwen::Controls::Base* control);
 	void setE(Gwen::Controls::Base* control);
+	void setFrequencyRatio(Gwen::Controls::Base* control);
 	void setFy(Gwen::Controls::Base* control);
 	void setBlockSteelScale(Gwen::Controls::Base* control);
 	void setMaxPlasticStrain(Gwen::Controls::Base* control);
@@ -418,13 +425,14 @@ public:
 		return btScalar(lsx*lsy*lsz*getDensity(blockSteelScale));
 	}
 	btScalar getEquilibriumPoint(){
-		bool useZero = true;
+		bool useZero = false;
 		if (useZero){
 			return 0;
 		}
 		btScalar f = -10 * getMass();
 		btScalar k = axisMapper->getStiffness(1);
 		btScalar val(f / k);
+		b3Printf("getEquilibriumPoint: %.4f",val);
 		return val;
 	}
 	void updateDashboards(){
@@ -475,6 +483,17 @@ public:
 		dumpPngGc = gc;
 		gy += gyInc;
 		gc->onCheckChanged.Add(pPage, &BlockDemo::setDumpPng);
+	}
+	void addFrequencyRatio(){
+		addLabel("frequencyRatio");
+		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
[…]

### 2016-07-29 pdocs/thesis/article-section-3.tex
article-section-3 continues

@@
+% !TeX root = article.tex
 \section{Processing plasticity}
@@
 
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (O) at (0,0,0);
+
+% axes
+\draw[->] (O) -- (1,0,0) node[anchor=north east]{$x$};
+\draw[->] (O) -- (0,1,0) node[anchor=north west]{$y$};
+\draw[->] (O) -- (0,0,1) node[anchor=south]{$z$};
+
+% 1,3,1 block, base at y=1.5
+\draw (1,1.5,0) -- ++(0,3,0) -- ++(-1,0,0);
+\draw (0,1.5,1) -- ++(1,0,0) -- ++(0,3,0) -- ++(-1,0,0) -- ++(0,-3,0);
+\draw (1,1.5,0) -- ++(0,0,1);
+\draw (0,4.5,0) -- ++(0,0,1);
+\draw (1,4.5,0) -- ++(0,0,1);
+
+\draw[->] (0.5,3,0.5) -- ++(0,1.5,0);
+\end{tikzpicture}
+\caption{Single body model for plasticy processing demonstration.}
+\label{fig:tensionModel}
+\end{figure}
+
 System has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
 Cross section is one square meter.
+Constraint is set so that connecting frame is at top of block. 
+In this case frame could be anywhere but if multiple bodies are involved, connecting frame must be defined
+so that it reflects wanted scenario.
 Concrete density is $2000 \frac{kg}{m^3}$ and steel density id $7800 \frac{kg}{m^3}$. 
@@
 Constraint activation of older spring and fixed constraints is implemented so that 
-it can be activated explicitly at start of simulation or it can be automatically activated when constraint is violated.
-In this case constraint is not activated explicitly and thus block drops freely during first simulation step.
+they are activated after constraint is violated.
+Thus body drops freely during first simulation step and 
+gains enough kinetic energy so that higher impulses are needed in few following steps.
 
@@
 \item[rhs] is calculated by velError jInv + posError jInv
-\item[jInv] is calculated using masses and constraint geometry. In this case, it is mass of body (6000).
+\item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
+In this case, it is mass of body (6000).
 \end{description}

### 2016-08-01 examples/CharpyDemo/BlockDemo.cpp
upper/lower limits are set via setMaxPlasticStrain

@@ -425,7 +425,7 @@ public:
 		return btScalar(lsx*lsy*lsz*getDensity(blockSteelScale));
 	}
 	btScalar getEquilibriumPoint(){
-		bool useZero = false;
+		bool useZero = true;
 		if (useZero){
 			return 0;
 		}
@@ -856,7 +856,7 @@ public:
 			sc->setMaxForce(i, axisMapper->getMaxForce(i));
 		}
 		sc->setEquilibriumPoint(1, getEquilibriumPoint());
-		b3Printf("fpLimit[1]=",sc->getFpsLimit(1));
+		b3Printf("fpsLimit[1]=%.1f",sc->getFpsLimit(1));
 		sc->setFrequencyRatio(frequencyRatio);
 		m_dynamicsWorld->addAction(sc);
 		return sc;

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp

@@ -1221,6 +1221,14 @@ void bt6DofElasticPlastic2Constraint::setMaxForce(int index, btScalar maxForce)
 }
 void bt6DofElasticPlastic2Constraint::setMaxPlasticStrain(btScalar value){
 	m_maxPlasticStrain = value;
+	/*
+	upper/lower limits cannot be used at same time as plasticity
+	They are set to +/- maxPlasticStrain to avoid conflict in most cases
+	*/
+	btScalar ul(value);
+	btScalar ll(-value);
+	setLinearUpperLimit(btVector3(ul, ul, ul));
+	setLinearLowerLimit(btVector3(ll, ll, ll));
 }
 btScalar bt6DofElasticPlastic2Constraint::getMaxPlasticStrain(){
 	return m_maxPlasticStrain;

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp

@@ -261,9 +261,12 @@ void bt6DofElasticPlasticConstraint::setMaxPlasticStrain(btScalar value){
 	m_maxPlasticStrain = value;
 	/*
 	upper/lower limits cannot be used at same time as plasticity
+	They are set to +/- maxPlasticStrain to avoid conflict in most cases
 	*/
-	setLinearUpperLimit(btVector3(2 * value, 2 * value, 2 * value));
-	setLinearLowerLimit(btVector3(-2 * value, -2 * value, -2 * value));
+	btScalar ul(value);
+	btScalar ll(-value);
+	setLinearUpperLimit(btVector3(ul,ul,ul));
+	setLinearLowerLimit(btVector3(ll,ll,ll));
 }
 btScalar bt6DofElasticPlasticConstraint::getMaxPlasticStrain(){
 	return m_maxPlasticStrain;

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlasticConstraint.h

@@ -97,7 +97,7 @@ public:
 	virtual btScalar getCurrentPlasticStrain();
 	virtual btScalar getCurrentPlasticRotation();
 	virtual btScalar getMaxRatio(){ return m_maxRatio; }
-	btScalar getFpsLimit(int index){ return m_fpsLimit[i]; }
+	btScalar getFpsLimit(int index){ return m_fpsLimit[index]; }
 	virtual int getMaxRatioDof(){ return m_maxRatioDof; }
 	void updatePlasticity(btJointFeedback& forces, btCollisionWorld* collisionWorld, btScalar step);
 	void calculateFpsLimit(int index);

### 2016-08-01 examples/CharpyDemo/BlockDemo.cpp
debugdraw refactored

@@ -36,6 +36,7 @@ Based on DemolisherDemo by Simo Nikula 2016-
 #include "../plasticity/PlasticityExampleBrowser.h"
 #include "../plasticity/PlasticityData.h"
 #include "../plasticity/PlasticityStatistics.h"
+#include "../plasticity/PlasticityDebugDrawer.h"
 
 #include "BulletDynamics/ConstraintSolver/btSliderConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btGeneric6DofSpringConstraint.h"
@@ -1346,6 +1347,7 @@ void BlockDemo::physicsDebugDraw(int debugFlags)
 	{
 		m_dynamicsWorld->getDebugDrawer()->setDebugMode(debugFlags);
 		m_dynamicsWorld->debugDrawWorld();
+		PlasticityDebugDrawer::drawPlasticConstraints(m_dynamicsWorld);
 	}
 }
 

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp

@@ -46,6 +46,7 @@ http://gimpact.sf.net
 #include "LinearMath/btQuickprof.h"
 #include <new>
 #include "../plasticity/PlasticityData.h"
+#include "../plasticity/PlasticityDebugDrawer.h"
 
 
 
@@ -81,6 +82,7 @@ void bt6DofElasticPlastic2Constraint::resetIdCounter(){
 void bt6DofElasticPlastic2Constraint::initPlasticity()
 {
 	id = ++idCounter;
+	m_objectType = btPlasticConstraintType::BPT_EP2;
 	for (int i = 0; i < 6; i++)
 	{
 		m_maxForce[i] = btScalar(SIMD_INFINITY);
@@ -1245,6 +1247,18 @@ btScalar bt6DofElasticPlastic2Constraint::getMaxPlasticRotation(){
 btScalar bt6DofElasticPlastic2Constraint::getCurrentPlasticRotation(){
 	return m_currentPlasticRotation;
 }
+btTransform & bt6DofElasticPlastic2Constraint::getFrameA(){
+	return m_frameInA;
+}
+btTransform & bt6DofElasticPlastic2Constraint::getFrameB(){
+	return m_frameInB;
+}
+btTransform & bt6DofElasticPlastic2Constraint::getTransformA(){
+	return m_calculatedTransformA;
+}
+btTransform & bt6DofElasticPlastic2Constraint::getTransformB(){
+	return m_calculatedTransformB;
+}
 void bt6DofElasticPlastic2Constraint::scalePlasticity(btScalar scale){
 	m_maxPlasticStrain *= scale;
 	m_maxPlasticRotation *= scale;
@@ -1335,25 +1349,11 @@ void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
 		setEnabled(false);
 	}
 }
-#define BLEN 6 
 void bt6DofElasticPlastic2Constraint::debugDraw(btIDebugDraw* debugDrawer)
 {
-	if (!isEnabled()){
-		return;
-	}
-	char buffer[BLEN];
-	sprintf_s(buffer, BLEN, "%d", id);
-	btVector3 color(1, 0, 0);
-	btVector3 from = m_rbA.getCenterOfMassPosition();
-	btTransform tr = m_rbA.getCenterOfMassTransform();
-	btVector3 to = tr*m_frameInA.getOrigin();
-	debugDrawer->draw3dText(to, buffer);
[…]

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.h

@@ -300,6 +300,10 @@ public:
 	virtual btScalar getCurrentPlasticRotation();
 	virtual btScalar getMaxRatio();
 	virtual int getMaxRatioDof();
+	virtual btTransform & getFrameA();
+	virtual btTransform & getFrameB();
+	virtual btTransform & getTransformA();
+	virtual btTransform & getTransformB();
 	void initPlasticity();
 	void updatePlasticity(btJointFeedback& forces);
 	///btActionInterface interface

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp

@@ -18,6 +18,8 @@ subject to the following restrictions:
 #include "LinearMath/btTransformUtil.h"
 #include "LinearMath/btIDebugDraw.h"
 #include "LinearMath/btQuickprof.h"
+#include "../plasticity/PlasticityDebugDrawer.h"
+#include "../plasticity/PlasticityData.h"
 
 
 bt6DofElasticPlasticConstraint::bt6DofElasticPlasticConstraint(btRigidBody& rbA, btRigidBody& rbB, const btTransform& frameInA, const btTransform& frameInB ,bool useLinearReferenceFrameA)
@@ -41,7 +43,7 @@ void bt6DofElasticPlasticConstraint::resetIdCounter(){
 
 void bt6DofElasticPlasticConstraint::init()
 {
-	m_objectType = D6_SPRING_CONSTRAINT_TYPE;
+	m_objectType = btPlasticConstraintType::BPT_EP;
 
 	for(int i = 0; i < 6; i++)
 	{
@@ -288,6 +290,19 @@ btScalar bt6DofElasticPlasticConstraint::getMaxPlasticRotation(){
 btScalar bt6DofElasticPlasticConstraint::getCurrentPlasticRotation(){
 	return m_currentPlasticRotation;
 }
+btTransform & bt6DofElasticPlasticConstraint::getFrameA(){
+	return m_frameInA;
+}
+btTransform & bt6DofElasticPlasticConstraint::getFrameB(){
+	return m_frameInB;
+}
+btTransform & bt6DofElasticPlasticConstraint::getTransformA(){
+	return m_calculatedTransformA;
+}
+btTransform & bt6DofElasticPlasticConstraint::getTransformB(){
+	return m_calculatedTransformB;
+}
+
 void bt6DofElasticPlasticConstraint::scalePlasticity(btScalar scale){
 	m_maxPlasticStrain *= scale;
 	m_maxPlasticRotation *= scale;
@@ -375,24 +390,10 @@ void bt6DofElasticPlasticConstraint::updatePlasticity(btJointFeedback& forces,
 	}
 }
 
-#define BLEN 6 
 void bt6DofElasticPlasticConstraint::debugDraw(btIDebugDraw* debugDrawer)
 {
-	if (!isEnabled()){
-		return;
-	}
-	char buffer[BLEN];
-	sprintf_s(buffer, BLEN, "%d", id);
-	btVector3 color(1, 0, 0);
[…]

### 2016-08-01 examples/CharpyDemo/bt6DofElasticPlasticConstraint.h

@@ -97,6 +97,10 @@ public:
 	virtual btScalar getCurrentPlasticStrain();
 	virtual btScalar getCurrentPlasticRotation();
 	virtual btScalar getMaxRatio(){ return m_maxRatio; }
+	virtual btTransform & getFrameA();
+	virtual btTransform & getFrameB();
+	virtual btTransform & getTransformA();
+	virtual btTransform & getTransformB();
 	btScalar getFpsLimit(int index){ return m_fpsLimit[index]; }
 	virtual int getMaxRatioDof(){ return m_maxRatioDof; }
 	void updatePlasticity(btJointFeedback& forces, btCollisionWorld* collisionWorld, btScalar step);

### 2016-08-01 examples/CharpyDemo/btElasticPlasticConstraint.h

@@ -17,6 +17,11 @@ public:
 	virtual btScalar getCurrentPlasticStrain() = 0;
 	virtual btScalar getMaxRatio() = 0;
 	virtual int getMaxRatioDof() = 0;
+	virtual int getId() = 0;
+	virtual btTransform & getFrameA() = 0;
+	virtual btTransform & getFrameB() = 0;
+	virtual btTransform & getTransformA() = 0;
+	virtual btTransform & getTransformB() = 0;
 };
 
 #endif //_BT_ELASTIC_PLASTIC_CONSTRAINT_INTERFACE_H

### 2016-08-01 examples/plasticity/PlasticityDebugDrawer.cpp

@@ -1,6 +1,7 @@
 #include "../ExampleBrowser/OpenGLGuiHelper.h"
 #include "btBulletDynamicsCommon.h"
 #include "PlasticityDebugDrawer.h"
+#include "PlasticityData.h"
 
 PlasticityDebugDrawer::PlasticityDebugDrawer(CommonGraphicsApp* app)
 		: m_glApp(app)
@@ -73,5 +74,69 @@ void PlasticityDebugDrawer::flushLines()
 			1);
 		m_linePoints.clear();
 		m_lineIndices.clear();
+	}
+}
+
+#define BLEN 6 
+void PlasticityDebugDrawer::drawPlasticConstraint(btElasticPlasticConstraint* epc,
+	btTypedConstraint* c, btIDebugDraw* debugDrawer){
+	if (!c->isEnabled()){
+		return;
+	}
+	char buffer[BLEN];
+	sprintf_s(buffer, BLEN, "%d", epc->getId());
+	btVector3 color(1, 0, 0);
+	btRigidBody & rbA = c->getRigidBodyA(), &rbB = c->getRigidBodyB();
+	btScalar dbgDrawSize = btScalar(0.3);
+	bool drawTransforms = true;
+	if (!rbA.isStaticObject()){
+		btVector3 from = rbA.getCenterOfMassPosition();
+		btTransform mtra = rbA.getCenterOfMassTransform();
+		btVector3 to = mtra*epc->getFrameA().getOrigin();
+		debugDrawer->draw3dText(to, buffer);
+		debugDrawer->drawLine(from, to, color);
+		if (drawTransforms){
+			dbgDrawSize *= from.distance(to);
+		}
+	}
+	if (!rbB.isStaticObject()){
+		btVector3 from = rbB.getCenterOfMassPosition();
+		btTransform mtrb = rbB.getCenterOfMassTransform();
+		btVector3 to = mtrb*epc->getFrameB().getOrigin();
+		debugDrawer->draw3dText(to, buffer);
+		debugDrawer->drawLine(from, to, color);
+		if (drawTransforms){
+			dbgDrawSize *= from.distance(to);
+		}
+	}
+	if (drawTransforms){
+		btTransform cta = epc->getTransformA();
+		debugDrawer->drawTransform(cta, dbgDrawSize);
+		btTransform ctb = epc->getTransformB();
+		debugDrawer->drawTransform(ctb, dbgDrawSize);
+	}
+}
[…]

### 2016-08-01 examples/plasticity/PlasticityDebugDrawer.h

@@ -3,6 +3,7 @@
 
 #include "../ExampleBrowser/OpenGLGuiHelper.h"
 #include "../CommonInterfaces/CommonGraphicsAppInterface.h"
+#include "../CharpyDemo/btElasticPlasticConstraint.h"
 #include "btBulletDynamicsCommon.h"
 
 #define BT_LINE_BATCH_SIZE 512
@@ -47,5 +48,8 @@ public:
 	virtual void	setDebugMode(int debugMode);
 	virtual void flushLines();
 	virtual int		getDebugMode() const;
+	static void drawPlasticConstraint(btElasticPlasticConstraint* epc, 
+		btTypedConstraint* c, btIDebugDraw* debugDrawer);
+	static void drawPlasticConstraints(btDiscreteDynamicsWorld* dw);
 };
 #endif // PLASTICITY_DEBUG_DRAWER

### 2016-08-01 examples/plasticity/PlasticityExampleBrowser.cpp

@@ -580,7 +580,7 @@ bool PlasticityExampleBrowser::init(int argc, char* argv[])
 #endif
 #define BLEN 256
 	char buf[BLEN];
-	sprintf_s(buf, BLEN, "0.9.1 - %d-%02d-%02d - %s",
+	sprintf_s(buf, BLEN, "0.9.2 - %d-%02d-%02d - %s",
 		tm.tm_year + 1900, tm.tm_mon + 1, tm.tm_mday,
 		btScalarType);
 #if defined (_DEBUG) || defined (DEBUG)

### 2016-08-02 examples/CharpyDemo/BlockDemo.cpp
timeSeries added to BlockDemo

@@ -37,7 +37,6 @@ Based on DemolisherDemo by Simo Nikula 2016-
 #include "../plasticity/PlasticityData.h"
 #include "../plasticity/PlasticityStatistics.h"
 #include "../plasticity/PlasticityDebugDrawer.h"
-
 #include "BulletDynamics/ConstraintSolver/btSliderConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btGeneric6DofSpringConstraint.h"
 #include "BulletDynamics/ConstraintSolver/btGeneric6DofSpring2Constraint.h"
@@ -51,6 +50,7 @@ Based on DemolisherDemo by Simo Nikula 2016-
 #include "../CommonInterfaces/CommonGraphicsAppInterface.h"
 #include "../ExampleBrowser/GwenGUISupport/gwenUserInterface.h"
 #include "../ExampleBrowser/GwenGUISupport/gwenInternalData.h"
+#include "../RenderingExamples/TimeSeriesCanvas.h"
 const char * PROFILE_TENSION_SLEEP = "BlockDemo::Sleep";
 
 class BlockDemo : public Gwen::Event::Handler, public CommonRigidBodyBase
@@ -94,7 +94,8 @@ public:
 	btScalar damping=0.9; // 1= no damping, 0=full damping
 	btScalar damping2 = 0.1; // 0= no damping, 1=critically damped, >1 overdamped
 	bool hasFullGravity;
-	btScalar m_fixedTimeStep = btScalar(1) / btScalar(60);
+	int intFreq = 60;
+	btScalar m_fixedTimeStep = btScalar(1) / btScalar(intFreq);
 	btScalar blockSteelScale;
 	btScalar density;
 	btScalar maxPlasticRotation;
@@ -229,35 +230,20 @@ public:
 			PlasticityExampleBrowser::getRenderer()->getActiveCamera();
 		camera->setCameraTargetPosition(scl.x(), scl.y(), scl.z());
 	}
-	/** update fps and kmph after interval seconds have passed */
-	float speedometerUpdated;
-	float updateInterval = 0.3;
-	int updateViewCount = 0;
 	void resetClocks(){
 		driveClock.reset();
-		speedometerUpdated = 0;
 	}
 	void updateView(){
-		updateViewCount++;
[…]

### 2016-08-02 examples/plasticity/PlasticityExampleBrowser.cpp

@@ -491,7 +491,7 @@ struct QuickCanvas : public Common2dCanvasInterface
 		:m_myTexLoader(myTexLoader),
 		m_curNumGraphWindows(0)
 	{
-		for (int i=0;i<MAX_GRAPH_WINDOWS;i++)
+		for (int i = 0; i<MAX_GRAPH_WINDOWS; i++)
 		{
 			m_gw[i] = 0;
 			m_gt[i] = 0;
@@ -503,45 +503,53 @@ struct QuickCanvas : public Common2dCanvasInterface
 		if (m_curNumGraphWindows<MAX_GRAPH_WINDOWS)
 		{
 			//find a slot
-			int slot = 0;//hardcoded for now
+			int slot = m_curNumGraphWindows;
+			btAssert(slot<MAX_GRAPH_WINDOWS);
+			if (slot >= MAX_GRAPH_WINDOWS)
+				return 0;//don't crash
+
 			m_curNumGraphWindows++;
 
 			MyGraphInput input(gui->getInternalData());
-			input.m_width=width;
-			input.m_height=height;
-			input.m_xPos = 300;
-			input.m_yPos = height-input.m_height;
-			input.m_name=canvasName;
+			input.m_width = width;
+			input.m_height = height;
+			input.m_xPos = 10000;//GUI will clamp it to the right//300;
+			input.m_yPos = 10000;//GUI will clamp it to bottom
+			input.m_name = canvasName;
 			input.m_texName = canvasName;
 			m_gt[slot] = new GraphingTexture;
-			m_gt[slot]->create(width,height);
+			m_gt[slot]->create(width, height);
 			int texId = m_gt[slot]->getTextureId();
 			m_myTexLoader->m_hashMap.insert(canvasName, texId);
 			m_gw[slot] = setupTextureWindow(input);
-			
+
 			return slot;
 		}
 		return -1;
 	}
 	virtual void destroyCanvas(int canvasId)
 	{
-		btAssert(canvasId==0);//hardcoded to zero for now, only a single canvas
-		btAssert(m_curNumGraphWindows==1);
+		btAssert(canvasId >= 0);
+		delete m_gt[canvasId];
+		m_gt[canvasId] = 0;
 		destroyTextureWindow(m_gw[canvasId]);
+		m_gw[canvasId] = 0;
 		m_curNumGraphWindows--;
 	}
[…]

### 2016-08-03 pdocs/thesis/article-section-2.tex
velocity based formulation basics added to article using erleben as source

@@
+% !TeX root = article.tex
 \section{Adding plasticity to physics engine}
@@
 Acceleration-based formulations need to stop at collision and switch to an impulse based method.
+
+Background
+for velocity based formulation shown here is based on  \citet[p.~45-50]{erleben.thesis}.
+% pdf page 64
+In following section, these formulations will be clarified by simple example using \bullet\ implementation.
+Assuming that force $\vec{f}_{true}(t)$, is known the impulse $\vec{J}$
+in the time interval $\Delta t $ can we written as
+\begin{equation} \label{eq:impulseIntegral}
+\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt
+\end{equation}
+Using Newton's second law of motion $F=ma$ one can solve for the velocity, $\vec{v}^{\Delta t}$
+\begin{equation} \label{eq:impulseIntegraWithNewton}
+\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
+\end{equation}
+\begin{equation} \label{eq:impulse}
+m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J}
+\end{equation}
+where superscripts denote time, i.e. ${v}^{\Delta t}=v(\Delta t)$. Next position can be found
+by integrating the velocity.
+Derivation of equations of motion is not shown here in detail but results 
+for simple contact can be summarized as
+equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
+Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom}.
+\begin{equation} \label{eq:eomL} % pdf page 69
+\vec{s}^{\, t+\Delta t} = s^{t}+\Delta t S \vec{u}^{\, t+\Delta t}
+\end{equation}
+
+\begin{equation} \label{eq:eomV}
+\vec{u}^{\, t+\Delta t} = u^{t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
+\end{equation}
+
+Friction in contacts and joint constraints can be handled in unified way by refactoring
+equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}.
+Jacobian terms are derived by taking time derivates of kinematic constraints.
+
+\begin{equation} \label{eq:eomV2}
+\vec{u}^{t+\Delta t} = u^{t}+\Delta t M^{-1}(
+ J_{contact}^T \vec{\lambda}_{contact}
++ J_{joint}^T \vec{\lambda}_{joint}
++ \vec{f}_{ext})
+\end{equation}
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (O1) at(2,1);
+\coordinate (O2) at(2,4);
+\coordinate (C) at(2,2);
+\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
+\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
+\node at (3.5,0.4) {$B_1$};
+\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
+\node at (2.8,4.5) {$B_2$};
+\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
+\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
+\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
+\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
+\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
+\end{tikzpicture}
+\caption{Illustration of nomenclature for equations of motion.}
+\label{fig:eom}
+\end{figure}
+
+% pdf page 33, notation in typical ODEs
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l| l|}
+\hline
+{\bf Symbol} & {\bf Description} \\  \hline
+$\vec{r}_i$ & position of center of mass for body i  \\ \hline
+$\vec{q}_i$ & orientation as a quaternion for body i  \\ \hline
+$\vec{p}_i$ & contact or joint point i  \\ \hline
+$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
+$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
+$S$ & generalized transformation matrix  \\ \hline
+$\vec{v}_i$ & linear velocity of  center of mass for body i  \\ \hline
+$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
+$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
+$M$ & generalized mass matrix  \\ \hline
+$C$ & contact condition matrix  \\ \hline
+$N$ & contact normal matrix  \\ \hline
+$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
+$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
+$J_{joint}$ & Jacobian matrix for joints  \\ \hline
+$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
+\end {tabular}
+\end{center}
+\caption {Nomenclature for equations of motion} \label{tab:eom}
+\end {table}
+
 In structural analysis, a formulation and associated numerical solution procedure are selected 
@@
 
-Strain hardening is taken into account in this work mainly by assuming that plasticity in bending expands, \citet[p.~672]{dowling}.
+Strain hardening is taken into account in this work mainly by assuming that plasticity in bending
+expands, \citet[p.~672]{dowling}.
 Material that starts to yield first is hardened and yielding moves slightly.
@@
 {\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
+$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
 $J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline

### 2016-08-04 pdocs/thesis/article-section-3.tex
article-section-3 typo fixes
Tension:None documented to article

@@
 
-In this section, changes to constraint formulation are clarified by doing few simulation steps of rigid and
+In this section, changes to constraint formulation are clarified by doing few simulation steps of non constrained, rigid and
 elastic-plastic examples.
@@
 
-\draw[->] (0.5,3,0.5) -- ++(0,1.5,0);
+\draw[-Stealth] (0.5,3,0.5) -- ++(0,1.5,0);
 \end{tikzpicture}
@@
 so that it reflects wanted scenario.
-Concrete density is $2000 \frac{kg}{m^3}$ and steel density id $7800 \frac{kg}{m^3}$. 
+Concrete density is $2000 \frac{kg}{m^3}$ and steel density is $7800 \frac{kg}{m^3}$. 
 Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
@@
 iterations are done for single step.
+Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t=0.548 s 
+($\sqrt{\frac{2 \cdot 1.5}{10}}$).
 
@@
 \begin{enumerate}
-\item Apply gravity
-\item Predict unconstrained motion
-\item Predict contacts
-\item Perform collision detection
+\item Apply gravity to each non static body. 
+This step just makes programming easier as otherwise each program should add forces due to gravity in each step. 
+In this case, $\vec{f}_{ext}$ gets added by $\lbrace{0,-60000,0}\rbrace$.
+\item Predict unconstrained motion for each non static body. Prediction is done based on bodys current linear and angular velocity.
+\item Predict contacts.
+\item Perform collision detection. All body pairs created in previous step are processed and manifoldPoints are created for
+each detected contact. In unconstrained case, contact is detected after t=0.55 s and manifolds distance gets negative value (-0.058 m).
 \item Calculate simulation islands(groups). Bodies that are near each other or connected with constraints are grouped in same group.
 \item Solve constraints. Both contact and other constraints are processed in this step.
-\item Integrate transforms
+\item Integrate transforms using velocities modified in previous step.
 \item Update actions (callbacks) are called. 
@@
 Equation \ref{eq:constraintEquation} is simplified
-to \ref{eq:fixedConstraint} in these cases.
+to \ref{eq:fixedConstraint} in constrained cases.
 \begin{itemize}
@@
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ was used to pick up values. Actual values for fixed constraint are shown in table
-\ref{tab:fixedBlockValues} and values for elastic-plastic case are shown in tables \ref{tab:epBlockValues} and 
- \ref{tab:ep2BlockValues}. There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
-in this work elastic-plastic versions of both of them were developed. 
-Constraint activation of older spring and fixed constraints is implemented so that 
-they are activated after constraint is violated.
-Thus body drops freely during first simulation step and 
-gains enough kinetic energy so that higher impulses are needed in few following steps.
+in \bullet\ was used to pick up values for internal variables.
 
@@
 \item[velError] is calculated using velocities and external impulses of connected bodies. 
- In this case, it is dominant contributor for constraint's impulse.
+ \begin{description}
+\item[In constraint cases,] it is dominant contributor for constraint's impulse.
+\item[In penetrating contact case,] main contributor is bodies relative speed at point of contact. 
+\end{description}
 \item[posError] is calculated by constraint. It is significant factor in designing stable constraints.
@@
 spring simulation cannot be done in stable way.
+ \item[In contact case,] value is zero if there is no penetration. For penetration cases it is 
+$\frac{-penetration\, erp}{timeStep}$
  \end{description}
@@
 \item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
-In this case, it is mass of body (6000).
+ \begin{description}
+\item[In constraint cases,] it is mass of body (6000).
+\item[In contact case,] it varies below mass of body.
 \end{description}
+\item[Impulse]
+ \begin{description}
+\item[In constraint cases,] it is obtained from btJointFeedback structure.
+\item[In contact case,] it is obtained by summing applied impulses from active manifolds.
+\end{description}
+\end{description}
+
+Actual values for unconstrained case 
+are shown in \ref{tab:freeBlockValues}.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l| l|l| l|l|l|l|l|}
+\hline
+{\bf Time} & 
+{\bf Location} &
+{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
+{\bf Velocity} & 
+{\bf Impulse} \\  \hline
+0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
+0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
+0.567 &  -1.511 & 5.66 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
+0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
+0.600 &  -1.496 & -0.38&-0.002 & 0.1  & -1000& 0.38 & 0 \\  \hline
+0.617 &  -1.492 &-0.44 & 0.004 & 0     & -1600& 0.22 & 0 \\  \hline
+0.717 &  -1.497 &0.004-0.08  &-0.0003-0.001 &0-0.01 & 10-400 & -0.01 & 400 \\  \hline
+0.817 &  -1.499 & & & & & -0.08 & 700 \\  \hline
+0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
+\end {tabular}
+\end{center}
+\caption {Simulation values for unconstrained case. 
+For internal contact values typical values are shown
+as number of contacts and detailed values differ.} 
+\label{tab:freeBlockValues} 
+\end {table}
+
+Values for fixed constraint are shown in table
+\ref{tab:fixedBlockValues}.
 
@@
 {\bf Time} & 
-{\bf velError} & {\bf posError} & {\bf rhs} &
-{\bf velocity} & 
-{\bf impulse} \\  \hline
+{\it velError} & {\it posError} & {\it rhs} &
+{\bf Velocity} & 
+{\bf Impulse} \\  \hline
 0.017 &  & & & -0.17 & 0 \\  \hline
@@
 
+Values for elastic-plastic case are shown in tables \ref{tab:epBlockValues} and 
+ \ref{tab:ep2BlockValues}. 
+There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
+in this work elastic-plastic versions of both of them were developed. 
+Constraint activation of older spring and fixed constraints is implemented so that 
+they are activated after constraint is violated.
+Thus body drops freely during first simulation step and 
+gains enough kinetic energy so that higher impulses are needed in few following steps.
+
 \begin {table}[htb!]
@@
 {\bf Time} & 
-{\bf velError} & {\bf posError} & {\bf rhs} &
-{\bf velocity} & 
-{\bf impulse} & 
-{\bf plastic strain} \\  \hline
+{\it velError} & {\it posError} & {\it rhs} &
+{\it velocity} & 
+{\bf Impulse} & 
+{\bf Plastic strain} \\  \hline
 0.017 &          &    &          & -0.17 & 0 & 0 \\  \hline
@@
 {\bf Time} & 
-{\bf velError} & {\bf posError} & {\bf rhs} &
-{\bf velocity} & 
-{\bf impulse} & 
-{\bf plastic strain} \\  \hline
+{\it velError} & {\it posError} & {\it rhs} &
+{\bf Velocity} & 
+{\bf Impulse} & 
+{\bf Plastic strain} \\  \hline
 0.017... & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline

### 2016-08-04 examples/CharpyDemo/BlockDemo.cpp

@@ -233,11 +233,31 @@ public:
 	void resetClocks(){
 		driveClock.reset();
 	}
+	btScalar sumContactImpulses(){
+		btScalar totalImpact = btScalar(0);
+		btDispatcher *dispatcher = m_dynamicsWorld->getDispatcher();
+		int numManifolds = dispatcher->getNumManifolds();
+		for (int i = 0; i < numManifolds; i++)
+		{
+			btPersistentManifold* contactManifold = dispatcher->getManifoldByIndexInternal(i);
+			int numContacts = contactManifold->getNumContacts();
+			for (int j = 0; j < numContacts; j++)
+			{
+				btManifoldPoint& pt = contactManifold->getContactPoint(j);
+				totalImpact += pt.m_appliedImpulse;
+			}
+		}
+		return totalImpact;
+	}
 	void updateView(){
 		if (m_constraint){
 			c_y_force = jf.m_appliedForceBodyB[1];
 			c_impulse = btFabs(c_y_force*m_fixedTimeStep);
 			c_percent = 100.*c_impulse/maxImpulse;
+		}
+		else{
+			c_impulse = sumContactImpulses();
+			c_y_force = c_impulse / m_fixedTimeStep;
 		}
 		if (m_body){
 			b_y_location = m_body->getWorldTransform().getOrigin().y() - yStart;
@@ -1055,7 +1075,7 @@ public:
 			tsYVelocity = new TimeSeriesCanvas
 				(app->m_2dCanvasInterface, tsWidth, tsHeight, "Y velocity [m/s]");
 		}
-		tsYVelocity->setupTimeSeries(5, intFreq, 0);
+		tsYVelocity->setupTimeSeries(6, intFreq, 0);
 		tsYVelocity->addDataSource("",0 , 255, 0);
 		btScalar maxForce = axisMapper->getMaxForce(1);
 		if (0 == tsYForce){

### 2016-08-05 examples/CharpyDemo/BlockDemo.cpp
ccd studies

@@ -993,6 +993,11 @@ public:
 		btVector3 pos = btVector3(0, yStart, 0);
 		loadTrans.setOrigin(pos);
 		m_body=localCreateRigidBody(mass, loadTrans, loadShape);
+		bool useCcd = true;
+		if (useCcd){
+			m_body->setCcdMotionThreshold(lsy / 10000);
+			m_body->setCcdSweptSphereRadius(lsy / 10);
+		}
 		axisMapper=new AxisMapper(lsx, lsy, lsz, cpos);
 		axisMapper->setE(E*blockSteelScale);
 		axisMapper->setFy(fy*blockSteelScale);
@@ -1356,14 +1361,14 @@ void BlockDemo::initPhysics()
 	stepTime = 0;
 	int upAxis = 1;
 	m_guiHelper->setUpAxis(upAxis);
-	btVector3 groundExtents(200,200,200);
+	btVector3 groundExtents(2*lsx,2*lsy,2*lsz);
 	groundExtents[upAxis]=3;
 	btCollisionShape* groundShape = new btBoxShape(groundExtents);
 	m_collisionShapes.push_back(groundShape);
 	m_collisionConfiguration = new btDefaultCollisionConfiguration();
 	m_dispatcher = new btCollisionDispatcher(m_collisionConfiguration);
-	btVector3 worldMin(-1000,-1000,-1000);
-	btVector3 worldMax(1000,1000,1000);
+	btVector3 worldMin(-3*lsx,-3*lsy,-3*lsz);
+	btVector3 worldMax(3 * lsx, 3 * lsy, 3 * lsz);
 	m_overlappingPairCache = new btAxisSweep3(worldMin,worldMax);
 	if (useMCLPSolver)
 	{

### 2016-08-11 pdocs/thesis/article-section-3.tex
ccd parameters added to BlockDemo UI

@@
 \item Predict unconstrained motion for each non static body. Prediction is done based on bodys current linear and angular velocity.
-\item Predict contacts.
-\item Perform collision detection. All body pairs created in previous step are processed and manifoldPoints are created for
-each detected contact. In unconstrained case, contact is detected after t=0.55 s and manifolds distance gets negative value (-0.058 m).
-\item Calculate simulation islands(groups). Bodies that are near each other or connected with constraints are grouped in same group.
+\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
+Each rigid body is represented by sphere and contact prediction is done if body moves more than given threshold value during simulation step.
+\item Perform discrete collision detection. All overlapping body pairs are processed and manifoldPoints are created for
+each detected contact at end of simulation step. 
+In unconstrained case, contact is detected after t=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
+\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
+ or connected with constraints are grouped in same group.
 \item Solve constraints. Both contact and other constraints are processed in this step.
@@
 
-Actual values for unconstrained case 
+Actual values for unconstrained case without continuous collision detection
 are shown in \ref{tab:freeBlockValues}.

### 2016-08-11 examples/CharpyDemo/BlockDemo.cpp

@@ -109,6 +109,8 @@ public:
 	void shoot();
 	btBoxShape* ammoShape = 0;
 	btVector4 ammoColor = btVector4(0.5, 0.5, 0.5, 0.5);
+	btScalar loadCcdSweptSphereRadius,loadCcdMotionThreshold;
+	btScalar ammoCcdSweptSphereRadius, ammoCcdMotionThreshold;
 	char *logDir = _strdup("d:/wrk");
 	int gx = 10; // for labels
 	int gxi = 120; // for inputs elements
@@ -419,6 +421,10 @@ public:
 	void setDensity(Gwen::Controls::Base* control);
 	void setE(Gwen::Controls::Base* control);
 	void setFrequencyRatio(Gwen::Controls::Base* control);
+	void setLoadCcdSweptSphereRadius(Gwen::Controls::Base* control);
+	void setLoadCcdMotionThreshold(Gwen::Controls::Base* control);
+	void setAmmoCcdSweptSphereRadius(Gwen::Controls::Base* control);
+	void setAmmoCcdMotionThreshold(Gwen::Controls::Base* control);
 	void setFy(Gwen::Controls::Base* control);
 	void setBlockSteelScale(Gwen::Controls::Base* control);
 	void setMaxPlasticStrain(Gwen::Controls::Base* control);
@@ -515,6 +521,35 @@ public:
 		useCcdGc = gc;
 		gy += gyInc;
 		gc->onCheckChanged.Add(pPage, &BlockDemo::setUseCcd);
+	}
+	template <typename T>
+	void addScalar(string label, const string & tooltip, btScalar value, T f){
+		addLabel(label);
+		Gwen::Controls::TextBoxNumeric* gc = 
+			new Gwen::Controls::TextBoxNumeric(pPage);
+		string text;
+		if (value < 10e-3){
+			text = uif(value, "%.6f");
+		}
+		else{
+			text = uif(value);
+		}
+		gc->SetToolTip(tooltip);
+		gc->SetText(text);
+		gc->SetPos(gxi, gy);
+		gc->SetWidth(wxi);
+		gy += gyInc;
+		gc->onReturnPressed.Add(pPage, f);
+	}
+	void addCcdValues(){
+		addScalar("loadCcdSphere","Size of ccd sphere for load",
+			loadCcdSweptSphereRadius,&BlockDemo::setLoadCcdSweptSphereRadius);
[…]

### 2016-08-12 pdocs/thesis/article-section-3.tex
article section 3 tuning

@@
 Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t=0.548 s 
-($\sqrt{\frac{2 \cdot 1.5}{10}}$).
+($\sqrt{2 \cdot 1.5 / 10}$).
 
@@
 In this case, $\vec{f}_{ext}$ gets added by $\lbrace{0,-60000,0}\rbrace$.
-\item Predict unconstrained motion for each non static body. Prediction is done based on bodys current linear and angular velocity.
+\item Predict unconstrained motion for each non static body. Prediction is done based on current linear and angular velocities of the body.
 \item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
 Each rigid body is represented by sphere and contact prediction is done if body moves more than given threshold value during simulation step.
+Continuous collision detection is configured for each body and as default it is not active.
+Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
 \item Perform discrete collision detection. All overlapping body pairs are processed and manifoldPoints are created for
@@
  \end{description}
-\item[rhs] is calculated by velError jInv + posError jInv
+\item[rhs (c)] is calculated by velError jInv + posError jInv
 \item[jInv] is calculated using masses and inertias of connected bodies and constraint geometry. 
@@
 \end{description}
-\item[Impulse]
+\item[Impulse] is impulse applied to body during timestep.
  \begin{description}
@@
 Actual values for unconstrained case without continuous collision detection
-are shown in \ref{tab:freeBlockValues}.
+are shown in \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s
+{\it velError}
 

### 2016-08-13 pdocs/thesis/article-section-3.tex
article section 3 tuning continues

@@
 In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
-\item Activation state of bodies is updated. Bodies that are not active
- are typically marked as sleeping and they are not processed.
+\item Activation state of bodies is updated. 
+To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
+given threshold values (default 0.8 and 1.0) more than set limit (2.0).
 \end{enumerate}
@@
 \item[In constraint cases,] it is dominant contributor for constraint's impulse.
-\item[In penetrating contact case,] main contributor is bodies relative speed at point of contact. 
+\item[In contact case,] main contributor is bodies relative speed at point of contact. 
+Bodies are not allowed to penetrate each other.
 \end{description}
@@
  \item[In fixed case,] value is about 12 times actual position error. Factor 12 is based on time step (60) 
- and default value of error reduction parameter (0.2).
+ and default value of error reduction parameter (erp) which has value of 0.2 in this context.
  \item[In elastic plastic case,]  value is set to zero if impulse would be larger than maximum impulse or
@@
 \end{description}
+\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. object drifting. 
+Setting erp to 1 would in theory eliminate error but in practice value of 0.8 is used in most cases.
 \end{description}
@@
 Actual values for unconstrained case without continuous collision detection
-are shown in \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s
-{\it velError}
+are shown in \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
+{\it velError} is 5.67 (5.5+0.167) and
+{\it posError} is 2.8 (0.058*0.8/0.0167). Impulse is 34000 and contact force is thus about 2 MN (34000/0.0167).
+After few steps location and position stabilize although internally calculation is needed for each time step
+until body is deactivated.
 
@@
 0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
-0.567 &  -1.511 & 5.66 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
+0.567 &  -1.511 & 5.67 &-0.058 &2.8 &  21270 & 0.01 & 34000 \\  \hline
 0.583 &  -1.502 & 0.14 &-0.011 & 0.54& 2570  & 0.55 & 420 \\  \hline
@@
 
-Values for fixed constraint are shown in table
-\ref{tab:fixedBlockValues}.
+Actual values for unconstrained case with continuous collision detection (ccd) using 1.5 
+as radius of ccd sphere and 0.001 as ccd motion threshold
+are shown in \ref{tab:freeBlockValuesWithCcd}. Collision is detected at time 0.550 s when
+{\it velError} is  3.5 (5.34+0.167)-0.033/0.0167 and
+{\it posError} is  0 as collision is detected before penetration. 
+It should be noted that in general ccd sphere should not extend actual body as 
+premature contacts are created if collision takes place in those regions.
 
@@
 \begin{center}
-\begin{tabular}{|l| l| l|l|l|l|}
+\begin{tabular}{|l| l|l| l|l|l|l|l|}
 \hline
 {\bf Time} & 
+{\bf Location} &
+{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
+{\bf Velocity} & 
+{\bf Impulse} \\  \hline
+0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
+0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline
+0.567 &  -1.500 & 2.17 & 0 & 0  &  8100 & 0.01 & 13000 \\  \hline
+0.583 &  -1.500 & 0.15 & 0 & 0 & 600  & 0 & 937 \\  \hline
+0.600 &  -1.500 & 0.17 & 0 & 0 & 600  & 0 & 1040 \\  \hline
+\end {tabular}
+\end{center}
+\caption {Simulation values for unconstrained case with continuous collision detection. } 
+\label{tab:freeBlockValuesWithCcd} 
+\end {table}
+
+Values for fixed constraint are shown in table
+\ref{tab:fixedBlockValues}. Constraint is activated in second step and positional error is corrected
+about 20 \% in each step as requested by using erp value 0.2.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l|l| l| l|l|l|l|}
+\hline
+{\bf Time} & 
+{\bf Location} &
 {\it velError} & {\it posError} & {\it rhs} &
@@
 {\bf Impulse} \\  \hline
-0.017 &  & & & -0.17 & 0 \\  \hline
-0.033 &  0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
-0.050 &  -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
-0.067 &  -0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
-0.35... &  -0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
+0.017 & -0.0028 &  & & & -0.17 & 0 \\  \hline
+0.033 & -0.0022 & 0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
+0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline
+0.067 & -0.0014 &-0.14 & -0.021 & -970 & 0.021 & 970 \\  \hline
+0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
 \end {tabular}
@@
 
-Values for elastic-plastic case are shown in tables \ref{tab:epBlockValues} and 
- \ref{tab:ep2BlockValues}. 
 There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
 in this work elastic-plastic versions of both of them were developed. 
-Constraint activation of older spring and fixed constraints is implemented so that 
-they are activated after constraint is violated.
-Thus body drops freely during first simulation step and 
+Values for elastic-plastic case are shown in tables \ref{tab:epBlockValues}.
+Body drops freely during first simulation step and 
 gains enough kinetic energy so that higher impulses are needed in few following steps.
+This causes plastic strain during next three steps.
 
@@
 \begin{center}
-\begin{tabular}{|l| l| l|l|l|l|l|}
+\begin{tabular}{|l|l| l| l|l|l|l|l|}
 \hline
 {\bf Time} & 
+{\bf Location} &
 {\it velError} & {\it posError} & {\it rhs} &
@@
 {\bf Plastic strain} \\  \hline
-0.017 &          &    &          & -0.17 & 0 & 0 \\  \hline
-0.033 &  -0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
-0.050 &  -0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
-0.067 &  -0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
-0.083... &  -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
+0.017 & -0.0028 &        &    &          & -0.17 & 0 & 0 \\  \hline
+0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
+0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
+0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
+0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
 \end {tabular}
@@
 
+Six-dof-spring constraint 2 has feature to avoid unstability by automatically softening constraint
+spring if needed.
+If this feature is active for elastic-plastic constraint 2.
+ \ref{tab:ep2BlockValues}.
+
 \begin {table}[htb!]
 \begin{center}
-\begin{tabular}{|l| l| l|l|l|l|l|}
+\begin{tabular}{|l|l| l| l|l|l|l|l|}
 \hline
 {\bf Time} & 
+{\bf Location} &
 {\it velError} & {\it posError} & {\it rhs} &
@@
 {\bf Plastic strain} \\  \hline
-0.017... & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
+0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
 \end {tabular}

### 2016-08-13 examples/CharpyDemo/BlockDemo.cpp

@@ -103,7 +103,7 @@ public:
 	btScalar maxPlasticStrain;
 	bool disableCollisionsBetweenLinkedBodies = true;
 	bool dumpPng = false;
-	bool useCcd;
+	bool useCcd, limitIfNeeded;
 	int shootCount = 0;
 	btScalar ammoVelocity;
 	void shoot();
@@ -439,6 +439,7 @@ public:
 	void setLogDir(Gwen::Controls::Base* control);
 	void setDumpPng(Gwen::Controls::Base* control);
 	void setUseCcd(Gwen::Controls::Base* control);
+	void setLimitIfNeeded(Gwen::Controls::Base* control);
 	Gwen::Controls::Base* pPage;
 	Gwen::Controls::Button* pauseButton, *singleStepButton;
 	Gwen::Controls::Label *db11, *db12, *db13, *db14, *db15, *db16;
@@ -512,15 +513,26 @@ public:
 		gy += gyInc;
 		gc->onCheckChanged.Add(pPage, &BlockDemo::setDumpPng);
 	}
-	Gwen::Controls::CheckBox* useCcdGc;
-	void addUseCcd(){
-		Gwen::Controls::Label* label = addLabel("useCcd");
+	template <typename T>
+	void addCheckBox(string label, 
+			const string & tooltip, bool value, T f){
+		addLabel(label);
 		Gwen::Controls::CheckBox* gc = new Gwen::Controls::CheckBox(pPage);
 		gc->SetPos(gxi, gy);
-		gc->SetChecked(useCcd);
-		useCcdGc = gc;
+		gc->SetChecked(value);
 		gy += gyInc;
-		gc->onCheckChanged.Add(pPage, &BlockDemo::setUseCcd);
+		gc->onCheckChanged.Add(pPage,f);
+		gc->SetToolTip(tooltip);
+	}
+	void addUseCcd(){
+		addCheckBox("useCcd", 
+			"Use continous collision detection",
+			useCcd, &BlockDemo::setUseCcd);
+	}
+	void addLimitIfNeeded(){
+		addCheckBox("limitIfNeeded", 
+			"limit spring stiffness if needed",
+			limitIfNeeded, &BlockDemo::setLimitIfNeeded);
 	}
 	template <typename T>
 	void addScalar(string label, const string & tooltip, btScalar value, T f){
@@ -823,6 +835,9 @@ public:
 			addFrequencyRatio();
 			// nobreak
[…]

### 2016-08-14 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp
article section 3 tuning continues

@@ -1290,7 +1290,7 @@ void bt6DofElasticPlastic2Constraint::updatePlasticity(btJointFeedback& forces){
 				m_maxRatio = ratio;
 				m_maxRatioDof = i;
 			}
-			if (absForce>m_maxForce[i]){
+			if (absForce>=m_maxForce[i]){
 				btScalar elasticPart = m_maxForce[i] / m_linearLimits.m_springStiffness[i];
 				btScalar newVal = currPos;
 				if (btFabs(elasticPart)<btFabs(currPos)){

### 2016-08-16 pdocs/thesis/article-section-3.tex
article section 3 tuning continues
article section 3 tuning
article section 3 tuning - constraint parameters

@@
 in this work elastic-plastic versions of both of them were developed.
+
+Table  \ref{tab:ep2Parameters} summarizes most significant parameters for this study.
+There are also e.g. parameters for controlling joint motors.
+Additional equations are created if needed. Enabled springs and motors add one row and
+limits add one if upper and lower are same and two if both upper and lower limits are defined.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l|l| l| l|l|l|l|l|}
+\hline
+{\bf Parameter} & 
+{\bf Description} 
+lowerLimit &  
+minimum allowed translation or rotation
+ \\  \hline
+upperLimit &  
+maximum allowed translation or rotation
+ \\  \hline
+springStiffness & elastic spring stiffness
+ \\  \hline
+enableSpring & defines if spring is active
+ \\  \hline
+springStiffnessLimited & should elastic behaviour be tuned to avoid instability
+ \\  \hline
+equilibriumPoint & should elastic behaviour be tuned to avoid instability
+ \\  \hline
+currentLimit & describes state of constraint - 0: not limited,
+ 3: loLimit=hiLimit, 4: current value is between loLimit and HiLimit
+ \\ \hline
+
+\end {tabular}
+\end{center}
+\caption {Selected constraint parameters for elastic-plastic constraint 2} \label{tab:ep2Parameters} 
+\end {table}
+
 Values for elastic-plastic case are shown in tables \ref{tab:epBlockValues}.
@@
 spring if needed.
-If this feature is active for elastic-plastic constraint 2.
+If this feature is active for elastic-plastic constraint 2 constraint it behaves as shown in table
  \ref{tab:ep2BlockValues}.
@@
 {\bf Plastic strain} \\  \hline
-0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
+0.017... &  0 & -0.17  & 2.34 & 13000 & 0       & 1000 & 0 \\  \hline
 \end {tabular}

### 2016-08-16 examples/CharpyDemo/bt6DofElasticPlastic2Constraint.cpp
free linear limits low>max in setMaxPlasticStrain

@@ -1229,8 +1229,8 @@ void bt6DofElasticPlastic2Constraint::setMaxPlasticStrain(btScalar value){
 	*/
 	btScalar ul(value);
 	btScalar ll(-value);
-	setLinearUpperLimit(btVector3(ul, ul, ul));
-	setLinearLowerLimit(btVector3(ll, ll, ll));
+	setLinearUpperLimit(btVector3(ll, ll, ll));
+	setLinearLowerLimit(btVector3(ul, ul, ul));
 }
 btScalar bt6DofElasticPlastic2Constraint::getMaxPlasticStrain(){
 	return m_maxPlasticStrain;

### 2016-08-18 pdocs/thesis/article-section-3.tex
article section 3 tuning - explain ep2 behaviour

@@
 Single simulation step is done using following substeps. 
-In this work, changes are made only to constraing solving and update actions.
+In this work, changes are made only to constraint setup and update actions.
 \begin{enumerate}
@@
 This step just makes programming easier as otherwise each program should add forces due to gravity in each step. 
-In this case, $\vec{f}_{ext}$ gets added by $\lbrace{0,-60000,0}\rbrace$.
+In this case, $\vec{f}_{ext}$ in  \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
 \item Predict unconstrained motion for each non static body. Prediction is done based on current linear and angular velocities of the body.
@@
 \item Solve constraints. Both contact and other constraints are processed in this step. 
+ This step is subdivided to setup, iterations and finish phases. 
+In setup phase constraints are queried for positional errors for calculation of c in
+equation \ref{eq:constraintEquation} and maximum and minimum impulses in equations \ref{eq:lambdaLow} and 
+\ref{eq:lambdaHigh}.
+In finish phase constraint forces are calculated if requested and velocities of bodies are updated.
 \item Integrate transforms using velocities modified in previous step.
@@
 To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
-given threshold values (default 0.8 and 1.0) more than set limit (2.0).
+given threshold values (default 0.8 and 1.0) longer than set time limit (2.0).
 \end{enumerate}
@@
 \item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. object drifting. 
-Setting erp to 1 would in theory eliminate error but in practice value of 0.8 is used in most cases.
+Setting erp to 1 would in theory eliminate error in one step but in practice value of 0.2 - 0.8 is used in most cases.
 \end{description}
@@
 Actual values for unconstrained case without continuous collision detection
-are shown in \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
+are shown in table \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
 {\it velError} is 5.67 (5.5+0.167) and
@@
 \begin{center}
-\begin{tabular}{|l|l| l| l|l|l|l|l|}
+\begin{tabular}{|l|p{12cm}|}
 \hline
@@
 {\bf Description} 
+\\ \hline
 lowerLimit &  
@@
  \\  \hline
-currentLimit & describes state of constraint - 0: not limited,
- 3: loLimit=hiLimit, 4: current value is between loLimit and HiLimit
+currentLimit & describes state of constraint \newline
+ 0: not limited \newline
+ 3: loLimit=hiLimit \newline 4: current value is between loLimit and HiLimit
  \\ \hline
@@
 {\bf Plastic strain} \\  \hline
-0.017 & -0.0028 &        &    &          & -0.17 & 0 & 0 \\  \hline
+0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
 0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
@@
 
-Six-dof-spring constraint 2 has feature to avoid unstability by automatically softening constraint
-spring if needed.
-If this feature is active for elastic-plastic constraint 2 constraint it behaves as shown in table
- \ref{tab:ep2BlockValues}.
+Six-dof-spring constraint 2 has optional feature to avoid unstability by automatically softening constraint
+spring if needed. Feature is activated if current timeStep is too large for spring-mass system simulation.
+Equation \ref{eq:frequencyLimited} is used for decision, where $k$ is elastic spring stiffness between bodies,
+$m_{min}$ is smaller of connected masses or inertias and $dt$ is integration timestep.
+\begin{equation} \label{eq:frequencyLimited}
+\sqrt{ k /  m_{min}} dt > 0.25 
+\end{equation}
+If this feature is active for elastic-plastic constraint 2 constraint  
+it does not report positional error but it allows full plastic capacity to be used for correcting
+velocity based error instead of force provided by spring.
+In this scenario object behaves as shown in table  \ref{tab:ep2BlockValues} i.e. it does not move at all.
 
@@
 {\bf Plastic strain} \\  \hline
-0.017... &  0 & -0.17  & 2.34 & 13000 & 0       & 1000 & 0 \\  \hline
+0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
 \end {tabular}

### 2016-08-19 pdocs/thesis/article-section-3.tex
article sections 2-3 finetuning

@@
  \begin{description}
-\item[In constraint cases,] it is dominant contributor for constraint's impulse.
+\item[In constraint cases,] main contibutor is bodies relative speed at joint point.
 \item[In contact case,] main contributor is bodies relative speed at point of contact. 
-Bodies are not allowed to penetrate each other.
+Bodies are not allowed to penetrate each other. Restitution increases velError.
+If contact is not penetrating velError is reduced by $penetration * timeStep$.
 \end{description}
@@
  \item[In contact case,] value is zero if there is no penetration. For penetration cases it is 
-$\frac{-penetration\, erp}{timeStep}$
+$\frac{-penetration\, erp}{timeStep}$. In contact cases default value for erp is 0.8.
  \end{description}
@@
 \begin{center}
-\begin{tabular}{|l|p{12cm}|}
+\begin{tabular}{|l|p{10cm}|}
 \hline
@@
  0: not limited \newline
- 3: loLimit=hiLimit \newline 4: current value is between loLimit and HiLimit
+ 3: loLimit=hiLimit \newline 
+ 4: current value is between loLimit and HiLimit
  \\ \hline
@@
 \end {table}
+
+In following sections, results for few selected scenarios are presented.
+Simulation of Charpy impact test shows that plausible results can be obtained even for this quite difficult scenario.

### 2016-08-20 examples/CharpyDemo/CharpyDemo.cpp
fixed damping for spring and elastic plastic in Charpy

@@ -1304,7 +1304,7 @@ public:
 			}
 			for (int i = 0; i < 6; i++)
 			{
-				sc->setDamping(i, damping);
+				sc->setDamping(i, 1-damping);
 			}
 			sc->setEquilibriumPoint();
 		}
@@ -1483,7 +1483,7 @@ public:
 			}
 			for (int i = 0; i < 6; i++)
 			{
-				sc->setDamping(i, damping);
+				sc->setDamping(i, 1-damping);
 			}
 			sc->setFrequencyRatio(frequencyRatio);
 			sc->setEquilibriumPoint();

### 2016-08-20 examples/CharpyDemo/bt6DofElasticPlasticConstraint.cpp
searching way to avoid drifting in elasticPlastic

@@ -156,6 +156,13 @@ btScalar bt6DofElasticPlasticConstraint::getFrequencyRatio(){
 
 void bt6DofElasticPlasticConstraint::internalUpdateSprings(btConstraintInfo2* info)
 {
+	/**
+	setting useOld to false allows experimenting while
+	trying to find stable solution.
+	Currently visible drifting occurs in Charpy and Bending
+	*/
+	bool useOld = true;
+	btScalar maxForce;
 	int i;
 	for(i = 0; i < 3; i++)
 	{
@@ -170,20 +177,44 @@ void bt6DofElasticPlasticConstraint::internalUpdateSprings(btConstraintInfo2* in
 			bool overMaxForce=false;
 			// bcc
 			if (btFabs(force)>=m_maxForce[i]){
-				force = (delta > 0 ? m_maxForce[i] : -m_maxForce[i]);
+				maxForce = m_maxForce[i];
+				force = (delta > 0 ?  maxForce: -maxForce);
 				overMaxForce = true;
+			}
+			else{
+				maxForce = force;
 			}
 			btScalar ratio(info->fps/m_fpsLimit[i]);
 			btScalar velFactor;
-			if (!overMaxForce && ratio>m_frequencyRatio){
-				velFactor = m_springDamping[i] /
-					btScalar(info->m_numIterations);
+			if (useOld){
+				if (!overMaxForce && ratio > m_frequencyRatio){
+					velFactor = m_springDamping[i] /
+						btScalar(info->m_numIterations);
+				}
+				else{
+					velFactor = 0;
+				}
 			}
 			else{
-				velFactor = 0;
+				if (overMaxForce){
+					velFactor = 0;
+				}
+				else{
+					velFactor = m_springDamping[i] /
+						btScalar(info->m_numIterations);
+					if (ratio < m_frequencyRatio){
+						velFactor *= ratio / m_frequencyRatio;
+						maxForce = m_maxForce[i];
+					}
+				}
 			}
 			m_linearLimits.m_targetVelocity[i] = info->fps*velFactor * force;
-			m_linearLimits.m_maxMotorForce[i] = btFabs(force) / info->fps;
+			if (useOld){
+				m_linearLimits.m_maxMotorForce[i] = btFabs(force) / info->fps;
+			}
+			else{
+				m_linearLimits.m_maxMotorForce[i] = maxForce / info->fps;
+			}
 		}
 	}
 	for(i = 0; i < 3; i++)
[…]

### 2016-08-21 examples/CharpyDemo/DemolisherDemo.cpp
Demolisher: updateView moved to renderScene, after graphicsSync to avoid flicker

@@ -2058,6 +2058,7 @@ void DemolisherDemo::renderScene()
 		BT_PROFILE("m_guiHelper::syncPhysicsToGraphics");
 		syncWheelsToGraphics();
 		m_guiHelper->syncPhysicsToGraphics(m_dynamicsWorld);
+		updateView();
 		syncedStep = stepCount;
 	}
 	btScalar idleTime = idleClock.getTimeSeconds();
@@ -2128,7 +2129,6 @@ void DemolisherDemo::stepSimulation(float deltaTime)
 			PlasticityExampleBrowser::setPauseSimulation(true);
 		}
 		stepTime += timeStep;
-		updateView();
 	}
 	else{
 		kmph = 0;

### 2016-08-23 pdocs/thesis/article-section-2.tex
added erleben as source for joints in article
article sections 2-3 finetuning
more math details in article section 2

@@
 $\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
-$S$ & generalized transformation matrix  \\ \hline
+$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$  \\ \hline
+$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$  \\ \hline
 $\vec{v}_i$ & linear velocity of  center of mass for body i  \\ \hline
@@
 $\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
-$M$ & generalized mass matrix  \\ \hline
-$C$ & contact condition matrix  \\ \hline
-$N$ & contact normal matrix  \\ \hline
+$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$  \\ \hline
+$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
+$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$\\ \hline
 $J_{contact}$ & Jacobian matrix for contacts  \\ \hline
@@
 \end {table}
+
+\begin{equation} \label{eq:vectorqi}
+q_i = \lbrack s_i, x_i, y_i, z_i \rbrack ^T
+\end{equation}
+
+\begin{equation} \label{eq:matrixQi}
+Q_i = \frac{1}{2} \left[ \begin{array}{ccc}
+-x_i & -y_i & -z_i \\
+s_i & z_i & -y_i \\
+-z_i & s_i & x_i \\
+y_i & -x_i & s_i
+\end{array} \right]
+\end{equation}
+
+\begin{equation} \label{eq:matrixS}
+S =  \left[ \begin{array}{ccccc}
+1 &  &  & & 0 \\
+ & Q_i  \\
+ & & \ddots  \\
+ & & & 1 \\
+0 & & & & Q_n 
+\end{array} \right]
+\end{equation}
+
+\begin{equation} \label{eq:matrixM}
+M =  \left[ \begin{array}{ccccc}
+m_i 1 &  &  & & 0 \\
+ & I_1  \\
+ & & \ddots  \\
+ & & & m_n 1 \\
+0 & & & & I_n 
+\end{array} \right]
+\end{equation}
+
+\begin{equation} \label{eq:matrixN}
+N =  \left[ \begin{array}{cccc}
+\vec{n}_1 &  & & 0 \\
+ & \vec{n}_2  \\
+ & & \ddots  \\
+0  & & & \vec{n}_k 
+\end{array} \right]
+\end{equation}
 
@@
 Further enhancements are material nonlinearity and dynamic analysis.
-Physics engine provides dynamic analysis with large large reference translations and rotations.
+Physics engine provides dynamic analysis with large reference translations and rotations.
 
@@
 Mathematical background and detailed examples are available by \cite{ode.joints}.
+Joints are also discussed in detail in  \citet[p.~60-90]{erleben.thesis}.
 In following section, these equations will be clarified by simple example.

### 2016-08-25 pdocs/thesis/article-section-2.tex
fixed vector notations in article section 2
clarifying notations in article section 2

@@
 \begin{equation} \label{eq:eomL} % pdf page 69
-\vec{s}^{\, t+\Delta t} = s^{t}+\Delta t S \vec{u}^{\, t+\Delta t}
+\vec{s}^{\, t+\Delta t} = \vec{s}^{\, t}+\Delta t S \vec{u}^{\, t+\Delta t}
 \end{equation}
@@
 \begin{equation} \label{eq:eomV}
-\vec{u}^{\, t+\Delta t} = u^{t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
+\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
 \end{equation}
@@
 \begin{equation} \label{eq:eomV2}
-\vec{u}^{t+\Delta t} = u^{t}+\Delta t M^{-1}(
+\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
  J_{contact}^T \vec{\lambda}_{contact}
@@
 \draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
+\node[anchor=west] at (4.5,4.5) {
+$\vec{r}_i$  =  orientation of body i 
+};
+\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point k};
+\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
+\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point k };
 \end{tikzpicture}
-\caption{Illustration of nomenclature for equations of motion.}
-\label{fig:eom}
+\caption{Illustration of nomenclature for equations of motion for contact.}
+\label{fig:eom-contact}
+\end{figure}
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (O1) at(1,1);
+\coordinate (O2) at(1,3);
+\coordinate (C) at(1,2);
+\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
+\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
+\node at (0.5,0.4) {$B_1$};
+\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
+\node at (0.5,3.5) {$B_2$};
+\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
+\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
+\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
+\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
+\node[anchor=west] at (4.5,4.5) {
+$\vec{r}_i$  =  orientation of body i 
+};
+\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of joint point k};
+\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
+\end{tikzpicture}
+\caption{Illustration of nomenclature for equations of motion for joint.}
+\label{fig:eom-joint}
 \end{figure}
@@
 $\vec{r}_i$ & position of center of mass for body i  \\ \hline
-$\vec{q}_i$ & orientation as a quaternion for body i  \\ \hline
+$\vec{q}_i$ & orientation as a quaternion for body i  (\ref{eq:vectorqi})  \\ \hline
 $\vec{p}_i$ & contact or joint point i  \\ \hline
@@
 $\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
-$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$  \\ \hline
-$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$  \\ \hline
+$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$ (\ref{eq:matrixQi})  \\ \hline
+$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ (\ref{eq:matrixS}) \\ \hline
 $\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
@@
 $\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
-$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$  \\ \hline
+$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ (\ref{eq:matrixM})\\ \hline
+$I_i$ & inertia tensor for body i  \\ \hline
 $C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
-$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$\\ \hline
+$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$  (\ref{eq:matrixN}) \\ \hline
 $J_{contact}$ & Jacobian matrix for contacts  \\ \hline
@@
 
-\begin{equation} \label{eq:matrixQi}
+\begin{equation} \label{eq:matrixQi} % 4.13 p 50
 Q_i = \frac{1}{2} \left[ \begin{array}{ccc}

### 2016-08-27 pdocs/thesis/article-section-2.tex
added maximum forces in article section 2

@@
 equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
-Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom}.
+Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom-contact}.
+With this formulation, simplified scenario can be simulated. 
+Bodies do not penetrate each other and simulation is guided
+by external forces like gravity, air drag and shooting.
 \begin{equation} \label{eq:eomL} % pdf page 69
@@
 \vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
-\end{equation}
-
-Friction in contacts and joint constraints can be handled in unified way by refactoring
-equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}.
-Jacobian terms are derived by taking time derivates of kinematic constraints.
-
-\begin{equation} \label{eq:eomV2}
-\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
- J_{contact}^T \vec{\lambda}_{contact}
-+ J_{joint}^T \vec{\lambda}_{joint}
-+ \vec{f}_{ext})
 \end{equation}
@@
 
-\begin{figure}[htb!]
-\centering
-\begin{tikzpicture}
-\coordinate (O1) at(1,1);
-\coordinate (O2) at(1,3);
-\coordinate (C) at(1,2);
-\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
-\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
-\node at (0.5,0.4) {$B_1$};
-\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
-\node at (0.5,3.5) {$B_2$};
-\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
-\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
-\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
-\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
-\node[anchor=west] at (4.5,4.5) {
-$\vec{r}_i$  =  orientation of body i 
-};
-\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of joint point k};
-\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
-\end{tikzpicture}
-\caption{Illustration of nomenclature for equations of motion for joint.}
-\label{fig:eom-joint}
-\end{figure}
-
 % pdf page 33, notation in typical ODEs
@@
 $N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$  (\ref{eq:matrixN}) \\ \hline
-$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
-$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
-$J_{joint}$ & Jacobian matrix for joints  \\ \hline
-$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
 \end {tabular}
@@
 \end{equation}
+
+Friction in contacts and joint constraints can be handled in unified way by refactoring
+equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}.
+Jacobian terms are derived by taking time derivates of kinematic constraints.
+Added symbols are summarized in table 
+\ref{tab:eom-g} and figure \ref{fig:eom-contact}.
+
+\begin{equation} \label{eq:eomV2}
+\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
+ J_{contact}^T \vec{\lambda}_{contact}
++ J_{joint}^T \vec{\lambda}_{joint}
++ \vec{f}_{ext})
+\end{equation}
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (O1) at(1,1);
+\coordinate (O2) at(1,3);
+\coordinate (C) at(1,2);
+\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
+\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
+\node at (0.5,0.4) {$B_1$};
+\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
+\node at (0.5,3.5) {$B_2$};
+\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
+\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{anc}^{\,1}$};
+\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{anc}^{\,2}$};
+\node[anchor=west] at (4.5,3.5) {
+$\vec{r}_i$  =  orientation of body i 
+};
+\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = anchor point for body i};
+\end{tikzpicture}
+\caption{Illustration of nomenclature for equations of motion for joint.}
+\label{fig:eom-joint}
+\end{figure}
+
+% pdf page 33, notation in typical ODEs
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|l| l|}
+\hline
+{\bf Symbol} & {\bf Description} \\  \hline
+$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
+$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
+$J_{joint}$ & Jacobian matrix for joints  \\ \hline
+$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
+\end {tabular}
+\end{center}
+\caption {Additional terms  for generalized equations of motion} \label{tab:eom-g}
+\end {table}
+
+Constraint processing in \bullet\ is based on ODE, \cite{ode}.
+Mathematical background and detailed examples are available by \cite{ode.joints}.
+Joints are also discussed in detail in  \citet[p.~60-90]{erleben.thesis}.
+In following section, these equations will be clarified by simple example.
+Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
+\ref{eq:lambdaHigh} 
+are created for each constraint.
+
+\begin{equation} \label{eq:constraintEquation}
+J_1 \vec{v}_1 + \Omega_1 \vec{\omega}_1 + 
+J_2 \vec{v}_2 + \Omega_2 \vec{\omega}_2 = \vec{c} + C \vec{\lambda}
+\end{equation}
+
+\begin{equation} \label{eq:lambdaLow}
+\vec{\lambda} \geq \vec{l}
+\end{equation}
+
+\begin{equation} \label{eq:lambdaHigh}
+\vec{\lambda} \leq \vec{h}
+\end{equation}
+
+Main parameters  and corresponding fields in \bullet\  
+ are described in table \ref{tab:constraintParameters}.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{|c| l| l|}
+\hline
+{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
+$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
+$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
+$\vec{v}$ & linear velocity & \\ \hline
+$\vec{\omega}$ & angular velocity & \\ \hline
+$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
+$C$  & constraint force mixing & cfm \\  \hline
+$\vec{\lambda}$ & constraint force &  \\ \hline
+$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
+$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
+\end {tabular}
+\end{center}
+\caption {Constraint parameters} \label{tab:constraintParameters} 
+\end {table}
 
@@
 
+For rectangular section lagrange multiplier limits
+in equations \ref{eq:lambdaLow} and  
+\ref{eq:lambdaHigh}
+i.e. maximum forces 
+are summarized in figure \ref{fig:maxForces}.
+Yield shear stress is assumed here to be $ 0.5\, f_y$ using Tresca yield critetion.
+If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$).
+These are not exact values in multiaxial stress state but they
+should be quite usable in most cases.
+For rotation around y-axis there is closed form solution only for
+circular cross sections.
+Given approximation is 
+best suited for cases where b and h are quite similar.
+Better approximation for any given b and h can be obtained 
+by integrating distance from center of joint over cross section and
+multiplying it with yield shear stress e.g. using octave.
+
+\begin{lstlisting}
+b=0.01
+h=0.01
+fy=200e6
+wpy=fy/2*dblquad(@(x,y) sqrt(x.*x+y.*y),-b/2,b/2,-h/2,h/2)
+38.2
+\end{lstlisting}
+
+\lstset{language=octave}
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (O) at (0,0,0);
+
+% axes
+\draw[->] (O) -- (1,0,0) node[anchor=north east]{$x$};
+\draw[->] (O) -- (0,1,0) node[anchor=north west]{$y$};
+\draw[->] (O) -- (0,0,1) node[anchor=south]{$z$};
+
+% 1,3,1 block, base at y=1.5
+\draw (1,1.5,0) -- ++(0,3,0) -- ++(-1,0,0);
+\draw (0,1.5,1) -- ++(1,0,0) -- ++(0,3,0) -- ++(-1,0,0) -- ++(0,-3,0);
+\draw (1,1.5,0) -- ++(0,0,1);
+\draw (0,4.5,0) -- ++(0,0,1);
+\draw (1,4.5,0) -- ++(0,0,1);
+
+% l is not significant
+%\draw[Stealth-Stealth] (2,1.5,0) -- ++(0,3,0);
+%\node[] at (2.5,3) [anchor=west]{l};
+
+\draw[Stealth-Stealth] (0,5,0) -- ++(1,0,0);
+\node[] at (0.5,5) [anchor=south]{b};
+
+\draw[Stealth-Stealth] (0,5,0) -- ++(0,0,1);
+\node[] at (0,5,0.5) [anchor=south east]{h};
+
+\draw[ultra thin](0,4.5,0) -- ++(0,0.6);
+\draw[ultra thin](1,4.5,0) -- ++(0,0.6);
+\draw[ultra thin](0,4.5,1) -- ++(0,0.6);
+
+\draw[-Stealth] (0.5,3,0.5) -- ++(0,1.5,0);
+\node at (0,3.8) {$\vec{r}_{anc}$};
+
+\node[anchor=west] at (4.0,3) {
+\begin{tabular}{l}
+maximum force in x direction = $0.5\, b\, h f_y$ \\
+maximum force in y direction = $b\, h\, f_y$  \\
+maximum force in z direction = $0.5\, b\, h\, f_y$  \\
+\\
+maximum moment around x-axis = $0.25\, b\, h^2 \, f_y$  \\
+maximum moment around y-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
+maximum moment around z-axis = $0.25\, b^2\, h\, f_y$ 
+\end{tabular}
+};
+
+\end{tikzpicture}
+\caption{Maximum forces in plastic joint for rectangular cross section.}
+\label{fig:maxForces}
+\end{figure}
+
 Basic idea in this work can be tested with any framework having motors and hinge constraints.
@@
 and adding new constraints \cite{pbullet}.
-Constraint processing in \bullet\ is based on ODE, \cite{ode}.
-Mathematical background and detailed examples are available by \cite{ode.joints}.
-Joints are also discussed in detail in  \citet[p.~60-90]{erleben.thesis}.
-In following section, these equations will be clarified by simple example.
-Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
-\ref{eq:lambdaHigh} 
-are created for each constraint.
-
-\begin{equation} \label{eq:constraintEquation}
-J_1 v_1 + \Omega_1 \omega_1 + J_2 v_2 + \Omega_2 \omega_2 = c + C \lambda
-\end{equation}
-
-\begin{equation} \label{eq:lambdaLow}
-\lambda \geq l
-\end{equation}
-
-\begin{equation} \label{eq:lambdaHigh}
-\lambda \leq h
-\end{equation}
-
-Main parameters  and corresponding fields in \bullet\  
- are described in table \ref{tab:constraintParameters}.
-
-\begin {table}[htb!]
-\begin{center}
-\begin{tabular}{|c| l| l|}
-\hline
-{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
-$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
-$v$ & linear velocity & \\ \hline
-$\omega$ & angular velocity & \\ \hline
-$c$        &  right side vector   & m\_constraintError \\ \hline
-$C$  & constraint force mixing & cfm \\  \hline
-$\lambda$ & constraint force &  \\ \hline
-$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
-$h$ & high limit for constraint force & m\_upperLimit \\ \hline
-\end {tabular}
-\end{center}
-\caption {Constraint parameters} \label{tab:constraintParameters} 
-\end {table}

### 2016-08-28 pdocs/thesis/article-section-2.tex
added more details about maximum forces in article section 2

@@
 $\vec{r}_i$ & position of center of mass for body i  \\ \hline
-$\vec{q}_i$ & orientation as a quaternion for body i  (\ref{eq:vectorqi})  \\ \hline
+$\vec{q}_i$ & orientation for body i as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
+\hline
 $\vec{p}_i$ & contact or joint point i  \\ \hline
@@
 $\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
-$Q_i$ & rotation matrix $Q_i \in  \mathbb{R}^{4 \times 3}$ (\ref{eq:matrixQi})  \\ \hline
-$S$ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ (\ref{eq:matrixS}) \\ \hline
-$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
-$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
-$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
-$M$ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ (\ref{eq:matrixM})\\ \hline
-$I_i$ & inertia tensor for body i  \\ \hline
-$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
-$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$  (\ref{eq:matrixN}) \\ \hline
-\end {tabular}
-\end{center}
-\caption {Nomenclature for equations of motion} \label{tab:eom}
-\end {table}
-
-\begin{equation} \label{eq:vectorqi}
-q_i = \lbrack s_i, x_i, y_i, z_i \rbrack ^T
-\end{equation}
-
-\begin{equation} \label{eq:matrixQi} % 4.13 p 50
-Q_i = \frac{1}{2} \left[ \begin{array}{ccc}
+$Q_i$ & rotation matrix  $
+\frac{1}{2} \left[ \begin{array}{ccc}
 -x_i & -y_i & -z_i \\
@@
 \end{array} \right]
-\end{equation}
-
-\begin{equation} \label{eq:matrixS}
-S =  \left[ \begin{array}{ccccc}
+$
+\\ \hline
+ & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ \\ 
+$S$ & $
+  \left[ \begin{array}{ccccc}
 1 &  &  & & 0 \\
@@
 \end{array} \right]
-\end{equation}
-
-\begin{equation} \label{eq:matrixM}
-M =  \left[ \begin{array}{ccccc}
+$
+\\ \hline
+$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
+$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
+$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
+ & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ \\ 
+$M$ &
+$
+\left[ \begin{array}{ccccc}
 m_i 1 &  &  & & 0 \\
@@
 \end{array} \right]
-\end{equation}
-
-\begin{equation} \label{eq:matrixN}
-N =  \left[ \begin{array}{cccc}
-\vec{n}_1 &  & & 0 \\
- & \vec{n}_2  \\
- & & \ddots  \\
-0  & & & \vec{n}_k 
-\end{array} \right]
-\end{equation}
+$
+\\ \hline
+$I_i$ & inertia tensor for body i  \\ \hline
+$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
+$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
+\end {tabular}
+\end{center}
+\caption {Nomenclature for equations of motion} \label{tab:eom}
+\end {table}
 
@@
 };
-\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = anchor point for body i};
+\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector i};
 \end{tikzpicture}
@@
 \end{tikzpicture}
-\caption{Stress distribution under elastic and plastic loads.}
+\caption{Stress distribution for bending under elastic and plastic loads.}
 \label{fig:wp}
@@
 
-For rectangular section lagrange multiplier limits
+Plasticity is handled by defining maximum forces
 in equations \ref{eq:lambdaLow} and  
-\ref{eq:lambdaHigh}
-i.e. maximum forces 
+\ref{eq:lambdaHigh} using plastic capasities as maximum forces.
+
+Maximum force acting in same direction as $\vec{r}_{anc}^{\,i} $
+is product of area and yield stress as given in equation \ref{eq:fN}.
+
+\begin{equation} \label{eq:fN}
+N_{max}= \int_A f_y
+\end{equation}
+
+Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
+are product of area and shear yield stress $\tau_y$ as given in equation \ref{eq:fQ}.
+\begin{equation} \label{eq:fQ}
+Q_{max}= \int_A \tau_y
+\end{equation}
+
+Maximum moments acting around axis perpendicular to $\vec{r}_{anc}^{\,i} $
+are integrals of perpdendicular distance 
+and yield stress $f_y$ as given in equations \ref{eq:Mx} for moment around
+x-axis and \ref{eq:Mz} for moment around z-axis.
+
+\begin{equation} \label{eq:Mx}
+M_{max}^x= \int_A z f_y
+\end{equation}
+
+\begin{equation} \label{eq:Mz}
+M_{max}^z= \int_A x f_y
+\end{equation}
+
+Maximum moment around $\vec{r}_{anc}^{\,i} $
+is integral of distance $d$ from joint point
+and shear yield stress $\tau_y$ as given in equations \ref{eq:My}.
+
+\begin{equation} \label{eq:My}
+M_{max}^y= \int_A d \tau_y
+\end{equation}
+
+Maximum forces and moments for 
+rectangular section using constant yield stress
 are summarized in figure \ref{fig:maxForces}.
@@
 should be quite usable in most cases.
+
 For rotation around y-axis there is closed form solution only for
@@
 
+\lstset{language=octave}
 \begin{lstlisting}
-b=0.01
-h=0.01
-fy=200e6
+b=0.01; h=0.01; fy=200e6;
 wpy=fy/2*dblquad(@(x,y) sqrt(x.*x+y.*y),-b/2,b/2,-h/2,h/2)
@@
 
-\lstset{language=octave}
 \begin{figure}[htb!]
@@
 \begin{tikzpicture}
-\coordinate (O) at (0,0,0);
+\coordinate (O) at (2,2,0);
 
 % axes
-\draw[->] (O) -- (1,0,0) node[anchor=north east]{$x$};
-\draw[->] (O) -- (0,1,0) node[anchor=north west]{$y$};
-\draw[->] (O) -- (0,0,1) node[anchor=south]{$z$};
+\draw[->] (O) -- ++(1,0,0) node[anchor=north east]{$x$};
+\draw[->] (O) -- ++(0,1,0) node[anchor=north west]{$y$};
+\draw[->] (O) -- ++(0,0,1) node[anchor=south]{$z$};
 
@@
 \begin{tabular}{l}
-maximum force in x direction = $0.5\, b\, h f_y$ \\
+maximum force in x and z directions = $0.5\, b\, h f_y$ \\
 maximum force in y direction = $b\, h\, f_y$  \\
-maximum force in z direction = $0.5\, b\, h\, f_y$  \\
 \\

### 2016-09-14 examples/CharpyDemo/BlockDemo.cpp
BlockDemo/Hinge: create bodyA yourself to get proper reference later

@@ -945,12 +945,22 @@ public:
 		sc->setTargetLinMotorVelocity(0);
 		return sc;
 	}
-	btTypedConstraint* addHingeConstraint(btRigidBody* rb, btVector3& pivot, btVector3 & axis){
+	/**
+	create bodyA yourself to get proper reference later
+	*/
+	btTypedConstraint* addHingeConstraint(btRigidBody* rbB, btVector3& pivot, btVector3 & axis){
+		btTransform trA;
+		trA.setIdentity();
+		trA.setOrigin(rbB->getCenterOfMassPosition() + pivot);
+		btCollisionShape* fixedShape = new btSphereShape(lsz/10);
+		m_collisionShapes.push_back(fixedShape);
+		btRigidBody* rbA = localCreateRigidBody(0, trA, fixedShape);
+		btVector3 pivotInA(0,0,0);
 		btHingeConstraint *sc =
-			new btHingeConstraint(*rb, pivot, axis);
+			new btHingeConstraint(*rbA, *rbB, pivotInA, pivot, axis, axis);
 		m_dynamicsWorld->addConstraint(sc, disableCollisionsBetweenLinkedBodies);
 		sc->enableAngularMotor(true,0,axisMapper->getMaxForce(5)*m_fixedTimeStep);
-		sc->setLimit(-SIMD_PI, SIMD_PI);
+		sc->setLimit(1, -1); // free
 		return sc;
 	}
 	btTypedConstraint* addSpring2Constraint(btRigidBody* rb, btVector3& cpos){

### 2016-09-14 examples/plasticity/PlasticityExampleBrowser.cpp
Save GraphWindow positions and reuse if graph has same name

@@ -479,12 +479,17 @@ void quitCallback()
 
 #define MAX_GRAPH_WINDOWS 5
 
+struct GraphWindowInfo {
+	const char * name;
+	int x, y;
+};
+
 struct QuickCanvas : public Common2dCanvasInterface
 {
 	GL3TexLoader* m_myTexLoader;
-
 	MyGraphWindow* m_gw[MAX_GRAPH_WINDOWS];
 	GraphingTexture* m_gt[MAX_GRAPH_WINDOWS];
+	GraphWindowInfo m_gwi[MAX_GRAPH_WINDOWS];
 	int m_curNumGraphWindows;
 
 	QuickCanvas(GL3TexLoader* myTexLoader)
@@ -495,9 +500,26 @@ struct QuickCanvas : public Common2dCanvasInterface
 		{
 			m_gw[i] = 0;
 			m_gt[i] = 0;
+			m_gwi[i].name = "";
 		}
 	}
 	virtual ~QuickCanvas() {}
+	/**
+	Use saved position if there is one with same name
+	*/
+	void findPosition(MyGraphInput* input){
+		for (int i = 0; i<MAX_GRAPH_WINDOWS; i++)
+		{
+			GraphWindowInfo& gwi = m_gwi[i];
+			if (strcmp(input->m_name, gwi.name) == 0){
+				input->m_xPos = gwi.x;
+				input->m_yPos = gwi.y;
+				return;
+			}
+		}
+		input->m_xPos = 10000;//GUI will clamp it to the right if too high (first)
+		input->m_yPos = 10000;//GUI will clamp it to bottom
+	}
 	virtual int createCanvas(const char* canvasName, int width, int height)
 	{
 		if (m_curNumGraphWindows<MAX_GRAPH_WINDOWS)
@@ -513,16 +535,15 @@ struct QuickCanvas : public Common2dCanvasInterface
 			MyGraphInput input(gui->getInternalData());
 			input.m_width = width;
 			input.m_height = height;
-			input.m_xPos = 10000;//GUI will clamp it to the right//300;
-			input.m_yPos = 10000;//GUI will clamp it to bottom
 			input.m_name = canvasName;
 			input.m_texName = canvasName;
+			findPosition(&input);
 			m_gt[slot] = new GraphingTexture;
 			m_gt[slot]->create(width, height);
 			int texId = m_gt[slot]->getTextureId();
[…]

### 2016-09-14 examples/CharpyDemo/BlockDemo.cpp
BlockDemo: number formatting tuning

@@ -566,6 +566,12 @@ public:
 		if (value < 10e-3){
 			text = uif(value, "%.6f");
 		}
+		else if (value>10e3){
+			text = uif(value, "%.0f");
+		}
+		else if (value>10e2){
+			text = uif(value, "%.1f");
+		}
 		else{
 			text = uif(value);
 		}

### 2016-09-16 pdocs/thesis/article-section-2.tex
working on Aki's comments on article

@@
 % !TeX root = article.tex
-\section{Adding plasticity to physics engine}
+\section{Description of plasticity in the framework of physics engine}
 
@@
 
-Velocity-based formulation is not often used in structural or mechanical engineering while, 
-however, it is popular within physics based game developers and film production teams.
+Velocity-based formulation is  popular within physics based game
+developers and film production teams.
  \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details on why 
@@
 Main reason is that collision handling can be done without additional procedures.
-Acceleration-based formulations need to stop at collision and switch to an impulse based method.
 
@@
 \end{equation}
-Using Newton's second law of motion $F=ma$ one can solve for the velocity, $\vec{v}^{\Delta t}$
+Using Newton's second law of motion $\vec{F}=m\vec{a}$ one can solve for the velocity,
+$\vec{v}^{\Delta t}$
 \begin{equation} \label{eq:impulseIntegraWithNewton}
@@
 \end{equation}
-where superscripts denote time, i.e. ${v}^{\Delta t}=v(\Delta t)$. Next position can be found
+where superscripts denote time, i.e. ${\vec{v}}^{\Delta t}=\vec{v}(\Delta t)$.
+Next position can be found
 by integrating the velocity.
-Derivation of equations of motion is not shown here in detail but results 
-for simple contact can be summarized as
-equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
-Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom-contact}.
-With this formulation, simplified scenario can be simulated. 
-Bodies do not penetrate each other and simulation is guided
-by external forces like gravity, air drag and shooting.
+
 \begin{equation} \label{eq:eomL} % pdf page 69
@@
 \end{equation}
+
+Symbols are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
+Figure \ref{fig:eom-contact} shows terms needed for contact processing.
+For each contact  normal of contact point and
+orientation of both bodies and contact point are needed.
+With this formulation, simplified scenario can be simulated.
 
@@
 $\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
-$Q_i$ & rotation matrix  $
+$Q_i$ & \begin{tabular}{@{}c}
+rotation of quaternion $\vec{q}_i$
+as matrix where \\
+$\frac{1}{2}\vec{\omega}_i q_i=Q_i \vec{\omega}_i$
+\end{tabular}
+   $
 \frac{1}{2} \left[ \begin{array}{ccc}
@@
 Jacobian terms are derived by taking time derivates of kinematic constraints.
-Added symbols are summarized in table 
-\ref{tab:eom-g} and figure \ref{fig:eom-contact}.
+Added symbols are summarized in Table
+\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
+Figure \ref{fig:eom-joint} shows terms needed for joint processing.
+New term $\vec{r}_{anc}^{\,i}$ is used to define at which point
+joint constraint is applied.
 
@@
 Main parameters  and corresponding fields in \bullet\  
- are described in table \ref{tab:constraintParameters}.
+ are described in Table \ref{tab:constraintParameters}.
 
@@
 based on needed features.
-For most complex scenarios finite element method is used.
-In most cases, static solution with assumption of linear strain-displacement relation small 
-displacement solution using displacement based boundary conditions is used.
+Often,  finite element method is used.
+In most cases, static solution with assumption of linear strain-displacement relation
+using displacement based boundary conditions is used.
 \citet{bathe-1975} provides description for handling of various nonlinearities.
@@
 Further enhancements are material nonlinearity and dynamic analysis.
-Physics engine provides dynamic analysis with large reference translations and rotations.
+Physics engine provides dynamic analysis with large reference translations and rotations
+while assuming bodies to be undeformable.
 
-Material plasticity has typically taken into account in games by using suitable coefficient of restitution.
+Material plasticity can be accounted in games by using suitable coefficient of restitution.
 This provides reasonable means to simulate loss of energy in collisions.
@@
 to multiple bodies which are connected by energy absorbing joints.
-Typical engineering stress-strain curve of ductile steel is shown in \ref{fig:sscurve}.
-E.g. \citet{dowling} provides detailed descriptions of engineering and true stress-strain curves.
+Typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.
+Engineering stress and strain mean that original dimensions are used in stress calculation,
+\citet[p.~108]{dowling}.
 Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
@@
 In this work elastic-fully plastic material model is used in most scenarios.
-It allows realistic simulations for most scenarios.
 Having elastic part allows elastic displacements for slender structures. 
-Elastic part is ignored in method suggested in this work if deformation is related
-to higher frequency than integration stability would allow.
-Strain hardening part of stress-strain curve could probably be taken into account but it has not been tried in this work.
+Elastic material behavior is ignored in approach introduced in this work provided
+that deformation is related to higher frequency
+than integration stability would allow.
 It should be noted that geometry
 of objects is not updated during analysis and thus engineering stress-strain properties should
-be used even with strain hardening.
+be used.
 
@@
 
-Difference between elastic and plastic section modulus is shown in \ref{fig:wp}. 
+Difference between elastic and plastic section modulus is shown in Figure \ref{fig:wp}.
 If stress is below yield limit, stress and strain are linear within material.
@@
 \end{tikzpicture}
-\caption{Stress distribution for bending under elastic and plastic loads.}
+\caption{Axial stress distribution for bending under elastic and plastic loads.}
 \label{fig:wp}
@@
 
-Maximum force acting in same direction as $\vec{r}_{anc}^{\,i} $
-is product of area and yield stress as given in equation \ref{eq:fN}.
+Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
+is product of area and yield stress as follows
 
@@
 Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
-are product of area and shear yield stress $\tau_y$ as given in equation \ref{eq:fQ}.
+are product of area and shear yield stress $\tau_y$ as follows
 \begin{equation} \label{eq:fQ}
@@
 are integrals of perpdendicular distance 
-and yield stress $f_y$ as given in equations \ref{eq:Mx} for moment around
-x-axis and \ref{eq:Mz} for moment around z-axis.
+and yield stress $f_y$ as given for the moment around x-axis and,
+correspondingly, moment around z-axis..
 
@@
 rectangular section using constant yield stress
-are summarized in figure \ref{fig:maxForces}.
-Yield shear stress is assumed here to be $ 0.5\, f_y$ using Tresca yield critetion.
+are summarized in Figure \ref{fig:maxForces}.
+Yield shear stress is assumed to be $ 0.5\, f_y$ using Tresca yield critetion.
 If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$).
@@
 Given approximation is 
-best suited for cases where b and h are quite similar.
+best suited for cases where $b$ and $h$ are quite similar.
 Better approximation for any given b and h can be obtained 
@@
 \draw[Stealth-Stealth] (0,5,0) -- ++(1,0,0);
-\node[] at (0.5,5) [anchor=south]{b};
+\node[] at (0.5,5) [anchor=south]{$b$};
 
 \draw[Stealth-Stealth] (0,5,0) -- ++(0,0,1);
-\node[] at (0,5,0.5) [anchor=south east]{h};
+\node[] at (0,5,0.5) [anchor=south east]{$h$};
 

### 2016-09-16 pdocs/thesis/article-section-2.tex
working on Aki's comments on article

@@
 \node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
+\node[anchor=west] at (-2,3) {
+\begin{tabular}{l}
+Under elastic load\\
+stress increases\\
+linearly from zero\\
+at neutral axis to\\
+maximum value at \\
+surface of body
+\end{tabular}
+};
 \end{tikzpicture}
@@
 \node[above] at (S) [align=center] {\large{$\sigma=f_y$}};
+\node[anchor=west] at (3,3) {
+\begin{tabular}{l}
+Under plastic load\\
+stress is at yield\\
+level over full\\
+cross section
+\end{tabular}
+};
 \end{tikzpicture}
-\caption{Axial stress distribution for bending under elastic and plastic loads.}
+\caption{Axial stress distribution over cross section for bending under elastic and plastic loads.}
 \label{fig:wp}

### 2016-09-17 pdocs/thesis/article-section-2.tex
working on Aki's comments on article

@@
 by integrating the velocity.
+Updates after each step can be summarized as Equation
+ \ref{eq:eomL} for locations and  Equation 
+\ref{eq:eomV} for velocities.
 
@@
 rotation of quaternion $\vec{q}_i$
-as matrix where \\
-$\frac{1}{2}\vec{\omega}_i q_i=Q_i \vec{\omega}_i$
+as matrix \\ where 
+$\frac{1}{2}\vec{\omega}_i \vec{q}_i=Q_i \vec{\omega}_i$
 \end{tabular}
@@
 \\ \hline
- & generalized transformation matrix $ S \in \mathbb{R}^{7n \times 6n}$ \\ 
-$S$ & $
-  \left[ \begin{array}{ccccc}
+$S$ & 
+\begin{tabular}{@{}c}
+generalized transformation matrix \\
+$ S \in \mathbb{R}^{7n \times 6n}$
+\end{tabular}
+ $ \left[ \begin{array}{ccccc}
 1 &  &  & & 0 \\
@@
 $\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
- & generalized mass matrix $ M \in \mathbb{R}^{6n \times 6n}$ \\ 
 $M$ &
+\begin{tabular}{@{}c}
+ generalized mass matrix \\
+$ M \in \mathbb{R}^{6n \times 6n}$  
+\end{tabular}
 $
@@
 Plasticity is handled by defining maximum forces
-in equations \ref{eq:lambdaLow} and  
+in Equations \ref{eq:lambdaLow} and  
 \ref{eq:lambdaHigh} using plastic capasities as maximum forces.
@@
 is integral of distance $d$ from joint point
-and shear yield stress $\tau_y$ as given in equations \ref{eq:My}.
+and shear yield stress $\tau_y$ as
 
@@
 best suited for cases where $b$ and $h$ are quite similar.
-Better approximation for any given b and h can be obtained 
+Better approximation for any given $b$ and $h$ can be obtained 
 by integrating distance from center of joint over cross section and

### 2016-09-17 pdocs/thesis/article-section-3.tex
working on Aki's comments on article

@@
 % !TeX root = article.tex
-\section{Processing plasticity}
+\section{Numerical examples}
 
-In this section, changes to constraint formulation are clarified by doing few simulation steps of non constrained, rigid and
-elastic-plastic examples.
+In this section, changes to constraint formulation are clarified by doing few simulation steps
+of non constrained, rigid and
+elastic-plastic examples using numerical values.
 
@@
 In this case frame could be anywhere but if multiple bodies are involved, connecting frame must be defined
-so that it reflects wanted scenario.
+so that it reflects scenario under investigation.
 Concrete density is $2000 \frac{kg}{m^3}$ and steel density is $7800 \frac{kg}{m^3}$. 
@@
 \item Apply gravity to each non static body. 
-This step just makes programming easier as otherwise each program should add forces due to gravity in each step. 
-In this case, $\vec{f}_{ext}$ in  \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
-\item Predict unconstrained motion for each non static body. Prediction is done based on current linear and angular velocities of the body.
+This step makes programming easier as otherwise each program should add forces due to gravity in each step. 
+In this case, $\vec{f}_{ext}$ in  Equation \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
+\item Predict unconstrained motion for each body controlled by physics simulation.
+Static and kinematic bodies are not processed in this step.
+Prediction is done based on current linear and angular velocities of the body.
 \item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
-Each rigid body is represented by sphere and contact prediction is done if body moves more than given threshold value during simulation step.
-Continuous collision detection is configured for each body and as default it is not active.
+Each rigid body is represented by sphere geometry and contact prediction is done if body moves more than given threshold value during simulation step.
+Continuous collision detection is configured for each body.
 Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
-\item Perform discrete collision detection. All overlapping body pairs are processed and manifoldPoints are created for
+\item Perform discrete collision detection. All overlapping bodies are processed and manifoldPoints are created for
 each detected contact at end of simulation step. 
-In unconstrained case, contact is detected after t=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
+In unconstrained case, contact is detected after $t$=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
 \item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
@@
  This step is subdivided to setup, iterations and finish phases. 
-In setup phase constraints are queried for positional errors for calculation of c in
-equation \ref{eq:constraintEquation} and maximum and minimum impulses in equations \ref{eq:lambdaLow} and 
+In setup phase constraints are queried for positional errors for calculation of $c$ in
+Equation \ref{eq:constraintEquation} and maximum and minimum impulses in Equations \ref{eq:lambdaLow} and 
 \ref{eq:lambdaHigh}.
@@
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ was used to pick up values for internal variables.
+in \bullet\ was used to pick up values for internal variables described in Figure \ref{fig:solVars}.
 
+\begin{figure}[htb!]
 \begin{description}
@@
 \end{description}
+\caption{Internal variables used in \bullet\ in constraint solving.}
+\label{fig:solVars}
+\end{figure}
 
@@
 There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
-in this work elastic-plastic versions of both of them were developed.
+in this work elastic-plastic versions of both of them are developed.
 
@@
 There are also e.g. parameters for controlling joint motors.
-Additional equations are created if needed. Enabled springs and motors add one row and
+Additional equation is created if for each additional constraint. 
+Enabled springs and motors add one row and
 limits add one if upper and lower are same and two if both upper and lower limits are defined.
@@
 Six-dof-spring constraint 2 has optional feature to avoid unstability by automatically softening constraint
-spring if needed. Feature is activated if current timeStep is too large for spring-mass system simulation.
-Equation \ref{eq:frequencyLimited} is used for decision, where $k$ is elastic spring stiffness between bodies,
-$m_{min}$ is smaller of connected masses or inertias and $dt$ is integration timestep.
+spring. Feature is activated if current timeStep is too large for spring-mass system simulation.
+The feature is trigged based on the equation below 
 \begin{equation} \label{eq:frequencyLimited}
@@
 \end{equation}
+where $k$ is elastic spring stiffness between bodies,
+$m_{min}$ is smaller of connected masses or inertias and $dt$ is integration timestep.
+
 If this feature is active for elastic-plastic constraint 2 constraint  
@@
 velocity based error instead of force provided by spring.
-In this scenario object behaves as shown in table  \ref{tab:ep2BlockValues} i.e. it does not move at all.
+In this scenario object behaves as depicted in table \ref{tab:ep2BlockValues} i.e. it does not move at all.
 

### 2016-09-19 pdocs/thesis/article-conclusion.tex
skeleton for conclusion

--- 2016-06-18 pdocs/thesis/article-introduction.tex
@@
-\section{Introduction}
+\section{Conclusion}
 
-The application of modern computer game techniques enables the description of complex dynamic systems 
-such as military vehicles with a high level of detail while still solving the equations in real-time.
-Film production and war games, in particular, is a key area that have benefited from simulation technology. 
-In practice, games are often accomplished using an open-source platform such 
-like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
-
-Computational methods used in physics engines are divided to modules that handle collision detection and 
-contact description and modules that handle solution of equations in real-time. Equations need to be 
-solved can further be subdivided to be associated to motion, constraints and collisions. 
-Velocity-based formulation is typically used in constraint based rigid body simulation. 
-Friction is typically taken into account and mechanical joints are handled by constraint equations.
-Detailed description of various components can be found in e.g. \cite{erleben.thesis}.
-
-Plasticity is not typically taken into account in gaming solutions. 
-Breaking of various objects typically takes place based on collision or impulse.
-Nevertheless, breaking of steel or reinforced concrete structures using this approach 
-is not appropriate making a simulation to look unrealistic. Theory for handling of plasticity 
-has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} 
-present a method for modeling and animating of elastic and plastic objects in real-time using 
-point based animation. This approach is not been widely used in simulation applications.  
-On major issue is collision handling of deformable objects.
-
-This study will introduce an approach to account plastic deformation in game applications.   
+This study has presented an approach to account plastic deformation in 
+velocity based formulation.
 In the introduced method, the plastic deformation takes place if force or moment exceeds given 
-limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced. 
-The approach is based on using joint motors to model plasticity. \citet[p.~90]{erleben.thesis} 
-suggests similar method for modelling friction in joints. Adjacent objects are connected by motors. 
-Motor power production limits are estimated based on plastic section modulus. 
-Joint breaking is accounted by summing plastic deformation and comparing it to 
-predefined material based limit. Elastic part of deformation is modelled by employing 
-spring description which is based on modification of existing constraint in \bullet.
-
+limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
 Approach presented in this work can be used in gaming industry to provide more realistic 
@@
 methods to modern simulation frameworks.
+As already noted in \cite{cg1988}, the modeling of inelastic deformation
+remains open for further exploration in the context of computer graphics.

### 2016-09-19 pdocs/thesis/article-introduction.tex
skeleton for conclusion

@@
 In the introduced method, the plastic deformation takes place if force or moment exceeds given 
-limit, deformation absorbs energy and joint breaks if plastic capacity is exceeced. 
+limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
 The approach is based on using joint motors to model plasticity. \citet[p.~90]{erleben.thesis} 

### 2016-09-24 pdocs/thesis/article-section-2.tex
tuning of article

@@
 In following section, these formulations will be clarified by simple example using \bullet\ implementation.
-Assuming that force $\vec{f}_{true}(t)$, is known the impulse $\vec{J}$
+Impulse $\vec{J}$
 in the time interval $\Delta t $ can we written as
@@
 \end{equation}
+where $\vec{f}_{true}(t)$ is force.
+
 Using Newton's second law of motion $\vec{F}=m\vec{a}$ one can solve for the velocity,
@@
 Symbols are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
-Figure \ref{fig:eom-contact} shows terms needed for contact processing.
-For each contact  normal of contact point and
-orientation of both bodies and contact point are needed.
-With this formulation, simplified scenario can be simulated.
+Figure \ref{fig:eom-contact} describes collision of two bodies $B_1$ and $B_2$.
 
@@
 \end{figure}
+
+$\vec{r}_i$ is orientation of body i.
+$\vec{p}_k $ orientation of contact point k. 
+$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point k.
+$\vec{n}_{k}$ is contact normal for contact point k.
+With this formulation, simplified scenario can be simulated.
 
@@
 Friction in contacts and joint constraints can be handled in unified way by refactoring
-equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}.
-Jacobian terms are derived by taking time derivates of kinematic constraints.
-Added symbols are summarized in Table
-\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
-Figure \ref{fig:eom-joint} shows terms needed for joint processing.
-New term $\vec{r}_{anc}^{\,i}$ is used to define at which point
-joint constraint is applied.
-
+equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}
 \begin{equation} \label{eq:eomV2}
@@
 \end{equation}
+
+where jacobian terms $J_{contact}^T$ for joints are 
+derived by taking time derivates of kinematic constraints.
+
+Added symbols are summarized in Table
+\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
+Figure \ref{fig:eom-joint} shows terms needed for joint processing.
 
@@
 
+$\vec{r}_{anc}^{\,i}$ is used to define at which point
+joint constraint is applied.
+
 % pdf page 33, notation in typical ODEs
@@
 Joints are also discussed in detail in  \citet[p.~60-90]{erleben.thesis}.
-In following section, these equations will be clarified by simple example.
 Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
@@
 
+In following section, these equations will be clarified by simple example.
 Main parameters  and corresponding fields in \bullet\  
@@
 Typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.
+
+\begin{figure}[htb!]
+\centering
+\begin{tikzpicture}
+\coordinate (Y) at (1,4);
+\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
+\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
+\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
+\draw[dashed](0,4) -- (Y);
+\node at (-0.2,4) [align=right] {$f_y$};
+\draw(0.25,1) -- (0.5,1) -- (0.5,2);
+\node at (0.75,1.5) {$E$};
+\node at (0.8,2.5) [anchor=west] {$\sigma = E \epsilon$ if $\sigma \le f_y$};
+\end{tikzpicture}
+\caption{Engineering stress-strain curve of ductile steel (not to scale).}
+\label{fig:sscurve}
+\end{figure}
+
+where $\sigma$ is stress, $E$ is Youngs modulus and $f_y$ is yield stress.
 Engineering stress and strain mean that original dimensions are used in stress calculation,
@@
 
-\begin{figure}[htb!]
-\centering
-\begin{tikzpicture}
-\coordinate (Y) at (0.2,4);
-\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
-\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
-\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
-\draw[dashed](0,4) -- (Y);
-\node at (-0.2,4) [align=right] {$f_y$};
-\end{tikzpicture}
-\caption{Engineering stress-strain curve of ductile steel (not to scale).}
-\label{fig:sscurve}
-\end{figure}
-
 Difference between elastic and plastic section modulus is shown in Figure \ref{fig:wp}.
-If stress is below yield limit, stress and strain are linear within material.
-If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
-so plastic section modulus is higher than elastic section modulus.
 
@@
 
+If stress is below yield limit $f_y$, stress and strain are linear within material.
+If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
+so plastic section modulus is higher than elastic section modulus.
+
 Plasticity is handled by defining maximum forces
@@
 and yield stress $f_y$ as given for the moment around x-axis and,
-correspondingly, moment around z-axis..
+correspondingly, moment around z-axis.
 
@@
 should be quite usable in most cases.
-
-For rotation around y-axis there is closed form solution only for
-circular cross sections.
-Given approximation is 
-best suited for cases where $b$ and $h$ are quite similar.
-Better approximation for any given $b$ and $h$ can be obtained 
-by integrating distance from center of joint over cross section and
-multiplying it with yield shear stress e.g. using octave.
-
-\lstset{language=octave}
-\begin{lstlisting}
-b=0.01; h=0.01; fy=200e6;
-wpy=fy/2*dblquad(@(x,y) sqrt(x.*x+y.*y),-b/2,b/2,-h/2,h/2)
-38.2
-\end{lstlisting}
 
@@
 \begin{tabular}{l}
-maximum force in x and z directions = $0.5\, b\, h f_y$ \\
-maximum force in y direction = $b\, h\, f_y$  \\
+maximum force in $x$ and $z$ directions = $0.5\, b\, h f_y$ \\
+maximum force in $y$ direction = $b\, h\, f_y$  \\
 \\
-maximum moment around x-axis = $0.25\, b\, h^2 \, f_y$  \\
-maximum moment around y-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
-maximum moment around z-axis = $0.25\, b^2\, h\, f_y$ 
+maximum moment around $x$-axis = $0.25\, b\, h^2 \, f_y$  \\
+maximum moment around $y$-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
+maximum moment around $z$-axis = $0.25\, b^2\, h\, f_y$ 
 \end{tabular}
@@
 
+For rotation around y-axis there is closed form solution only for
+circular cross sections.
+Given approximation is 
+best suited for cases where $b$ and $h$ are quite similar.
+Better approximation for any given $b$ and $h$ can be obtained 
+by integrating distance from center of joint over cross section and
+multiplying it with yield shear stress e.g. using octave.
+Example of calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $
+is shown in Figure \ref{fig:octave-mp}
+
+\begin{figure}[htb!]
+\centering
+\lstset{language=octave}
+\begin{lstlisting}
+b=0.01; h=0.01; fy=200e6;
+wpy=fy/2*dblquad(@(x,z) sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
+38.2
+\end{lstlisting}
+
+\caption{Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using octave.}
+\label{fig:octave-mp}
+\end{figure}
+
 Basic idea in this work can be tested with any framework having motors and hinge constraints.
@@
 Further enhancements were created and tested by forking \bullet\ source code
-and adding new constraints \cite{pbullet}.
+and adding new constraints, \cite{pbullet}.

### 2016-09-24 pdocs/thesis/article-section-3.tex
charpy results in article

@@
 
+\subsection{Block under tension}
 In this section, changes to constraint formulation are clarified by doing few simulation steps
@@
 
-In following sections, results for few selected scenarios are presented.
-Simulation of Charpy impact test shows that plausible results can be obtained even for this quite difficult scenario.
+\subsection{Charpy impact test}
+Simulation of Charpy impact test was used to benchmark approach in more complicated scenario. 
+Material is steel. Density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
+Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
+Support anvils initially have 40 mm open space between them. Their width is 40 mm. 
+Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m. Mass is 19.5 kg.
+Hammer has 40 mm draft.
+If specimen bends about 1.9 radians (108 degrees) it will go between anvils.
+Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
+yield stress of specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa expected energy 
+loss is 122 J.
+
+Usually \bullet\ simulations are done using fixed time step of 1/60 s i.e. 16.67 ms. 
+For this case that is too large. 
+Default time step was selected to be 5 ms outside impact time and 0.1 ms during impact. 
+Automatic time stepping routine changes timestep so that at angles higher than 0.2 5 ms time step is
+used and adjusts it linearly to selected time step between angles of 0.2 and 0.05.
+
+Table \ref{tab:ep2ts} shows energy loss for few timesteps as example to demonstrate sensitivity of solution
+using  elastic-plastic constraint 2.
+
+\begin {table}[htb!]
+\begin{center}
+\begin{tabular}{| c| c|l|}
+\hline
+{\bf Timestep[ms]} & {\bf Energy loss [J]} & {\bf Notes} \\ \hline
+ 0.2 &  170 & specimen penetrates hammer  \\ \hline
+ 0.1 &  120 & \\ \hline
+ 0.05 &  150 & \\ \hline
+\end {tabular}
+\end{center}
+\caption {Energy loss for few timesteps} \label{tab:ep2ts} 
+\end {table}
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=8cm]{figs/article-charpy}
+\caption{Simulation of Charpy impact test}
+\label{fig:f1-ok-results}
+\end{figure}

### 2016-09-24 pdocs/thesis/bullet-plasticity.tex
charpy results in article

@@
 Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
-ultimate tensile strength of specimen. For hinge angle of 1.9 radians and yields stress of 400 MPa expected energy 
+yield stress of specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa expected energy 
 loss is 122 J. Mass is about 40 g and in most cases it is modelled as two 20 g items. In laboratory tests variation for similar 

### 2016-09-24 examples/CharpyDemo/CharpyDemo.cpp
maxSimSubSteps added

@@ -59,6 +59,8 @@ btScalar margin(0.001);
 btScalar floorHE(0.1);
 btScalar defaultTimeStep(0.005);
 btScalar timeStep(defaultTimeStep);
+int initialMaxSimSubSteps = 1;
+int maxSimSubSteps=initialMaxSimSubSteps;
 btScalar initialTimeStep = 1e-4;
 btScalar setTimeStep = initialTimeStep;
 int initialNumIterations = 10;
@@ -623,6 +625,10 @@ public:
 		}
 		dropFocus = true;
 	}
+	void setUiMaxSimSubSteps(Gwen::Controls::Base* control){
+		setInt(control, &maxSimSubSteps);
+		dropFocus = true;
+	}
 
 	Gwen::Controls::Base* pPage;
 	Gwen::Controls::Label* addLabel(string txt){
@@ -885,6 +891,17 @@ public:
 		gy += gyInc;
 		gc->onReturnPressed.Add(pPage, &CharpyDemo::setUiNumIterations);
 	}
+	void addMaxSimSubSteps(){
+		addLabel("maxSimSubSteps");
+		Gwen::Controls::TextBoxNumeric* gc = new Gwen::Controls::TextBoxNumeric(pPage);
+		string text = std::to_string(maxSimSubSteps);
+		gc->SetText(text);
+		gc->SetToolTip("Number of simulation steps for render");
+		gc->SetPos(gxi, gy);
+		gc->SetWidth(wxi);
+		gy += gyInc;
+		gc->onReturnPressed.Add(pPage, &CharpyDemo::setUiMaxSimSubSteps);
+	}
 	bool isDampingUsed(){
 		switch (m_mode){
 		case 2:
@@ -943,6 +960,7 @@ public:
 		if (solverType != BT_MLCP_SOLVER){
 			addNumIterations();
 		}
+		addMaxSimSubSteps();
 		addPauseSimulationButton();
 		addSingleStepButton();
 		addRestartButton();
@@ -2015,7 +2033,12 @@ void CharpyDemo::stepSimulation(float deltaTime){
 	}
 	if (m_dynamicsWorld)	{
 		PlasticityData::setTime(currentTime);
-		stepCount += m_dynamicsWorld->stepSimulation(timeStep, 30, timeStep);
+		btScalar maxSimulationStep(timeStep*maxSimSubSteps);
+		if (deltaTime > maxSimulationStep){
+			deltaTime = maxSimulationStep;
+		}
[…]

### 2016-09-25 examples/CharpyDemo/CharpyDemo.cpp
optional ccd added to charpy - does not work well

@@ -54,7 +54,6 @@ btScalar startAngle(initialStartAngle);
 long  initialDisplayWait = 0;
 long displayWait = initialDisplayWait;
 long setDisplayWait = displayWait;
-btScalar ccdMotionThreshHold(0.001);
 btScalar margin(0.001);
 btScalar floorHE(0.1);
 btScalar defaultTimeStep(0.005);
@@ -124,6 +123,8 @@ btScalar initialFrequencyRatio(10);
 btScalar frequencyRatio(initialFrequencyRatio);
 bool initialLimitIfNeeded = true;
 bool limitIfNeeded=initialLimitIfNeeded;
+bool initialUseCcd = false;
+bool useCcd = initialUseCcd;
 float energy = 0;
 float maxEnergy;
 btDynamicsWorld* dw;
@@ -281,23 +282,6 @@ void setInt(Gwen::Controls::Base* control, int * vp){
 	*vp = fv;
 }
 
-/**
-http://www.bulletphysics.org/mediawiki-1.5.8/index.php?title=Anti_tunneling_by_Motion_Clamping
-*/
-void resetCcdMotionThreshHold()
-{
-	btScalar radius(ccdMotionThreshHold / 5.f);
-	for (int i = dw->getNumCollisionObjects() - 1; i >= 0; i--)
-	{
-		btCollisionObject* obj = dw->getCollisionObjectArray()[i];
-		btRigidBody* body = btRigidBody::upcast(obj);
-		if (body && body->getMotionState())
-		{
-			body->setCcdMotionThreshold(ccdMotionThreshHold);
-			body->setCcdSweptSphereRadius(radius);
-		}
-	}
-}
 float getRotationEnergy(btScalar invInertia, btScalar angularVelocity){
 	if (invInertia == 0){
 		return 0.f;
@@ -602,6 +586,12 @@ public:
 		limitIfNeeded = cb->IsChecked();
 		restartHandler(control);
 	}
+	void setUseCcd(Gwen::Controls::Base* control){
+		Gwen::Controls::CheckBox* cb =
+			static_cast<Gwen::Controls::CheckBox*>(control);
+		useCcd = cb->IsChecked();
+		restartHandler(control);
+	}
 	void setVariableTimeStep(Gwen::Controls::Base* control){
 		Gwen::Controls::CheckBox* cb =
[…]

### 2016-09-27 pdocs/thesis/article-section-3.tex
charpy results tuning in article
demolisher results in article

@@
 
+Figure \ref{fig:charpy} shows screenshot from simulation.
 \begin{figure}[htb!]
@@
 \caption{Simulation of Charpy impact test}
-\label{fig:f1-ok-results}
+\label{fig:charpy}
 \end{figure}
+
+\subsection{Demolisher}
+Demolisher is  collection of possible scenarios that can be added to games to provide ductile joints between
+rigid bodies.  Simulation can be done in real time on modest hardware. 
+Figure \ref{fig:demolisher} shows three screenshots from simulation.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=5cm]{figs/demolisher-pre}
+\includegraphics[height=5cm]{figs/demolisher-wip}
+\includegraphics[height=5cm]{figs/demolisher-done}
+\caption{Simulation of Demolisher scenario}
+\label{fig:demolisher}
+\end{figure}
+
+Vehicle is modelled as single box btRaycastVehicle with four driving wheels. Wheels are rendered for visual feedback.
+In left picture, initial state is shown. Gate in front is modelled to be so slender that it has visible elastic deflection.
+Gate support is modelled as heavy box.
+Bridge has rigid ramps and supports at both ends.
+Gravity is applied gradually to avoid collapse of bridge at start of simulation.
+In middle picture,  bridge is shown in state where car hs driven once over bridge without stopping and
+other bodies have taken light damage.
+In right picture, some of joints have been broken.

### 2016-09-29 pdocs/thesis/article-section-2.tex
demolisher results in article
article tuning

@@
 Symbols are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
-Figure \ref{fig:eom-contact} describes collision of two bodies $B_1$ and $B_2$.
+Figure \ref{fig:eom-contact} describes collision of two bodies $B_1$ and $B_2$
 
@@
 
-$\vec{r}_i$ is orientation of body i.
-$\vec{p}_k $ orientation of contact point k. 
-$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point k.
+where $\vec{r}_i$ is orientation of body i,
+$\vec{p}_k $ is orientation of contact point k,
+$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point k
+and
 $\vec{n}_{k}$ is contact normal for contact point k.
@@
 
-$\vec{r}_{anc}^{\,i}$ is used to define at which point
+where $\vec{r}_{anc}^{\,i}$ is used to define at which point
 joint constraint is applied.
@@
 and adding new constraints, \cite{pbullet}.
+Instructions for using  windows executable and  source code are available, \cite{bp}.

### 2016-09-29 pdocs/thesis/article-section-3.tex
article tuning

@@
 For internal contact values typical values are shown
-as number of contacts and detailed values differ.} 
+as number of contacts and values at contact points are different.} 
 \label{tab:freeBlockValues} 
@@
 There are also e.g. parameters for controlling joint motors.
-Additional equation is created if for each additional constraint. 
+Additional equation is created for each additional constraint. 
 Enabled springs and motors add one row and
@@
 Gravity is applied gradually to avoid collapse of bridge at start of simulation.
-In middle picture,  bridge is shown in state where car hs driven once over bridge without stopping and
+In middle picture,  bridge is shown in state where car has driven once over bridge without stopping and
 other bodies have taken light damage.

### 2016-10-01 pdocs/thesis/article-conclusion.tex
demolisher results in article
article tuning

@@
 Approach presented in this work can be used in gaming industry to provide more realistic 
-simulations without significant extra work. For gaming purposes presented method works 
+simulations without significant extra work. 
+For gaming purposes presented method works 
 best in scenarios where connected parts are relatively heavy. This allows normal 
 integration timestep to be used without stability issues.
-This kind of metodology also opens large area of combining old structural analysis
-methods to modern simulation frameworks.
+
+This kind of metodology also opens large area of combining  structural analysis
+methods to realtime simulation frameworks.
+Handling of significant axial and shear forces and moments at same time
+is one possible area of further studies.
+Handling of more complex structures is already being studied by author.
+
+There is also work to be done in the area of performance optimization.
+One possible area is reusing once calculated values if memory usage is not highly limited.
+This could done e.g. when dealing with high frequency modes.
+Another area is usage of graphics processing unit (GPU) for calculation.
+\bullet\ already has already experimental GPU pipeline but most constraint types are not 
+yet supported. GPU support was one reason for selecting
+it to be used in this study
+
+Integration to 3D modelling and animation software products 
+should also be solved before we can expect to
+see realistic plasticity in main stream games. 
+\bullet\ is already integrated into several ones and it was another significant reason for selecting
+it to be used in this study.
+
 As already noted in \cite{cg1988}, the modeling of inelastic deformation

### 2016-10-01 examples/plasticity/PlasticityExampleBrowser.cpp
0.9.3 release binary

@@ -613,7 +613,7 @@ bool PlasticityExampleBrowser::init(int argc, char* argv[])
 #endif
 #define BLEN 256
 	char buf[BLEN];
-	sprintf_s(buf, BLEN, "0.9.2 - %d-%02d-%02d - %s",
+	sprintf_s(buf, BLEN, "0.9.3 - %d-%02d-%02d - %s",
 		tm.tm_year + 1900, tm.tm_mon + 1, tm.tm_mday,
 		btScalarType);
 #if defined (_DEBUG) || defined (DEBUG)

### 2016-10-02 pdocs/thesis/article-section-2.tex
Fixes in article based on Timo's comments

@@
 \begin {table}[htb!]
+\caption {Nomenclature for equations of motion}
+\label{tab:eom}
 \begin{center}
@@
 \end{center}
-\caption {Nomenclature for equations of motion} \label{tab:eom}
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Additional terms  for generalized equations of motion}
+\label{tab:eom-g}
 \begin{center}
@@
 \end{center}
-\caption {Additional terms  for generalized equations of motion} \label{tab:eom-g}
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Constraint parameters}
+\label{tab:constraintParameters} 
 \begin{center}
@@
 \end{center}
-\caption {Constraint parameters} \label{tab:constraintParameters} 
 \end {table}
@@
 \begin{tabular}{l}
-Under plastic load\\
+Under fully plastic load\\
 stress is at yield\\
@@
 \end{tikzpicture}
-\caption{Axial stress distribution over cross section for bending under elastic and plastic loads.}
+\caption{Axial stress distribution over cross section for bending under elastic and fully plastic loads.}
 \label{fig:wp}
@@
 are integrals of perpdendicular distance 
-and yield stress $f_y$ as given for the moment around x-axis and,
-correspondingly, moment around z-axis.
+and yield stress $f_y$ as given for the moment around $x$-axis 
+and moment around $z$-axis, respectively.
 

### 2016-10-02 pdocs/thesis/article-section-3.tex
article tuning
Fixes in article based on Timo's comments

@@
 so that it reflects scenario under investigation.
-Concrete density is $2000 \frac{kg}{m^3}$ and steel density is $7800 \frac{kg}{m^3}$. 
+Concrete density is $2000\ \frac{kg}{m^3}$ and steel density is $7800\ \frac{kg}{m^3}$. 
 Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
-Gravity is $10 \frac{m}{s^2}$. Simulation step is $\frac{1}{60} s$ and 10 
+Gravity is $10\ \frac{m}{s^2}$. Simulation step is $\frac{1}{60}\ s$ and 10 
 iterations are done for single step.
-Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t=0.548 s 
+Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t = 0.548 s 
 ($\sqrt{2 \cdot 1.5 / 10}$).
 
-Single simulation step is done using following substeps. 
+Single simulation step is performed using following substeps. 
 In this work, changes are made only to constraint setup and update actions.
@@
 \begin {table}[htb!]
+\caption {Simulation values for unconstrained case. 
+For internal contact values typical values are shown
+as number of contacts and values at contact points are different.} 
+\label{tab:freeBlockValues} 
 \begin{center}
@@
 \end{center}
-\caption {Simulation values for unconstrained case. 
-For internal contact values typical values are shown
-as number of contacts and values at contact points are different.} 
-\label{tab:freeBlockValues} 
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Simulation values for unconstrained case with continuous collision detection. } 
+\label{tab:freeBlockValuesWithCcd} 
 \begin{center}
@@
 \end{center}
-\caption {Simulation values for unconstrained case with continuous collision detection. } 
-\label{tab:freeBlockValuesWithCcd} 
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Constraint parameter values for fixed constraint} 
+\label{tab:fixedBlockValues} 
 \begin{center}
@@
 \end{center}
-\caption {Constraint parameter values for fixed constraint} \label{tab:fixedBlockValues} 
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Selected constraint parameters for elastic-plastic constraint 2}
+\label{tab:ep2Parameters} 
 \begin{center}
@@
 \end{center}
-\caption {Selected constraint parameters for elastic-plastic constraint 2} \label{tab:ep2Parameters} 
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Constraint parameter values for elastic-plastic constraint} 
+\label{tab:epBlockValues} 
 \begin{center}
@@
 \end{center}
-\caption {Constraint parameter values for elastic-plastic constraint} \label{tab:epBlockValues} 
 \end {table}
@@
 \begin {table}[htb!]
+\caption {Constraint parameter values for elastic-plastic constraint 2} 
+\label{tab:ep2BlockValues} 
 \begin{center}
@@
 \end{center}
-\caption {Constraint parameter values for elastic-plastic constraint 2} \label{tab:ep2BlockValues} 
 \end {table}
@@
 Simulation of Charpy impact test was used to benchmark approach in more complicated scenario. 
-Material is steel. Density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
+Material is steel and density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
 Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
@@
 \begin {table}[htb!]
+\caption {Energy loss for few timesteps}
+\label{tab:ep2ts} 
 \begin{center}
@@
 \end{center}
-\caption {Energy loss for few timesteps} \label{tab:ep2ts} 
 \end {table}
@@
 Demolisher is  collection of possible scenarios that can be added to games to provide ductile joints between
-rigid bodies.  Simulation can be done in real time on modest hardware. 
+rigid bodies.  Rigid concrete bodies are connected by steel enforcement.
+Concrete density is 2000 $\frac{kg}{m^3}$ and steel density is 7800 $\frac{kg}{m^3}$. 
+Steel is assumed to handle load. Yield stress of steel is 200 MPa.
+Gravity is 10  $\frac{m}{s^2}$. Simulation step is $\frac{1}{60} $ s and 10 
+iterations are done for single step. Multiple simulation steps are done for each render if needed.
+
+Simulation can be done in real time on modest hardware. 
 Figure \ref{fig:demolisher} shows three screenshots from simulation.

### 2016-10-05 pdocs/thesis/article-introduction.tex
tuning of article
Aki's comments on article

@@
+% !TeX root = article.tex
 \section{Introduction}
@@
 The application of modern computer game techniques enables the description of complex dynamic systems 
-such as military vehicles with a high level of detail while still solving the equations in real-time.
-Film production and war games, in particular, is a key area that have benefited from simulation technology. 
+such as military vehicles with a high level of detail while still solving the equations of motion in real-time.
+Film production and war games, in particular, is a key area that have benefited from rapidly enhanced simulation technology. 
 In practice, games are often accomplished using an open-source platform such 
@@
 Computational methods used in physics engines are divided to modules that handle collision detection and 
-contact description and modules that handle solution of equations in real-time. Equations need to be 
+contact description and modules that handle solution of equations of motion in real-time. Equations of motion need to be 
 solved can further be subdivided to be associated to motion, constraints and collisions. 
-Velocity-based formulation is typically used in constraint based rigid body simulation. 
-Friction is typically taken into account and mechanical joints are handled by constraint equations.
-Detailed description of various components can be found in e.g. \cite{erleben.thesis}.
+Velocity-based formulation is typically used in constraint based rigid body simulation, \cite{erleben.thesis}. 
+Friction is typically taken into account and mechanical joints are handled by constraint equations,
+\cite{erleben.thesis}.
 
 Plasticity is not typically taken into account in gaming solutions. 
-Breaking of various objects typically takes place based on collision or impulse.
+Breaking of various objects often takes place based on collision or impulse.
 Nevertheless, breaking of steel or reinforced concrete structures using this approach 
-is not appropriate making a simulation to look unrealistic. Theory for handling of plasticity 
+is not appropriate making a simulation to look realistic. Theory for handling of plasticity 
 has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} 
 present a method for modeling and animating of elastic and plastic objects in real-time using 
-point based animation. This approach is not been widely used in simulation applications.  
+point based animation. This approach is not been widely used in computer games.  
 On major issue is collision handling of deformable objects.
@@
 This study will introduce an approach to account plastic deformation in game applications.   
-In the introduced method, the plastic deformation takes place if force or moment exceeds given 
+In the introduced method, the plastic deformation takes place if force or moment exceeds predefined 
 limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
-The approach is based on using joint motors to model plasticity. \citet[p.~90]{erleben.thesis} 
-suggests similar method for modelling friction in joints. Adjacent objects are connected by motors. 
+The approach is based on using joint motors to model plasticity. 
+This study extends a method introduced by
+\citet{erleben.thesis} 
+%\citet[p.~90]{erleben.thesis} 
+which is originally proposed for modelling friction in joints. 
+In the introduced method adjacent objects are connected by motors. 
 Motor power production limits are estimated based on plastic section modulus. 

### 2016-10-05 pdocs/thesis/article-section-2.tex
Aki's comments on article

@@
 developers and film production teams.
- \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details on why 
+%\citet[p.~45]{erleben.thesis} 
+\citet{erleben.thesis} 
+provides reasoning and theoretical details on why 
 velocity-based formulation is  popular in constraint-based rigid body simulation. 
@@
 Background
-for velocity based formulation shown here is based on  \citet[p.~45-50]{erleben.thesis}.
+for velocity based formulation shown here is based on  
+\citet{erleben.thesis}.
+%\citet[p.~45-50]{erleben.thesis}.
 % pdf page 64
 In following section, these formulations will be clarified by simple example using \bullet\ implementation.
+
 Impulse $\vec{J}$
@@
 
-Using Newton's second law of motion $\vec{F}=m\vec{a}$ one can solve for the velocity,
-$\vec{v}^{\Delta t}$
+Newton's second law of motion $\vec{F}=m\vec{a}$ one can solve for the velocity,
+$\vec{v}^{\Delta t}$ as
 \begin{equation} \label{eq:impulseIntegraWithNewton}
@@
 by integrating the velocity.
-Updates after each step can be summarized as Equation
- \ref{eq:eomL} for locations and  Equation 
-\ref{eq:eomV} for velocities.
+Updates after each step can be summarized  for locations and  
+for velocities correspondingly as follows.
 
@@
 
-Symbols are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
+Symbols used in Equations \ref{eq:eomL} and \ref{eq:eomV}
+are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
 Figure \ref{fig:eom-contact} describes collision of two bodies $B_1$ and $B_2$
@@
 
-where $\vec{r}_i$ is orientation of body i,
-$\vec{p}_k $ is orientation of contact point k,
-$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point k
+where $\vec{r}_i$ is orientation of body $i$,
+$\vec{p}_k $ is orientation of contact point $k$,
+$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point $k$
 and
-$\vec{n}_{k}$ is contact normal for contact point k.
-With this formulation, simplified scenario can be simulated.
+$\vec{n}_{k}$ is contact normal for contact point $k$.
+By employing this formulation, simplified scenario can be simulated.
 
@@
 Friction in contacts and joint constraints can be handled in unified way by refactoring
-equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}
+equation \ref{eq:eomV} as,  
+\citet{erleben.thesis}
+%\citet[p.~66-67]{erleben.thesis}
 \begin{equation} \label{eq:eomV2}
@@
 
-Added symbols are summarized in Table
+Symbols used in Equation \ref{eq:eomV2} are summarized in Table
 \ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
@@
 Mathematical background and detailed examples are available by \cite{ode.joints}.
-Joints are also discussed in detail in  \citet[p.~60-90]{erleben.thesis}.
+Joints are also discussed in detail in  
+\citet{erleben.thesis}.
+%\citet[p.~60-90]{erleben.thesis}.
 Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
@@
 Engineering stress and strain mean that original dimensions are used in stress calculation,
-\citet[p.~108]{dowling}.
+\citet{dowling}.
+%\citet[p.~108]{dowling}.
 Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
@@
 Strain hardening is taken into account in this work mainly by assuming that plasticity in bending
-expands, \citet[p.~672]{dowling}.
+expands, 
+\citet{dowling}.
+%\citet[p.~672]{dowling}.
 Material that starts to yield first is hardened and yielding moves slightly.

### 2016-10-05 pdocs/thesis/article.tex
pdocs/article
pdocs/article changes based on Aki's comments
first draft of section 2 for article
article tuning for section 2
article section 3 started
velocity based formulation basics added to article using erleben as source
more math details in article section 2
added maximum forces in article section 2
working on Aki's comments on article
skeleton for conclusion
Aki's comments on article

--- 2016-01-02 pdocs/thesis/main.tex
@@
 % based on latex template at http://www.lut.fi/web/en/library/dissertations1 
-\def\mytitle{Handling of plasticity in physics engine}
+\def\mytitle{Description of plasticity in physics engine}
 \def\myisbn{978-952-265-XXX-X}
@@
 \def\myname{Simo Nikula}
-\documentclass[a4paper,twoside,12pt,notitlepage,openright]{article}
+\def\lut{Lappeenranta University of Technology}
+\def\bullet{Bullet Physics}
+\def\demolisher{DemolisherDemo}
+\def\charpy{CharpyDemo}
+\documentclass[a4paper,twoside,12pt,notitlepage]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{amsmath}
+\usepackage{listings}
+% \usepackage{mathdots}  % improved \ddots and \iidots
 \usepackage{amssymb}
@@
 \usepackage{ifthen}
-\usepackage[hmargin=3.0cm,vmargin=3.6cm]{geometry} % setting marginals
+\usepackage[hmargin=2.0cm,vmargin=1.6cm]{geometry} % setting marginals
 \usepackage{fancyhdr,extramarks}  % header ja footer manipulation
 \usepackage{times}  % to change font to times
-\usepackage{setspace} % for linespacing
-
+\usepackage{setspace} % for linespacingx
+\usepackage{tikz}
+\usetikzlibrary{arrows}
+\usetikzlibrary{arrows.meta}
+\usepackage{caption}
+\usepackage{varwidth}
+\usepackage{enumitem} % 
 \newcommand{\nomunit}[1]{\renewcommand{\nomentryend}{\hspace*{\fill}#1}} % Inserts units on the right at symbol list
@@
 \pagestyle{empty}
-\include{nimio}
-\include{abstract}
-\addtocontents{toc}{\contentsline {section}{Abstract}{}}
-
-\include{acknowledgments}
-\addtocontents{toc}{\contentsline {section}{Acknowledgments}{}}%
-\input{preface}
-\cleardoublepage%
-\thispagestyle{empty}%
-\addtocontents{toc}{\contentsline{section}{Contents}{}}%
-
-\tableofcontents%
-\thispagestyle{empty}%
-\cleardoublepage%
-\pagestyle{fancy}%
-
-\cleardoublepage%
-\printnomenclature[2.0cm]%
-
-\include{introduction}
-
-%\def\refname{what ever}% to change References title
-\addcontentsline{toc}{section}{References}%
+\title{\mytitle}
+\author{Simo Nikula, Aki Mikkola and Timo Björk}
+\author{}
+\maketitle
+\include{article-introduction}
+\include{article-section-2}
+\include{article-section-3}
+\include{article-conclusion}
 
@@
 \extramarks{References}{References}
+\bibliographystyle{LUTapa}% LUTapa
 \bibliography{ref}%
-\bibliographystyle{LUTapa}%
 

### 2016-10-07 pdocs/thesis/article-section-2.tex
Finetuning of article - italic font usage

@@
 \node[anchor=west] at (4.5,4.5) {
-$\vec{r}_i$  =  orientation of body i 
+$\vec{r}_i$  =  orientation of body $i$
 };
-\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point k};
+\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point $k$};
 \node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
-\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point k };
+\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point $k$};
 \end{tikzpicture}
@@
 $\vec{p}_k $ is orientation of contact point $k$,
-$\vec{r}_{ik} $ is vector  between center of gravity of body i and contact point $k$
+$\vec{r}_{ik} $ is vector  between center of gravity of body $i$ and contact point $k$
 and
@@
 {\bf Symbol} & {\bf Description} \\  \hline
-$\vec{r}_i$ & position of center of mass for body i  \\ \hline
-$\vec{q}_i$ & orientation for body i as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
+$\vec{r}_i$ & position of center of mass for body $i$  \\ \hline
+$\vec{q}_i$ & orientation for body $i$ as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
 \hline
-$\vec{p}_i$ & contact or joint point i  \\ \hline
+$\vec{p}_h$ & contact or joint point $k$  \\ \hline
 $\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
@@
 \\ \hline
-$\vec{v}_i$ & linear velocity of  center of mass for body i   \\ \hline
-$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
+$\vec{v}_i$ & linear velocity of  center of mass for body $i$   \\ \hline
+$\vec{\omega}_i$ & angular velocity of center of mass for body $i$  \\ \hline
 $\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
@@
 \\ \hline
-$I_i$ & inertia tensor for body i  \\ \hline
+$I_i$ & inertia tensor for body $i$  \\ \hline
 $C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
@@
 \node[anchor=west] at (4.5,3.5) {
-$\vec{r}_i$  =  orientation of body i 
+$\vec{r}_i$  =  orientation of body $i$ 
 };
-\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector i};
+\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector $i$};
 \end{tikzpicture}

### 2016-10-08 pdocs/thesis/abstract.tex
pdocs and debugDraw for ElasticPlastic2Constraint
Aki's comments on article and in general

@@
 In this work suitable methods for handling of plasticity in physics engine are presented.
-Handling of plasticity basically means that objects involved may absorb energy and possibly break.
+Handling of plasticity basically means that involved bodies may absorb energy and possibly break.
 
@@
 
-Keywords: Physics engine, simulation, dynamics, plasticity, deformation.
+Keywords: Physics engine, simulation, dynamics, plasticity, inelastic, multibody dynamics,  physical simulation, deformation.

### 2016-10-08 pdocs/thesis/article-conclusion.tex
Aki's comments on article and in general

@@
 
-This study has presented an approach to account plastic deformation in 
+This study presents an approach to account plastic deformation in 
 velocity based formulation.
@@
 limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
-Approach presented in this work can be used in gaming industry to provide more realistic 
-simulations without significant extra work. 
+Approach presented in this work can be used in gaming industry to provide realistic 
+simulations. 
 For gaming purposes presented method works 
-best in scenarios where connected parts are relatively heavy. This allows normal 
+best in scenarios where connected parts are heavy. This allows normal 
 integration timestep to be used without stability issues.
@@
 is one possible area of further studies.
-Handling of more complex structures is already being studied by author.
 
 There is also work to be done in the area of performance optimization.
-One possible area is reusing once calculated values if memory usage is not highly limited.
+One possible area is reusing once calculated values if memory usage is not limited.
 This could done e.g. when dealing with high frequency modes.

### 2016-10-08 pdocs/thesis/article-introduction.tex
Aki's comments on article and in general

@@
 Plasticity is not typically taken into account in gaming solutions. 
-Breaking of various objects often takes place based on collision or impulse.
+Breaking of various bodies often takes place based on collision or impulse.
 Nevertheless, breaking of steel or reinforced concrete structures using this approach 
@@
 has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} 
-present a method for modeling and animating of elastic and plastic objects in real-time using 
+present a method for modeling and animating of elastic and plastic bodies in real-time using 
 point based animation. This approach is not been widely used in computer games.  
-On major issue is collision handling of deformable objects.
+On major issue is collision handling of deformable bodies.
 
@@
 which is originally proposed for modelling friction in joints. 
-In the introduced method adjacent objects are connected by motors. 
+In the introduced method adjacent bodies are connected by motors. 
 Motor power production limits are estimated based on plastic section modulus. 
@@
 simulations without significant extra work. For gaming purposes presented method works 
-best in scenarios where connected parts are relatively heavy. This allows normal 
+best in scenarios where connected parts are heavy. This allows normal 
 integration timestep to be used without stability issues. 

### 2016-10-08 pdocs/thesis/article-section-2.tex
Aki's comments on article and in general

@@
 
-where $\vec{r}_{anc}^{\,i}$ is used to define at which point
+In Figure \ref{fig:eom-joint}, $\vec{r}_{anc}^{\,i}$ is used to define at which point
 joint constraint is applied.
@@
 Constraint processing in \bullet\ is based on ODE, \cite{ode}.
-Mathematical background and detailed examples are available by \cite{ode.joints}.
 Joints are also discussed in detail in  
@@
 are created for each constraint. 
+Derivation for terms in Equation \ref{eq:constraintEquation}
+can be done using position and orientation of connected bodies.
+E.g. for ball joint formulation is based on both joint points having same position.
+In contact cases formulation is easier if it is done using velocities, \cite{ode.joints}.
 
@@
 This provides reasonable means to simulate loss of energy in collisions.
-Simulation of breaking of objects made of ductile material can be made more realistic by splitting rigid bodies
+Simulation of breaking of bodies made of ductile material can be made more realistic by splitting rigid body
 to multiple bodies which are connected by energy absorbing joints.
@@
 
-where $\sigma$ is stress, $E$ is Youngs modulus and $f_y$ is yield stress.
+In Figure \ref{fig:sscurve}, $\sigma$ is stress, $E$ is Youngs modulus and $f_y$ is yield stress.
 Engineering stress and strain mean that original dimensions are used in stress calculation,
@@
 It should be noted that geometry
-of objects is not updated during analysis and thus engineering stress-strain properties should
+of bodies is not updated during analysis and thus engineering stress-strain properties should
 be used.
 
-Strain hardening is taken into account in this work mainly by assuming that plasticity in bending
+In this work, strain hardening is taken into account by assuming that plasticity in bending
 expands, 
@@
 %\citet[p.~672]{dowling}.
-Material that starts to yield first is hardened and yielding moves slightly.
+Material that starts to yield first is hardened and as a result of which yielding moves.
 This can be seen e.g. by bending paperclip. It does not break at low angles but can take few full bends.
 
-Difference between elastic and plastic section modulus is shown in Figure \ref{fig:wp}.
+Difference between elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.
 
@@
 
-If stress is below yield limit $f_y$, stress and strain are linear within material.
-If cross section is fully plastic, stress is assumed to be at yield level over whole cross section and 
-so plastic section modulus is higher than elastic section modulus.
+As shown in Figure 2.4, if stress is below yield limit $f_y$, stress and strain are linear within material.
+If cross section is fully plastic, stress is assumed to be at yield level over whole cross section such that 
+plastic section modulus is higher than elastic section modulus.
 
-Plasticity is handled by defining maximum forces
+In this work, plasticity is handled by defining maximum forces
 in Equations \ref{eq:lambdaLow} and  
-\ref{eq:lambdaHigh} using plastic capasities as maximum forces.
+\ref{eq:lambdaHigh} using plastic capasities which are defined below.
 
@@
 Yield shear stress is assumed to be $ 0.5\, f_y$ using Tresca yield critetion.
-If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$).
+If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$), \cite{dowling}.
+% p. 262, p. 268
 These are not exact values in multiaxial stress state but they
-should be quite usable in most cases.
+are usable in most cases.
 
@@
 
-For rotation around y-axis there is closed form solution only for
+For rotation around $y$-axis there is closed form solution only for
 circular cross sections.
 Given approximation is 
-best suited for cases where $b$ and $h$ are quite similar.
+best suited for cases where $b$ and $h$ are similar.
 Better approximation for any given $b$ and $h$ can be obtained 
 by integrating distance from center of joint over cross section and
-multiplying it with yield shear stress e.g. using octave.
+multiplying it with yield shear stress e.g. using octave, \cite{octave}.
 Example of calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $
@@
 
-Basic idea in this work can be tested with any framework having motors and hinge constraints.
+Basic idea introduced in this study can be tested with any framework having motors and hinge constraints.
 This can be done by setting target velocity of motor to zero and limiting maximum motor impulse to plastic moment 

### 2016-10-08 pdocs/thesis/article-section-3.tex
Aki's comments on article and in general

@@
 
-System has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
+In the first example, system under investigation
+has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
 Cross section is one square meter.
@@
 Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
-Gravity is $10\ \frac{m}{s^2}$. Simulation step is $\frac{1}{60}\ s$ and 10 
+Gravity is $10\ \frac{m}{s^2}$ to keep numbers simple. 
+Simulation step is $\frac{1}{60}\ s$ and 10 
 iterations are done for single step.
-Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at t = 0.548 s 
+Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at $t$ = 0.548 s 
 ($\sqrt{2 \cdot 1.5 / 10}$).
@@
 \begin{enumerate}
-\item Apply gravity to each non static body. 
+\item Apply gravity to each moveable body. 
 This step makes programming easier as otherwise each program should add forces due to gravity in each step. 
@@
 Prediction is done based on current linear and angular velocities of the body.
-\item Predict contacts. In this phase, continuous collision detection is done based on simplified objects. 
+\item Predict contacts. In this phase, continuous collision detection is done based on simplified bodies. 
 Each rigid body is represented by sphere geometry and contact prediction is done if body moves more than given threshold value during simulation step.
@@
 
-Equations \ref{eq:lambdaLow} and \ref{eq:lambdaHigh} are not active for fixed case.
-For elastic-plastic case maximum impulse is set to product of yield stress, area of steel enforcement and time step (1330).
+Equations \ref{eq:lambdaLow} and \ref{eq:lambdaHigh} are not active for fixed case as
+there are no upper nor lower limit for forces.
+For elastic-plastic case maximum impulse is set to product of yield stress, 
+area of steel enforcement and time step (1330).
 
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ was used to pick up values for internal variables described in Figure \ref{fig:solVars}.
+in \bullet\ is used to pick up values for internal variables described in Figure \ref{fig:solVars}.
 
@@
 \end{description}
-\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. object drifting. 
+\item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. body drifting. 
 Setting erp to 1 would in theory eliminate error in one step but in practice value of 0.2 - 0.8 is used in most cases.
@@
 Six-dof-spring constraint 2 has optional feature to avoid unstability by automatically softening constraint
-spring. Feature is activated if current timeStep is too large for spring-mass system simulation.
+spring. Feature is activated if current timeStep is overly large for spring-mass system simulation.
 The feature is trigged based on the equation below 
@@
 velocity based error instead of force provided by spring.
-In this scenario object behaves as depicted in table \ref{tab:ep2BlockValues} i.e. it does not move at all.
+In this scenario body behaves as depicted in table \ref{tab:ep2BlockValues} i.e. it does not move at all.
 
@@
 \subsection{Charpy impact test}
-Simulation of Charpy impact test was used to benchmark approach in more complicated scenario. 
+In a second numerical example, simulation of Charpy impact test is
+used to benchmark approach in more complicated scenario. 
 Material is steel and density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
 Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
-Support anvils initially have 40 mm open space between them. Their width is 40 mm. 
-Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m. Mass is 19.5 kg.
+Support anvils initially have 40 mm open space between them and their width is 40 mm. 
+Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m and mass is 19.5 kg.
 Hammer has 40 mm draft.
@@
 \subsection{Demolisher}
+Third numerical examples introduces a demolisher.
 Demolisher is collection of possible scenarios that can be added to games to provide ductile joints between
@@
 
-Simulation can be done in real time on modest hardware. 
+Simulation can be done in real time on modest hardware, e.g. 
+on laptop having 2.0 GHz Intel i3 CPU with integrated graphics. 
 Figure \ref{fig:demolisher} shows three screenshots from simulation.
@@
 other bodies have taken light damage.
-In right picture, some of joints have been broken.
+In right picture, two joints have been broken and bridge has fallen from support due to bending.

### 2016-10-08 pdocs/thesis/bullet-plasticity.tex
Aki's comments on article and in general

@@
 \begin{itemize}
-\item Avoid very small and very larger collision shapes
+\item Avoid small and large collision shapes
 \item Avoid large mass ratios
@@
 
-\subsubsection{Breaking of constraints between objects}
+\subsubsection{Breaking of constraints between bodies}
 For standard \bullet constraints breakingImpulseThreshold (BITh) can be defined. If impulse is larger than set limit constraint is activated. 
-This allows object breaking to be simulated with very cheap calculation. 
+This allows breaking to be simulated with fast calculation. 
 This method is too simplified for ductile material and developing more precise and realistic methods is main objective of this work.
@@
 Constraint is made breakable by setting breakingImpulseThreshold.
-This is common technique to provide breakable objects in games and provides visually accetable results for brittle materials.
+This is common technique to provide breakable bodies in games and provides visually accetable results for brittle materials.
 
@@
 This shows how btGeneric6DofSpringConstraints act in case like this. 
-Specimen can absorb very little elastic energy so rotational spring constants are calculated using plastic state i.e.
+Specimen can absorb little elastic energy so rotational spring constants are calculated using plastic state i.e.
 $W=bh^2/4$ where h=b-0.002m and yield stress is used instead of Young’s modulus for rotational springs.
@@
 \label{sec:bullet-charpy-results}
-Results are quite sensitive and obtained in February 2016 using following processes
+Results are sensitive and they were obtained in February 2016 using following processes
 \begin{itemize}
@@
 \subsection{Results for spring2Constraint}
-Results are quite similar as for springConstraint but automatic parameter tuning prevents unstable simulations.
+Results are similar as for springConstraint but automatic parameter tuning prevents unstable simulations.
 Energy loss is about 40 J.
@@
 \subsection{Results for plasticHingeConstraint}
-Results are quite similar as for hinge constraint with motors. 
+Results are similar as for hinge constraint with motors. 
 Breaking point can be configured based on rotation angle which makes material behaviour more realistic.
@@
 \subsection{Results for elasticPlasticConstraint}
-Results are quite similar as for plasticHingeConstraint as this case is dominated by plastic bending.
+Results are similar as for plasticHingeConstraint as this case is dominated by plastic bending.
 Additional logic helps to avoid stablity issues.
@@
 \subsection{Results for elasticPlastic2Constraint}
-Results are quite similar as for elasticPlasticConstraint and this constraint seems to best candidate for further development.
+Results are similar as for elasticPlasticConstraint and this constraint seems to best candidate for further development.
 Table \ref{tab:ep2ts} shows energy loss for few timesteps as example to demonstrate sensitivity of solution.
@@
 as it is based on widely used motor-codes. 
-Additional features are quite simple to implement and make spring based constraints even more usable than currently by 
+Additional features are simple to implement and make spring based constraints even more usable than currently by 
 allowing stiff constraints to be used without instability.

### 2016-10-08 pdocs/thesis/introduction.tex
pdocs/article
Aki's comments on article and in general

--- 2016-02-14 pdocs/thesis/introduction.tex
@@
 rigid body dynamics can be handled in real time in most scenarios.
-Soft body dynamics in which  object are deformable is in many cases still too complicated for real time applications.
-Finite element method will probably replace other methods by providing general solution at some time but not in very near future.
+Soft body dynamics in which  bodies are deformable is in many cases still too complicated for real time applications.
+Finite element method will probably replace other methods by providing general solution at some time but not in near future.
 In this work focus is on method where rigid bodies are connected by constraints.
@@
 
-In this work all theories in area of structural analysis are quite old and e.g. 
+In this work all theories in area of structural analysis are old and e.g. 
 \cite{timoshenko} is typical reference. Main focus is on elastic and 
@@
 E.g. \cite{Obrien:1999:GMA} used finite element model to analyze crack initiation and propagation.
-Already \cite{cg1988} used quite similar principles as in this work to handle plasticity.
+
+Already \cite{cg1988} used similar principles as in this work to handle plasticity and there are a few further
+papers that cite \cite{cg1988}. E.g. following papers provide alternative solutions to similar kind of simulations as 
+are handled in this work 
+\begin{itemize}
+\item \cite{gibson1997survey} presents extensive study on deformable modelling in computer graphics
+\item \cite{o2002graphical} provides excellent paper on ductile behaviour of material
+\item \cite{muller2004point} presents a method for modeling and animating of 
+elastic and plastic bodies in realtime using point based animation
+\item \cite{irving2004invertible} introduces algortithm for  finite element simulation of elastoplastic solids with
+large deformations in reasonable time 
+\end{itemize}
 
@@
 \item plastic deformation takes place if force or moment exceeds given limit
-\item plastic deformation absorbs enegry
+\item plastic deformation absorbs energy
 \item after predefined limit of displacement is exceeded constraint is deactivated 
@@
 Few decades ago software engineering and structural analysis were highly coupled
-and e.g. \cite{bathe} and \cite{cook} contain very detailed instructions on
+and e.g. \cite{bathe} and \cite{cook} contain detailed instructions on
 computer implementation. Those books provided good baseline for many computer 
@@
 thin-walled structures \cite{agifap,fsm,vtb}.  
-Getting and compiling complicated software packages is currently quite easy. 
+Getting and compiling complicated software packages is currently easy. 
 Understanding and getting enough knowledge to extend them requires a fair amount of hard work.
@@
 in commonly used physics engine \bullet. 
-Handling of plasticity basically means that objects involved may absorb kinetic energy and break.
+Handling of plasticity basically means that involved bodies may absorb kinetic energy and break.
 
@@
 
-Methods presented in this work can be used in gaming industry to provide objects that behave more realistically than currently without significant extra work.
+Methods presented in this work can be used in gaming industry to provide 
+bodies that behave more realistically than currently without significant extra work. 
 Efficient simulation algorithms can also help in saving energy. 
-Saving even few wats or percent for single user makes big difference as gaming PCs are estimated to have consumed 75 TWh/year of electricity globally in 2012, \cite{gaming.energy}.
+Saving even few wats or percent for single user makes big difference as gaming PCs are estimated to have consumed 
+75 TWh/year of electricity globally in 2012, \cite{gaming.energy}.
 

### 2016-10-09 pdocs/thesis/article-section-3.tex
Finetuning of article

@@
 Actual values for unconstrained case without continuous collision detection
-are shown in table \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
+are shown in Table \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
 {\it velError} is 5.67 (5.5+0.167) and
@@
 
-Values for fixed constraint are shown in table
+Values for fixed constraint are shown in Table
 \ref{tab:fixedBlockValues}. Constraint is activated in second step and positional error is corrected
@@
 
-Values for elastic-plastic case are shown in tables \ref{tab:epBlockValues}.
+Values for elastic-plastic case are shown in Tables \ref{tab:epBlockValues}.
 Body drops freely during first simulation step and 
@@
 velocity based error instead of force provided by spring.
-In this scenario body behaves as depicted in table \ref{tab:ep2BlockValues} i.e. it does not move at all.
+In this scenario body behaves as depicted in Table \ref{tab:ep2BlockValues} i.e. it does not move at all.
 

### 2016-10-09 pdocs/thesis/article-section-3.tex
Finetuning of article -new pictures to charpy

@@
 
+Figure \ref{fig:charpy-series} shows three detailed pictures from simulation.
+Left picture shows early phase of simulation. In middle picture hammer has not yet collided with specimen.
+In right picture specimen is bending.
+
+\begin{figure}[htb!]
+\centering
+\includegraphics[height=3cm]{figs/article-charpy-1}
+\includegraphics[height=3cm]{figs/article-charpy-2}
+\includegraphics[height=3cm]{figs/article-charpy-3}
+\caption{Detailed pictures from simulation of Charpy impact test}
+\label{fig:charpy-series}
+\end{figure}
+
 Table \ref{tab:ep2ts} shows energy loss for few timesteps as example to demonstrate sensitivity of solution
@@
 
-Figure \ref{fig:charpy} shows screenshot from simulation.
+Figure \ref{fig:charpy} shows overview from simulation as specimen is going between supports.
+
 \begin{figure}[htb!]
@@
 \includegraphics[height=8cm]{figs/article-charpy}
-\caption{Simulation of Charpy impact test}
+\caption{Overview of simulation of Charpy impact test}
 \label{fig:charpy}

### 2016-10-10 pdocs/thesis/article-introduction.tex
refences with article style in article

@@
 is not appropriate making a simulation to look realistic. Theory for handling of plasticity 
-has been presented already in \cite{cg1988}. \cite{muller2004point} and \cite{muller2005meshless} 
+has been presented already in \cite{cg1988}. \cite{muller2005meshless} 
 present a method for modeling and animating of elastic and plastic bodies in real-time using 

### 2016-10-10 pdocs/thesis/article-section-2.tex
refences with article style in article

@@
 Maximum forces and moments for 
-rectangular section using constant yield stress
-are summarized in Figure \ref{fig:maxForces}.
+rectangular section with width $b$ and height $h$ using constant yield stress
+are summarized in Table \ref{tab:maxForces}.
 Yield shear stress is assumed to be $ 0.5\, f_y$ using Tresca yield critetion.
@@
 
-\begin{figure}[htb!]
-\centering
-\begin{tikzpicture}
-\coordinate (O) at (2,2,0);
+\begin {table}[htb!]
+\caption {Maximum forces and moments for 
+rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
+\label{tab:maxForces} 
+\begin{center}
+\begin{tabular}{| c| c|}
+\hline
+{\bf Direction} & {\bf Maximum value}  \\ \hline
+maximum shear force & $0.5\, b\, h f_y$ \\ \hline
+maximum normal force & $b\, h\, f_y$  \\ \hline
+maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
+maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
+maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
+\end{tabular}
+\end{center}
+\end {table}
 
-% axes
-\draw[->] (O) -- ++(1,0,0) node[anchor=north east]{$x$};
-\draw[->] (O) -- ++(0,1,0) node[anchor=north west]{$y$};
-\draw[->] (O) -- ++(0,0,1) node[anchor=south]{$z$};
-
-% 1,3,1 block, base at y=1.5
-\draw (1,1.5,0) -- ++(0,3,0) -- ++(-1,0,0);
-\draw (0,1.5,1) -- ++(1,0,0) -- ++(0,3,0) -- ++(-1,0,0) -- ++(0,-3,0);
-\draw (1,1.5,0) -- ++(0,0,1);
-\draw (0,4.5,0) -- ++(0,0,1);
-\draw (1,4.5,0) -- ++(0,0,1);
-
-% l is not significant
-%\draw[Stealth-Stealth] (2,1.5,0) -- ++(0,3,0);
-%\node[] at (2.5,3) [anchor=west]{l};
-
-\draw[Stealth-Stealth] (0,5,0) -- ++(1,0,0);
-\node[] at (0.5,5) [anchor=south]{$b$};
-
-\draw[Stealth-Stealth] (0,5,0) -- ++(0,0,1);
-\node[] at (0,5,0.5) [anchor=south east]{$h$};
-
-\draw[ultra thin](0,4.5,0) -- ++(0,0.6);
-\draw[ultra thin](1,4.5,0) -- ++(0,0.6);
-\draw[ultra thin](0,4.5,1) -- ++(0,0.6);
-
-\draw[-Stealth] (0.5,3,0.5) -- ++(0,1.5,0);
-\node at (0,3.8) {$\vec{r}_{anc}$};
-
-\node[anchor=west] at (4.0,3) {
-\begin{tabular}{l}
-maximum force in $x$ and $z$ directions = $0.5\, b\, h f_y$ \\
-maximum force in $y$ direction = $b\, h\, f_y$  \\
-\\
-maximum moment around $x$-axis = $0.25\, b\, h^2 \, f_y$  \\
-maximum moment around $y$-axis $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\
-maximum moment around $z$-axis = $0.25\, b^2\, h\, f_y$ 
-\end{tabular}
-};
-
-\end{tikzpicture}
-\caption{Maximum forces in plastic joint for rectangular cross section.}
-\label{fig:maxForces}
-\end{figure}
-
-For rotation around $y$-axis there is closed form solution only for
+For torque there is closed form solution only for
 circular cross sections.

### 2016-10-10 pdocs/thesis/article-section-3.tex
refences with article style in article

@@
 {\bf Impulse} \\  \hline
-0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
+0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
+0.533 &  -1.467 & & & & & -5.33 & 0 \\  \hline
 0.550 &  -1.558 & & & & & -5.5 & 0 \\  \hline
@@
 \end {table}
+
+\include{octave/uc-fig}
 
@@
 {\bf Impulse} \\  \hline
-0.017 &  0 & & & &  &-0.17 & 0 \\  \hline
+0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
 0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline

### 2016-10-11 pdocs/thesis/article-section-3.tex
added octave plot to section 3

@@
 Actual values for unconstrained case without continuous collision detection
-are shown in Table \ref{tab:freeBlockValues}. Penetration is detected at time 0.567 s when
+are shown in Table \ref{tab:freeBlockValues} and 
+location, velocity, posError and velError are depicted in Figure \ref{fig:uc}. 
+Penetration is detected at time 0.567 s when
 {\it velError} is 5.67 (5.5+0.167) and
@@
 \begin {table}[htb!]
-\caption {Simulation values for unconstrained case. 
+\caption {Simulation values for unconstrained case without continuous collision detection. 
 For internal contact values typical values are shown
@@
 
-\include{octave/uc-fig}
+\begin{figure}[htb!]
+\centering
+\input{figs/uc-fig.tikz}
+\caption{Graph for selected values for unconstrained case simulation without continuous collision detection.}
+\label{fig:uc}
+\end{figure}
 

### 2016-10-11 pdocs/thesis/octave/article.m

@@ -1,0 +1,37 @@
+%{
+For octave, initially with version 4.0.3
+unconstrained case
+1: time 
+2: location at end of step 
+3: velError
+4: penetration
+5: posError
+6: rhs
+7: velocity
+8: impulse
+%}
+uc=[
+0.533 ,  -1.467 ,0,0 ,0,0,-5.33 , 0   ;
+0.550 ,  -1.558 ,0,0,0,0, -5.5 , 0   ;
+0.567 ,  -1.511 , 5.67 ,-0.058 ,2.8 ,  21270 , 0.01 , 34000   ;
+0.583 ,  -1.502 , 0.14 ,-0.011 , 0.54, 2570  , 0.55 , 420   ;
+0.600 ,  -1.496 , -0.38,-0.002 , 0.1  , -1000, 0.38 , 0   ;
+0.617 ,  -1.492 ,-0.44 , 0.004 , 0     , -1600, 0.22 , 0   ;
+0.717 ,  -1.497 ,0  ,-0 ,0-0.01 , 50 , -0.01 , 400   ;
+0.817 ,  -1.499 ,0,0,0,0, -0.08 , 700   ;
+0.917 ,  -1.500 ,0,0,0,0, -0.001 , 1000   ;
+];
+
+subplot(2,2,1)
+plot(uc(:,1),uc(:,2),"color","black");
+title("Location");
+subplot(2,2,3)
+plot(uc(:,1),uc(:,5),"color","black");
+title("posError");
+subplot(2,2,2)
+plot(uc(:,1),uc(:,7),"color","black");
+title("Velocity");
+subplot(2,2,4)
+plot(uc(:,1),uc(:,3),"color","black");
+title("velError");
+%print -dtikz "-S600,400" "uc-fig"

### 2016-10-22 pdocs/thesis/article-section-3.tex
missing minus sign added to velError in result for fixed constraint at t=0.033

@@
 0.017 & -0.0028 & & & 	 & -0.17 & 0 \\  \hline
-0.033 & -0.0022 & 0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
+0.033 & -0.0022 & -0.33 & -0.033 & -2200 & 0.033 & 2200 \\  \hline
 0.050 & -0.0018 & -0.13 & -0.027 & -960 & 0.027 & 960 \\  \hline

### 2016-10-22 examples/CharpyDemo/BlockDemo.cpp
velError and posError visualization in BlockDemo

@@ -90,7 +90,7 @@ public:
 	btJointFeedback jf;
 	AxisMapper* axisMapper=0;
 	long stepCount,maxStepCount,syncedStep;
-	btScalar stepTime,gravityRampUpTime;
+	btScalar stepTime,targetTime,gravityRampUpTime;
 	btVector3 m_gravity;
 	btScalar frequencyRatio;
 	btScalar E,fy,G;
@@ -476,10 +476,11 @@ public:
 		b_location=0, b_velocity=0, yStart=0;
 	TimeSeriesCanvas *tsYLocation=0, *tsYVelocity=0, *tsYForce=0;
 	TimeSeriesCanvas *tsZRLocation = 0, *tsZRVelocity = 0, *tsZMoment = 0, *tsXForce=0;
-	TimeSeriesCanvas *tsError = 0;
+	TimeSeriesCanvas *tsVelError = 0,*tsPosError=0;
 #ifdef CDBG_CALLBACK
 #define C_ERROR_SIZE 6
-	float cError[C_ERROR_SIZE];
+	float velError[C_ERROR_SIZE];
+	float posError[C_ERROR_SIZE];
 #endif
 	btScalar getMass(){
 		return btScalar(lsx*lsy*lsz*getDensity(blockSteelScale));
@@ -1309,51 +1310,103 @@ public:
 			tsXForce->nextTick();
 		}
 	}
-	float maxError = 0.1, currentMaxError;
-	void updateMaxError(float offer){
-		offer = btFabs(offer);
-		if (offer > maxError){
-			maxError = offer;
+	float maxVelError = 0.1, currentMaxVelError;
+	void updateMaxVelError(float offer){
+		float absValue = btFabs(offer);
+		if (absValue > maxVelError){
+			maxVelError = absValue;
 		}
+	}
+	float maxPosError = 0.1, currentMaxPosError;
+	void updateMaxPosError(float offer){
+		float absValue = btFabs(offer);
+		if (absValue > maxPosError){
+			maxPosError = absValue;
+		}
+	}
+	/**
+	*/
+	unsigned char getRGBVal(int k1, int b1, int i, int k2, int b2){
+		int v = 255*b1 + 255*k1 * i *3/ C_ERROR_SIZE;
+		if (v>255){
+			v = 255*b2+255*k2*i*3/C_ERROR_SIZE;
+		}
+		if (v < 0){
+			v = 0;
+		}
[…]

### 2016-10-22 examples/plasticity/PlasticityExampleBrowser.cpp
print example number to make bulletDemo.txt direct editing easier

@@ -268,7 +268,7 @@ void selectDemo(int demoIndex)
 				bool isLeft = true;
 				gui->setStatusBarMessage("Status: OK", false);
 			}
-			b3Printf("Selected demo: %s",gAllExamples->getExampleName(demoIndex));
+			b3Printf("Selected demo: %s (%d)",gAllExamples->getExampleName(demoIndex),demoIndex);
 			gui->setExampleDescription(gAllExamples->getExampleDescription(demoIndex));
 			
 			sCurrentDemo->initPhysics();

### 2016-10-23 examples/CharpyDemo/BlockDemo.cpp
velError and posError visualization in BlockDemo modified to use impulses

@@ -690,9 +690,9 @@ public:
 	void addPauseSimulationButton(){
 		Gwen::Controls::Button* gc = new Gwen::Controls::Button(pPage);
 		pauseButton = gc;
-		gc->SetText(L"Pause");
 		gc->SetPos(gx, gy);
 		gc->SetSize(wxi - 4, gyInc - 4);
+		updatePauseButtonText();
 		gc->onPress.Add(pPage, &BlockDemo::handlePauseSimulation);
 	}
 	void addSingleStepButton(){
@@ -1310,16 +1310,17 @@ public:
 			tsXForce->nextTick();
 		}
 	}
-	float maxVelError = 0.1, currentMaxVelError;
-	void updateMaxVelError(float offer){
-		float absValue = btFabs(offer);
+	bool errorScaled = false;
+	float maxVelError = 1000, currentMaxVelError;
+	void updateMaxVelError(int index){
+		float absValue = btFabs(velError[index]);
 		if (absValue > maxVelError){
 			maxVelError = absValue;
 		}
 	}
-	float maxPosError = 0.1, currentMaxPosError;
-	void updateMaxPosError(float offer){
-		float absValue = btFabs(offer);
+	float maxPosError = 1000, currentMaxPosError;
+	void updateMaxPosError(int index){
+		float absValue = btFabs(posError[index]);
 		if (absValue > maxPosError){
 			maxPosError = absValue;
 		}
@@ -1350,22 +1351,25 @@ public:
 	provide clean scaling value
 	*/
 	float getScale(float in){
-		float s = 1000;
+		float s = 1;
 		float v=ceilf(s*in);
 		return v;
 	}
 	void addErrorTimeSeries(){
 #ifdef CDBG_CALLBACK
+		if (0==m_constraint){
+			return;
+		}
 		const char *label[] = {"0","1","2","3","4","5"};
 		int tsWidth = 300, tsHeight = 300;
 		CommonGraphicsApp * app = PlasticityExampleBrowser::getApp();
 		if (0 == tsVelError){
 			tsVelError = new TimeSeriesCanvas
-				(app->m_2dCanvasInterface, tsWidth, tsHeight, "velError [mm or mrad/s]");
+				(app->m_2dCanvasInterface, tsWidth, tsHeight, "constraint's velImpulse [Ns]");
 		}
 		if (0 == tsPosError){
 			tsPosError = new TimeSeriesCanvas
[…]

### 2016-10-23 examples/CharpyDemo/BlockDemo.cpp
z-rotation visualization in BlockDemo modified to use yaw/pitch/roll

@@ -281,8 +281,13 @@ public:
 		if (m_body){
 			switch (orientationType){
 			case X:
-				b_location = m_body->getWorldTransform().getRotation().getAngle();
+			{
+				btTransform tr = m_body->getWorldTransform();
+				btScalar yaw, pitch, roll;
+				tr.getBasis().getEulerYPR(yaw, pitch, roll);
+				b_location = yaw;
 				b_velocity = m_body->getAngularVelocity().z();
+			}
 				break;
 			case Y:
 				b_location = m_body->getWorldTransform().getOrigin().y() - yStart;
@@ -1215,7 +1220,7 @@ public:
 		m_dynamicsWorld->setConstraintSolver(m_constraintSolver);
 	}
 	void addYTimeSeries(){
-		int tsWidth = 300, tsHeight = 200;
+		int tsWidth = 300, tsHeight = 300;
 		CommonGraphicsApp * app = PlasticityExampleBrowser::getApp();
 		if (0 == tsYLocation){
 			tsYLocation = new TimeSeriesCanvas
@@ -1248,13 +1253,13 @@ public:
 		}
 	}
 	void addXTimeSeries(){
-		int tsWidth = 300, tsHeight = 200;
+		int tsWidth = 300, tsHeight = 300;
 		CommonGraphicsApp * app = PlasticityExampleBrowser::getApp();
 		if (0 == tsZRLocation){
 			tsZRLocation = new TimeSeriesCanvas
 				(app->m_2dCanvasInterface, tsWidth, tsHeight, "Z rotation");
 		}
-		tsZRLocation->setupTimeSeries(1, intFreq, 0);
+		tsZRLocation->setupTimeSeries(3.2, intFreq, 0);
 		tsZRLocation->addDataSource("", 255, 0, 0);
 		if (0 == tsZRVelocity){
 			tsZRVelocity = new TimeSeriesCanvas
@@ -1627,7 +1632,7 @@ void BlockDemo::reinit(){
 	resetClocks();
 	useCcd = false;
 	limitIfNeeded = true;
-	ammoVelocity = 30;
+	ammoVelocity = 300;
 	ammoCcdMotionThreshold=minDist/2;
 	ammoCcdSweptSphereRadius = minDist / 10;
 	loadCcdMotionThreshold = 1e-3;

### 2016-10-24 examples/CharpyDemo/BlockDemo.cpp
more stats in showMessage of BlockDemo

@@ -1115,10 +1115,38 @@ public:
 			infoMsg(buf);
 		}
 		if (m_constraint != 0 && m_constraint->isEnabled()){
-			sprintf_s(buf, B_LEN * 2, "FX=%8.3g FY=%8.3g MZ=%8.3g",
+			sprintf_s(buf, B_LEN * 2, "B: FX=%8.3g FY=%8.3g MZ=%8.3g",
 				jf.m_appliedForceBodyB.getX(),
 				jf.m_appliedForceBodyB.getY(),
 				jf.m_appliedTorqueBodyB.getZ());
+			infoMsg(buf);
+			sprintf_s(buf, B_LEN * 2, "A: FX=%8.3g FY=%8.3g MZ=%8.3g",
+				jf.m_appliedForceBodyA.getX(),
+				jf.m_appliedForceBodyA.getY(),
+				jf.m_appliedTorqueBodyA.getZ());
+			infoMsg(buf);
+		}
+		if (m_body != 0){
+			btVector3 loc = m_body->getCenterOfMassPosition();
+			sprintf_s(buf, B_LEN * 2, "x=%8.3g y=%8.3g z=%8.3g",
+				loc.x(),
+				loc.y(),
+				loc.z());
+			infoMsg(buf);
+			btTransform tr = m_body->getCenterOfMassTransform();
+			btQuaternion q = tr.getRotation();
+			sprintf_s(buf, B_LEN * 2, "q={%8.3g, %8.3g, %8.3g, %8.3g}",
+				q.x(),
+				q.y(),
+				q.z(),
+				q.w());
+			infoMsg(buf);
+			btScalar yaw, pitch, roll;
+			tr.getBasis().getEulerYPR(yaw, pitch, roll);
+			sprintf_s(buf, B_LEN * 2, "yaw=%8.3g pitch=%8.3g roll=%8.3g",
+				yaw,
+				pitch,
+				roll);
 			infoMsg(buf);
 		}
 		PlasticityData::setData(&pData);
@@ -1609,7 +1637,7 @@ void BlockDemo::reinit(){
 	gravityRampUpTime = 0;
 	switch (orientationType){
 	case X:
-		lsx = 5;
+		lsx = 6;
 		lsy = 1;
 		blockSteelScale = 0.004;
 		break;

### 2016-10-28 pdocs/thesis/article-introduction.tex
working on Peter's comments

@@
 
-The application of modern computer game techniques enables the description of complex dynamic systems 
-such as military vehicles with a high level of detail while still solving the equations of motion in real-time.
-Film production and war games, in particular, is a key area that have benefited from rapidly enhanced simulation technology. 
-In practice, games are often accomplished using an open-source platform such 
+Computational techniques used in modern computer games are able to describe
+complex dynamic systems such as those found in military and other motorized vehicles
+with a high level of accuracy while still being able to solve the equations of motion in real time.
+Film production and the computer game genre of war games
+are two areas that have greatly benefited from rapidly improving simulation technology. 
+In practice, software for such games is often written using open-source physics engine platforms  
 like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
 
-Computational methods used in physics engines are divided to modules that handle collision detection and 
-contact description and modules that handle solution of equations of motion in real-time. Equations of motion need to be 
-solved can further be subdivided to be associated to motion, constraints and collisions. 
+Computational methods used in physics engines are divided into modules that handle collision detection and 
+contact description and modules that handle solution of equations of motion in real time. 
+The equations of motion that need to be 
+solved can be subdivided into equations related to motion, constraints and collisions. 
 Velocity-based formulation is typically used in constraint based rigid body simulation, \cite{erleben.thesis}. 
@@
 
-Plasticity is not typically taken into account in gaming solutions. 
-Breaking of various bodies often takes place based on collision or impulse.
+Plasticity is not typically taken into account in gaming solutions and 
+destruction of various bodies often takes place based on collision or impulse exceeding predefined limit.
 Nevertheless, breaking of steel or reinforced concrete structures using this approach 
-is not appropriate making a simulation to look realistic. Theory for handling of plasticity 
+is not appropriate if the simulation is to look realistic. Theory for handling of plasticity 
 has been presented already in \cite{cg1988}. \cite{muller2005meshless} 
-present a method for modeling and animating of elastic and plastic bodies in real-time using 
-point based animation. This approach is not been widely used in computer games.  
-On major issue is collision handling of deformable bodies.
+present a method for modeling and animating of elastic and plastic bodies in real time using 
+point based animation. This approach has not been widely used in computer games.  
+One major issue is collision handling of deformable bodies.
 
-This study will introduce an approach to account plastic deformation in game applications.   
-In the introduced method, the plastic deformation takes place if force or moment exceeds predefined 
+This study will introduce an approach to account for plastic deformation in game applications.   
+In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
 limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
 The approach is based on using joint motors to model plasticity. 
-This study extends a method introduced by
+The study extends a method introduced by
 \citet{erleben.thesis} 
 %\citet[p.~90]{erleben.thesis} 
-which is originally proposed for modelling friction in joints. 
+which was originally proposed for modelling friction in joints. 
 In the introduced method adjacent bodies are connected by motors. 
-Motor power production limits are estimated based on plastic section modulus. 
-Joint breaking is accounted by summing plastic deformation and comparing it to 
-predefined material based limit. Elastic part of deformation is modelled by employing 
-spring description which is based on modification of existing constraint in \bullet.
+Motor power production limits are estimated based on the plastic section modulus. 
+Joint breaking is based on summing plastic deformation and comparing it to a
+predefined material based limit. The elastic part of deformation is modelled by employing 
+a spring based on modification of an existing constraint in \bullet.
 
-Approach presented in this work can be used in gaming industry to provide more realistic 
-simulations without significant extra work. For gaming purposes presented method works 
-best in scenarios where connected parts are heavy. This allows normal 
+The approach presented in this work can be used in the gaming industry to provide more realistic 
+simulations without significant extra work. For gaming purposes, the presented method works 
+best in scenarios where the connected parts are heavy. This allows a normal 
 integration timestep to be used without stability issues. 
-This kind of metodology also opens large area of combining old structural analysis
-methods to modern simulation frameworks.
+This methodological approach enbles established structural analysis
+methods to be combined with modern simulation frameworks.

### 2016-10-28 pdocs/thesis/article-section-2.tex
working on Peter's comments

@@
 % !TeX root = article.tex
-\section{Description of plasticity in the framework of physics engine}
+\section{Description of plasticity in the framework of physics engines}
 
-In this section, key concepts related to the introduced model are explained. Main differences between 
-traditional structural analysis and physics engines are reviewed and discussed.
+In this section, key concepts related to the introduced model are explained. The main differences between 
+traditional structural analysis and physics engines based approaches are reviewed and discussed.
 
-Velocity-based formulation is  popular within physics based game
+Velocity-based formulation is  commonly used by physics based game
 developers and film production teams.
@@
 \citet{erleben.thesis} 
-provides reasoning and theoretical details on why 
-velocity-based formulation is  popular in constraint-based rigid body simulation. 
-Main reason is that collision handling can be done without additional procedures.
+provides reasoning and theoretical details for the popularity of 
+velocity-based formulation in constraint-based rigid body simulation. 
+The main reason is that collision handling can be done without the use of additional procedures.
 
-Background
-for velocity based formulation shown here is based on  
-\citet{erleben.thesis}.
+Work presented by  
+\citet{erleben.thesis} provides the basis for the velocity-based formulation discussed in this work.
 %\citet[p.~45-50]{erleben.thesis}.
@@
 \begin{equation} \label{eq:impulseIntegral}
-\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt
+\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt,
 \end{equation}

### 2016-10-29 pdocs/thesis/article-section-2.tex
working on Peter's comments

@@
 
-Velocity-based formulation is  commonly used by physics based game
+Velocity-based formulation of constraint based rigid body simulation 
+is  commonly used by physics based game
 developers and film production teams.
@@
 provides reasoning and theoretical details for the popularity of 
-velocity-based formulation in constraint-based rigid body simulation. 
+velocity-based formulation in constraint-based rigid body simulation instead of accelaration based. 
 The main reason is that collision handling can be done without the use of additional procedures.
@@
 % pdf page 64
-In following section, these formulations will be clarified by simple example using \bullet\ implementation.
+In the following section, these formulations will be clarified by a simple example using \bullet\ implementation.
 
 Impulse $\vec{J}$
-in the time interval $\Delta t $ can we written as
+in the time interval $\Delta t $ can be written as
 \begin{equation} \label{eq:impulseIntegral}
@@
 
-Newton's second law of motion $\vec{F}=m\vec{a}$ one can solve for the velocity,
-$\vec{v}^{\Delta t}$ as
+Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
+$\vec{v}^{\Delta t}$ can be solved as
 \begin{equation} \label{eq:impulseIntegraWithNewton}
@@
 \begin{equation} \label{eq:impulse}
-m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J}
+m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J},
 \end{equation}
@@
 Updates after each step can be summarized  for locations and  
-for velocities correspondingly as follows.
+for velocities respectively as follows:
 
@@
 \begin{equation} \label{eq:eomV}
-\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
+\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) .
 \end{equation}
 
-Symbols used in Equations \ref{eq:eomL} and \ref{eq:eomV}
-are summarized in Table \ref{tab:eom} and Figure \ref{fig:eom-contact}.
-Figure \ref{fig:eom-contact} describes collision of two bodies $B_1$ and $B_2$
+The symbols used in Equations \ref{eq:eomL} and \ref{eq:eomV}
+are summarized in Table \ref{tab:eom}.
+Figure \ref{fig:eom-contact} describes the collision of two bodies, $B_1$ and $B_2$
+where $\vec{r}_i$ is orientation of body $i$,
+$\vec{p}_k $ is orientation of contact point $k$,
+$\vec{r}_{ik} $ is a vector  between the center of gravity of body $i$ and contact point $k$, 
+and
+$\vec{n}_{k}$ is the contact normal for contact point $k$.
 
-\begin{figure}[htb!]
+\begin{figure}[tb!]
 \centering
@@
 \end{tikzpicture}
-\caption{Illustration of nomenclature for equations of motion for contact.}
+\caption{Illustration of nomenclature for equations of motion for collision.}
 \label{fig:eom-contact}
 \end{figure}
-
-where $\vec{r}_i$ is orientation of body $i$,
-$\vec{p}_k $ is orientation of contact point $k$,
-$\vec{r}_{ik} $ is vector  between center of gravity of body $i$ and contact point $k$
-and
-$\vec{n}_{k}$ is contact normal for contact point $k$.
-By employing this formulation, simplified scenario can be simulated.
 
@@
 
-Friction in contacts and joint constraints can be handled in unified way by refactoring
+Friction in contacts and joint constraints can be handled in a unified way by refactoring
 equation \ref{eq:eomV} as,  
@@
 + J_{joint}^T \vec{\lambda}_{joint}
-+ \vec{f}_{ext})
++ \vec{f}_{ext}),
 \end{equation}
-
-where jacobian terms $J_{contact}^T$ for joints are 
-derived by taking time derivates of kinematic constraints.
+where Jacobian terms $J_{contact}^T$ for joints are 
+derived by taking time derivatives of the kinematic constraints.
 
@@
 \ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
-Figure \ref{fig:eom-joint} shows terms needed for joint processing.
+Figure \ref{fig:eom-joint} presents terms needed for joint processing.
 
@@
 Derivation for terms in Equation \ref{eq:constraintEquation}
-can be done using position and orientation of connected bodies.
-E.g. for ball joint formulation is based on both joint points having same position.
-In contact cases formulation is easier if it is done using velocities, \cite{ode.joints}.
+can be done using the position and orientation of connected bodies
+e.g. for ball joint formulation is based on both joint points having the same position.
+In contact cases, formulation is easier if it is done using velocities, \cite{ode.joints}.
 
@@
 
-In following section, these equations will be clarified by simple example.
-Main parameters  and corresponding fields in \bullet\  
- are described in Table \ref{tab:constraintParameters}.
+In the following section, these equations will be explained by a simple example.
+The main parameters  and corresponding fields in \bullet\  
+are given in Table \ref{tab:constraintParameters}.
 
@@
 {\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
+$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\
 $J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
@@
 Often,  finite element method is used.
-In most cases, static solution with assumption of linear strain-displacement relation
+In most cases, a static solution with an assumption of linear strain-displacement relation
 using displacement based boundary conditions is used.
-\citet{bathe-1975} provides description for handling of various nonlinearities.
+\citet{bathe-1975} provides a description for handling of various nonlinearities.
 In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
@@
 
-Material plasticity can be accounted in games by using suitable coefficient of restitution.
-This provides reasonable means to simulate loss of energy in collisions.
-Simulation of breaking of bodies made of ductile material can be made more realistic by splitting rigid body
-to multiple bodies which are connected by energy absorbing joints.
-Typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.
+Material plasticity can be accounted for in simulations by using a suitable coefficient of restitution.
+This provides a reasonable means to simulate loss of energy in collisions.
+In this work simulation of breaking of bodies made of ductile material is made more realistic 
+by splitting the rigid body
+to multiple bodies that are connected by energy absorbing joints.
+A typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.
 
@@
 %\citet[p.~108]{dowling}.
-Stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 0.001 to 0.005.
+The stress-strain curve is not drawn to scale as elastic strain could not be seen as it is typically 
+0.001 to 0.005 and fracture strain can be 100 times larger.
 
-In this work elastic-fully plastic material model is used in most scenarios.
+In this work, an elastic-fully plastic material model is used in most scenarios.
 Having elastic part allows elastic displacements for slender structures. 
-Elastic material behavior is ignored in approach introduced in this work provided
-that deformation is related to higher frequency
+Elastic material behavior is ignored in approach introduced in this work if
+the deformation is related to ahigher frequency
 than integration stability would allow.
 It should be noted that geometry
-of bodies is not updated during analysis and thus engineering stress-strain properties should
-be used.
+of bodies is not updated during analysis and thus engineering stress-strain properties are used.
 
-In this work, strain hardening is taken into account by assuming that plasticity in bending
+In this work, strain hardening is taken into account by assuming that plastic volume in bending
 expands, 
@@
 Material that starts to yield first is hardened and as a result of which yielding moves.
-This can be seen e.g. by bending paperclip. It does not break at low angles but can take few full bends.
 
-Difference between elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.
+The difference between the elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.
 
@@
 \end{tikzpicture}
-\caption{Axial stress distribution over cross section for bending under elastic and fully plastic loads.}
+\caption{Axial stress distribution over a cross section for bending under elastic and fully plastic loads.}
 \label{fig:wp}
@@
 
-As shown in Figure 2.4, if stress is below yield limit $f_y$, stress and strain are linear within material.
-If cross section is fully plastic, stress is assumed to be at yield level over whole cross section such that 
-plastic section modulus is higher than elastic section modulus.
+As shown in Figure 2.4, if stress is below yield limit $f_y$, stress and strain are linear within the material.
+If cross section is fully plastic, stress is assumed to be at yield level over the whole cross section such that 
+the plastic section modulus is higher than the elastic section modulus.
 
@@
 in Equations \ref{eq:lambdaLow} and  
-\ref{eq:lambdaHigh} using plastic capasities which are defined below.
+\ref{eq:lambdaHigh} using plastic capacities, which are defined below.
 
@@
 Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
-are product of area and shear yield stress $\tau_y$ as follows
+are a product of area and shear yield stress $\tau_y$ as follows
 \begin{equation} \label{eq:fQ}
@@
 
-Maximum moments acting around axis perpendicular to $\vec{r}_{anc}^{\,i} $
-are integrals of perpdendicular distance 
-and yield stress $f_y$ as given for the moment around $x$-axis 
-and moment around $z$-axis, respectively.
+Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
+are integrals of the perpdendicular distance 
+and yield stress $f_y$ as given for the moment around the $x$-axis 
+and moment around the $z$-axis, respectively.
 
@@
 Maximum moment around $\vec{r}_{anc}^{\,i} $
-is integral of distance $d$ from joint point
+is an integral of distance $d$ from the joint point
 and shear yield stress $\tau_y$ as
@@
 
-Maximum forces and moments for 
+Maximum forces and moments for a
 rectangular section with width $b$ and height $h$ using constant yield stress
-are summarized in Table \ref{tab:maxForces}.
-Yield shear stress is assumed to be $ 0.5\, f_y$ using Tresca yield critetion.
-If von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$), \cite{dowling}.
+are given in Table \ref{tab:maxForces}.
+Yield shear stress is assumed to be $ 0.5\, f_y$ using the Tresca yield critetion.
+If the von Mises yield criterion is used 0.5 is replaced by 0.58 ($1/\sqrt{3}$), \cite{dowling}.
 % p. 262, p. 268
-These are not exact values in multiaxial stress state but they
-are usable in most cases.
+These are not exact values in a multiaxial stress state but they
+should be acceptable in most gaming scenarios.
 
@@
 
-For torque there is closed form solution only for
+For torque there is a closed form solution only for
 circular cross sections.
@@
 Better approximation for any given $b$ and $h$ can be obtained 
-by integrating distance from center of joint over cross section and
-multiplying it with yield shear stress e.g. using octave, \cite{octave}.
-Example of calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $
+by integrating distance from the center of the joint over cross section and
+multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
+An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
 is shown in Figure \ref{fig:octave-mp}
@@
 
-\caption{Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using octave.}
+\caption{Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}
 \label{fig:octave-mp}
@@
 
-Basic idea introduced in this study can be tested with any framework having motors and hinge constraints.
-This can be done by setting target velocity of motor to zero and limiting maximum motor impulse to plastic moment 
-multiplied by timestep.
+The basic idea introduced in this study can be tested with any framework having motors and hinge constraints.
+This can be done by setting the target velocity of the motor to zero and limiting 
+the maximum motor impulse to plastic moment multiplied by a timestep.
 

### 2016-10-29 pdocs/thesis/article-section-3.tex
working on Peter's comments

@@
 \subsection{Block under tension}
-In this section, changes to constraint formulation are clarified by doing few simulation steps
+In this section, changes to the constraint formulation are exemplified with some simulation steps
 of non constrained, rigid and
@@
 \end{tikzpicture}
-\caption{Single body model for plasticy processing demonstration.}
+\caption{Single body model for plasticity processing.}
 \label{fig:tensionModel}
@@
 
-In the first example, system under investigation
-has only one dynamic rigid body which is three meters high block of 0.04 \% steel enforced concrete. 
-Cross section is one square meter.
-Constraint is set so that connecting frame is at top of block. 
-In this case frame could be anywhere but if multiple bodies are involved, connecting frame must be defined
-so that it reflects scenario under investigation.
+In the first example, the system under investigation
+has only one dynamic rigid body which is a three meters high block of 0.04 \% steel enforced concrete. 
+The cross section is one square meter.
+A constraint is set so that the connecting frame is at the top of the block. 
+In this case, the frame could be anywhere but if multiple bodies are involved, the connecting frame must be defined
+so that it reflects the scenario under investigation.
 Concrete density is $2000\ \frac{kg}{m^3}$ and steel density is $7800\ \frac{kg}{m^3}$. 
-Steel is assumed to handle load in elastic-plastic case. Yield stress of steel is 200 MPa.
+Steel is assumed to handle load in the elastic-plastic case. The yield stress of steel is 200 MPa.
 Gravity is $10\ \frac{m}{s^2}$ to keep numbers simple. 
 Simulation step is $\frac{1}{60}\ s$ and 10 
-iterations are done for single step.
-Body is 1.5 meters above rigid ground. In unconstrained case body will hit hit ground at $t$ = 0.548 s 
+iterations are done for a single step.
+The body is 1.5 meters above rigid ground. In the unconstrained case the body will hit hit the ground at $t$ = 0.548 s 
 ($\sqrt{2 \cdot 1.5 / 10}$).
 
-Single simulation step is performed using following substeps. 
-In this work, changes are made only to constraint setup and update actions.
+A single simulation step is performed using the following substeps. 
+In this work, changes are made only to the constraint setup and the update actions.
 \begin{enumerate}
@@
 This step makes programming easier as otherwise each program should add forces due to gravity in each step. 
-In this case, $\vec{f}_{ext}$ in  Equation \ref{eq:eomV2} gets added by $\lbrace{0,-60000,0}\rbrace$.
-\item Predict unconstrained motion for each body controlled by physics simulation.
+In this case, $\vec{f}_{ext}$ in  Equation \ref{eq:eomV2} is added by $\lbrace{0,-60000,0}\rbrace$.
+\item Predict unconstrained motion for each body controlled by the physics simulation.
 Static and kinematic bodies are not processed in this step.
@@
 \item Predict contacts. In this phase, continuous collision detection is done based on simplified bodies. 
-Each rigid body is represented by sphere geometry and contact prediction is done if body moves more than given threshold value during simulation step.
+Each rigid body is represented by sphere geometry and contact prediction is done if 
+the body moves more than a given threshold value during the simulation step. 
 Continuous collision detection is configured for each body. 
-Typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
+A typical scenario that requires continuous collision detection is fast moving bodies that would otherwise go through walls.
 \item Perform discrete collision detection. All overlapping bodies are processed and manifoldPoints are created for
-each detected contact at end of simulation step. 
-In unconstrained case, contact is detected after $t$=0.55 s if continuous collision detection is not used and manifolds distance gets negative value (-0.058 m).
-\item Calculate simulation islands(groups). Bodies that are near each other based on contact prediction or discrete collision detection
- or connected with constraints are grouped in same group.
+each detected contact at the end of the simulation step. 
+In the unconstrained case, contact is detected after $t$=0.55 s if 
+continuous collision detection is not used and manifold distance gets a negative value (-0.058 m).
+\item Calculate simulation islands(groups). Bodies that are near each other based 
+on contact prediction or discrete collision detection
+ or connected with constraints are grouped in the same group.
 \item Solve constraints. Both contact and other constraints are processed in this step. 
  This step is subdivided to setup, iterations and finish phases. 
-In setup phase constraints are queried for positional errors for calculation of $c$ in
+In teh setup phase constraints are queried for positional errors for calculation of $c$ in
 Equation \ref{eq:constraintEquation} and maximum and minimum impulses in Equations \ref{eq:lambdaLow} and 
 \ref{eq:lambdaHigh}.
-In finish phase constraint forces are calculated if requested and velocities of bodies are updated.
-\item Integrate transforms using velocities modified in previous step.
+In the finish phase constraint forces are calculated if requested and velocities of bodies are updated.
+\item Integrate transforms using velocities modified in the previous step.
 \item Update actions (callbacks) are called. 
-In elastic-plastic case, plasticity is summed and equilibrium point of elastic part is updated if maximum force or moment is exceeded.
+In the elastic-plastic case, plasticity is summed and the equilibrium
+point of the elastic part is updated if the maximum force or moment is exceeded.
 \item Activation state of bodies is updated. 
-To avoid extra calculation bodies are as default put to sleeping state if linear and angular velocities of body are less than
-given threshold values (default 0.8 and 1.0) longer than set time limit (2.0).
+To avoid extra calculation bodies are as default put into sleeping state if linear and angular velocities of
+the body are less than
+given threshold values (default 0.8 and 1.0) or longer than the set time limit (2.0).
 \end{enumerate}
@@
 \begin{itemize}
-\item No rotation takes place. $\omega_1$ and $\omega_2$ are zeros.
+\item No rotation takes place. $\omega_1$ and $\omega_2$ are zero.
 \item Constraint force mixing can be ignored.
 \item Only vertical velocity is handled.
-\item Other involved body is rigid and it does not move.
+\item The other involved body is rigid and it does not move.
 \end{itemize}
@@
 
-Equations \ref{eq:lambdaLow} and \ref{eq:lambdaHigh} are not active for fixed case as
-there are no upper nor lower limit for forces.
-For elastic-plastic case maximum impulse is set to product of yield stress, 
+Equations \ref{eq:lambdaLow} and \ref{eq:lambdaHigh} are not active for fixed cases as
+there are no upper or lower limit for forces.
+For the elastic-plastic case maximum impulse is set to the product of the yield stress, 
 area of steel enforcement and time step (1330).
@@
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ is used to pick up values for internal variables described in Figure \ref{fig:solVars}.
+in \bullet\ is used to get values for the internal variables described in Figure \ref{fig:solVars}.
 
@@
 
-Actual values for unconstrained case without continuous collision detection
+Actual values for an unconstrained case without continuous collision detection
 are shown in Table \ref{tab:freeBlockValues} and 
@@
 {\it posError} is 2.8 (0.058*0.8/0.0167). Impulse is 34000 and contact force is thus about 2 MN (34000/0.0167).
-After few steps location and position stabilize although internally calculation is needed for each time step
-until body is deactivated.
+After a few steps the location and position stabilize although internally calculation is needed for each time step
+until a body is deactivated.
 
 \begin {table}[htb!]
-\caption {Simulation values for unconstrained case without continuous collision detection. 
-For internal contact values typical values are shown
-as number of contacts and values at contact points are different.} 
+\caption {Simulation values for the unconstrained case without continuous collision detection. 
+Typical values are shown for internal contact values
+as the number of contacts and the values at contact points differ.} 
 \label{tab:freeBlockValues} 

### 2016-10-30 pdocs/thesis/article-conclusion.tex
working on Peter's comments

--- 2016-09-19 pdocs/thesis/article-conclusion.tex
@@
 
-This study has presented an approach to account plastic deformation in 
-velocity based formulation.
-In the introduced method, the plastic deformation takes place if force or moment exceeds given 
-limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
-Approach presented in this work can be used in gaming industry to provide more realistic 
-simulations without significant extra work. For gaming purposes presented method works 
-best in scenarios where connected parts are relatively heavy. This allows normal 
+This study presents an approach to account for plastic deformation in 
+a velocity based formulation.
+In the introduced method, the plastic deformation takes place if the force or moment exceeds the given 
+limit, the deformation absorbs energy and the joint breaks if plastic capacity is exceeded. 
+The approach presented in this work can be used in the gaming industry to provide realistic 
+simulations. 
+For gaming purposes the presented method works 
+best in scenarios where the connected parts are heavy. This allows a normal 
 integration timestep to be used without stability issues.
-This kind of metodology also opens large area of combining old structural analysis
-methods to modern simulation frameworks.
+
+This kind of methodology also has considerable potential for the combination of
+the structural analysis methods with real-time simulation frameworks.
+Handling of significant axial and shear forces and moments at the same time
+is one possible area of further studies.
+
+There is also work to be done in the area of performance optimization.
+One possible area is the reuse of once calculated values if memory capacity is not limited.
+This could done, e.g., when dealing with high frequency modes.
+Another area is usage of the graphics processing unit (GPU) for calculation.
+\bullet\ already has an experimental GPU pipeline but most constraint types are not 
+yet supported. GPU support was one reason for selecting
+\bullet\ for use in this study
+
+Integration to 3D modelling and animation software products 
+should also be solved before we can expect to
+see realistic plasticity in main stream games. 
+\bullet\ is already integrated into several ones like Blender 
+and it was another significant reason for selecting
+\bullet\  to be used in this study.
+
 As already noted in \cite{cg1988}, the modeling of inelastic deformation

### 2016-10-30 pdocs/thesis/article-section-3.tex
working on Peter's comments

@@
 \begin{description}
-\item[velError] is calculated using velocities and external impulses of connected bodies.  
+\item[velError] is calculated using velocities and external impulses of the connected bodies.  
  \begin{description}
-\item[In constraint cases,] main contibutor is bodies relative speed at joint point.
-\item[In contact case,] main contributor is bodies relative speed at point of contact. 
+\item[In constraint cases,] the main contibutor is the relative speed of the bodies at the joint point.
+\item[In contact cases,] the main contributor is the relative speed of the bodies at the point of contact. 
 Bodies are not allowed to penetrate each other. Restitution increases velError.
@@
 \end{description}
-\item[posError] is calculated by constraint. It is significant factor in designing stable constraints.
+\item[posError] is provided by the constraint. It is a significant factor in designing stable constraints.
  \begin{description}
- \item[In fixed case,] value is about 12 times actual position error. Factor 12 is based on time step (60) 
- and default value of error reduction parameter (erp) which has value of 0.2 in this context.
- \item[In elastic plastic case,]  value is set to zero if impulse would be larger than maximum impulse or
-spring simulation cannot be done in stable way.
- \item[In contact case,] value is zero if there is no penetration. For penetration cases it is 
-$\frac{-penetration\, erp}{timeStep}$. In contact cases default value for erp is 0.8.
+ \item[In fixed cases,] the value is about 12 times the actual position error. Factor 12 is based on the time step (60) 
+ and the default value of error reduction parameter (erp), which has a value of 0.2 in this context.
+ \item[In elastic plastic cases,]  the value is set to zero if the impulse would be larger the than maximum impulse or
+the spring simulation cannot be done in a stable way.
+ \item[In contact cases,] the value is zero if there is no penetration. For penetration cases it is 
+$\frac{-penetration\, erp}{timeStep}$. In contact cases the default value for erp is 0.8.
  \end{description}
@@
  \begin{description}
-\item[In constraint cases,] it is mass of body (6000).
-\item[In contact case,] it varies below mass of body.
+\item[In constraint cases,] it is the mass of the body (6000).
+\item[In contact cases,] it varies below the mass of the body.
 \end{description}
-\item[Impulse] is impulse applied to body during timestep.
+\item[Impulse] is impulse applied to the body during a timestep.
  \begin{description}
-\item[In constraint cases,] it is obtained from btJointFeedback structure.
-\item[In contact case,] it is obtained by summing applied impulses from active manifolds.
+\item[In constraint cases,] it is obtained from the btJointFeedback structure.
+\item[In contact cases,] it is obtained by summing the applied impulses from the active manifolds.
 \end{description}
 \item[erp] Error reduction parameter (0...1) is used to handle numerical issues e.q. body drifting. 
-Setting erp to 1 would in theory eliminate error in one step but in practice value of 0.2 - 0.8 is used in most cases.
+Setting erp to 1 would in theory eliminate error in one step but in practice a value of 0.2 - 0.8 is used in most cases.
 \end{description}
@@
 \input{figs/uc-fig.tikz}
-\caption{Graph for selected values for unconstrained case simulation without continuous collision detection.}
+\caption{Graph for selected values for the unconstrained case without continuous collision detection.}
 \label{fig:uc}
@@
 
-Actual values for unconstrained case with continuous collision detection (ccd) using 1.5 
-as radius of ccd sphere and 0.001 as ccd motion threshold
-are shown in \ref{tab:freeBlockValuesWithCcd}. Collision is detected at time 0.550 s when
+Actual values for an unconstrained case with continuous collision detection (ccd) using 1.5 
+as the radius of the ccd sphere and 0.001 as the ccd motion threshold
+are shown in Table \ref{tab:freeBlockValuesWithCcd}. Collision is detected at time 0.550 s when
 {\it velError} is  3.5 (5.34+0.167)-0.033/0.0167 and
 {\it posError} is  0 as collision is detected before penetration. 
-It should be noted that in general ccd sphere should not extend actual body as 
-premature contacts are created if collision takes place in those regions.
+It should be noted that in general the ccd sphere should not extend the actual body as 
+premature contacts are created if collision takes place in regions where the ccd sphere extends out of the actual body.
 
 \begin {table}[htb!]
-\caption {Simulation values for unconstrained case with continuous collision detection. } 
+\caption {Simulation values for the unconstrained case with continuous collision detection. } 
 \label{tab:freeBlockValuesWithCcd} 
@@
 
-Values for fixed constraint are shown in Table
-\ref{tab:fixedBlockValues}. Constraint is activated in second step and positional error is corrected
-about 20 \% in each step as requested by using erp value 0.2.
+Values for a fixed constraint are shown in Table
+\ref{tab:fixedBlockValues}. The constraint is activated in the second step and positional error is corrected
+about 20 \% in each step as requested by using an erp value of 0.2.
 
 \begin {table}[htb!]
-\caption {Constraint parameter values for fixed constraint} 
+\caption {Constraint parameter values for the fixed constraint} 
 \label{tab:fixedBlockValues} 
@@
 
-Table  \ref{tab:ep2Parameters} summarizes most significant parameters for this study.
-There are also e.g. parameters for controlling joint motors.
-Additional equation is created for each additional constraint. 
+Table  \ref{tab:ep2Parameters} the summarizes most significant parameters for this study.
+There are also other parameters, e.g. for controlling joint motors.
+An additional equation is created for each additional constraint. 
 Enabled springs and motors add one row and
-limits add one if upper and lower are same and two if both upper and lower limits are defined.
+limits add one if the upper and lower limits are the same and two if both the 
+upper and the lower limits are defined.
 
@@
 
-Values for elastic-plastic case are shown in Tables \ref{tab:epBlockValues}.
-Body drops freely during first simulation step and 
-gains enough kinetic energy so that higher impulses are needed in few following steps.
-This causes plastic strain during next three steps.
+Values for an elastic-plastic case are shown in Table \ref{tab:epBlockValues}.
+The body drops freely during first simulation step and 
+gains enough kinetic energy so that higher impulses are needed in the following steps.
+This causes plastic strain during the next three steps.
 
@@
 
-Six-dof-spring constraint 2 has optional feature to avoid unstability by automatically softening constraint
-spring. Feature is activated if current timeStep is overly large for spring-mass system simulation.
-The feature is trigged based on the equation below 
+Six-dof-spring constraint 2 has an optional feature to avoid unstability by automatically softening the constraint
+spring. The feature is activated if the current timeStep is overly large for the spring-mass system simulation.
+The feature is triggered based on the equation below 
 \begin{equation} \label{eq:frequencyLimited}
-\sqrt{ k /  m_{min}} dt > 0.25 
+\sqrt{ k /  m_{min}} dt > 0.25, 
 \end{equation}
-where $k$ is elastic spring stiffness between bodies,
-$m_{min}$ is smaller of connected masses or inertias and $dt$ is integration timestep.
+where $k$ is the elastic spring stiffness between the bodies,
+$m_{min}$ is the smaller of the connected masses or inertias and $dt$ is the integration timestep.
 
-If this feature is active for elastic-plastic constraint 2 constraint  
-it does not report positional error but it allows full plastic capacity to be used for correcting
-velocity based error instead of force provided by spring.
-In this scenario body behaves as depicted in Table \ref{tab:ep2BlockValues} i.e. it does not move at all.
+If this feature is active for elastic-plastic constraint 2   
+it does not report positional error but  allows full plastic capacity to be used for correcting
+the velocity based error instead of the force provided by the spring.
+In this scenario the body behaves as depicted in Table \ref{tab:ep2BlockValues}, i.e., it does not move at all.
 
@@
 \subsection{Charpy impact test}
-In a second numerical example, simulation of Charpy impact test is
-used to benchmark approach in more complicated scenario. 
-Material is steel and density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
-Specimen dimensions are 10x10x55 mm with 2 mm notch in middle which is taken into account in calculations.
+In a second numerical example, simulation of a Charpy impact test is
+used to benchmark the approach in a more complicated scenario. 
+The material is steel and density is 7800 $\frac{kg}{m^{3}}$. Young’s modulus is 200 GPa.
+Specimen dimensions are 10x10x55 mm with a 2 mm notch in the middle which is taken into account in the calculations.
 Support anvils initially have 40 mm open space between them and their width is 40 mm. 
-Hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m and mass is 19.5 kg.
-Hammer has 40 mm draft.
-If specimen bends about 1.9 radians (108 degrees) it will go between anvils.
-Expected energy loss is product of plastic moment of section, hinge angle needed for specimen to go through supports and 
-yield stress of specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa expected energy 
+The hammer is 0.5 m wide and 0.25 m high, thickness is 0.02 m and mass is 19.5 kg.
+The hammer has 40 mm draft.
+If the specimen bends about 1.9 radians (108 degrees) it can go between the anvils.
+The expected energy loss is the product of the plastic moment of section, 
+hinge angle needed for the specimen to go through the supports and 
+the yield stress of the specimen. For hinge angle of 1.9 radians and yield stress of 400 MPa, expected energy 
 loss is 122 J.
 
-Usually \bullet\ simulations are done using fixed time step of 1/60 s i.e. 16.67 ms. 
+\bullet\ simulations are usually done using a fixed time step of 1/60 s, i.e., 16.67 ms. 
 For this case that is too large. 
-Default time step was selected to be 5 ms outside impact time and 0.1 ms during impact. 
-Automatic time stepping routine changes timestep so that at angles higher than 0.2 5 ms time step is
-used and adjusts it linearly to selected time step between angles of 0.2 and 0.05.
+The default time step was selected to be 5 ms outside impact  and 0.1 ms during impact. 
+An automatic time stepping routine changes the timestep so that at angles higher than 0.2 radians 5 ms time step is
+used and adjusts it linearly to a selected time step between the angles of 0.2 and 0.05 radians.
 
-Figure \ref{fig:charpy-series} shows three detailed pictures from simulation.
-Left picture shows early phase of simulation. In middle picture hammer has not yet collided with specimen.
-In right picture specimen is bending.
+Figure \ref{fig:charpy-series} shows three detailed pictures from the simulation.
+The left picture shows early phase of simulation. 
+In the middle picture the hammer has not yet collided with the specimen.
+In the right picture the specimen is bending.
 
@@
 \includegraphics[height=3cm]{figs/article-charpy-3}
-\caption{Detailed pictures from simulation of Charpy impact test}
+\caption{Detailed pictures from simulation of a Charpy impact test}
 \label{fig:charpy-series}
@@
 
-Table \ref{tab:ep2ts} shows energy loss for few timesteps as example to demonstrate sensitivity of solution
+Table \ref{tab:ep2ts} shows energy loss for a few timesteps to demonstrate the sensitivity of the solution
 using  elastic-plastic constraint 2.
@@
 
-Figure \ref{fig:charpy} shows overview from simulation as specimen is going between supports.
+Figure \ref{fig:charpy} shows an overview from simulation as the specimen passes between the supports.
 
@@
 \subsection{Demolisher}
-Third numerical examples introduces a demolisher.
-Demolisher is collection of possible scenarios that can be added to games to provide ductile joints between
+The third numerical example introduces the demolisher.
+The demolisher is a collection of possible scenarios that can be added to games to provide ductile joints between
 rigid bodies.  Rigid concrete bodies are connected by steel enforcement.
@@
 Gravity is 10  $\frac{m}{s^2}$. Simulation step is $\frac{1}{60} $ s and 10 
-iterations are done for single step. Multiple simulation steps are done for each render if needed.
+iterations are done for a single step. Multiple simulation steps are done for each render if needed.
 
-Simulation can be done in real time on modest hardware, e.g. 
+The simulation can be done in real time on modest hardware, e.g. 
 on laptop having 2.0 GHz Intel i3 CPU with integrated graphics. 
-Figure \ref{fig:demolisher} shows three screenshots from simulation.
+Figure \ref{fig:demolisher} shows three screenshots from the simulation.
 
@@
 
-Vehicle is modelled as single box btRaycastVehicle with four driving wheels. Wheels are rendered for visual feedback.
-In left picture, initial state is shown. Gate in front is modelled to be so slender that it has visible elastic deflection.
-Gate support is modelled as heavy box.
-Bridge has rigid ramps and supports at both ends.
-Gravity is applied gradually to avoid collapse of bridge at start of simulation.
-In middle picture,  bridge is shown in state where car has driven once over bridge without stopping and
+The vehicle is modelled as a single box btRaycastVehicle with four driving wheels. Wheels are rendered for visual feedback.
+In the left picture, the initial state is shown. 
+The gate in front is modelled to be so slender that it has visible elastic deflection.
+The gate support is modelled as a heavy box.
+The bridge has rigid ramps and supports at both ends.
+Gravity is applied gradually to avoid collapse of bridge at the start of the simulation.
+In the middle picture, the bridge is shown in a state where a car has driven once over the bridge without stopping and
 other bodies have taken light damage.
-In right picture, two joints have been broken and bridge has fallen from support due to bending.
+In the right picture, two joints have been broken and the bridge has fallen from teh support due to bending.

### 2016-10-30 pdocs/thesis/article.tex
working on Peter's comments
acm tog styling of article started

@@
 % Simo.Nikula@gmail.com
-% based on latex template at http://www.lut.fi/web/en/library/dissertations1 
-\def\mytitle{Description of plasticity in physics engine}
-\def\myisbn{978-952-265-XXX-X}
-\def\mypdfisbn{978-952-265-XXX-X}
-\def\myyear{2017}
-\def\mymonth{May}
-\def\myname{Simo Nikula}
+\def\mytitle{Elastic-plastic Constraint for Rigid Body Simulation}
 \def\lut{Lappeenranta University of Technology}
@@
 \def\charpy{CharpyDemo}
-\documentclass[a4paper,twoside,12pt,notitlepage]{article}
+\def\myvolume{VV}
+\def\mynumber{N}
+\def\myyear{YYYY}
+\def\mymonth{Month}
+\def\myaan{XXX}
+\def\mydoi{XXXXXXX.YYYYYYY}
+
+\documentclass{acmtog}
+\acmVolume{\myvolume}
+\acmNumber{\mynumber}
+\acmYear{\myyear}
+\acmMonth{\mymon}
+\acmArticleNum{\myaan}  
+\acmdoi{10.1145/\mydoi}
+
 \usepackage[utf8]{inputenc}
 \usepackage{graphicx}
-\usepackage{natbib}
-\usepackage{lastpage}   % page count
-\usepackage{amsmath}
+%\usepackage{natbib}
+%\usepackage{lastpage}   % page count
+%\usepackage{amsmath}
 \usepackage{listings}
 % \usepackage{mathdots}  % improved \ddots and \iidots
-\usepackage{amssymb}
-\usepackage[english]{babel}
-\usepackage[intoc]{nomencl} % including table of content
-\usepackage{ifthen}
-\usepackage[hmargin=2.0cm,vmargin=1.6cm]{geometry} % setting marginals
-\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
-\usepackage{times}  % to change font to times
-\usepackage{setspace} % for linespacingx
+%\usepackage{amssymb}
+%\usepackage[english]{babel}
+%\usepackage[intoc]{nomencl} % including table of content
+%\usepackage{ifthen}
+%\usepackage[hmargin=2.0cm,vmargin=1.6cm]{geometry} % setting marginals
+%\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
+%\usepackage{times}  % to change font to times
+%\usepackage{setspace} % for linespacingx
 \usepackage{tikz}
@@
 \usepackage{enumitem} % 
-\newcommand{\nomunit}[1]{\renewcommand{\nomentryend}{\hspace*{\fill}#1}} % Inserts units on the right at symbol list
-\renewcommand{\nomgroup}[1]{%
- \ifthenelse{\equal{#1}{C}}{\item[\textbf{Latin alphabet}]\item}{%
- \ifthenelse{\equal{#1}{G}}{\item\item[\textbf{Greek alphabet}]\item}{}}{%
- \ifthenelse{\equal{#1}{L}}{\item\item[\textbf{Subscripts}]\item}{}}{%
- \ifthenelse{\equal{#1}{H}}{\item\item[\textbf{Superscripts}]\item}{}}{%
- \ifthenelse{\equal{#1}{W}}{\item\item[\textbf{Abbreviations}]\item}{}}{}}
-
-\singlespacing
-
-\pagestyle{fancy}
-\fancyhf{}% clearing the header and footer
-\fancyhead[LE,RO]{\bfseries\thepage}    % page number to header
-\fancyhead[LO]{\nouppercase{\bfseries\rightmark}}     %
-\fancyhead[RE]{\nouppercase{\bfseries\leftmark}}%
-
-\renewcommand\sectionmark[1]
- {\markboth{\thesection\ #1}{}}         % section name to header
-\renewcommand\subsectionmark[1]
- {\markright{\thesubsection\ #1}}       % subsection name to header
-\renewcommand{\headrulewidth}{0.5pt}    % ruler thickness between head and body
-\renewcommand{\footrulewidth}{0pt}      % no ruler between body and footer
-
-\setlength{\nomitemsep}{-\parsep}   % removing default extra skip between entries at nomenclature
-
-\numberwithin{equation}{section}    % equation numbers with section numbers
-\numberwithin{table}{section}       % table numbers with section numbers
-\numberwithin{figure}{section}      % figure numbers with section numbers
-
-% makeindex command needs to run at command prompt to create nomenclature list file
-\makenomenclature % makeindex main_v2.nlo -s nomencl.ist -o main_v2.nls
-%\bibpunct{(}{)}{;}{a}{,}{,}%
-
 \begin{document}
 
-\pagestyle{empty}
 \title{\mytitle}
-\author{Simo Nikula, Aki Mikkola and Timo Björk}
-\author{}
+markboth{S. Nikula et al.}{\mytitle}
+
+\title{\mytitle}
+
+\author{SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK
+\affil{\lut}}
+
+\category{I.3.7}{Computer Graphics}{Three-Dimensional Graphics and Realism}[Animation]
+\category{I.6.3}{Computer Graphics}{Simulation and Modelling}[Applications]
+
+\keywords{Modeling, Animation, Deformation, Elasticity, Plasticity, Dynamics, Simulation}
+
+\acmformat{Nikula, S., Mikkola, A., Björk, T.,  G. V. G. \myyear. 
+\mytitle.  {ACM Trans. Graph.} \myvolume, \mynumber, Article \myaan (\mymon \myyear), # pages.\newline  DOI $=$
+10.1145/\mydoi\newline http://doi.acm.org/10.1145/\mydoi}
+
 \maketitle
+
+\begin{bottomstuff}
+
+Authors' addresses: Simo.Nikula@gmail.com
+\end{bottomstuff}
+
 \include{article-introduction}
@@
 
-\fancyhead[LO]{\nouppercase{\bfseries\firstleftxmark}}%
-\fancyhead[RE]{\nouppercase{\bfseries\lastrightxmark}}%
-\extramarks{}{References}   % this will remove the "References" to be appearing at the header at first page of bibliography
-\extramarks{References}{References}
-\bibliographystyle{LUTapa}% LUTapa
-\bibliography{ref}%
-
-%\cleardoublepage %
+\bibliographystyle{acmtog}
+\bibliography{ref}
 

### 2016-10-31 pdocs/thesis/article-section-2.tex
acm tog styling of article continues

@@
 %\citet[p.~45]{erleben.thesis} 
-\citet{erleben.thesis} 
+\cite{erleben.thesis} 
 provides reasoning and theoretical details for the popularity of 
@@
 Work presented by  
-\citet{erleben.thesis} provides the basis for the velocity-based formulation discussed in this work.
+\cite{erleben.thesis} provides the basis for the velocity-based formulation discussed in this work.
 %\citet[p.~45-50]{erleben.thesis}.
@@
 Impulse $\vec{J}$
-in the time interval $\Delta t $ can be written as
+in the time interval $\Delta t $ can be written as:
 \begin{equation} \label{eq:impulseIntegral}
@@
 % pdf page 33, notation in typical ODEs
-\begin {table}[htb!]
-\caption {Nomenclature for equations of motion}
-\label{tab:eom}
-\begin{center}
+\begin{table}
+\tbl{Nomenclature for equations of motion}{
 \begin{tabular}{|l| l|}
@@
 $N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
-\end {tabular}
-\end{center}
-\end {table}
+\end {tabular}}
+\label{tab:eom}
+\end{table}
 
@@
 equation \ref{eq:eomV} as,  
-\citet{erleben.thesis}
+\cite{erleben.thesis}
 %\citet[p.~66-67]{erleben.thesis}
@@
 \end{equation}
-where Jacobian terms $J_{contact}^T$ for joints are 
+where Jacobian terms $J_{joint}^T$ for joints are 
 derived by taking time derivatives of the kinematic constraints.
@@
 
-\begin{figure}[htb!]
+\begin{figure}
 \centering
@@
 % pdf page 33, notation in typical ODEs
-\begin {table}[htb!]
+\begin{table}
 \caption {Additional terms  for generalized equations of motion}
@@
 Joints are also discussed in detail in  
-\citet{erleben.thesis}.
+\cite{erleben.thesis}.
 %\citet[p.~60-90]{erleben.thesis}.
@@
 
-\begin {table}[htb!]
+\begin {table}
 \caption {Constraint parameters}
@@
 using displacement based boundary conditions is used.
-\citet{bathe-1975} provides a description for handling of various nonlinearities.
+\cite{bathe-1975} provides a description for handling of various nonlinearities.
 In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
@@
 Engineering stress and strain mean that original dimensions are used in stress calculation,
-\citet{dowling}.
+\cite{dowling}.
 %\citet[p.~108]{dowling}.
@@
 expands, 
-\citet{dowling}.
+\cite{dowling}.
 %\citet[p.~672]{dowling}.
@@
 \node[above] at (S) [align=center] {\large{$\sigma<f_y$}};
-\node[anchor=west] at (-2,3) {
+\node[anchor=west] at (3,3) {
 \begin{tabular}{l}
@@
 
-\begin {table}[htb!]
+\begin {table}
 \caption {Maximum forces and moments for 
@@
 b=0.01; h=0.01; fy=200e6;
-wpy=fy/2*dblquad(@(x,z) sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
+wpy=fy/2*dblquad(@(x,z)...
+ sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
 38.2

### 2016-11-01 pdocs/thesis/article-section-2.tex
acm tog styling of article continues

@@
 Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
-$\vec{v}^{\Delta t}$ can be solved as
+$\vec{v}^{\Delta t}$ can be solved for the velocity as:
 \begin{equation} \label{eq:impulseIntegraWithNewton}
@@
 derived by taking time derivatives of the kinematic constraints.
-
 Symbols used in Equation \ref{eq:eomV2} are summarized in Table
-\ref{tab:eom-g} and Figure \ref{fig:eom-joint}.
-Figure \ref{fig:eom-joint} presents terms needed for joint processing.
+\ref{tab:eom-g} and Figure \ref{fig:eom-joint},
+where $\vec{r}_{anc}^{\,i}$ is used to define at which point
+joint constraint is applied.
 
@@
 
-In Figure \ref{fig:eom-joint}, $\vec{r}_{anc}^{\,i}$ is used to define at which point
-joint constraint is applied.
-
 % pdf page 33, notation in typical ODEs
 \begin{table}
-\caption {Additional terms  for generalized equations of motion}
-\label{tab:eom-g}
-\begin{center}
+\tbl{Additional terms  for generalized equations of motion}{
 \begin{tabular}{|l| l|}
@@
 $\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
-\end {tabular}
-\end{center}
+\end {tabular}}
+\label{tab:eom-g}
 \end {table}
@@
 \begin {table}
-\caption {Constraint parameters}
-\label{tab:constraintParameters} 
-\begin{center}
+\tbl {Constraint parameters}{
 \begin{tabular}{|c| l| l|}
@@
 $\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
-\end {tabular}
-\end{center}
+\end {tabular}}
+\label{tab:constraintParameters} 
 \end {table}
@@
 \coordinate (Y) at (1,4);
-\draw[->] (0,0) -- (10,0) node[right] {\large{$\epsilon$}};
-\draw[->] (0,0) -- (0,6) node[above] {\large{$\sigma$}};
-\draw(0,0) -- (Y) -- (2,4) .. controls (7,6) .. (10,5);
+\draw[->] (0,0) -- (6,0) node[right] {\large{$\epsilon$}};
+\draw[->] (0,0) -- (0,5) node[above] {\large{$\sigma$}};
+\draw(0,0) -- (Y) -- (2,4) .. controls (5,5) .. (6,4);
 \draw[dashed](0,4) -- (Y);
@@
 Maximum force acting in a direction of $\vec{r}_{anc}^{\,i} $
-is product of area and yield stress as follows
+is product of area and yield stress as follows:
 
 \begin{equation} \label{eq:fN}
-N_{max}= \int_A f_y
+N_{max}= \int_A f_y.
 \end{equation}
@@
 Maximum forces acting perpendicular to $\vec{r}_{anc}^{\,i} $
-are a product of area and shear yield stress $\tau_y$ as follows
+are a product of area and shear yield stress $\tau_y$ as follows:
 \begin{equation} \label{eq:fQ}
-Q_{max}= \int_A \tau_y
+Q_{max}= \int_A \tau_y.
 \end{equation}
@@
 and yield stress $f_y$ as given for the moment around the $x$-axis 
-and moment around the $z$-axis, respectively.
+and moment around the $z$-axis, respectively:
 
 \begin{equation} \label{eq:Mx}
-M_{max}^x= \int_A z f_y
+M_{max}^x= \int_A z f_y,
 \end{equation}
@@
 \begin{equation} \label{eq:Mz}
-M_{max}^z= \int_A x f_y
+M_{max}^z= \int_A x f_y.
 \end{equation}
@@
 is an integral of distance $d$ from the joint point
-and shear yield stress $\tau_y$ as
+and shear yield stress $\tau_y$ as:
 
 \begin{equation} \label{eq:My}
-M_{max}^y= \int_A d \tau_y
+M_{max}^y= \int_A d \tau_y.
 \end{equation}
@@
 \begin {table}
-\caption {Maximum forces and moments for 
-rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
-\label{tab:maxForces} 
-\begin{center}
+\tbl {Maximum forces and moments for 
+rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}{
 \begin{tabular}{| c| c|}
@@
 maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
-\end{tabular}
-\end{center}
+\end{tabular}}
+\label{tab:maxForces} 
 \end {table}
@@
 An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
-is shown in Figure \ref{fig:octave-mp}
+is shown in Figure \ref{fig:octave-mp}.
 

### 2016-11-01 pdocs/thesis/article-section-3.tex
acm tog styling of article continues

@@
 
-\begin{figure}[htb!]
+\begin{figure}
 \begin{description}
@@
 
-\begin {table}[htb!]
-\caption {Simulation values for the unconstrained case without continuous collision detection. 
+\begin {table*}
+\tbl {Simulation values for the unconstrained case without continuous collision detection. 
 Typical values are shown for internal contact values
-as the number of contacts and the values at contact points differ.} 
-\label{tab:freeBlockValues} 
-\begin{center}
+as the number of contacts and the values at contact points differ.} {
 \begin{tabular}{|l| l|l| l|l|l|l|l|}
@@
 0.917 &  -1.500 & & & & & -0.001 & 1000 \\  \hline
-\end {tabular}
-\end{center}
-\end {table}
+\end {tabular}}
+\label{tab:freeBlockValues} 
+\end {table*}
 
-\begin{figure}[htb!]
-\centering
+\begin{figure}
 \input{figs/uc-fig.tikz}
-\caption{Graph for selected values for the unconstrained case without continuous collision detection.}
+\caption{Graph for location, velocity, positional error and velocity error
+for the unconstrained case without continuous collision detection.}
 \label{fig:uc}
@@
 {\it velError} is  3.5 (5.34+0.167)-0.033/0.0167 and
-{\it posError} is  0 as collision is detected before penetration. 
+{\it posError} is  zero all the time as collision is detected before penetration. 
 It should be noted that in general the ccd sphere should not extend the actual body as 
@@
 
-\begin {table}[htb!]
-\caption {Simulation values for the unconstrained case with continuous collision detection. } 
-\label{tab:freeBlockValuesWithCcd} 
-\begin{center}
+\begin {table}
+\tbl {Simulation values for the unconstrained case with continuous collision detection.}{ 
 \begin{tabular}{|l| l|l| l|l|l|l|l|}
@@
 {\bf Location} &
-{\it velError} & {\it penetration} & {\it posError} & {\it rhs} &
+{\it velError} & {\it pene-} & {\it rhs} &
 {\bf Velocity} & 
-{\bf Impulse} \\  \hline
-0.017 &  -0.003 & & & &  &-0.17 & 0 \\  \hline
-0.550 &  -1.500 & 3.5 & 0.033 & 0 & 21000& -2 & 21000 \\  \hline
-0.567 &  -1.500 & 2.17 & 0 & 0  &  8100 & 0.01 & 13000 \\  \hline
-0.583 &  -1.500 & 0.15 & 0 & 0 & 600  & 0 & 937 \\  \hline
-0.600 &  -1.500 & 0.17 & 0 & 0 & 600  & 0 & 1040 \\  \hline
-\end {tabular}
-\end{center}
+{\bf Impulse} \\ 
+ &  & & {\it tration} &  &
+ &  \\ 
+ \hline
+0.017 &  -0.003 & & &  &-0.17 & 0 \\  \hline
+0.550 &  -1.500 & 3.5 & 0.033 & 21000& -2 & 21000 \\  \hline
+0.567 &  -1.500 & 2.17 & 0 &  8100 & 0.01 & 13000 \\  \hline
+0.583 &  -1.500 & 0.15 & 0 & 600  & 0 & 937 \\  \hline
+0.600 &  -1.500 & 0.17 & 0 & 600  & 0 & 1040 \\  \hline
+\end {tabular}}
 \end {table}
@@
 
-\begin {table}[htb!]
-\caption {Constraint parameter values for the fixed constraint} 
-\label{tab:fixedBlockValues} 
-\begin{center}
+\begin{table}
+\tbl {Constraint parameter values for the fixed constraint} {
 \begin{tabular}{|l|l| l| l|l|l|l|}
@@
 0.35... & 0 &-0.17 & $\approx$ 0 & -1000 &0.0 & 1000 \\  \hline
-\end {tabular}
-\end{center}
+\end {tabular}}
+\label{tab:fixedBlockValues} 
 \end {table}
@@
 
-\begin {table}[htb!]
-\caption {Selected constraint parameters for elastic-plastic constraint 2}
-\label{tab:ep2Parameters} 
-\begin{center}
-\begin{tabular}{|l|p{10cm}|}
+\begin{table}
+\tbl {Selected constraint parameters for elastic-plastic constraint 2}{
+\begin{tabular}{|l|p{5cm}|}
 \hline
@@
  \\ \hline
-
-\end {tabular}
-\end{center}
+\end {tabular}}
+\label{tab:ep2Parameters} 
 \end {table}
@@
 This causes plastic strain during the next three steps. 
+Positional error is reported as zero at start of simulation because force exceeds
+maximum force.  At later phases positional error is not reported to avoid instability.
 
-\begin {table}[htb!]
-\caption {Constraint parameter values for elastic-plastic constraint} 
-\label{tab:epBlockValues} 
-\begin{center}
+\begin{table}
+\tbl{Constraint parameter values for elastic-plastic constraint}{ 
 \begin{tabular}{|l|l| l| l|l|l|l|l|}
@@
 {\bf Location} &
-{\it velError} & {\it posError} & {\it rhs} &
+{\it velError} & 
+{\it rhs} &
 {\it velocity} & 
 {\bf Impulse} & 
-{\bf Plastic strain} \\  \hline
-0.017 & -0.0028 &-0.17 & 0 & -1000 & -0.17 & 0 & 0 \\  \hline
-0.033 & -0.0046 &-0.33 & 0 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
-0.050 & -0.0056 &-0.28 & 0 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
-0.067 & -0.0056 &-0.22 & 0 & -1340 &  -0.001&  1330 & 0.004\\  \hline
-0.083... & -0.0056  & -0.17 & 0 & -1000 &  0.0&  1000 & 0.004\\  \hline
-\end {tabular}
-\end{center}
+{\bf Plastic } \\ 
+& & & & & &{\bf strain}\\
+\hline
+0.017 & -0.0028 &-0.17 & -1000 & -0.17 & 0 & 0 \\  \hline
+0.033 & -0.0046 &-0.33 & -2000 & -0.11 &  1330 & 0.001 \\  \hline
+0.050 & -0.0056 &-0.28 & -1670 & -0.056 &  1330 & 0.003 \\  \hline
+0.067 & -0.0056 &-0.22 & -1340 &  -0.001&  1330 & 0.004\\  \hline
+0.083... & -0.0056  & -0.17 & -1000 &  0.0&  1000 & 0.004\\  \hline
+\end {tabular}}
+\label{tab:epBlockValues} 
 \end {table}
@@
 
-\begin {table}[htb!]
-\caption {Constraint parameter values for elastic-plastic constraint 2} 
+\begin{table}
+\tbl {Constraint parameter values for elastic-plastic constraint 2} {
+\begin{tabular}{|l| l| l|l|l|l|}
+\hline
+{\bf Location} &
+{\it velError} &  {\it rhs} &
+{\bf Impulse} & 
+{\bf Plastic } \\ 
+ & & & &{\bf strain}\\
+\hline
+ 0 & -0.17  & -1000  & 1000 & 0 \\  \hline
+\end {tabular}}
 \label{tab:ep2BlockValues} 
-\begin{center}
-\begin{tabular}{|l|l| l| l|l|l|l|l|}
-\hline
-{\bf Time} & 
-{\bf Location} &
-{\it velError} & {\it posError} & {\it rhs} &
-{\bf Velocity} & 
-{\bf Impulse} & 
-{\bf Plastic strain} \\  \hline
-0.017... &  0 & -0.17  & 0 & -1000 & 0       & 1000 & 0 \\  \hline
-\end {tabular}
-\end{center}
 \end {table}
@@
 Figure \ref{fig:charpy-series} shows three detailed pictures from the simulation.
-The left picture shows early phase of simulation. 
-In the middle picture the hammer has not yet collided with the specimen.
-In the right picture the specimen is bending.
+The first picture shows early phase of simulation. 
+In the second picture the hammer has not yet collided with the specimen.
+In the last picture the specimen is bending. 
+After few more time steps specimen goes between anvils in one peace.
+If maximum rotation is set lower than 1.9 radians, joint is deactivated.
 
-\begin{figure}[htb!]
-\centering
-\includegraphics[height=3cm]{figs/article-charpy-1}
-\includegraphics[height=3cm]{figs/article-charpy-2}
-\includegraphics[height=3cm]{figs/article-charpy-3}
+\begin{figure}
+\includegraphics[height=2.3cm]{figs/article-charpy-1}
+\includegraphics[height=2.3cm]{figs/article-charpy-2}
+\includegraphics[width=8.65cm]{figs/article-charpy-3}
 \caption{Detailed pictures from simulation of a Charpy impact test}
@@
 
-\begin {table}[htb!]
-\caption {Energy loss for few timesteps}
-\label{tab:ep2ts} 
-\begin{center}
+\begin {table}
+\tbl {Energy loss for few timesteps}{
 \begin{tabular}{| c| c|l|}
@@
  0.05 &  150 & \\ \hline
-\end {tabular}
-\end{center}
+\end {tabular}}
+\label{tab:ep2ts} 
 \end {table}
@@
 
-\begin{figure}[htb!]
+\begin{figure}
 \centering
-\includegraphics[height=8cm]{figs/article-charpy}
+\includegraphics[height=4.5cm]{figs/article-charpy}
 \caption{Overview of simulation of Charpy impact test}
@@
 
-\begin{figure}[htb!]
+\begin{figure}
 \centering
-\includegraphics[height=5cm]{figs/demolisher-pre}
-\includegraphics[height=5cm]{figs/demolisher-wip}
-\includegraphics[height=5cm]{figs/demolisher-done}
+\includegraphics[width=8.65cm]{figs/demolisher-pre}
+\includegraphics[height=4.3cm]{figs/demolisher-wip}
+\includegraphics[height=4.3cm]{figs/demolisher-done}
 \caption{Simulation of Demolisher scenario}

### 2016-11-01 pdocs/thesis/article.tex
acm tog styling of article continues

@@
 \def\charpy{CharpyDemo}
-\def\myvolume{VV}
-\def\mynumber{N}
-\def\myyear{YYYY}
-\def\mymonth{Month}
-\def\myaan{XXX}
-\def\mydoi{XXXXXXX.YYYYYYY}
 
 \documentclass{acmtog}
-\acmVolume{\myvolume}
-\acmNumber{\mynumber}
-\acmYear{\myyear}
-\acmMonth{\mymon}
-\acmArticleNum{\myaan}  
-\acmdoi{10.1145/\mydoi}
+\acmVolume{VV}
+\acmNumber{N}
+\acmYear{201Y}
+\acmMonth{Month}
+\acmArticleNum{AN}  
+\acmdoi{10.1145/xxx.yy}
 
@@
 \usepackage{graphicx}
-%\usepackage{natbib}
-%\usepackage{lastpage}   % page count
-%\usepackage{amsmath}
+\usepackage{amsmath}
 \usepackage{listings}
-% \usepackage{mathdots}  % improved \ddots and \iidots
-%\usepackage{amssymb}
-%\usepackage[english]{babel}
-%\usepackage[intoc]{nomencl} % including table of content
-%\usepackage{ifthen}
-%\usepackage[hmargin=2.0cm,vmargin=1.6cm]{geometry} % setting marginals
-%\usepackage{fancyhdr,extramarks}  % header ja footer manipulation
-%\usepackage{times}  % to change font to times
-%\usepackage{setspace} % for linespacingx
+\usepackage{amssymb}
 \usepackage{tikz}
@@
 \usetikzlibrary{arrows.meta}
-\usepackage{caption}
 \usepackage{varwidth}
@@
 \begin{document}
-
-\title{\mytitle}
-markboth{S. Nikula et al.}{\mytitle}
+\markboth{S. Nikula, A. Mikkola and T. Björk}{\mytitle}
 
@@
 
-\acmformat{Nikula, S., Mikkola, A., Björk, T.,  G. V. G. \myyear. 
-\mytitle.  {ACM Trans. Graph.} \myvolume, \mynumber, Article \myaan (\mymon \myyear), # pages.\newline  DOI $=$
-10.1145/\mydoi\newline http://doi.acm.org/10.1145/\mydoi}
+\acmformat{Nikula, S., Mikkola, A., Björk, T.,  G. V. G. 201x. 
+\mytitle.  {ACM Trans. Graph.} VV, NN, Article AN (Mon 201x), nb pages.\newline  DOI $=$
+10.1145/xx.yy\newline http://doi.acm.org/10.1145/XXXXXXX.YYYYYYY}
 
@@
 
-Authors' addresses: Simo.Nikula@gmail.com
+%Authors' addresses: Simo.Nikula@gmail.com
 \end{bottomstuff}
 
-\include{article-introduction}
-\include{article-section-2}
-\include{article-section-3}
-\include{article-conclusion}
+\input{article-introduction}
+\input{article-section-2}
+\input{article-section-3}
+\input{article-conclusion}
 

### 2016-11-01 pdocs/thesis/octave/article.m

@@ -21,7 +21,6 @@ uc=[
 0.817 ,  -1.499 ,0,0,0,0, -0.08 , 700   ;
 0.917 ,  -1.500 ,0,0,0,0, -0.001 , 1000   ;
 ];
-
 subplot(2,2,1)
 plot(uc(:,1),uc(:,2),"color","black");
 title("Location");
@@ -34,4 +33,4 @@ title("Velocity");
 subplot(2,2,4)
 plot(uc(:,1),uc(:,3),"color","black");
 title("velError");
-%print -dtikz "-S600,400" "uc-fig"
+print -dtikz "-S300,200" "uc-fig"

### 2016-11-02 pdocs/thesis/article-conclusion.tex
bullet def changed to newcommand cbullet

@@
 Another area is usage of the graphics processing unit (GPU) for calculation.
-\bullet\ already has an experimental GPU pipeline but most constraint types are not 
+\cbullet\ already has an experimental GPU pipeline but most constraint types are not 
 yet supported. GPU support was one reason for selecting
-\bullet\ for use in this study
+\cbullet\ for use in this study
 
@@
 see realistic plasticity in main stream games. 
-\bullet\ is already integrated into several ones like Blender 
+\cbullet\ is already integrated into several ones like Blender 
 and it was another significant reason for selecting
-\bullet\  to be used in this study.
+\cbullet\  to be used in this study.
 

### 2016-11-02 pdocs/thesis/article-introduction.tex
acm tog styling of article started
acm tog styling of article continues
bullet def changed to newcommand cbullet

@@
 % !TeX root = article.tex
+\begin{abstract}
+We introduce efficient method to simulate ductile fracture in existing physics engines.
+\end{abstract}
+
 \section{Introduction}
@@
 In practice, software for such games is often written using open-source physics engine platforms  
-like ODE - \cite{ode}, \bullet\ - \cite{bullet} and Box2D - \cite{box2d}.
+like ODE - \cite{ode}, \cbullet\ - \cite{bullet} and Box2D - \cite{box2d}.
 
@@
 The study extends a method introduced by
-\citet{erleben.thesis} 
+\cite{erleben.thesis} 
 %\citet[p.~90]{erleben.thesis} 
@@
 predefined material based limit. The elastic part of deformation is modelled by employing 
-a spring based on modification of an existing constraint in \bullet.
+a spring based on modification of an existing constraint in \cbullet.
 

### 2016-11-02 pdocs/thesis/article-section-2.tex
bullet def changed to newcommand cbullet

@@
 % pdf page 64
-In the following section, these formulations will be clarified by a simple example using \bullet\ implementation.
+In the following section, these formulations will be clarified by a simple example using \cbullet\ implementation.
 
@@
 
-Constraint processing in \bullet\ is based on ODE, \cite{ode}.
+Constraint processing in \cbullet\ is based on ODE, \cite{ode}.
 Joints are also discussed in detail in  
@@
 In the following section, these equations will be explained by a simple example.
-The main parameters  and corresponding fields in \bullet\  
+The main parameters  and corresponding fields in \cbullet\  
 are given in Table \ref{tab:constraintParameters}.
@@
 
-Further enhancements were created and tested by forking \bullet\ source code
+Further enhancements were created and tested by forking \cbullet\ source code
 and adding new constraints, \cite{pbullet}.

### 2016-11-02 pdocs/thesis/article-section-3.tex
bullet def changed to newcommand cbullet

@@
 Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
-in \bullet\ is used to get values for the internal variables described in Figure \ref{fig:solVars}.
+in \cbullet\ is used to get values for the internal variables described in Figure \ref{fig:solVars}.
 
@@
 \end{description}
-\caption{Internal variables used in \bullet\ in constraint solving.}
+\caption{Internal variables used in \cbullet\ in constraint solving.}
 \label{fig:solVars}
@@
 
-There are currently two alternative six-dof-spring constraint implementations in \bullet\ and 
+There are currently two alternative six-dof-spring constraint implementations in \cbullet\ and 
 in this work elastic-plastic versions of both of them are developed.
@@
 
-\bullet\ simulations are usually done using a fixed time step of 1/60 s, i.e., 16.67 ms. 
+\cbullet\ simulations are usually done using a fixed time step of 1/60 s, i.e., 16.67 ms. 
 For this case that is too large. 

### 2016-11-02 pdocs/thesis/bullet-plasticity.tex
bullet def changed to newcommand cbullet

@@
-\section{Simulation of Charpy impact test using \bullet}
+\section{Simulation of Charpy impact test using \cbullet}
 \label{sec:bullet-charpy}
@@
 well known Charpy impact test was selected to be used as primary scenario
-although it is against first two tips in \bullet Manual.
+although it is against first two tips in \cbullet Manual.
 
@@
 
-\subsection{Introduction to constraint processing in \bullet}
-Constraint processing in \bullet is based on ODE, \cite{ode}.
+\subsection{Introduction to constraint processing in \cbullet}
+Constraint processing in \cbullet is based on ODE, \cite{ode}.
 Mathematical background and detailed examples are available by \cite{ode.joints}.
@@
 
-Main parameters  and corresponding fields in \bullet\  
+Main parameters  and corresponding fields in \cbullet\  
  are described in table \ref{tab:constraintParameters}.
@@
 \subsubsection{Breaking of constraints between bodies}
-For standard \bullet constraints breakingImpulseThreshold (BITh) can be defined. If impulse is larger than set limit constraint is activated. 
+For standard \cbullet constraints breakingImpulseThreshold (BITh) can be defined. If impulse is larger than set limit constraint is activated. 
 This allows breaking to be simulated with fast calculation. 
@@
 This basically same as SpringConstraint but uses newer btGeneric6DofSpring2Constraint which has recently appeared in
-\bullet. 
+\cbullet. 
 Automatic stiffness limitation helps to avoid instability but may make constraint too soft.
@@
 
-\section{Charpy impact test results using \bullet}
+\section{Charpy impact test results using \cbullet}
 \label{sec:bullet-charpy-results}

### 2016-11-02 pdocs/thesis/discussion.tex
Aki's comments on article and in general
bullet def changed to newcommand cbullet

@@
 \subsubsection{Implementation using GPU}
-\bullet\ can already utilize GPU for many operations. 
+\cbullet\ can already utilize GPU for many operations. 
 Analysis on possible issues in used interfaces should be done before integrations.
@@
 
-Current implementation is done based on \bullet\ but it still needs refactoring and testing before it can be 
+Current implementation is done based on \cbullet\ but it still needs refactoring and testing before it can be 
 accepted to main stream. Interfaces should be reafctored so that handling of additional features like large 
@@
 to solver core are needed.
-In \bullet\ there are already three per step callbacks.
+In \cbullet\ there are already three per step callbacks.
 
@@
 Just in time activation could avoid extra overhead from multiple objects and constraints.
-Impacts causing plasticity are usually quite brief and they happen at distinct times. 
+Impacts causing plasticity are usually brief and they happen at distinct times. 
 Simulation could allocate processing units for handling these and release them after impact has been processed.

### 2016-11-02 pdocs/thesis/article-introduction.tex
acm tog styling of article and polishing

@@
 \begin{abstract}
-We introduce efficient method to simulate ductile fracture in existing physics engines.
+We introduce simple and efficient method to simulate ductile fracture in existing physics engines.
+Method is based on common technique of splitting bodies to multiple pieces.
 \end{abstract}
@@
 is not appropriate if the simulation is to look realistic. Theory for handling of plasticity 
-has been presented already in \cite{cg1988}. \cite{muller2005meshless} 
+has been presented already in \cite{cg1988}. 
+\cite{Jones:2016:EPD} provides extensive listing of related work during past decades.
+\cite{muller2005meshless} 
 present a method for modeling and animating of elastic and plastic bodies in real time using 
@@
 One major issue is collision handling of deformable bodies.
+Recently there has been multiple studies which have same target of bringing plasticity 
+into main stream games as this work, \cite{Jones:2016:EPD,Patkar:2015:EDB,Budsberg:2014:EPD}.
 

### 2016-11-05 pdocs/thesis/article-introduction.tex
article polishing

@@
 We introduce simple and efficient method to simulate ductile fracture in existing physics engines.
-Method is based on common technique of splitting bodies to multiple pieces.
+Method is based on common technique of splitting bodies to multiple pieces and joining them with constraints.
+Constraint behaviour is modelled using body dimensions and material parameters.
+Sample program with full source code is made available to allow developers already using \cbullet\
+to add plasticity into their simulations with minimal effort.
 \end{abstract}
@@
 Recently there has been multiple studies which have same target of bringing plasticity 
-into main stream games as this work, \cite{Jones:2016:EPD,Patkar:2015:EDB,Budsberg:2014:EPD}.
+into wider usage, \cite{Jones:2016:EPD,Patkar:2015:EDB,Budsberg:2014:EPD}.
+Suggested methods seem best suited for film production and possible also for high end games.
+Methodology presented in this study does not require significant software development efforts from
+game vendors and is thus easily adoptable.
 

### 2016-11-06 pdocs/thesis/article-section-3.tex
article polishing

@@
 
-Method btSequentialImpulseConstraintSolver::solveGroupCacheFriendlySetup
+Method \mbox{btSequentialImpulseConstraintSolver::} \mbox{solveGroupCacheFriendlySetup}
 in \cbullet\ is used to get values for the internal variables described in Figure \ref{fig:solVars}.
@@
 \end {tabular}}
+\label{tab:freeBlockValuesWithCcd} 
 \end {table}

### 2016-11-22 examples/CharpyDemo/BlockDemo.cpp
release translation in BlockDemo

@@ -999,6 +999,8 @@ public:
 			sc->setStiffness(i, axisMapper->getStiffness(i), limitIfNeeded);
 			sc->setDamping(i, damping2);
 		}
+		sc->setLinearLowerLimit(btVector3(1, 1, 1));
+		sc->setLinearUpperLimit(btVector3(-1, -1, -1));
 		sc->setEquilibriumPoint(1, getEquilibriumPoint());
 		return sc;
 	}

### 2016-11-22 pdocs/thesis/article-introduction.tex
article polishing
work on Aki's comments

--- 2016-11-02 pdocs/thesis/article-introduction.tex
@@
 \begin{abstract}
-We introduce efficient method to simulate ductile fracture in existing physics engines.
+This paper introduces simple and efficient method to simulate ductile fracture in existing physics engines.
+Method is based on technique of splitting bodies to multiple pieces and joining them with constraints.
+Constraints are described using body dimensions and material parameters.
+Sample program with source code is made available to allow developers already using \cbullet\
+to add plasticity into their simulations.
 \end{abstract}
@@
 solved can be subdivided into equations related to motion, constraints and collisions.
+
 Velocity-based formulation is typically used in constraint based rigid body simulation, \cite{erleben.thesis}. 
-Friction is typically taken into account and mechanical joints are handled by constraint equations,
+Contacts and mechanical joints are handled by constraint equations 
+and friction is typically taken into account,
 \cite{erleben.thesis}.
+Impulse-based rigid body simulation was introduced by \cite{Hahn:1998}, 
+where time integrals of contact forces were used to model the interactions.
+\cite{mirtich} suggested that all contact interactions
+whether colliding, rolling, sliding or resting are affected
+via trains of collisions.
 
-Plasticity is not typically taken into account in gaming solutions and 
-destruction of various bodies often takes place based on collision or impulse exceeding predefined limit.
+Plasticity can be  accounted in gaming solutions by destructing various bodies based on collision
+or impulse exceeding predefined limit.
 Nevertheless, breaking of steel or reinforced concrete structures using this approach 
 is not appropriate if the simulation is to look realistic. Theory for handling of plasticity 
-has been presented already in \cite{cg1988}. \cite{muller2005meshless} 
+has been presented already in \cite{cg1988}. 
+\cite{Jones:2016:EPD} provides extensive listing of related work during past decades.
+\cite{Obrien:2002:GMA} extends previously introduced finite element based method \cite{Obrien:1999:GMA}
+to ductile materials.
+\cite{muller2005meshless} 
 present a method for modeling and animating of elastic and plastic bodies in real time using 
@@
 One major issue is collision handling of deformable bodies.
+Recently there has been multiple studies which have same target of bringing plasticity 
+into wider usage.
+\cite{Jones:2016:EPD} introduces an plasticity model which
+maps impulses computed by rigid body solved to local example-based deformations.
+\cite{Patkar:2015:EDB} presents a method for denting and bending of rigid bodies
+
+\cite{Budsberg:2014:EPD}
+Suggested methods are best suited for film production and high end games.
 
 This study will introduce an approach to account for plastic deformation in game applications.   
+Presented methodology does not require significant software development efforts from
+game vendors and is thus easily adoptable. 
 In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 

### 2016-11-23 pdocs/thesis/article-introduction.tex
work on Aki's comments

@@
 maps impulses computed by rigid body solved to local example-based deformations.
-\cite{Patkar:2015:EDB} presents a method for denting and bending of rigid bodies
-
-\cite{Budsberg:2014:EPD}
-Suggested methods are best suited for film production and high end games.
+\cite{Patkar:2015:EDB} presents methods for denting and bending of rigid bodies.
+Denting is simulated by creating surface on the target body based on the projectile body
+and bending is modelled by augmenting each rigid body with an articulated skeleton.
+\cite{Budsberg:2014:EPD} presents a method of simulating flexible materials so
+that rigid body fragments are connected by ball and socket joints and plasticity is
+modelled by updating constraint parameters after elastic limit has been passed.
 
@@
 The approach presented in this work can be used in the gaming industry to provide more realistic 
-simulations without significant extra work. For gaming purposes, the presented method works 
-best in scenarios where the connected parts are heavy. This allows a normal 
+simulations. For gaming purposes, the presented method works 
+best in scenarios where the connected parts are heavy as this allows a normal 
 integration timestep to be used without stability issues. 

### 2016-11-23 pdocs/thesis/article-section-2.tex
work on Aki's comments

@@
 are summarized in Table \ref{tab:eom}.
-Figure \ref{fig:eom-contact} describes the collision of two bodies, $B_1$ and $B_2$
-where $\vec{r}_i$ is orientation of body $i$,
-$\vec{p}_k $ is orientation of contact point $k$,
+Figure \ref{fig:eom-contact} describes the collision of two
+bodies, rectangular body $B_1$ and triangular body $B_2$
+where $\vec{r}_i$ is position of body $i$,
+$\vec{p}_k $ is position of contact point $k$,
 $\vec{r}_{ik} $ is a vector between the center of gravity of body $i$ and contact point $k$, 
@@
 $\vec{n}_{k}$ is the contact normal for contact point $k$.
+It is common convention that the contact normal points
+from the body with the smallest index to the body with the largest index, \cite{erleben.thesis}.
+In case of point or edge contacts, averaging the normals of neighboring polygons can be used, \cite{Hahn:1998}.
 
@@
 \node[anchor=west] at (4.5,4.5) {
-$\vec{r}_i$  =  orientation of body $i$
+$\vec{r}_i$  =  position of body $i$
 };
-\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point $k$};
+\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = position of contact point $k$};
 \node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
@@
 where $\vec{r}_{anc}^{\,i}$ is used to define at which point
-joint constraint is applied.
+joint constraint is applied relative to body $i$.
 
@@
 \node[anchor=west] at (4.5,3.5) {
-$\vec{r}_i$  =  orientation of body $i$ 
+$\vec{r}_i$  =  position of body $i$ 
 };

### 2017-06-10 pdocs/thesis/epj-jcgt.tex
jcgt article started

--- 2016-11-01 pdocs/thesis/article-section-2.tex
@@
-% !TeX root = article.tex
+% Simo.Nikula@gmail.com
+\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation}
+\def\lut{Lappeenranta University of Technology}
+\def\
+
+\documentclass{jcgt}
+
+\setciteauthor{Simo Nikula, Aki Mikkola and Timo Björk}
+\setcitetitle{\\mytitle}
+
+% Mark submissions with the date of submission using the following line:
+\submitted{\today}
+
+% Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
+%\accepted{2014-02-07}{2014-02-07}{2014-02-07}{Editor Name}{3}{1}{1}{1}{2014}
+%\seturl{http://jcgt.org/published/0003/01/01/}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\begin{document}
+\newcommand{\cbullet}[0]{Bullet Physics}
+\title{\\mytitle}
+
+\author
+       {SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK\\\lut
+       }
+
+\teaser{
+  \includegraphics[width=2in]{figs/demolisher-done.PNG}
+  \label{fig:teaser}
+  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D
+allows more realistic simulation of many scenarios.
+}
+}
+
+\maketitle
+\thispagestyle{firstpagestyle}
+
+\begin{abstract}
+\small
+This paper introduces simple and efficient method to simulate ductile fracture in existing 
+impulse-based physics engines.
+Method is based on technique of splitting bodies to multiple pieces and joining them with constraints.
+Constraints are described using body dimensions and material parameters.
+
+Sample programs with source code are made available to allow developers already using \cbullet\
+and Box2D to add plasticity into their simulations.
+\end{abstract}
+
+%-------------------------------------------------------------------------
+\section{Introduction}
+\label{sec:introduction}
+
+Theory for handling of plasticity in computer graphics has been presented already 1988, \cite{cg1988}. 
+Recent paper by Jones  provides extensive listing of related work during past decades, \cite{Jones:2016:EPD}.
+
+In impulse-based phycis engines most breakable scenarios  are implemented by destructing various bodies based on collision
+or impulse exceeding predefined limit.
+Nevertheless, breaking of steel or reinforced concrete structures using this approach 
+is not appropriate if the simulation is to look realistic.
+
+More realistic simulation of ductile destructable bodies is possible but in many cases it would require selecting of new physics engine.
+
+\paragraph{Our approach}
+This study will introduce an approach to account for plastic deformation in impulse-based phycis engines.   
+Presented methodology does not require significant software development efforts from
+game vendors and is thus easily adoptable. 
+In the introduced method, plastic deformation takes place if the force or moment exceeds a predefined 
+limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
+Maximum forces and moments are estimated based on the plastic section modulus.
+Joint breaking is based on summing plastic deformation and comparing it to a
+predefined material based limit. The elastic part of deformation is modelled by employing 
+modification of an existing constraints.
+
+\paragraph{Limitations}
+Forces are calculated by dividing impulse by timestep and maximum impulse is calculated
+by multiplying maximum force by timestep. 
+This means that simulation of quick interactions requires small timestep.
+This limitation is not introduced by this approach but is build in feature of impulse-based simulation.
+
+Multiaxial stress is not taken into account.
+
 \section{Description of plasticity in the framework of physics engines}
 
-In this section, key concepts related to the introduced model are explained. The main differences between 
-traditional structural analysis and physics engines based approaches are reviewed and discussed.
-
-Velocity-based formulation of constraint based rigid body simulation 
-is  commonly used by physics based game
-developers and film production teams.
-%\citet[p.~45]{erleben.thesis} 
-\cite{erleben.thesis} 
-provides reasoning and theoretical details for the popularity of 
-velocity-based formulation in constraint-based rigid body simulation instead of accelaration based. 
-The main reason is that collision handling can be done without the use of additional procedures.
-
-Work presented by  
-\cite{erleben.thesis} provides the basis for the velocity-based formulation discussed in this work.
-%\citet[p.~45-50]{erleben.thesis}.
-% pdf page 64
-In the following section, these formulations will be clarified by a simple example using \bullet\ implementation.
-
-Impulse $\vec{J}$
-in the time interval $\Delta t $ can be written as:
-\begin{equation} \label{eq:impulseIntegral}
-\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt,
-\end{equation}
-where $\vec{f}_{true}(t)$ is force.
-
-Using Newton's second law of motion $\vec{F}=m\vec{a}$ ,
-$\vec{v}^{\Delta t}$ can be solved for the velocity as:
-\begin{equation} \label{eq:impulseIntegraWithNewton}
-\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
-\end{equation}
-\begin{equation} \label{eq:impulse}
-m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J},
-\end{equation}
-where superscripts denote time, i.e. ${\vec{v}}^{\Delta t}=\vec{v}(\Delta t)$.
-Next position can be found
-by integrating the velocity.
-Updates after each step can be summarized  for locations and  
-for velocities respectively as follows:
-
-\begin{equation} \label{eq:eomL} % pdf page 69
-\vec{s}^{\, t+\Delta t} = \vec{s}^{\, t}+\Delta t S \vec{u}^{\, t+\Delta t}
-\end{equation}
-
-\begin{equation} \label{eq:eomV}
-\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) .
-\end{equation}
-
-The symbols used in Equations \ref{eq:eomL} and \ref{eq:eomV}
-are summarized in Table \ref{tab:eom}.
-Figure \ref{fig:eom-contact} describes the collision of two bodies, $B_1$ and $B_2$
-where $\vec{r}_i$ is orientation of body $i$,
-$\vec{p}_k $ is orientation of contact point $k$,
-$\vec{r}_{ik} $ is a vector  between the center of gravity of body $i$ and contact point $k$, 
-and
-$\vec{n}_{k}$ is the contact normal for contact point $k$.
-
-\begin{figure}[tb!]
-\centering
-\begin{tikzpicture}
-\coordinate (O1) at(2,1);
-\coordinate (O2) at(2,4);
-\coordinate (C) at(2,2);
-\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
-\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
-\node at (3.5,0.4) {$B_1$};
-\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
-\node at (2.8,4.5) {$B_2$};
-\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
-\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
-\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
-\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
-\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
-\node[anchor=west] at (4.5,4.5) {
-$\vec{r}_i$  =  orientation of body $i$
-};
-\node[anchor=west] at (4.5,4) {$\vec{p}_k $ = orientation of contact point $k$};
-\node[anchor=west] at (4.5,3.5) {$\vec{r}_{ik} $ = $\vec{p}_k - \vec{r}_i $};
-\node[anchor=west] at (4.5,3) {$\vec{n}_{k} $ = normal for contact point $k$};
-\end{tikzpicture}
-\caption{Illustration of nomenclature for equations of motion for collision.}
-\label{fig:eom-contact}
-\end{figure}
-
-% pdf page 33, notation in typical ODEs
-\begin{table}
-\tbl{Nomenclature for equations of motion}{
-\begin{tabular}{|l| l|}
-\hline
-{\bf Symbol} & {\bf Description} \\  \hline
-$\vec{r}_i$ & position of center of mass for body $i$  \\ \hline
-$\vec{q}_i$ & orientation for body $i$ as quaternion $\lbrack s_i, x_i, y_i, z_i \rbrack ^T $ \\
-\hline
-$\vec{p}_h$ & contact or joint point $k$  \\ \hline
-$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
-$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
-$Q_i$ & \begin{tabular}{@{}c}
-rotation of quaternion $\vec{q}_i$
-as matrix \\ where 
-$\frac{1}{2}\vec{\omega}_i \vec{q}_i=Q_i \vec{\omega}_i$
-\end{tabular}
-   $
-\frac{1}{2} \left[ \begin{array}{ccc}
--x_i & -y_i & -z_i \\
-s_i & z_i & -y_i \\
--z_i & s_i & x_i \\
-y_i & -x_i & s_i
-\end{array} \right]
-$
-\\ \hline
-$S$ & 
-\begin{tabular}{@{}c}
-generalized transformation matrix \\
-$ S \in \mathbb{R}^{7n \times 6n}$
-\end{tabular}
- $ \left[ \begin{array}{ccccc}
-1 &  &  & & 0 \\
- & Q_i  \\
- & & \ddots  \\
- & & & 1 \\
-0 & & & & Q_n 
-\end{array} \right]
-$
-\\ \hline
-$\vec{v}_i$ & linear velocity of  center of mass for body $i$   \\ \hline
-$\vec{\omega}_i$ & angular velocity of center of mass for body $i$  \\ \hline
-$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
-$M$ &
-\begin{tabular}{@{}c}
- generalized mass matrix \\
-$ M \in \mathbb{R}^{6n \times 6n}$  
-\end{tabular}
-$
-\left[ \begin{array}{ccccc}
-m_i 1 &  &  & & 0 \\
- & I_1  \\
- & & \ddots  \\
- & & & m_n 1 \\
-0 & & & & I_n 
-\end{array} \right]
-$
-\\ \hline
-$I_i$ & inertia tensor for body $i$  \\ \hline
-$C$ & contact condition matrix  $ C \in \mathbb{R}^{6n \times 3K}$ \\ \hline
-$N$ & contact normal matrix  $ N \in \mathbb{R}^{3K \times K}$ \\ \hline
-\end {tabular}}
-\label{tab:eom}
-\end{table}
-
-Friction in contacts and joint constraints can be handled in a unified way by refactoring
-equation \ref{eq:eomV} as,  
-\cite{erleben.thesis}
-%\citet[p.~66-67]{erleben.thesis}
-\begin{equation} \label{eq:eomV2}
-\vec{u}^{\, t+\Delta t} = \vec{u}^{\, t}+\Delta t M^{-1}(
- J_{contact}^T \vec{\lambda}_{contact}
-+ J_{joint}^T \vec{\lambda}_{joint}
-+ \vec{f}_{ext}),
-\end{equation}
-where Jacobian terms $J_{joint}^T$ for joints are 
-derived by taking time derivatives of the kinematic constraints.
-Symbols used in Equation \ref{eq:eomV2} are summarized in Table
-\ref{tab:eom-g} and Figure \ref{fig:eom-joint},
-where $\vec{r}_{anc}^{\,i}$ is used to define at which point
-joint constraint is applied.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}
-\coordinate (O1) at(1,1);
-\coordinate (O2) at(1,3);
-\coordinate (C) at(1,2);
-\draw (0,0) -- (2,0) -- (2,2) -- (0,2) --(0,0);
-\draw (0,2) -- (2,2) -- (2,4) -- (0,4) --(0,2) ;
-\node at (0.5,0.4) {$B_1$};
-\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
-\node at (0.5,3.5) {$B_2$};
-\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
-\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{anc}^{\,1}$};
-\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{anc}^{\,2}$};
-\node[anchor=west] at (4.5,3.5) {
-$\vec{r}_i$  =  orientation of body $i$ 
-};
-\node[anchor=west] at (4.5,3) {$\vec{r}_{anc}^{\,i} $ = body frame vector $i$};
-\end{tikzpicture}
-\caption{Illustration of nomenclature for equations of motion for joint.}
-\label{fig:eom-joint}
-\end{figure}
-
-% pdf page 33, notation in typical ODEs
-\begin{table}
-\tbl{Additional terms  for generalized equations of motion}{
-\begin{tabular}{|l| l|}
-\hline
-{\bf Symbol} & {\bf Description} \\  \hline
-$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
-$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
-$J_{joint}$ & Jacobian matrix for joints  \\ \hline
-$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
-\end {tabular}}
-\label{tab:eom-g}
-\end {table}
-
-Constraint processing in \bullet\ is based on ODE, \cite{ode}.
-Joints are also discussed in detail in  
-\cite{erleben.thesis}.
-%\citet[p.~60-90]{erleben.thesis}.
-Equations \ref{eq:constraintEquation}, \ref{eq:lambdaLow} and
-\ref{eq:lambdaHigh} 
-are created for each constraint. 
-Derivation for terms in Equation \ref{eq:constraintEquation}
-can be done using the position and orientation of connected bodies
-e.g. for ball joint formulation is based on both joint points having the same position.
-In contact cases, formulation is easier if it is done using velocities, \cite{ode.joints}.
-
-\begin{equation} \label{eq:constraintEquation}
-J_1 \vec{v}_1 + \Omega_1 \vec{\omega}_1 + 
-J_2 \vec{v}_2 + \Omega_2 \vec{\omega}_2 = \vec{c} + C \vec{\lambda}
-\end{equation}
-
-\begin{equation} \label{eq:lambdaLow}
-\vec{\lambda} \geq \vec{l}
-\end{equation}
-
-\begin{equation} \label{eq:lambdaHigh}
-\vec{\lambda} \leq \vec{h}
-\end{equation}
-
-In the following section, these equations will be explained by a simple example.
-The main parameters  and corresponding fields in \bullet\  
-are given in Table \ref{tab:constraintParameters}.
-
-\begin {table}
-\tbl {Constraint parameters}{
-\begin{tabular}{|c| l| l|}
-\hline
-{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
-$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\
-$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
-$\vec{v}$ & linear velocity & \\ \hline
-$\vec{\omega}$ & angular velocity & \\ \hline
-$\vec{c}$        &  right side vector   & m\_constraintError \\ \hline
-$C$  & constraint force mixing & cfm \\  \hline
-$\vec{\lambda}$ & constraint force &  \\ \hline
-$\vec{l}$ & low limit for constraint force & m\_lowerLimit \\ \hline
-$\vec{h}$ & high limit for constraint force & m\_upperLimit \\ \hline
-\end {tabular}}
-\label{tab:constraintParameters} 
-\end {table}
-
-In structural analysis, a formulation and associated numerical solution procedure are selected 
-based on needed features.
-Often,  finite element method is used.
-In most cases, a static solution with an assumption of linear strain-displacement relation
-using displacement based boundary conditions is used.
-\cite{bathe-1975} provides a description for handling of various nonlinearities.
-In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
-Lagrangian formulation where initial configuration is used.
-Further enhancements are material nonlinearity and dynamic analysis.
-Physics engine provides dynamic analysis with large reference translations and rotations
-while assuming bodies to be undeformable.
-
-Material plasticity can be accounted for in simulations by using a suitable coefficient of restitution.
-This provides a reasonable means to simulate loss of energy in collisions.
 In this work simulation of breaking of bodies made of ductile material is made more realistic 
@@
 to multiple bodies that are connected by energy absorbing joints.
+
+\paragraph{Stress-strain behaviour of ductile steel}
 A typical engineering stress-strain curve of ductile steel is shown in Figure \ref{fig:sscurve}.
@@
 0.001 to 0.005 and fracture strain can be 100 times larger.
+In practice this means that elastic displacement due to axial forces is usually not visible.
+Visible elastic displacement is usually due to bending.
 
@@
 Elastic material behavior is ignored in approach introduced in this work if
-the deformation is related to ahigher frequency
+the deformation is related to a higher frequency
 than integration stability would allow.
@@
 
-As shown in Figure 2.4, if stress is below yield limit $f_y$, stress and strain are linear within the material.
+As shown in Figure \ref{fig:wp}, if stress is below yield limit $f_y$, stress and strain are linear within the material.
 If cross section is fully plastic, stress is assumed to be at yield level over the whole cross section such that 
@@
 
-In this work, plasticity is handled by defining maximum forces
-in Equations \ref{eq:lambdaLow} and  
-\ref{eq:lambdaHigh} using plastic capacities, which are defined below.
+\paragraph{Plastic capacities}
+In this work, plasticity is handled by defining maximum forces using plastic capacities, which are defined below.
 
@@
 
-\begin{figure}
-\centering
 \lstset{language=octave}
-\begin{lstlisting}
+\begin{lstlisting}[caption={Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}, 
+label={lst:octave-mp}, float]
 b=0.01; h=0.01; fy=200e6;
@@
 
-\caption{Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}
-\label{fig:octave-mp}
-\end{figure}
+\paragraph{Joint breaking}
 
-The basic idea introduced in this study can be tested with any framework having motors and hinge constraints.
-This can be done by setting the target velocity of the motor to zero and limiting 
-the maximum motor impulse to plastic moment multiplied by a timestep.
+\section{Implementation for Box2D}
 
-Further enhancements were created and tested by forking \bullet\ source code
-and adding new constraints, \cite{pbullet}.
-Instructions for using  windows executable and  source code are available, \cite{bp}.
+\section{Implementation for \cbullet\}
+
+\small
+\bibliographystyle{jcgt}
+\bibliography{paper}
+
+\section*{Index of Supplemental Materials}
+
+\section*{Author Contact Information}
+
+\hspace{-2mm}\begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
+Simo Nikula\newline
+\href{mailto:Simo.Nikula@gmail.com}{Simo.Nikula@gmail.com}
+&
+Aki Mikkola\newline
+\href{mailto:Aki.Mikkola@lut.fi}{Aki.Mikkola@lut.fi}
+&
+Timo Björk\newline
+\href{mailto:Timo.Bjork@lut.fi}{Timo.Bjork@lut.fi}
+
+\end{tabular}
+
+\afterdoc
+
+\end{document}

### 2017-06-11 pdocs/thesis/epj-jcgt.tex
jcgt article - first pdf

@@
 \def\lut{Lappeenranta University of Technology}
-\def\
 
 \documentclass{jcgt}
-
+\usepackage{tikz}
+\usepackage{verbatim}
+\usepackage[utf8]{inputenc}
+\usetikzlibrary{arrows}
+\usetikzlibrary{arrows.meta}
 \setciteauthor{Simo Nikula, Aki Mikkola and Timo Björk}
-\setcitetitle{\\mytitle}
+\setcitetitle{\mytitle}
 
 % Mark submissions with the date of submission using the following line:
-\submitted{\today}
+%\submitted{\today}
 
 % Once an article is accepted accepted, switch to the following line and comment the preceding one. The editor will supply the argument values.
-%\accepted{2014-02-07}{2014-02-07}{2014-02-07}{Editor Name}{3}{1}{1}{1}{2014}
-%\seturl{http://jcgt.org/published/0003/01/01/}
+\accepted{2014-02-07}{2014-02-07}{2014-02-07}{Editor Name}{3}{1}{1}{1}{2014}
+\seturl{http://jcgt.org/published/0003/01/01/}
 
@@
 \newcommand{\cbullet}[0]{Bullet Physics}
-\title{\\mytitle}
+\title{Breakable Elastic-plastic Constraint for\\Rigid Body Simulation}
 
 \author
-       {SIMO NIKULA {\upshape and} AKI MIKKOLA {\upshape and} TIMO BJÖRK\\\lut
+       {Simo Nikula  Aki Mikkola Timo Björk\\\lut
        }
@@
 \teaser{
-  \includegraphics[width=2in]{figs/demolisher-done.PNG}
+  \includegraphics[width=5cm]{figs/demolisher-done}
   \label{fig:teaser}
-  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D
-allows more realistic simulation of many scenarios.
-}
+  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D allows more realistic simulation of many scenarios.}
 }
@@
 This limitation is not introduced by this approach but is build in feature of impulse-based simulation.
-
-Multiaxial stress is not taken into account.
+Bending moment - normal force interactions and multiaxial stress are not taken into account. 
+Basically these limitations just state that results from these simulations cannot not be used
+for structural analysis.
 
@@
 Material that starts to yield first is hardened and as a result of which yielding moves.
-
+%
 The difference between the elastic and plastic section modulus is depicted in Figure \ref{fig:wp}.
@@
 \begin {table}
-\tbl {Maximum forces and moments for 
-rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}{
+\small
+\begin{center}
 \begin{tabular}{| c| c|}
@@
 maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
-\end{tabular}}
+\end{tabular}
+\end{center}
+\caption{Maximum forces and moments for 
+rectangular section with width $b$ and height $h$ using constant yield stress $f_y$}
 \label{tab:maxForces} 
@@
 
-\section{Implementation for \cbullet\}
+\section{Implementation for \cbullet}
 
@@
 \bibliographystyle{jcgt}
-\bibliography{paper}
+\bibliography{ref}
 
@@
 
-\hspace{-2mm}\begin{tabular}{p{0.5\textwidth}p{0.5\textwidth}}
+\hspace{-2mm}\begin{tabular}{p{0.3\textwidth}p{0.3\textwidth}p{0.3\textwidth}}
 Simo Nikula\newline
@@
 \href{mailto:Timo.Bjork@lut.fi}{Timo.Bjork@lut.fi}
-
 \end{tabular}

### 2017-07-16 pdocs/thesis/epj-jcgt.tex
jcgt article - eol
minor tuning jcgt article

@@
 \author
-       {Simo Nikula  Aki Mikkola Timo Björk\\\lut
+       {Simo Nikula,  Aki Mikkola, Timo Björk\\\lut
        }
@@
 Method is based on technique of splitting bodies to multiple pieces and joining them with constraints.
-Constraints are described using body dimensions and material parameters.
+Procedure to provide realistic parameters for constraints is described.
+Procedure is based on body dimensions and material parameters.
 
@@
 An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
-is shown in Figure \ref{fig:octave-mp}.
+is shown in Listing \ref{lst:octave-mp}.
 

### 2017-08-04 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 %  Simo.Nikula@gmail.com
-\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation}
+\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse based Physics}
 \def\lut{Lappeenranta University of Technology}
@@
 In impulse-based phycis engines most breakable scenarios  are implemented by destructing various bodies based on collision
-or impulse exceeding predefined limit.
+or impulse exceeding predefined limit. This provides simple means for visualization of explosions and high speed collisions 
+on brittle structures.  
 Nevertheless, breaking of steel or reinforced concrete structures using this approach 
-is not appropriate if the simulation is to look realistic.
+is not appropriate if the simulation is to look realistic or structures should not fail completely.
 
@@
 limit, deformation absorbs energy and joint breaks if plastic capacity is exceeded. 
-Maximum forces and moments are estimated based on the plastic section modulus.
+Maximum forces and moments can be estimated based on the plastic section modulus or
+by defining maximum elastic displacement.
 Joint breaking is based on summing plastic deformation and comparing it to a
-predefined material based limit. The elastic part of deformation is modelled by employing 
+predefined material and geometry based limit. The elastic part of deformation is modelled by employing 
 modification of an existing constraints.
@@
 \paragraph{Joint breaking}
+Joint breaking is based on summing of plastic displacement or rotation.
+Process can be subdivided few phases.
+
+\begin{enumerate}
+\item Hitting maximum impulse value triggers additional processing.
+\item New neutral position is calculated so that elastic part is subtracted from current position.
+\item Plastic strain or rotation is incremented.
+\item Current plastic strain or rotation is compared to maximum value and joint is broken if 
+current value exceeds predefined maximum value. This part can be finetuned e.g. so that joint is 
+not broken during compression if connected parts are made of material which can handle large
+strains due to compressive forces. 
+\end{enumerate}
 
 \section{Implementation for Box2D}
+b2ElasticPlastic joint implementation is based on b2WeldJoint which implements rigid and flexible joints as
+described in Catto's paper~\shortcite{ecsc} on soft constraints. Main idea is that game designer should not have
+find suitable values to get desired functionality but constraint is configured by giving desired  damping ratio and frequency.
+This principle leads in many cases to quite soft constraints to keep simulation stable. 
+b2WeldJoint supports also rigid joint in which case iterative solver creates some flexibility.
 
@@
 \section*{Index of Supplemental Materials}
+
+\begin{enumerate}
+\item Box2D source with plasticity extensions 
+ \href{https://github.com/simo-11/Box2D}{https://github.com/simo-11/Box2D}
+
+\item \cbullet source with plasticity extensions 
+ \href{https://github.com/simo-11/bullet3}{https://github.com/simo-11/bullet3}
+\item two minute video demonstrating \cbullet implementation 
+ \href{https://youtu.be/_QNbv1-f_ao}{https://youtu.be/_QNbv1-f_ao}
+\item 2 MB executable used for \cbullet video {https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}{https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}
+\end{enumerate}
 

### 2017-08-04 examples/CharpyDemo/CharpyDemo.cpp

@@ -867,6 +867,7 @@ public:
 	void addUseCcd(){
 		Gwen::Controls::Label* label = addLabel("useCcd");
 		Gwen::Controls::CheckBox* gc = new Gwen::Controls::CheckBox(pPage);
+		gc->SetText("");
 		gc->SetToolTip("use continous collision detection");
 		gc->SetPos(gxi, gy);
 		gc->SetChecked(useCcd);

### 2017-08-06 pdocs/thesis/epj-jcgt.tex
Box2D implementation in jcgt article

@@
 %  Simo.Nikula@gmail.com
-\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse based Physics}
+\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse-based Physics}
 \def\lut{Lappeenranta University of Technology}
@@
 
+\paragraph{Analysis of results}
+Unexpected results is typical issue when new kind of structures are analyzed or new methods are are
+used. In gaming results do not have to match real world results but main focus of implementations
+in this work is to provide more realistic simulations.
+
 \section{Implementation for Box2D}
-b2ElasticPlastic joint implementation is based on b2WeldJoint which implements rigid and flexible joints as
+b2ElasticPlasticJoint implementation is based on b2WeldJoint which implements rigid and flexible joints as
 described in Catto's paper~\shortcite{ecsc} on soft constraints. Main idea is that game designer should not have
@@
 b2WeldJoint supports also rigid joint in which case iterative solver creates some flexibility.
+If maximum forces and moment are set to very high value b2ElasticPlasticJoint behaves like b2WeldJoint but is slighly slower.
+
+\paragraph{Changes to define maximum forces}
+
+\paragraph{Neutral position update}
+
+\paragraph{Debugging tools}
+
+\subsection{Test cases}
+
+\paragraph{Cantilever beam}
+
+\paragraph{Multifloor frame}
+
+\begin{enumerate}
+\item H
+\item New neutral position is calculated so that elastic part is subtracted from current position.
+\item Plastic strain or rotation is incremented.
+\item Current plastic strain or rotation is compared to maximum value and joint is broken if 
+current value exceeds predefined maximum value. This part can be finetuned e.g. so that joint is 
+not broken during compression if connected parts are made of material which can handle large
+strains due to compressive forces. 
+\end{enumerate}
 

### 2017-08-07 pdocs/thesis/epj-jcgt.tex
Box2D implementation in jcgt article

@@
 
-\paragraph{Changes to define maximum forces}
+\paragraph{Changes to processing} 
+In addition to allowing definition of maximum forces and moment during initialization 
+b2ElasticPlasticJoint allows definition of maximum moment by giving maximum elastic rotation.
 
-\paragraph{Neutral position update}
+For each step maximum rotational impulse and linear impulses are calculated using step length and 
+current orientation of the joint.
+
+During solution phase impulse is clamped so that it does not too big.
+If clamping is activated additional processing is done at the beginnig of next step to update neutral position and
+increment plasticity values.
 
 \paragraph{Debugging tools} 
+Box2D testbed offers many useful tools and only few additions were made for testing of plasticity.
+\begin{enumerate}
+\item Tunable visualization of joint forces
+\item Numerical output of joint forces as absolute values and as percentage of maximum forces
+\item Numerical output of usage of plastic capacity
+\item Selection of joints for display of numerical values
+\end{enumerate}
 

### 2017-09-30 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 %  Simo.Nikula@gmail.com
-\def\mytitle{Breakable Elastic-plastic Constraint for Rigid Body Simulation with Impulse-based Physics}
+\def\mytitle{Breakable Elastic-plastic Constraint for Impulse-based Physics Engines}
 \def\lut{Lappeenranta University of Technology}
@@
 \begin{document}
-\newcommand{\cbullet}[0]{Bullet Physics}
 \title{Breakable Elastic-plastic Constraint for\\Rigid Body Simulation}
@@
   \label{fig:teaser}
-  \caption{Adding ductile breakable constraint to popular open source physics engines \cbullet\ and Box2D allows more realistic simulation of many scenarios.}
+  \caption{Adding ductile breakable constraint to popular open source physics engine Box2D allows more realistic simulation of many scenarios.}
 }
@@
 
-Sample programs with source code are made available to allow developers already using \cbullet\
-and Box2D to add plasticity into their simulations.
+Sample program with source code are made available to allow developers already using 
+Box2D to add plasticity into their simulations.
 \end{abstract}
@@
 
-More realistic simulation of ductile destructable bodies is possible but in many cases it would require selecting of new physics engine.
+More realistic simulation of ductile destructable bodies is possible but in many cases it would require 
+selecting of new physics engine.
 
@@
 Maximum forces and moments can be estimated based on the plastic section modulus or
-by defining maximum elastic displacement.
+by defining maximum elastic displacement or based on desired capacity.
 Joint breaking is based on summing plastic deformation and comparing it to a
@@
 modification of an existing constraints.
+
+This paper follows idea of
+allowing game designers to get desired behaviour without diving into details of structural engineering.
+Idea was presented by Catto, \cite{ecsc}. Basic idea is that game designer selects spring frequency so that
+integration is stable. Downside is that constraints will be quite soft. In out method game designers get new option
+to use rigid-plastic constraints.
 
@@
 
-Maximum moments acting around the axis perpendicular to $\vec{r}_{anc}^{\,i} $
-are integrals of the perpdendicular distance 
-and yield stress $f_y$ as given for the moment around the $x$-axis 
-and moment around the $z$-axis, respectively:
-
-\begin{equation} \label{eq:Mx}
-M_{max}^x= \int_A z f_y,
-\end{equation}
+Maximum moment is integral of the perpdendicular distance 
+and yield stress $f_y$ as given for the moment around  the $z$-axis:
 
@@
 M_{max}^z= \int_A x f_y.
-\end{equation}
-
-Maximum moment around $\vec{r}_{anc}^{\,i} $
-is an integral of distance $d$ from the joint point
-and shear yield stress $\tau_y$ as:
-
-\begin{equation} \label{eq:My}
-M_{max}^y= \int_A d \tau_y.
 \end{equation}
@@
 maximum bending moment in direction of $h$& $0.25\, b\, h^2 \, f_y$  \\ \hline
-maximum bending moment in direction of $b$ & $0.25\, b^2\, h\, f_y$  \\ \hline
-maximum torque & $ \approx 0.19\, b\, h\, \frac{b\, + h}{2} f_y$  \\ \hline
 \end{tabular}
@@
 \end {table}
-
-For torque there is a closed form solution only for
-circular cross sections.
-Given approximation is 
-best suited for cases where $b$ and $h$ are similar.
-Better approximation for any given $b$ and $h$ can be obtained 
-by integrating distance from the center of the joint over cross section and
-multiplying it with the yield shear stress e.g. using Octave, \cite{octave}.
-An example of calculation of the maximum moment  around $\vec{r}_{anc}^{\,i} $
-is shown in Listing \ref{lst:octave-mp}.
-
-\lstset{language=octave}
-\begin{lstlisting}[caption={Calculation of maximum moment  around $\vec{r}_{anc}^{\,i} $ using Octave.}, 
-label={lst:octave-mp}, float]
-b=0.01; h=0.01; fy=200e6;
-wpy=fy/2*dblquad(@(x,z)...
- sqrt(x.*x+z.*z),-b/2,b/2,-h/2,h/2)
-38.2
-\end{lstlisting}
 
@@
 
-\begin{enumerate}
-\item H
-\item New neutral position is calculated so that elastic part is subtracted from current position.
-\item Plastic strain or rotation is incremented.
-\item Current plastic strain or rotation is compared to maximum value and joint is broken if 
-current value exceeds predefined maximum value. This part can be finetuned e.g. so that joint is 
-not broken during compression if connected parts are made of material which can handle large
-strains due to compressive forces. 
-\end{enumerate}
-
-\section{Implementation for \cbullet}
+\paragraph{Car}
 
@@
 
-\item \cbullet source with plasticity extensions 
- \href{https://github.com/simo-11/bullet3}{https://github.com/simo-11/bullet3}
-\item two minute video demonstrating \cbullet implementation 
- \href{https://youtu.be/_QNbv1-f_ao}{https://youtu.be/_QNbv1-f_ao}
-\item 2 MB executable used for \cbullet video {https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}{https://github.com/simo-11/bullet3/blob/master/pbin/App_PlasticityExampleBrowser_vs2013.exe?raw=true}
 \end{enumerate}

### 2017-10-01 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 Bending moment - normal force interactions and multiaxial stress are not taken into account. 
-Basically these limitations just state that results from these simulations cannot not be used
-for structural analysis.
+Simulation of many bodies connected by constraints
 
@@
 
-\paragraph{Plastic capacities}
+\paragraph{Plastic capacities based on dimensions and material}
 In this work, plasticity is handled by defining maximum forces using plastic capacities, which are defined below.
@@
 
+\paragraph{Plastic capacities based on desired behaviour}
+Capacities can also be defined directly based on desired scenarios.
+E.g. joints for bridge can designed so that maximum shear force is larger than
+weight of heaviest vehicle so that all vehicles can drive to bridge. Maximum 
+bending moment could be set so that joint will start to rotate when heaviest vehicle is
+near the middle of bridge.
+
+For elastic-plastic cases maximum capacity can also be defined by setting maximum elastic 
+displacement or rotation.
+
 \paragraph{Joint breaking}
@@
 \end{enumerate}
-
-\paragraph{Analysis of results}
-Unexpected results is typical issue when new kind of structures are analyzed or new methods are are
-used. In gaming results do not have to match real world results but main focus of implementations
-in this work is to provide more realistic simulations.
 
@@
 
-\subsection{Test cases}
+\subsection{Examples}
 

### 2017-10-08 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 \teaser{
-  \includegraphics[width=5cm]{figs/demolisher-done}
+  \includegraphics[width=5cm]{figs/CarAtRamp}
   \label{fig:teaser}
-  \caption{Adding ductile breakable constraint to popular open source physics engine Box2D allows more realistic simulation of many scenarios.}
+  \caption{Adding ductile breakable constraint to popular open source physics engine Box2D allows more realistic simulation of many scenarios. In picture above ramp is under elastic loading.}
 }
@@
 Procedure to provide realistic parameters for constraints is described.
-Procedure is based on body dimensions and material parameters.
+Procedure may be based on body dimensions and material parameters or desired bahviour.
 
@@
 Idea was presented by Catto, \cite{ecsc}. Basic idea is that game designer selects spring frequency so that
-integration is stable. Downside is that constraints will be quite soft. In out method game designers get new option
+integration is stable. Downside is that constraints will be quite soft. In our method game designers get new option
 to use rigid-plastic constraints.
@@
 
-During solution phase impulse is clamped so that it does not too big.
+During solution phase impulse is clamped so that it does exceed predefined limit.
 If clamping is activated additional processing is done at the beginnig of next step to update neutral position and
@@
 \paragraph{Car} 
+Dynamic bodies are car (680 kg), wheels (50 kg for each axel), and ramp (2160 kg). 
+Ramp is joined to fixed part with b2ElasticPlasticJoint having frequency of 1 Hz.
+Maximum elastic rotation is set to 0.12 and maximum plastic rotation to 0.4.
+With this configuration own weight or ramp uses about 13 \% of plastic capacity.
+After first jump about 50 \% of  plastic capacity is used. Each additional jump uses about 
+7 \% of plastic capacity.
 

### 2017-10-09 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 
+\paragraph{Joint breaking in one simulation step} High velocity collisions may cause joints to break within one step.
+In such cases joints can cause too large impulse. Making simulation step smaller is not desired solution.
+In this paper following procedure is suggested.
+
+\begin{enumerate}
+\item 
+\end{enumerate}
+
 \section{Implementation for Box2D}
@@
 
-\subsection{Examples}
+\section{Examples}
 
-\paragraph{Cantilever beam}
+\subsection{Cantilever beam}
 
-\paragraph{Multifloor frame}
+\subsection{Multifloor frame}
 
-\paragraph{Car} 
+\subsection{Car} 
 Dynamic bodies are car (680 kg), wheels (50 kg for each axel), and ramp (2160 kg). 

### 2017-10-10 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 \paragraph{Joint breaking in one simulation step} High velocity collisions may cause joints to break within one step.
-In such cases joints can cause too large impulse. Making simulation step smaller is not desired solution.
+In such cases joints can cause too large impulse as shown in  Figure~\ref{fig:tippingCar}.
+
+\begin{figure}[htb]
+  \centering
+   \includegraphics[height=5cm]{figs/tippingCar1.png}
+   \includegraphics[height=5cm]{figs/tippingCar2.png}
+   \includegraphics[height=5cm]{figs/tippingCar3.png}
+   \caption{\label{fig:tippingCar}
+     Large impulse turns car over.}
+\end{figure}
+Making simulation step smaller is not desired solution.
 In this paper following procedure is suggested.

### 2017-10-29 pdocs/thesis/epj-jcgt.tex
work on jcgt article

@@
 This limitation is not introduced by this approach but is build in feature of impulse-based simulation.
+We discuss few methods to maximize realisticy of simulation of collisions but this issue will need further work.
 Bending moment - normal force interactions and multiaxial stress are not taken into account. 
-Simulation of many bodies connected by constraints
+Simulation of many bodies connected by constraints has convergence issues which will also need further work.
 
@@
 Making simulation step smaller is not desired solution.
-In this paper following procedure is suggested.
+In this paper following procedure suggested.
 
 \begin{enumerate}
-\item 
+\item At start of each step predict outcome of current step for each joint using current velocities and
+maximum and current strains and rotations.
+\item If current impulse is about to break constraint within current step scale down maximum impulse
+so that joint does not exceed planned capacity. 
 \end{enumerate}
+
+\paragraph{Impulse initialization}
+Joint impulses should be initialized using suitable methods to minimize errors.
 
@@
 described in Catto's paper~\shortcite{ecsc} on soft constraints. Main idea is that game designer should not have
-find suitable values to get desired functionality but constraint is configured by giving desired  damping ratio and frequency.
+to find suitable values to get desired functionality but constraint is configured by giving desired  damping ratio and frequency.
 This principle leads in many cases to quite soft constraints to keep simulation stable. 
@@
 Maximum elastic rotation is set to 0.12 and maximum plastic rotation to 0.4.
-With this configuration own weight or ramp uses about 13 \% of plastic capacity.
+With this configuration own weight of ramp uses about 13 \% of plastic capacity during initialization phase.
 After first jump about 50 \% of  plastic capacity is used. Each additional jump uses about 
